### 2010-02-26 papers/user2010/abstract.tex
initial version of useR abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@293 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[11pt]{article}
\usepackage{url}
\usepackage{vmargin}
\setpapersize{USletter}
\setmarginsrb{1in}{1in}{1in}{1in}{0pt}{0mm}{0pt}{0mm}
\usepackage{charter}

\newcommand{\proglang}[1]{\textsf{#1}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

\author{
Romain Fran\c{c}ois\\ {\small \url{romain@r-enthusiasts.com} } \and
Dirk Eddelbuettel\\ {\small \url{edd@debian.org} }
}

\title{\pkg{RProtoBuf}: Protocol Buffers for R }
\date{Abstract submitted for \textsl{useR! 2010}}

\begin{document}

\maketitle
\thispagestyle{empty}
\begin{abstract}
  \addtolength{\parskip}{\baselineskip} 	% add a little vertical space
  \noindent % no ident for first paragraph
  Protocol buffers are a flexible, efficient, 
  automated mechanism for serializing structured data---think XML, but 
  smaller, faster, and simpler. 
  You define how you want your data to be structured once, then you can use 
  special generated source code to easily write and read your structured 
  data to and from a variety of data streams and using a 
  variety of officially supported languages---
  \proglang{Java}, \proglang{C++}, or \proglang{Python}---or 
  third party implementations for languages such as \proglang{C\#}, 
  \proglang{Perl}, \proglang{Ruby}, \proglang{Visual Basic}, and now 
  \proglang{R} via the \pkg{RProtoBuf} package.

  The \pkg{RProtoBuf} package implements R bindings to the C++
  protobuf library from Google. It uses features of the protocol 
  buffer library to support creation, manipulation, parsing and serialization
  of protocol buffers messages. The \pkg{RProtoBuf} package uses S4 classes
  and external pointers to expose objects that look and feel like standard 
  R lists, but constrained by the language agnostic definition of the 
  message type.

  \noindent \textbf{Keywords:}  
  \proglang{R}, \proglang{C++}, Serialization, Data Interchange
\end{abstract}
\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2010-02-26 papers/user2010/abstract.tex
another iteration on the abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@295 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \title{\pkg{RProtoBuf}: Protocol Buffers for R }
-\date{Abstract submitted for \textsl{useR! 2010}}
+\date{Submitted to \textsl{useR! 2010}}
 
@@
   \noindent % no ident for first paragraph
-  Protocol buffers are a flexible, efficient, 
-  automated mechanism for serializing structured data---think XML, but 
-  smaller, faster, and simpler. 
-  You define how you want your data to be structured once, then you can use 
-  special generated source code to easily write and read your structured 
-  data to and from a variety of data streams and using a 
-  variety of officially supported languages---
-  \proglang{Java}, \proglang{C++}, or \proglang{Python}---or 
-  third party implementations for languages such as \proglang{C\#}, 
-  \proglang{Perl}, \proglang{Ruby}, \proglang{Visual Basic}, and now 
-  \proglang{R} via the \pkg{RProtoBuf} package.
+  %
+  Protocol buffers are a flexible, efficient, automated mechanism for
+  serializing structured data---think XML, but smaller, faster, and simpler.
+  Users define how they want the data to be structured once in a
+  \texttt{proto} file and then use special generated source code to easily
+  write and read structured data to and from a variety of data streams and
+  using a variety of officially supported languages--- \proglang{Java},
+  \proglang{C++}, or \proglang{Python}---or third party implementations for
+  languages such as \proglang{C\#}, \proglang{Perl}, \proglang{Ruby},
+  \proglang{Haskell}, and now \proglang{R} via the \pkg{RProtoBuf} package.
 
-  The \pkg{RProtoBuf} package implements R bindings to the C++
-  protobuf library from Google. It uses features of the protocol 
-  buffer library to support creation, manipulation, parsing and serialization
-  of protocol buffers messages. The \pkg{RProtoBuf} package uses S4 classes
-  and external pointers to expose objects that look and feel like standard 
-  R lists, but constrained by the language agnostic definition of the 
-  message type.
+  The \pkg{RProtoBuf} package implements \proglang{R} bindings to the
+  \proglang{C++} protobuf library from Google. It uses features of the
+  protocol buffer library to support creation, manipulation, parsing and
+  serialization of protocol buffers messages. Taking advantage of facilities
+  in the \pkg{Rcpp} package, \pkg{RProtoBuf} uses S4 classes and external
+  pointers to expose objects that look and feel like standard \proglang{R}
+  lists, but constrained by the language-agnostic definition of the message
+  type.
+  % [Dirk] what does that last half-sentence mean? constrained as in conforming
+
+  % [Dirk] ok to add 'forward-looking' statements?
+  As the protocal buffers library does not offer any built-in support
+  for networked access to protocol buffer streams, we intend to take
+  advantage on ongoing changes to the \proglang{R} system to expose a native
+  \proglang{R} server. This is work-in-progress that we hope to report on at
+  the conference.
 
   \noindent \textbf{Keywords:}  
-  \proglang{R}, \proglang{C++}, Serialization, Data Interchange
+  \proglang{R}, \proglang{C++}, Serialization, Data Interchange, Data Formats
 \end{abstract}

### 2010-02-27 papers/user2010/abstract.tex
ping/pong
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@296 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   pointers to expose objects that look and feel like standard \proglang{R}
-  lists, but constrained by the language-agnostic definition of the message
+  lists, but conforming to the language-agnostic definition of the message
   type.
   % [Dirk] what does that last half-sentence mean? constrained as in conforming
+  % [Romain] yes. updated. I also wanted to say that the actual memory of the object
+  %          is managed by the C++ library and not R, but I can't manage to spell it.
 
   % [Dirk] ok to add 'forward-looking' statements?
+  % [Romain] fine by me.
   As the protocal buffers library does not offer any built-in support

### 2010-02-27 papers/user2010/abstract.tex
one more ping-pong (and I now enabled commit messages at R-Forge so we may finally get email)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@297 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2010-02-26 papers/user2010/abstract.tex
@@
   pointers to expose objects that look and feel like standard \proglang{R}
-  lists, but constrained by the language-agnostic definition of the message
-  type.
+  lists, yet are managed by the underlying C++ library. These objects also
+  conform to the language-agnostic definition of the message type allowing
+  access to their content from other supported languages.
   % [Dirk] what does that last half-sentence mean? constrained as in conforming
+  % [Romain] yes. updated. I also wanted to say that the actual memory of the object
+  %          is managed by the C++ library and not R, but I can't manage to
+  %          spell it.  
+  % [Dirk] Ok, here is a try.  May be getting to detailed.
 
-  % [Dirk] ok to add 'forward-looking' statements?
   As the protocal buffers library does not offer any built-in support
@@
   advantage on ongoing changes to the \proglang{R} system to expose a native
-  \proglang{R} server. This is work-in-progress that we hope to report on at
-  the conference.
+  \proglang{R} server. This is work-in-progress, but we hope to be able to
+  report on this at the conference.
 

### 2013-12-16 papers/rjournal/eddelbuettel-francois-stokely.tex
Add boilerplate for an R Journal article.  Hadley's ggmap one from the last issue was nearly 20 pages, so I think we'll be fine length wise.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@553 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

% !TeX root = RJwrapper.tex
\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\author{by Dirk Eddelbuettel, Romain Fran\c{c}ois, and Murray Stokely}

\maketitle

\abstract{Modern data collection and analysis pipelines often involve
 a sophisticated mix of applications written in general purpose and
 specialized programming languages.  Protocol Buffers are a popular
 method of serializing structured data between applications.  The
 \textbf{RProtoBuf} package provides a complete interface to this
 library.
TODO keep it less than 150 words.
}

\section{Introduction}

Comparison with what people start with in R : CSV
comparison with what is only slightly better: JSON

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

Protocol buffers are a language-neutral, platform-neutral, extensible
way of serializing structured data for use in communications
protocols, data storage, and more.

Protocol Buffers offer key features such as an efficient data interchange
format that is both language- and operating system-agnostic yet uses a
lightweight and highly performant encoding, object serialization and
de-serialization as well data and configuration management. Protocol
buffers are also forward compatible: updates to the \texttt{proto}
files do not break programs built against the previous specification.

While benchmarks are not available, Google states on the project page that in
comparison to XML, protocol buffers are at the same time \textsl{simpler},
between three to ten times \textsl{smaller}, between twenty and one hundred
times \textsl{faster}, as well as less ambiguous and easier to program.

The protocol buffers code is released under an open-source (BSD) license. The
protocol buffer project (\url{http://code.google.com/p/protobuf/})
contains a C++ library and a set of runtime libraries and compilers for
C++, Java and Python.

With these languages, the workflow follows standard practice of so-called
Interface Description Languages (IDL)
(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
  on IDL}).  This consists of compiling a protocol buffer description file
(ending in \texttt{.proto}) into language specific classes that can be used
to create, read, write and manipulate protocol buffer messages. In other
words, given the 'proto' description file, code is automatically generated
for the chosen target language(s). The project page contains a tutorial for
each of these officially supported languages:
\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}

Besides the officially supported C++, Java and Python implementations, several projects have been
created to support protocol buffers for many languages. The list of known
languages to support protocol buffers is compiled as part of the
project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}

The protocol buffer project page contains a comprehensive
description of the language: \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}

%This section may contain a figure such as Figure~\ref{figure:rlogo}.
%
%\begin{figure}[htbp]
%  \centering
%  \includegraphics{Rlogo}
%  \caption{The logo of R.}
%  \label{figure:rlogo}
%\end{figure}

\section{Dynamic use: Protocol Buffers and R}

This section describes how to use the R API to create and manipulate
protocol buffer messages in R, and how to read and write the
binary \emph{payload} of the messages to files and arbitrary binary
R connections.

\subsection{Importing proto files}

In contrast to the other languages (Java, C++, Python) that are officially
supported by Google, the implementation used by the \texttt{RProtoBuf}
package does not rely on the \texttt{protoc} compiler (with the exception of
the two functions discussed in the previous section). This means that no
initial step of statically compiling the proto file into C++ code that is
then accessed by R code is necessary. Instead, \texttt{proto} files are
parsed and processed \textsl{at runtime} by the protobuf C++ library---which
is much more appropriate for a dynamic language.

The \texttt{readProtoFiles} function allows importing \texttt{proto}
files in several ways.

% Example code snippet.
% TODO(mstokely): Remove this.
\begin{example}
  x <- 1:10
  result <- myFunction(x)
\end{example}

\section{Related work on IDLs (greatly expanded from what you have)}

\section{Design tradeoffs: reflection vs proto compiler (not addressed
  at all in current vignettes)}

\subsection{Performance considerations}

TODO RProtoBuf is quite flexible and easy to use for interactive
analysis, but it is not designed for certain classes of operations one
might like to do with protocol buffers.  For example, taking a list of
10,000 protocol buffers, extracting a named field from each one, and
computing a aggregate statistics on those values would be extremely
slow with RProtoBuf, and while this is a useful class of operations,
it is outside of the scope of RProtoBuf.  We should be very clear
about this to clarify the goals and strengths of RProtoBuf and its
reflection and object mapping.

\subsection{Serialization comparison}

TODO comparison of protobuf serialization sizes/times for various vectors.  Compared to R's native serialization.  Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.

\section{Basic usage example - tutorial.Person}

\section{Application: distributed Data Collection with MapReduce}

We could describe a common MapReduce pattern of having the MR written
in another language output protocol buffers that are later pulled into
R.  There is some text about this in section 2 of
http://cran.r-project.org/web/packages/HistogramTools/vignettes/HistogramTools.pdf

\section{Application: Sending/receiving Interaction With Servers}

\section{Summary}

This file is only a basic article template. For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions for Authors}.

\bibliography{eddelbuettel-francois-stokely}

\address{Author One\\
  Affiliation\\
  Address\\
  Country}
\email{author1@work}

\address{Author Two\\
  Affiliation\\
  Address\\
  Country}
\email{author2@work}

\address{Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94043\\
  USA}
\email{mstokely@google.com}

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.tex
minor edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@555 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  specialized programming languages.  Protocol Buffers are a popular
- method of serializing structured data between applications.  The
+ method of serializing structured data between applications---while remaining
+ indendent of programming languages or operating system.  The
  \textbf{RProtoBuf} package provides a complete interface to this
  library.
-TODO keep it less than 150 words.
+ %TODO(ms) keep it less than 150 words.
 }
+
+%TODO(de) 'protocol buffers' or 'Protocol Buffers' ?
 
@@
 Comparison with what people start with in R : CSV
-comparison with what is only slightly better: JSON
+
+Comparison with what is only slightly better: JSON
+
+Maybe mention related, competing approaches such as BSON, Thrift, msgpack,
+though we get carried away.
 
@@
 
-\address{Author One\\
-  Affiliation\\
-  Address\\
-  Country}
-\email{author1@work}
+\address{Dirk Eddelbuettel\\
+  Debian and R Projects\\
+  711 Monroe Avenue, River Forest, IL 60305\\
+  USA}
+\email{edd@debian.org}
 

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.tex
Improve/flesh-out introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@556 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  indendent of programming languages or operating system.  The
- \textbf{RProtoBuf} package provides a complete interface to this
+ \CRANpkg{RProtoBuf} package provides a complete interface to this
  library.
@@
 
-Comparison with what people start with in R : CSV
+Modern data collection and analysis pipelines often involve a
+sophisticated mix of applications used for collecting, cleaning,
+analyzing, processing, and presenting data.  Each stage of the data
+analysis pipeline may involve storing intermediate results in a
+file or sending them over the network.  Programming langauges such as
+Java, Ruby, Python, and R include built-in serialization support, but
+these formats are tied to the specific programming language in use.
+CSV files can be read and written by many applications and so are
+often used for exporting tabular data.  However, CSV files have a
+number of disadvantages, such as a limitation of exporting only
+tabular datasets, lack of type-safety, inefficient text representation
+and parsing, and abiguities in the format involving special
+characters.  JSON is another widely supported format used mostly on
+the web that removes many of these disadvantages, but it too suffers
+from being too slow to parse and also does not provide strong typing
+between integers and floating point.  Large numbers of JSON messages
+would also be required to duplicate the field names with each message.
 
-Comparison with what is only slightly better: JSON
+This article describes the basics of Google's Protocol Buffers through
+an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
+basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
+several common use cases for protocol buffers in data analysis.
 
-Maybe mention related, competing approaches such as BSON, Thrift, msgpack,
-though we get carried away.
+\section{Protocol Bfufers}
+
+Once the data serialization needs get complex enough, application
+developers typically benefit from the use of an \emph{interface
+description language}, or \emph{IDL}.  IDLs like Google's Protocol
+Buffers and Apache Thrift provide a compact well-documented schema for
+cross-langauge data structures as well efficient binary interchange
+formats.  The schema can be used to generate model classes for
+statically typed programming languages such as C++ and Java, or can be
+used with reflection for dynamically typed programming languages.
+Since the schema is provided separately from the encoded data, the
+data can be efficiently encoded to minimize storage costs of the
+stored data when compared with simple ``schema-less'' binary
+interchange formats like BSON.
+
+%BSON, msgpack, Thrift, and Protocol Buffers take this latter approach,
+%with the
+
+% There are references comparing these we should use here.
+
+TODO Also mention Thrift and msgpack and the references comparing some
+of these tradeoffs.
 

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.tex
Remove latex temporary file from revision control. Add more citations and text to intro.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@557 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  method of serializing structured data between applications---while remaining
- indendent of programming languages or operating system.  The
+ independent of programming languages or operating system.  The
  \CRANpkg{RProtoBuf} package provides a complete interface to this
@@
 
-Modern data collection and analysis pipelines often involve a
-sophisticated mix of applications used for collecting, cleaning,
-analyzing, processing, and presenting data.  Each stage of the data
+Modern data collection and analysis pipelines are increasingly being
+built using collections of components to better manage software
+complexity through reusability, modularity, and fault
+isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  Different
+programming languages are often used for the different phases of data
+analysis -- collection, cleaning, analysis, post-processing, and
+presentation in order to take advantage of the unique combination of 
+performance, speed of development, and library support offered by
+different environments.  Each stage of the data
 analysis pipeline may involve storing intermediate results in a
@@
 
-\section{Protocol Bfufers}
+\section{Protocol Buffers}
 
@@
 \section{Dynamic use: Protocol Buffers and R}
+
+TODO(ms): random citations to work in:
+
+We make use of Object Tables \citep{RObjectTables} for lookup.
+Many sources compare data serialization formats and show protocol
+buffers very favorably to the alternatives, such
+as \citep{Sumaray:2012:CDS:2184751.2184810}
 

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Move the TeX file over to Rnw and update the makefile to run Sweave so we can more quickly insert example usage sections in the document.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@560 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.tex
@@
 would also be required to duplicate the field names with each message.
+
+TODO(ms): Also work in reference to Split-Apply-Combine pattern for
+data analysis \citep{wickham2011split}, since that is a great pattern
+but it seems overly optimistic to expect all of those phases to always
+be done in the same language.
 
@@
 TODO(ms): random citations to work in:
-
+q
 We make use of Object Tables \citep{RObjectTables} for lookup.

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Move over some sections from the introduction vignette that can be included here.  I've started moving the pieces over piecemeal because the intro vignette is too much like a reference manual and needs more of a narrative which I'm attempting to provide here.
Specifically, section 3 "Classes, Methods, and Pseudo-Methods" of the
intro vignette is 20 pages on its own of reference material that would
not be as appropriate for a writeup in R Journal or JSS or something.
It is good content, but we need to replace much of that with very
concise tables or similar.

The number of classes and methods i

@@
 
+\emph{TODO(mstokely): Remove this example code snippet}
+
+\begin{example}
+  x <- 1:10
+  result <- myFunction(x)
+\end{example}
+
 \subsection{Importing proto files}
@@
 
-% Example code snippet.
-% TODO(mstokely): Remove this.
-\begin{example}
-  x <- 1:10
-  result <- myFunction(x)
-\end{example}
+<<>>=
+library(RProtoBuf)
+args( readProtoFiles )
+@
+
+Using the \texttt{file} argument, one can specify one or several file
+paths that ought to be proto files.
+
+<<>>=
+proto.dir <- system.file( "proto", package = "RProtoBuf" )
+proto.file <- file.path( proto.dir, "addressbook.proto" )
+<<eval=FALSE>>=
+readProtoFiles( proto.file )
+@
+
+With the \texttt{dir} argument, which is
+ignored if the \texttt{file} is supplied, all files matching the
+\texttt{.proto} extension will be imported.
+
+<<>>=
+dir( proto.dir, pattern = "\\.proto$", full.names = TRUE )
+<<eval=FALSE>>=
+readProtoFiles( dir = proto.dir )
+@
+
+Finally, with the
+\texttt{package} argument (ignored if \texttt{file} or
+\texttt{dir} is supplied), the function will import all \texttt{.proto}
+files that are located in the \texttt{proto} sub-directory of the given
+package. A typical use for this argument is in the \texttt{.onLoad}
+function of a package.
+
+<<eval=FALSE>>=
+readProtoFiles( package = "RProtoBuf" )
+@
+
+Once the proto files are imported, all message descriptors are
+are available in the R search path in the \texttt{RProtoBuf:DescriptorPool}
+special environment. The underlying mechanism used here is
+described in more detail in section~\ref{sec-lookup}.
+
+<<>>=
+ls( "RProtoBuf:DescriptorPool" )
+@
+
+\subsection{Creating a message}
+
+The objects contained in the special environment are
+descriptors for their associated message types. Descriptors will be
+discussed in detail in another part of this document, but for the
+purpose of this section, descriptors are just used with the \texttt{new}
+function to create messages.
+
+<<>>=
+p <- new( tutorial.Person, name = "Romain", id = 1 )
+@
+
+\subsection{Access and modify fields of a message}
+
+Once the message is created, its fields can be queried
+and modified using the dollar operator of R, making protocol
+buffer messages seem like lists.
+
+<<>>=
+p$name
+p$id
+p$email <- "francoisromain@free.fr"
+@
+
+However, as opposed to R lists, no partial matching is performed
+and the name must be given entirely.
+
+The \verb|[[| operator can also be used to query and set fields
+of a mesages, supplying either their name or their tag number :
+
+<<>>=
+p[["name"]] <- "Romain Francois"
+p[[ 2 ]] <- 3
+p[[ "email" ]]
+@
+
+Protocol buffers include a 64-bit integer type, but R lacks native
+64-bit integer support.  A workaround is available and described in
+Section~\ref{sec:int64} for working with large integer values.
+
+% TODO(mstokely): Document extensions here.
+% There are none in addressbook.proto though.
+
+\subsection{Display messages}
+
+Protocol buffer messages and descriptors implement \texttt{show}
+methods that provide basic information about the message :
+
+<<>>=
+p
+@
+
+For additional information, such as for debugging purposes,
+the \texttt{as.character} method provides a more complete ASCII
+representation of the contents of a message.
+
+<<>>=
+writeLines( as.character( p ) )
+@
+
+\subsection{Serializing messages}
+
+However, the main focus of protocol buffer messages is
+efficiency. Therefore, messages are transported as a sequence
+of bytes. The \texttt{serialize} method is implemented for
+protocol buffer messages to serialize a message into the sequence of
+bytes (raw vector in R speech) that represents the message.
+
+<<>>=
+serialize( p, NULL )
+@
+
+The same method can also be used to serialize messages to files :
+
+<<>>=
+tf1 <- tempfile()
+tf1
+serialize( p, tf1 )
+readBin( tf1, raw(0), 500 )
+@
+
+Or to arbitrary binary connections:
+
+<<>>=
+tf2 <- tempfile()
+con <- file( tf2, open = "wb" )
+serialize( p, con )
+close( con )
+readBin( tf2, raw(0), 500 )
+@
+
+\texttt{serialize} can also be used in a more traditionnal
+object oriented fashion using the dollar operator :
+
+<<>>=
+# serialize to a file
+p$serialize( tf1 )
+# serialize to a binary connection
+con <- file( tf2, open = "wb" )
+p$serialize( con )
+close( con )
+@
+
+\subsection{Parsing messages}
+
+The \texttt{RProtoBuf} package defines the \texttt{read}
+function to read messages from files, raw vector (the message payload)
+and arbitrary binary connections.
+
+<<>>=
+args( read )
+@
+
+The binary representation of the message (often called the payload)
+does not contain information that can be used to dynamically
+infer the message type, so we have to provide this information
+to the \texttt{read} function in the form of a descriptor :
+
+<<>>=
+message <- read( tutorial.Person, tf1 )
+writeLines( as.character( message ) )
+@
+
+The \texttt{input} argument of \texttt{read} can also be a binary
+readable R connection, such as a binary file connection:
+
+<<>>=
+con <- file( tf2, open = "rb" )
+message <- read( tutorial.Person, con )
+close( con )
+writeLines( as.character( message ) )
+@
+
+Finally, the payload of the message can be used :
+
+<<>>=
+# reading the raw vector payload of the message
+payload <- readBin( tf1, raw(0), 5000 )
+message <- read( tutorial.Person, payload )
+@
+
+\texttt{read} can also be used as a pseudo method of the descriptor
+object :
+
+<<>>=
+# reading from a file
+message <- tutorial.Person$read( tf1 )
+# reading from a binary connection
+con <- file( tf2, open = "rb" )
+message <- tutorial.Person$read( con )
+close( con )
+# read from the payload
+message <- tutorial.Person$read( payload )
+@
 
@@
 
+\section{Descriptor lookup}
+\label{sec-lookup}
+
+The \texttt{RProtoBuf} package uses the user defined tables framework
+that is defined as part of the \texttt{RObjectTables} package available
+from the OmegaHat project.
+
+The feature allows \texttt{RProtoBuf} to install the
+special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
+The environment is special in that, instead of being associated with a
+static hash table, it is dynamically queried by R as part of R's usual
+variable lookup. In other words, it means that when the R interpreter
+looks for a binding to a symbol (foo) in its search path,
+it asks to our package if it knows the binding "foo", this is then
+implemented by the \texttt{RProtoBuf} package by calling an internal
+method of the \texttt{protobuf} C++ library.
+
+\section{64-bit integer issues}
+\label{sec:int64}
+
+R does not have native 64-bit integer support.  Instead, R treats
+large integers as doubles which have limited precision.  For example,
+it loses the ability to distinguish some distinct integers:
+
+<<>>=
+2^53 == (2^53 + 1)
+@
+
+Protocol Buffers are frequently used to pass data between different
+systems, however, and most other systems these days have support for
+64-bit integers.  To work around this, RProtoBuf allows users to get
+and set 64-bit integer types by treating them as characters.
+
+<<echo=FALSE,print=FALSE>>=
+if (!exists("protobuf_unittest.TestAllTypes",
+            "RProtoBuf:DescriptorPool")) {
+    unittest.proto.file <- system.file("unitTests", "data",
+                                       "unittest.proto",
+                                       package="RProtoBuf")
+    readProtoFiles(file=unittest.proto.file)
+}
+@
+
+If we try to set an int64 field in R to double values, we lose
+precision:
+
+<<>>=
+test <- new(protobuf_unittest.TestAllTypes)
+test$repeated_int64 <- c(2^53, 2^53+1)
+length(unique(test$repeated_int64))
+@
+
+However, we can specify the values as character strings so that the
+C++ library on which RProtoBuf is based can store a true 64-bit
+integer representation of the data.
+
+<<>>=
+test$repeated_int64 <- c("9007199254740992", "9007199254740993")
+@
+
+When reading the value back into R, numeric types are returned by
+default, but when the full precision is required a character value
+will be returned if the \texttt{RProtoBuf.int64AsString} option is set
+to \texttt{TRUE}.
+
+<<>>=
+options("RProtoBuf.int64AsString" = FALSE)
+test$repeated_int64
+length(unique(test$repeated_int64))
+options("RProtoBuf.int64AsString" = TRUE)
+test$repeated_int64
+length(unique(test$repeated_int64))
+@
+
+<<echo=FALSE,print=FALSE>>=
+options("RProtoBuf.int64AsString" = FALSE)
+@
+
+\section{Other approaches}
+
+Saptarshi Guha wrote another package that deals with integration
+of protocol buffer messages with R, taking a different angle :
+serializing any R object as a message, based on a single catch-all
+\texttt{proto} file.  Saptarshi's package is available at
+\url{http://ml.stat.purdue.edu/rhipe/doc/html/ProtoBuffers.html}.
+
+Jeroen Ooms took a similar approach influenced by Saptarshi in his
+\texttt{RProtoBufUtils} package.  Unlike Saptarshi's package,
+RProtoBufUtils depends on RProtoBuf for underlying message operations.
+This package is available at
+\url{https://github.com/jeroenooms/RProtoBufUtils}.
+
+% Phillip Yelland wrote another implementation, currently proprietary,
+% that has significant speed advantages when querying fields from a
+% large number of protocol buffers, but is less user friendly for the
+% basic cases documented here.
+
 \section{Basic usage example - tutorial.Person}
@@
 
+Unlike Apache Thrift, Protocol Buffers do not include a concrete RPC
+implementation.  However, serialized protocol buffers can trivially be
+sent over TCP or integrated with a proprietary RPC system.  Combined
+with an RPC system this means that one can interactively craft request
+messages, send the serialized message to a remote server, read back a
+response, and then parse the response protocol buffer interactively.
+
 \section{Summary}
 
-This file is only a basic article template.
+Its pretty useful.  Murray to see if he can get approval to talk a
+tiny bit about how much its used at Google.
+
+%This file is only a basic article template.
 For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions for Authors}.

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Include an example protocol buffer definition early in the file and example code accessing it. As with everything here, still very rough.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@562 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\begin{example}
+package tutorial;
+message Person {
+ required string name = 1;
+ required int32 id = 2;        // Unique ID number for person.
+ optional string email = 3;
+ enum PhoneType {
+   MOBILE = 0; HOME = 1; WORK = 2;
+ }
+ message PhoneNumber {
+   required string number = 1;
+   optional PhoneType type = 2 [default = HOME];
+ }
+ repeated PhoneNumber phone = 4;
+}
+\end{example}
+
+<<echo=TRUE>>=
+library(RProtoBuf)
+person <- new(tutorial.Person, id=1, name="Romain")
+person$id
+person$name
+person$name <- "Dirk"
+cat(as.character(person))
+@
+
 %This section may contain a figure such as Figure~\ref{figure:rlogo}.
@@
 TODO(ms): random citations to work in:
-q
-We make use of Object Tables \citep{RObjectTables} for lookup.
+
 Many sources compare data serialization formats and show protocol
@@
 files in several ways.
-
-<<>>=
-library(RProtoBuf)
-args( readProtoFiles )
-@
 
@@
 that is defined as part of the \texttt{RObjectTables} package available
-from the OmegaHat project.
+from the OmegaHat project \citep{RObjectTables}.
 
@@
 Jeroen Ooms took a similar approach influenced by Saptarshi in his
-\texttt{RProtoBufUtils} package.  Unlike Saptarshi's package,
+\pkg{RProtoBufUtils} package.  Unlike Saptarshi's package,
 RProtoBufUtils depends on RProtoBuf for underlying message operations.
@@
 
-Its pretty useful.  Murray to see if he can get approval to talk a
-tiny bit about how much its used at Google.
+%Its pretty useful.  Murray to see if he can get approval to talk a
+%tiny bit about how much its used at Google.
 

### 2013-12-17 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Add some new boilerplate for sections describing the basic abstractions and basic classes of RProtoBuf.  Also, make the first example more concise by putting the example .proto file and example R session using it side by side (before we go into full detail about all the accessors in RProtoBuf, this just gives a preview)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@563 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+\noindent
+\begin{tabular}{@{}p{.40\textwidth}|p{0.5\textwidth}@{}}
+\begin{minipage}{.35\textwidth}
 \begin{example}
@@
  required string name = 1;
- required int32 id = 2;        // Unique ID number for person.
+ required int32 id = 2;
  optional string email = 3;
  enum PhoneType {
-   MOBILE = 0; HOME = 1; WORK = 2;
+   MOBILE = 0; HOME = 1;
+   WORK = 2;
  }
@@
    required string number = 1;
-   optional PhoneType type = 2 [default = HOME];
+   optional PhoneType type = 2;
  }
@@
 \end{example}
-
+\end{minipage} & \begin{minipage}{.45\textwidth}
 <<echo=TRUE>>=
 library(RProtoBuf)
-person <- new(tutorial.Person, id=1, name="Romain")
-person$id
+person <- new(tutorial.Person, id=1,
+              name="Romain")
+person
 person$name
@@
 @ 
+\end{minipage}
+\end{tabular}
 
@@
 R connections.
-
-\emph{TODO(mstokely): Remove this example code snippet}
-
-\begin{example}
-  x <- 1:10
-  result <- myFunction(x)
-\end{example}
 
@@
 
-\section{Related work on IDLs (greatly expanded from what you have)}
+\section{Basic Abstractions: Messages, Descriptors, and
+  DescriptorPools}
 
-\section{Design tradeoffs: reflection vs proto compiler (not addressed
-  at all in current vignettes)}
+The three basic abstractions of \CRANpkg{RProtoBuf} are Messages,
+which encapsulate a data structure, Descriptors, which define the
+schema used by one or more messages, and DescriptorPools, which
+provide access to descriptors.
 
-\subsection{Performance considerations}
+\section{Under the hood: S4 Classes, Methods, and Pseudo Methods}
 
-TODO RProtoBuf is quite flexible and easy to use for interactive
-analysis, but it is not designed for certain classes of operations one
-might like to do with protocol buffers.  For example, taking a list of
-10,000 protocol buffers, extracting a named field from each one, and
-computing a aggregate statistics on those values would be extremely
-slow with RProtoBuf, and while this is a useful class of operations,
-it is outside of the scope of RProtoBuf.  We should be very clear
-about this to clarify the goals and strengths of RProtoBuf and its
-reflection and object mapping.
+The \CRANpkg{RProtoBuf} package uses the S4 system to store
+information about descriptors and messages, but the information stored
+in the R object is very minimal and mainly consists of an external
+pointer to a C++ variable that is managed by the \texttt{protobuf} C++
+library.
 
-\subsection{Serialization comparison}
+Using the S4 system allows the \texttt{RProtoBuf} package to dispatch
+methods that are not generic in the S3 sense, such as \texttt{new} and
+\texttt{serialize}.
 
-TODO comparison of protobuf serialization sizes/times for various vectors.
-Compared to R's native serialization.
-Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.
+The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
+of the form \verb|method( object, arguments)| and the more traditional
+object oriented notation \verb|object$method(arguments)|.
 
-\section{Descriptor lookup}
-\label{sec-lookup}
+TODO(ms): Perhaps a table here of the different S4 classes, how many
+methods they include, whether it dynamically does dispatch on other
+strings, whether/how it is available in the search path, etc.
 
-The \texttt{RProtoBuf} package uses the user defined tables framework
-that is defined as part of the \texttt{RObjectTables} package available
-from the OmegaHat project \citep{RObjectTables}.
+\subsection{Messages}
 
-The feature allows \texttt{RProtoBuf} to install the
-special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
-The environment is special in that, instead of being associated with a
-static hash table, it is dynamically queried by R as part of R's usual
-variable lookup. In other words, it means that when the R interpreter
-looks for a binding to a symbol (foo) in its search path,
-it asks to our package if it knows the binding "foo", this is then
-implemented by the \texttt{RProtoBuf} package by calling an internal
-method of the \texttt{protobuf} C++ library.
+The \texttt{Message} S4 class represents Protocol Buffer Messages and
+is the core abstraction of \CRANpkg{RProtoBuf}.  Each \texttt{Message}
+has a \texttt{Descriptor} S4 class which defines the schema of the
+data defined in the Message, as well as a number of
+\texttt{FieldDescriptors} for the individual fields of the message.
 
-\section{64-bit integer issues}
+represented in R using the \texttt{Message}
+S4 class. The class contains the slots \texttt{pointer} and \texttt{type} as
+described on the Table~\ref{Message-class-table}.
+
+\section{Type Coercion}
+
+\subsection{Booleans}
+Bools
+Int64s.
+
+\subsection{64-bit integers}
 \label{sec:int64}
@@
 @
+
+\section{Related work on IDLs (greatly expanded from what you have)}
+
+\section{Design tradeoffs: reflection vs proto compiler (not addressed
+  at all in current vignettes)}
+
+\subsection{Performance considerations}
+
+TODO RProtoBuf is quite flexible and easy to use for interactive
+analysis, but it is not designed for certain classes of operations one
+might like to do with protocol buffers.  For example, taking a list of
+10,000 protocol buffers, extracting a named field from each one, and
+computing a aggregate statistics on those values would be extremely
+slow with RProtoBuf, and while this is a useful class of operations,
+it is outside of the scope of RProtoBuf.  We should be very clear
+about this to clarify the goals and strengths of RProtoBuf and its
+reflection and object mapping.
+
+\subsection{Serialization comparison}
+
+TODO comparison of protobuf serialization sizes/times for various vectors.
+Compared to R's native serialization.
+Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.
+
+\section{Descriptor lookup}
+\label{sec-lookup}
+
+The \texttt{RProtoBuf} package uses the user defined tables framework
+that is defined as part of the \texttt{RObjectTables} package available
+from the OmegaHat project \citep{RObjectTables}.
+
+The feature allows \texttt{RProtoBuf} to install the
+special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
+The environment is special in that, instead of being associated with a
+static hash table, it is dynamically queried by R as part of R's usual
+variable lookup. In other words, it means that when the R interpreter
+looks for a binding to a symbol (foo) in its search path,
+it asks to our package if it knows the binding "foo", this is then
+implemented by the \texttt{RProtoBuf} package by calling an internal
+method of the \texttt{protobuf} C++ library.
 

### 2013-12-18 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Improve typography of table 1, with the addressbook schema side by side with an example R session creating and manipulating a person object.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@564 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % !TeX root = RJwrapper.tex
+% We don't want a left margin for Sinput or Soutput for our table 1.
+%\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=0em}
+%\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=0em}
+%\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}
+% Setting the topsep to 0 reduces spacing from input to output and
+% improves table 1.
+\fvset{listparameters={\setlength{\topsep}{0pt}}}
+\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
+
 \title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
@@
 \noindent
-\begin{tabular}{@{}p{.40\textwidth}|p{0.5\textwidth}@{}}
+\begin{table}
+\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .015\textwidth}|@{\hskip .015\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
+\hline
+Schema : \texttt{addressbook.proto} & Example R Session\\
+\hline
 \begin{minipage}{.35\textwidth}
+\vspace{2mm}
 \begin{example}
@@
 \end{example}
+\vspace{2mm}
 \end{minipage} & \begin{minipage}{.45\textwidth}
@@
 library(RProtoBuf)
-person <- new(tutorial.Person, id=1,
-              name="Romain")
+person <- new(tutorial.Person, id=1, name="Dirk")
 person
 person$name
-person$name <- "Dirk"
+person$name <- "Romain"
 cat(as.character(person))
+serialize(person, NULL)
 @ 
-\end{minipage}
+\end{minipage} \\
+\hline
 \end{tabular}
+\caption{The schema representation from a \texttt{.proto} file for the
+  \texttt{tutorial.Person} class (left) and simple R code for creating
+  an object of this class and accessing its fields (right).}
+\label{tab:proto}
+\end{table}
 

### 2013-12-18 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Add basic tables of methods for Messages and Descriptors and improve the basic usage section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@567 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 these formats are tied to the specific programming language in use.
+% TODO(ms): and they often don't support versioning among other faults.
 CSV files can be read and written by many applications and so are
@@
 interchange formats like BSON.
+
+% TODO(ms) Also talk about versioning and why its useful.
 
@@
 
-The protocol buffer project page contains a comprehensive
-description of the language: \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}
+\section{Basic Usage: Messages and Descriptors}
+
+This section describes how to use the R API to create and manipulate
+protocol buffer messages in R, and how to read and write the
+binary \emph{payload} of the messages to files and arbitrary binary
+R connections.
+
+The two fundamental building blocks of Protocol Buffers are Messages
+and Descriptors.  Messages provide a common abstract encapsulation of
+structured data fields of the type specified in a Message Descriptor.
+Message Descriptors are defined in \texttt{.proto} files and define a
+schema for a particular named class of messages.  This separation
+between schema and the message objects is in contrast to
+more verbose formats like JSON, and when combined with the efficient
+binary representation of any Message object explains a large part of
+the performance and storage-space advantage offered by Protocol
+Buffers. TODO(ms): we already said some of this above.  clean up.
+
+Table~\ref{tab:proto} shows an example \texttt{.proto} file which
+defines the \texttt{tutorial.Person} type.  The R code in the right
+column shows an example of creating a new message of this type and
+populating its fields.
+
+% lifted from protobuf page:
+%With Protocol Buffers you define how you want your data to be
+%structured once, and then you can read or write structured data to and
+%from a variety of data streams using a variety of different
+%languages.  The definition
 
@@
 
-\section{Dynamic use: Protocol Buffers and R}
+\subsection{Importing Message Descriptors from \texttt{.proto} files}
 
-TODO(ms): random citations to work in:
+%The three basic abstractions of \CRANpkg{RProtoBuf} are Messages,
+%which encapsulate a data structure, Descriptors, which define the
+%schema used by one or more messages, and DescriptorPools, which
+%provide access to descriptors.
 
-Many sources compare data serialization formats and show protocol
-buffers very favorably to the alternatives, such
-as \citep{Sumaray:2012:CDS:2184751.2184810}
+Before we can create a new Protocol Buffer Message or parse a
+serialized stream of bytes as a Message, we must read in the message
+type specification from a \texttt{.proto} file.
 
-This section describes how to use the R API to create and manipulate
-protocol buffer messages in R, and how to read and write the
-binary \emph{payload} of the messages to files and arbitrary binary
-R connections.
+New \texttt{.proto} files are imported with the \code{readProtoFiles}
+function, which can import a single file, all files in a directory, or
+all \texttt{.proto} files provided by another R package.
 
-\subsection{Importing proto files}
-
-In contrast to the other languages (Java, C++, Python) that are officially
-supported by Google, the implementation used by the \texttt{RProtoBuf}
-package does not rely on the \texttt{protoc} compiler (with the exception of
-the two functions discussed in the previous section). This means that no
-initial step of statically compiling the proto file into C++ code that is
-then accessed by R code is necessary. Instead, \texttt{proto} files are
-parsed and processed \textsl{at runtime} by the protobuf C++ library---which
-is much more appropriate for a dynamic language.
-
-The \texttt{readProtoFiles} function allows importing \texttt{proto}
-files in several ways.
-
-Using the \texttt{file} argument, one can specify one or several file
-paths that ought to be proto files.
-
-<<>>=
-proto.dir <- system.file( "proto", package = "RProtoBuf" )
-proto.file <- file.path( proto.dir, "addressbook.proto" )
-<<eval=FALSE>>=
-readProtoFiles( proto.file )
-@
-
-With the \texttt{dir} argument, which is
-ignored if the \texttt{file} is supplied, all files matching the
-\texttt{.proto} extension will be imported.
-
-<<>>=
-dir( proto.dir, pattern = "\\.proto$", full.names = TRUE )
-<<eval=FALSE>>=
-readProtoFiles( dir = proto.dir )
-@
-
-Finally, with the
-\texttt{package} argument (ignored if \texttt{file} or
-\texttt{dir} is supplied), the function will import all \texttt{.proto}
-files that are located in the \texttt{proto} sub-directory of the given
-package. A typical use for this argument is in the \texttt{.onLoad}
-function of a package.
-
-<<eval=FALSE>>=
-readProtoFiles( package = "RProtoBuf" )
-@
+The \texttt{.proto} file syntax for defining the structure of protocol
+buffer data is described comprehensively on Google Code:
+\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.
 
@@
 special environment. The underlying mechanism used here is
-described in more detail in section~\ref{sec-lookup}.
+described in more detail in Section~\ref{sec-lookup}.
 
@@
 
+%\subsection{Importing proto files}
+%In contrast to the other languages (Java, C++, Python) that are officially
+%supported by Google, the implementation used by the \texttt{RProtoBuf}
+%package does not rely on the \texttt{protoc} compiler (with the exception of
+%the two functions discussed in the previous section). This means that no
+%initial step of statically compiling the proto file into C++ code that is
+%then accessed by R code is necessary. Instead, \texttt{proto} files are
+%parsed and processed \textsl{at runtime} by the protobuf C++ library---which
+%is much more appropriate for a dynamic language.
+
 \subsection{Creating a message}
 
-The objects contained in the special environment are
-descriptors for their associated message types. Descriptors will be
-discussed in detail in another part of this document, but for the
-purpose of this section, descriptors are just used with the \texttt{new}
-function to create messages.
+New messages are created with the \texttt{new} function which accepts
+a Message Descriptor and optionally a list of ``name = value'' pairs
+to set in the message.
+%The objects contained in the special environment are
+%descriptors for their associated message types. Descriptors will be
+%discussed in detail in another part of this document, but for the
+%purpose of this section, descriptors are just used with the \texttt{new}
+%function to create messages.
 
 <<>>=
+p1 <- new( tutorial.Person )
 p <- new( tutorial.Person, name = "Romain", id = 1 )
@@
 of bytes. The \texttt{serialize} method is implemented for
-protocol buffer messages to serialize a message into the sequence of
-bytes (raw vector in R speech) that represents the message.
+protocol buffer messages to serialize a message into a sequence of
+bytes that represents the message.
+%(raw vector in R speech) that represents the message.
 
@@
 tf1 <- tempfile()
-tf1
 serialize( p, tf1 )
@@
 
-The \texttt{RProtoBuf} package defines the \texttt{read}
-function to read messages from files, raw vector (the message payload)
-and arbitrary binary connections.
-
-<<>>=
-args( read )
-@
+The \texttt{RProtoBuf} package defines the \texttt{read} and
+\texttt{readASCII} functions to read messages from files, raw vectors,
+or arbitrary connections.  \texttt{read} expects to read the message
+payload from binary files or connections and \texttt{readASCII} parses
+the human-readable ASCII output that is created with
+\code{as.character}.
 
@@
 <<>>=
-message <- read( tutorial.Person, tf1 )
-writeLines( as.character( message ) )
+msg <- read( tutorial.Person, tf1 )
+writeLines( as.character( msg ) )
 @
@@
 
-\section{Basic Abstractions: Messages, Descriptors, and
-  DescriptorPools}
-
-The three basic abstractions of \CRANpkg{RProtoBuf} are Messages,
-which encapsulate a data structure, Descriptors, which define the
-schema used by one or more messages, and DescriptorPools, which
-provide access to descriptors.
-
 \section{Under the hood: S4 Classes, Methods, and Pseudo Methods}
@@
 described on the Table~\ref{Message-class-table}.
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{has} & Indicates if a message has a given field.   \\
+\texttt{clone} & Creates a clone of the message \\
+\texttt{isInitialized} & Indicates if a message has all its required fields set\\
+\texttt{serialize} & serialize a message to a file, binary connection, or raw vector\\
+\texttt{clear} & Clear one or several fields of a message, or the entire message\\
+\texttt{size} & The number of elements in a message field\\
+\texttt{bytesize} & The number of bytes the message would take once serialized\\
+\hline
+\texttt{swap} & swap elements of a repeated field of a message\\
+\texttt{set} & set elements of a repeated field\\
+\texttt{fetch} & fetch elements of a repeated field\\
+\texttt{setExtension} & set an extension of a message\\
+\texttt{getExtension} & get the value of an extension of a message\\
+\texttt{add} & add elements to a repeated field \\
+\hline
+\texttt{str} & the R structure of the message\\
+\texttt{as.character} & character representation of a message\\
+\texttt{toString} & character representation of a message (same as \texttt{as.character}) \\
+\texttt{as.list} & converts message to a named R list\\
+\texttt{update} & updates several fields of a message at once\\
+\texttt{descriptor} & get the descriptor of the message type of this message\\
+\texttt{fileDescriptor} & get the file descriptor of this message's descriptor\\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{Message-methods-table}Description of methods for the \texttt{Message} S4 class}
+\end{table}
+
+\subsection{Descriptors}
+
+Message descriptors are represented in R with the
+\emph{Descriptor} S4 class. The class contains
+the slots \texttt{pointer} and \texttt{type} :
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{slot} & \textbf{description} \\
+\hline
+\texttt{pointer} & external pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
+\texttt{Descriptor} class is available from the protocol buffer project page:
+\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
+\hline
+\texttt{type} & fully qualified path of the message type. \\
+\hline
+\end{tabular}
+\caption{\label{Descriptor-class-table}Description of slots for the \texttt{Descriptor} S4 class}
+\end{table}
+
+Similarly to messages, the \verb|$| operator can be used to extract
+information from the descriptor, or invoke pseuso-methods.
+
+\subsubsection{Extracting descriptors}
+
+The \verb|$| operator, when used on a descriptor object retrieves
+descriptors that are contained in the descriptor.
+
+This can be a field descriptor (see section~\ref{subsec-field-descriptor} ),
+an enum descriptor (see section~\ref{subsec-enum-descriptor}) or a descriptor
+for a nested type
+
+<<>>=
+# field descriptor
+tutorial.Person$email
+
+# enum descriptor
+tutorial.Person$PhoneType
+
+# nested type descriptor
+tutorial.Person$PhoneNumber
+# same as
+tutorial.Person.PhoneNumber
+@
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{new} & Creates a prototype of a message described by this descriptor.\\
+\texttt{read} & Reads a message from a file or binary connection.\\
+\texttt{readASCII} & Read a message in ASCII format from a file or
+text connection.\\
+\hline
+\texttt{name} & Retrieve the name of the message type associated with
+this descriptor.\\
+\texttt{as.character} & character representation of a descriptor\\
+\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}) \\
+\hline
+\texttt{fileDescriptor} & Retrieve the file descriptor of this
+descriptor.\\
+\texttt{containing\_type} & Retrieve the descriptor describing the message type containing this descriptor.\\
+\texttt{field\_count} & Return the number of fields in this descriptor.\\
+\texttt{field} & Return the descriptor for the specified field in this descriptor.\\
+\texttt{nested\_type\_count} & The number of nested types in this descriptor.\\
+\texttt{nested\_type} & Return the descriptor for the specified nested
+type in this descriptor.\\
+\texttt{enum\_type\_count} & The number of enum types in this descriptor.\\
+\texttt{enum\_type} & Return the descriptor for the specified enum
+type in this descriptor.\\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{Descriptor-methods-table}Description of methods for the \texttt{Descriptor} S4 class}
+\end{table}
 
@@
 
+TODO(ms): random citations to work in:
+
+Many sources compare data serialization formats and show protocol
+buffers very favorably to the alternatives, such
+as \citep{Sumaray:2012:CDS:2184751.2184810}
+
 %Its pretty useful.  Murray to see if he can get approval to talk a

### 2013-12-21 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Flesh out the "Under the hood: S4 Classes, Methods, and Pseudo Methods" section by describing how we wrap over 100 functions of 6 main classes with Rcpp.  How this exercise partially motivated the development of Rcpp Modules, and add concise tables of the methods for descriptors and field descriptors rather than going through and providing examples of every function like we do in the other vignette.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@589 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The \CRANpkg{RProtoBuf} package uses the S4 system to store
-information about descriptors and messages, but the information stored
-in the R object is very minimal and mainly consists of an external
-pointer to a C++ variable that is managed by the \texttt{protobuf} C++
-library.
+information about descriptors and messages.  Using the S4 system
+allows the \texttt{RProtoBuf} package to dispatch methods that are not
+generic in the S3 sense, such as \texttt{new} and
+\texttt{serialize}.
 
-Using the S4 system allows the \texttt{RProtoBuf} package to dispatch
-methods that are not generic in the S3 sense, such as \texttt{new} and
-\texttt{serialize}.
+Each R object stores an external pointer to an object managed by
+the \texttt{protobuf} C++ library.
+The \CRANpkg{Rcpp} \citep{eddelbuettel2011rcpp} package is used to
+facilitate the integration of the R and C++ code for these objects.
+
+% Message, Descriptor, FieldDescriptor, EnumDescriptor,
+% FileDescriptor, EnumValueDescriptor
+%
+% grep RPB_FUNC * | grep -v define|wc -l
+% 84
+% grep RPB_ * | grep -v RPB_FUNCTION | grep METHOD|wc -l
+% 33
+
+There are over 100 C++ functions that provide the glue code between
+the member functions of the 6 primary Message and Descriptor classes
+in the protobuf library.  Wrapping each method individually allows us
+to add user friendly custom error handling, type coercion, and
+performance improvements at the cost of a more verbose
+implementation.  The RProtoBuf implementation in many ways motivated
+the development of Rcpp Modules \citep{eddelbuettel2010exposing},
+which provide a more concise way of wrapping C++ functions and classes
+in a single entity.
 
@@
 
-TODO(ms): Perhaps a table here of the different S4 classes, how many
+\emph{TODO(ms): Perhaps a table here of the different S4 classes, how many
 methods they include, whether it dynamically does dispatch on other
-strings, whether/how it is available in the search path, etc.
+strings, whether/how it is available in the search path, etc.}
 
@@
 The \texttt{Message} S4 class represents Protocol Buffer Messages and
-is the core abstraction of \CRANpkg{RProtoBuf}.  Each \texttt{Message}
-has a \texttt{Descriptor} S4 class which defines the schema of the
-data defined in the Message, as well as a number of
-\texttt{FieldDescriptors} for the individual fields of the message.
+is the core abstraction of \CRANpkg{RProtoBuf}.  The class contains
+the slots \texttt{pointer} and \texttt{type} as described on the
+Table~\ref{Message-class-table}.
 
-represented in R using the \texttt{Message}
-S4 class. The class contains the slots \texttt{pointer} and \texttt{type} as
-described on the Table~\ref{Message-class-table}.
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{Slot} & \textbf{Description} \\
+\hline
+\texttt{pointer} & External pointer to the \texttt{Message} object of the C++ proto library. Documentation for the
+\texttt{Message} class is available from the protocol buffer project page:
+\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message} \\
+\hline
+\texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
+has its \texttt{type} slot set to \texttt{tutorial.Person} \\
+\hline
+\end{tabular}
+\caption{\label{Message-class-table}Description of slots for the \texttt{Message} S4 class}
+\end{table}
+
+Each \texttt{Message} contains a pointer to a \texttt{Descriptor}
+which defines the schema of the data defined in the Message, as well
+as a number of \texttt{FieldDescriptors} for the individual fields of
+the message.  In addition to the field name extractors of
+\texttt{Messages} introduced in the previous section, a complete list
+of Message methods is available in Table~\ref{Message-methods-table}.
 
@@
 \hline
-\textbf{slot} & \textbf{description} \\
+\textbf{Slot} & \textbf{Description} \\
 \hline
-\texttt{pointer} & external pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
+\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
 \texttt{Descriptor} class is available from the protocol buffer project page:
@@
 \hline
-\texttt{type} & fully qualified path of the message type. \\
+\texttt{type} & Fully qualified path of the message type. \\
 \hline
@@
 
-Similarly to messages, the \verb|$| operator can be used to extract
-information from the descriptor, or invoke pseuso-methods.
-
-\subsubsection{Extracting descriptors}
-
-The \verb|$| operator, when used on a descriptor object retrieves
-descriptors that are contained in the descriptor.
-
-This can be a field descriptor (see section~\ref{subsec-field-descriptor} ),
-an enum descriptor (see section~\ref{subsec-enum-descriptor}) or a descriptor
-for a nested type
+Similarly to messages, the \verb|$| operator can be used to retrieve
+descriptors that are contained in the descriptor, or invoke
+pseudo-methods.  Thise can be used to extract field descriptors, enum
+descriptors, or descriptors for a nested type.
 
@@
 @
+
+Table~\ref{Descriptor-methods-table} provides a complete list of the
+avalailable methods for Descriptors.
 
@@
 \texttt{toString} & character representation of a descriptor (same as \texttt{as.character}) \\
+\texttt{as.list} & return a named
+list of the field, enum, and nested descriptors included in this descriptor.\\
+\texttt{asMessage} & return DescriptorProto message. \\
 \hline
@@
 \caption{\label{Descriptor-methods-table}Description of methods for the \texttt{Descriptor} S4 class}
+\end{table}
+
+\subsection{field descriptors}
+\label{subsec-field-descriptor}
+
+The class \emph{FieldDescriptor} represents field
+descriptor in R. This is a wrapper S4 class around the
+\texttt{google::protobuf::FieldDescriptor} C++ class.
+Table~\ref{fielddescriptor-methods-table} describes the methods
+defined for the \texttt{FieldDescriptor} class.
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{Slot} & \textbf{Description} \\
+\hline
+\texttt{pointer} & External pointer to the \texttt{FieldDescriptor} C++ variable \\
+\hline
+\texttt{name} & Simple name of the field \\
+\hline
+\texttt{full\_name} & Fully qualified name of the field \\
+\hline
+\texttt{type} & Name of the message type where the field is declared \\
+\hline
+\end{tabular}
+\caption{\label{FieldDescriptor-class-table}Description of slots for the \texttt{FieldDescriptor} S4 class}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\hline
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{as.character} & Character representation of a descriptor\\
+\texttt{toString} & Character
+representation of a descriptor (same as \texttt{as.character}) \\
+\texttt{asMessage} & Return FieldDescriptorProto message. \\
+\texttt{name} & Return the name of the field descriptor.\\
+\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
+\texttt{containing\_type} & Return the containing descriptor of this field.\\
+\texttt{is\_extension} & Return TRUE if this field is an extension.\\
+\texttt{number} & Gets the declared tag number of the field.\\
+\texttt{type} & Gets the type of the field.\\
+\texttt{cpp\_type} & Gets the C++ type of the field.\\
+\texttt{label} & Gets the label of a field (optional, required, or repeated).\\
+\texttt{is\_repeated} & Return TRUE if this field is repeated.\\
+\texttt{is\_required} & Return TRUE if this field is required.\\
+\texttt{is\_optional} & Return TRUE if this field is optional.\\
+\texttt{has\_default\_value} & Return TRUE if this field has a default value.\\
+\texttt{default\_value} & Return the default value.\\
+\texttt{message\_type} & Return the message type if this is a message type field.\\
+\texttt{enum\_type} & Return the enum type if this is an enum type field.\\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{fielddescriptor-methods-table}Description of methods for the \texttt{FieldDescriptor} S4 class}
 \end{table}

### 2013-12-21 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Add section on enumdescriptors.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@590 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
+% TODO(ms): Useful distinction to make -- FieldDescriptor does not do
+% separate '$' dispatch like Messages, Descriptors, and
+% EnumDescriptors do.  Should it?
+
+\subsection{enum descriptors}
+\label{subsec-enum-descriptor}
+
+The class \emph{EnumDescriptor} is an R wrapper
+class around the C++ class \texttt{google::protobuf::EnumDescriptor}.
+Table~\ref{enumdescriptor-methods-table} describes the methods
+defined for the \texttt{EnumDescriptor} class.
+
+The \verb|$| operator can be used to retrieve the value of enum
+constants contained in the EnumDescriptor, or to invoke
+pseudo-methods.
+
+<<>>=
+tutorial.Person$PhoneType
+tutorial.Person$PhoneType$WORK
+@
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{Slot} & \textbf{Description} \\
+\hline
+\texttt{pointer} & External pointer to the \texttt{EnumDescriptor} C++ variable \\
+\hline
+\texttt{name} & Simple name of the enum \\
+\hline
+\texttt{full\_name} & Fully qualified name of the enum \\
+\hline
+\texttt{type} & Name of the message type where the enum is declared \\
+\hline
+\end{tabular}
+\caption{\label{EnumDescriptor-class-table}Description of slots for the \texttt{EnumDescriptor} S4 class}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\hline
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{as.list} & return a named
+integer vector with the values of the enum and their names.\\
+\texttt{as.character} & character representation of a descriptor\\
+\texttt{toString} & character
+representation of a descriptor (same as \texttt{as.character}) \\
+\texttt{asMessage} & return EnumDescriptorProto message. \\
+\texttt{name} & Return the name of the enum descriptor.\\
+\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
+\texttt{containing\_type} & Return the containing descriptor of this field.\\
+\texttt{length} & Return the number of constants in this enum.\\
+\texttt{has} & Return TRUE if this enum contains the specified named constant string.\\
+\texttt{value\_count} & Return the number of constants in this enum (same as \texttt{length}).\\
+\texttt{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{enumdescriptor-methods-table}Description of methods for the \texttt{EnumDescriptor} S4 class}
+\end{table}
+
 \section{Type Coercion}

### 2013-12-26 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Document file descriptors and enumvalue descriptors, and add a summary table showing the number of slots, methods, and dynamically dispatched fields of each class.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@594 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\texttt{serialize} can also be used in a more traditionnal
+\texttt{serialize} can also be used in a more traditional
 object oriented fashion using the dollar operator :
@@
 object oriented notation \verb|object$method(arguments)|.
+Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
+(defined in the \texttt{utils} package) for all classes to enable tab
+completion.  Completion possibilities include pseudo method names for all
+classes, plus dynamic dispatch on names or types specific to a given object.
 
-\emph{TODO(ms): Perhaps a table here of the different S4 classes, how many
-methods they include, whether it dynamically does dispatch on other
-strings, whether/how it is available in the search path, etc.}
+% TODO(ms): Add column check box for doing dynamic dispatch based on type.
+\begin{table}[h]
+\centering
+\begin{tabular}{|l|c|c|l|}
+\hline
+\textbf{Class} & \textbf{Slots} & \textbf{Methods} & \textbf{Dynamic Dispatch}\\
+\hline
+\hline
+Message & 2 & 20 & yes (field names)\\
+\hline
+Descriptor & 2 & 16 & yes (field names, enum types, nested types)\\
+\hline
+FieldDescriptor & 4 & 18 & no\\
+\hline
+EnumDescriptor & 4 & 11 & yes (enum constant names)\\
+\hline
+FileDescriptor & 3 & 6 & yes (message/field definitions)\\
+\hline
+EnumValueDescriptor & 3 & 6 & no\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 
+\subsection{file descriptors}
+\label{subsec-file-descriptor}
+
+The class \emph{FileDescriptor} is an R wrapper
+class around the C++ class \texttt{google::protobuf::FileDescriptor}.
+Table~\ref{filedescriptor-methods-table} describes the methods
+defined for the \texttt{FileDescriptor} class.
+
+The \verb|$| operator can be used to retrieve named fields defined in
+the FileDescriptor, or to invoke pseudo-methods.
+
+<<>>=
+f <- tutorial.Person$fileDescriptor()
+f
+f$Person
+@
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{slot} & \textbf{description} \\
+\hline
+\texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the C++ proto library. Documentation for the
+\texttt{FileDescriptor} class is available from the protocol buffer project page:
+\url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
+\hline
+\texttt{filename} & fully qualified pathname of the \texttt{.proto} file.\\
+\hline
+\texttt{package} & package name defined in this \texttt{.proto} file.\\
+\hline
+\end{tabular}
+\caption{\label{FileDescriptor-class-table}Description of slots for the \texttt{FileDescriptor} S4 class}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\hline
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{name} & Return the filename for this FileDescriptorProto.\\
+\texttt{package} & Return the file-level package name specified in this FileDescriptorProto.\\
+\texttt{as.character} & character representation of a descriptor. \\
+\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
+\texttt{asMessage} & return FileDescriptorProto message. \\
+\texttt{as.list} & return named list of descriptors defined in this file descriptor.\\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{filedescriptor-methods-table}Description of methods for the \texttt{FileDescriptor} S4 class}
+\end{table}
+
+\subsection{enum value descriptors}
+\label{subsec-enumvalue-descriptor}
+
+The class \emph{EnumValueDescriptor} is an R wrapper
+class around the C++ class \texttt{google::protobuf::EnumValueDescriptor}.
+Table~\ref{EnumValueDescriptor-methods-table} describes the methods
+defined for the \texttt{EnumValueDescriptor} class.
+
+The \verb|$| operator can be used to invoke pseudo-methods.
+
+<<>>=
+tutorial.Person$PhoneType$value(1)
+tutorial.Person$PhoneType$value(name="HOME")
+tutorial.Person$PhoneType$value(number=1)
+@
+
+\begin{table}[h]
+\centering
+\begin{tabular}{|cp{10cm}|}
+\hline
+\textbf{slot} & \textbf{description} \\
+\hline
+\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} C++ variable \\
+\hline
+\texttt{name} & simple name of the enum value \\
+\hline
+\texttt{full\_name} & fully qualified name of the enum value \\
+\hline
+\end{tabular}
+\caption{\label{EnumValueDescriptor-class-table}Description of slots for the \texttt{EnumValueDescriptor} S4 class}
+\end{table}
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{l|l}
+\hline
+\textbf{Method} & \textbf{Description} \\
+\hline
+\hline
+\texttt{number} & return the number of this EnumValueDescriptor. \\
+\texttt{name} & Return the name of the enum value descriptor.\\
+\texttt{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
+\texttt{as.character} & character representation of a descriptor. \\
+\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
+\texttt{asMessage} & return EnumValueDescriptorProto message. \\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{enumvaluedescriptor-methods-table}Description of methods for the \texttt{EnumValueDescriptor} S4 class}
+\end{table}
+
 \section{Type Coercion}

### 2013-12-28 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Add a large new section showing how RProtoBuf can be used to serialize all of the example datasets that ship with R, and compare the sizes with R's built-in serialziation method.
Steal a section on applications/MapReduce from the HistogramTools
vignette.  To be improved further.

Combine each of the class slot and method tables into a single table
for each class.

Add a brief section on type coercion issues for e.g. bools and int64s.

Improve the wording in various places.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@624 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 several common use cases for protocol buffers in data analysis.
+
+XXX Related work on IDLs (greatly expanded )
+
+XXX Design tradeoffs: reflection vs proto compiler
 
@@
 The \texttt{Message} S4 class represents Protocol Buffer Messages and
-is the core abstraction of \CRANpkg{RProtoBuf}.  The class contains
-the slots \texttt{pointer} and \texttt{type} as described on the
-Table~\ref{Message-class-table}.
+is the core abstraction of \CRANpkg{RProtoBuf}. Each \texttt{Message}
+contains a pointer to a \texttt{Descriptor} which defines the schema
+of the data defined in the Message, as well as a number of
+\texttt{FieldDescriptors} for the individual fields of the message.  A
+complete list of the slots and methods for \texttt{Messages}
+is available in Table~\ref{Message-methods-table}.
 
@@
 \centering
-\begin{tabular}{|cp{10cm}|}
+\begin{small}
+\begin{tabular}{l|p{10cm}}
 \hline
@@
 \texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
-has its \texttt{type} slot set to \texttt{tutorial.Person} \\
-\hline
-\end{tabular}
-\caption{\label{Message-class-table}Description of slots for the \texttt{Message} S4 class}
-\end{table}
-
-Each \texttt{Message} contains a pointer to a \texttt{Descriptor}
-which defines the schema of the data defined in the Message, as well
-as a number of \texttt{FieldDescriptors} for the individual fields of
-the message.  In addition to the field name extractors of
-\texttt{Messages} introduced in the previous section, a complete list
-of Message methods is available in Table~\ref{Message-methods-table}.
-
-\begin{table}[h]
-\centering
-\begin{small}
-\begin{tabular}{l|l}
+has its \texttt{type} slot set to \texttt{tutorial.Person} \\[.3cm]
+\hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{Message-methods-table}Description of methods for the \texttt{Message} S4 class}
+\caption{\label{Message-methods-table}Description of slots and methods for the \texttt{Message} S4 class}
 \end{table}
@@
 
-Message descriptors are represented in R with the
-\emph{Descriptor} S4 class. The class contains
-the slots \texttt{pointer} and \texttt{type} :
-
-\begin{table}[h]
-\centering
-\begin{tabular}{|cp{10cm}|}
-\hline
-\textbf{Slot} & \textbf{Description} \\
-\hline
-\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
-\texttt{Descriptor} class is available from the protocol buffer project page:
-\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
-\hline
-\texttt{type} & Fully qualified path of the message type. \\
-\hline
-\end{tabular}
-\caption{\label{Descriptor-class-table}Description of slots for the \texttt{Descriptor} S4 class}
-\end{table}
-
-Similarly to messages, the \verb|$| operator can be used to retrieve
-descriptors that are contained in the descriptor, or invoke
-pseudo-methods.  Thise can be used to extract field descriptors, enum
-descriptors, or descriptors for a nested type.
+Message descriptors are represented in R with the \emph{Descriptor} S4
+class. The class contains the slots \texttt{pointer} and
+\texttt{type}.  Similarly to messages, the \verb|$| operator can be
+used to retrieve descriptors that are contained in the descriptor, or
+invoke pseudo-methods.
 
@@
 Table~\ref{Descriptor-methods-table} provides a complete list of the
-avalailable methods for Descriptors.
+slots and avalailable methods for Descriptors.
 
@@
 \begin{small}
-\begin{tabular}{l|l}
+\begin{tabular}{l|p{10cm}}
+\hline
+\textbf{Slot} & \textbf{Description} \\
+\hline
+\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
+\texttt{Descriptor} class is available from the protocol buffer project page:
+\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
+\hline
+\texttt{type} & Fully qualified path of the message type. \\[.3cm]
+\hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{Descriptor-methods-table}Description of methods for the \texttt{Descriptor} S4 class}
+\caption{\label{Descriptor-methods-table}Description of slots and methods for the \texttt{Descriptor} S4 class}
 \end{table}
 
-\subsection{field descriptors}
+\subsection{Field Descriptors}
 \label{subsec-field-descriptor}
@@
 \hline
-\texttt{type} & Name of the message type where the field is declared \\
-\hline
-\end{tabular}
-\caption{\label{FieldDescriptor-class-table}Description of slots for the \texttt{FieldDescriptor} S4 class}
-\end{table}
-
-\begin{table}[h]
-\centering
-\begin{small}
-\begin{tabular}{l|l}
+\texttt{type} & Name of the message type where the field is declared \\[.3cm]
 \hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{fielddescriptor-methods-table}Description of methods for the \texttt{FieldDescriptor} S4 class}
+\caption{\label{fielddescriptor-methods-table}Description of slots and
+  methods for the \texttt{FieldDescriptor} S4 class}
 \end{table}
@@
 
-\subsection{enum descriptors}
+\subsection{Enum Descriptors}
 \label{subsec-enum-descriptor}
@@
 \hline
-\texttt{type} & Name of the message type where the enum is declared \\
-\hline
-\end{tabular}
-\caption{\label{EnumDescriptor-class-table}Description of slots for the \texttt{EnumDescriptor} S4 class}
-\end{table}
-
-\begin{table}[h]
-\centering
-\begin{small}
-\begin{tabular}{l|l}
+\texttt{type} & Name of the message type where the enum is declared \\[.3cm]
 \hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{enumdescriptor-methods-table}Description of methods for the \texttt{EnumDescriptor} S4 class}
+\caption{\label{enumdescriptor-methods-table}Description of slots and methods for the \texttt{EnumDescriptor} S4 class}
 \end{table}
 
-\subsection{file descriptors}
+\subsection{File Descriptors}
 \label{subsec-file-descriptor}
@@
 \hline
-\texttt{package} & package name defined in this \texttt{.proto} file.\\
-\hline
-\end{tabular}
-\caption{\label{FileDescriptor-class-table}Description of slots for the \texttt{FileDescriptor} S4 class}
-\end{table}
-
-\begin{table}[h]
-\centering
-\begin{small}
-\begin{tabular}{l|l}
+\texttt{package} & package name defined in this \texttt{.proto} file.\\[.3cm]
 \hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{filedescriptor-methods-table}Description of methods for the \texttt{FileDescriptor} S4 class}
+\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \texttt{FileDescriptor} S4 class}
 \end{table}
 
-\subsection{enum value descriptors}
+\subsection{Enum Value Descriptors}
 \label{subsec-enumvalue-descriptor}
@@
 \hline
-\texttt{full\_name} & fully qualified name of the enum value \\
-\hline
-\end{tabular}
-\caption{\label{EnumValueDescriptor-class-table}Description of slots for the \texttt{EnumValueDescriptor} S4 class}
-\end{table}
-
-\begin{table}[h]
-\centering
-\begin{small}
-\begin{tabular}{l|l}
+\texttt{full\_name} & fully qualified name of the enum value \\[.3cm]
 \hline
 \textbf{Method} & \textbf{Description} \\
-\hline
 \hline
@@
 \end{small}
-\caption{\label{enumvaluedescriptor-methods-table}Description of methods for the \texttt{EnumValueDescriptor} S4 class}
+\caption{\label{EnumValueDescriptor-methods-table}Description of slots
+  and methods for the \texttt{EnumValueDescriptor} S4 class}
 \end{table}
@@
 
+One of the benefits of using an Interface Definition Language (IDL)
+like Protocol Buffers is that it provides a highly portable basic type
+system that different language and hardware implementations can map to
+the most appropriate type in different environments.
+Table~\ref{table-get-types} details the correspondance between the
+field type and the type of data that is retrieved by \verb|$| and \verb|[[|
+extractors.
+
+\begin{table}[h]
+\centering
+\begin{small}
+\begin{tabular}{|c|p{5cm}p{5cm}|}
+\hline
+field type & R type (non repeated) & R type (repeated) \\
+\hline
+\hline
+double	& \texttt{double} vector & \texttt{double} vector \\
+float	& \texttt{double} vector & \texttt{double} vector \\
+\hline
+int32	  & \texttt{integer} vector & \texttt{integer} vector \\
+uint32	  & \texttt{integer} vector & \texttt{integer} vector \\
+sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
+fixed32	  & \texttt{integer} vector & \texttt{integer} vector \\
+sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\
+\hline
+int64	  & \texttt{integer} or \texttt{character}
+vector \footnotemark & \texttt{integer} or \texttt{character} vector \\
+uint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
+sint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
+fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
+sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
+\hline
+bool	& \texttt{logical} vector & \texttt{logical} vector \\
+\hline
+string	& \texttt{character} vector & \texttt{character} vector \\
+bytes	& \texttt{character} vector & \texttt{character} vector \\
+\hline
+enum & \texttt{integer} vector & \texttt{integer} vector \\
+\hline
+message & \texttt{S4} object of class \texttt{Message} & \texttt{list} of \texttt{S4} objects of class \texttt{Message} \\
+\hline
+\end{tabular}
+\end{small}
+\caption{\label{table-get-types}Correspondance between field type and
+  R type retrieved by the extractors. \footnotesize{1. R lacks native
+  64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
+  available to return large integers as characters to avoid losing
+  precision.  This option is described in Section~\ref{sec:int64}}.}
+\end{table}
+
 \subsection{Booleans}
-Bools
-Int64s.
-
-\subsection{64-bit integers}
-\label{sec:int64}
-
-R does not have native 64-bit integer support.  Instead, R treats
-large integers as doubles which have limited precision.  For example,
-it loses the ability to distinguish some distinct integers:
-
-<<>>=
-2^53 == (2^53 + 1)
-@
-
-Protocol Buffers are frequently used to pass data between different
-systems, however, and most other systems these days have support for
-64-bit integers.  To work around this, RProtoBuf allows users to get
-and set 64-bit integer types by treating them as characters.
+
+R booleans can accept three values: \texttt{TRUE}, \texttt{FALSE}, and
+\texttt{NA}.  However, most other languages, including the protocol
+buffer schema, only accept \text{TRUE} or \text{FALSE}.  This means
+that we simply can not store R logical vectors that include all three
+possible values as booleans.  The library will refuse to store
+\texttt{NA}s in protocol buffer boolean fields, and users must instead
+choose another type (such as integers) capable of storing three
+distinct values.
 
@@
 
+<<>>=
+a <- new(protobuf_unittest.TestAllTypes)
+a$optional_bool <- TRUE
+a$optional_bool <- FALSE
+<<eval=F>>=
+a$optional_bool <- NA
+<<echo=FALSE,eval=TRUE,print=TRUE>>=
+try(a$optional_bool <- NA,silent=TRUE)
+@
+
+\subsection{64-bit integers}
+\label{sec:int64}
+
+R does not have native 64-bit integer support.  Instead, R treats
+large integers as doubles which have limited precision.  For example,
+it loses the ability to distinguish some distinct integers:
+
+<<>>=
+2^53 == (2^53 + 1)
+@
+
+Protocol Buffers are frequently used to pass data between different
+systems, however, and most other modern systems do have support for
+64-bit integers.  To work around this, RProtoBuf allows users to get
+and set 64-bit integer types by treating them as characters.
+
+<<echo=FALSE,print=FALSE>>=
+if (!exists("protobuf_unittest.TestAllTypes",
+            "RProtoBuf:DescriptorPool")) {
+    unittest.proto.file <- system.file("unitTests", "data",
+                                       "unittest.proto",
+                                       package="RProtoBuf")
+    readProtoFiles(file=unittest.proto.file)
+}
+@
+
 If we try to set an int64 field in R to double values, we lose
@@
 
-\section{Related work on IDLs (greatly expanded from what you have)}
-
-\section{Design tradeoffs: reflection vs proto compiler (not addressed
-  at all in current vignettes)}
+\section{Evaluation: data.frame to Protocol Buffer Serialization}
+
+Saptarshi Guha wrote the RHIPE package \citep{rhipe} which includes
+protocol buffer integration with R.  However, this implementation
+takes a different approach: any R object is serialized into a message
+based on a single catch-all \texttt{proto} schema.  Jeroen Ooms took a
+similar approach influenced by Saptarshi in his \pkg{RProtoBufUtils}
+package.  Unlike Saptarshi's package, however, RProtoBufUtils depends
+on RProtoBuf for underlying message operations.  This package is
+available at \url{https://github.com/jeroenooms/RProtoBufUtils}.
+
+The \textbf{RProtoBufUtils} package by Jereoen Ooms provides a
+\texttt{serialize\_pb} method to convert R objects into serialized
+protocol buffers in this format, and the \texttt{can\_serialize\_pb}
+method can be used to determine whether the given R object can safely
+be expressed in this way.  To show how how this method works, we
+attempt to convert all of the built-in datasets from R into this
+serialized protocol buffer representation.
+
+<<echo=TRUE>>=
+library(RProtoBufUtils)
+
+datasets <- subset(as.data.frame(data()$results), Package=="datasets")
+datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
+n <- nrow(datasets)
+@
+
+There are \Sexpr{n} standard data sets included in R.  We use the
+\texttt{can\_serialize\_pb} method to determine how many of those can
+be safely converted to a serialized protocol buffer representation.
+
+<<echo=TRUE>>=
+datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
+datasets <- subset(datasets, valid.proto==TRUE)
+m <- nrow(datasets)
+@
+
+\Sexpr{m} data sets could be converted to Protocol Buffers
+(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
+many bytes were usued to store the data sets under four different
+situations (1) normal R serialization, (2) R serialization followed by
+gzip, (3) normal protocol buffer serialization, (4) protocol buffer
+serialization followed by gzip.
+
+\subsection{Compression Performance}
+\label{sec:compression}
+
+<<echo=FALSE,print=FALSE>>=
+datasets$object.size <- unname(sapply(datasets$load.name, function(x) object.size(eval(as.name(x)))))
+
+datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
+
+datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
+
+datasets$R.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))
+
+datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))
+
+datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))
+
+clean.df <- data.frame(dataset=datasets$load.name,
+                       object.size=datasets$object.size,
+                       "serialized"=datasets$R.serialize.size,
+                       "gzipped serialized"=datasets$R.serialize.size.gz,
+                       "RProtoBuf"=datasets$RProtoBuf.serialize.size,
+                       "gzipped RProtoBuf"=datasets$RProtoBuf.serialize.size.gz,
+                       check.names=FALSE)
+@
+
+Table~\ref{tab:compression} shows the sizes of 50 sample R datasets as
+returned by object.size() compared to the serialized sizes.
+The summary compression sizes are listed below, and a full table for a
+sample of 50 datasets is included on the next page.  Sizes are comparable
+but protocol buffers provide simple getters and setters in multiple
+languages instead of requiring other programs to parse the R
+serialization format \citep{serialization}.  One takeaway from this
+table is that RProtoBuf does not in general provide any significant
+space-savings over R's normal serialization mechanism.  The benefit
+from RProtoBuf comes from its interoperability with other
+environments.
+
+TODO comparison of protobuf serialization sizes/times for various vectors.
+Compared to R's native serialization.
+Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.
+
+% N.B. see table.Rnw for how this table is created.
+%
+% latex table generated in R 3.0.2 by xtable 1.7-0 package
+% Fri Dec 27 17:00:03 2013
+\begin{table}[ht]
+\begin{center}
+\scalebox{0.9}{
+\begin{tabular}{l|r|r|r|r|r}
+  \hline
+Data Set & object.size & \multicolumn{2}{c|}{R Serialization} &
+\multicolumn{2}{c}{RProtoBuf Serialization} \\
+ & & Default & gzipped & Default & gzipped \\
+  \hline
+uspop & 584.00 & 268 & 172 & 211 & 148 \\ 
+  Titanic & 1960.00 & 633 & 257 & 481 & 249 \\ 
+  volcano & 42656.00 & 42517 & 5226 & 42476 & 4232 \\ 
+  euro.cross & 2728.00 & 1319 & 910 & 1207 & 891 \\ 
+  attenu & 14568.00 & 8234 & 2165 & 7771 & 2336 \\ 
+  ToothGrowth & 2568.00 & 1486 & 349 & 1239 & 391 \\ 
+  lynx & 1344.00 & 1028 & 429 & 971 & 404 \\ 
+  nottem & 2352.00 & 2036 & 627 & 1979 & 641 \\ 
+  sleep & 2752.00 & 746 & 282 & 483 & 260 \\ 
+  co2 & 4176.00 & 3860 & 1473 & 3803 & 1453 \\ 
+  austres & 1144.00 & 828 & 439 & 771 & 410 \\ 
+  ability.cov & 1944.00 & 716 & 357 & 589 & 341 \\ 
+  EuStockMarkets & 60664.00 & 59785 & 21232 & 59674 & 19882 \\ 
+  treering & 64272.00 & 63956 & 17647 & 63900 & 17758 \\ 
+  freeny.x & 1944.00 & 1445 & 1311 & 1372 & 1289 \\ 
+  Puromycin & 2088.00 & 813 & 306 & 620 & 320 \\ 
+  warpbreaks & 2768.00 & 1231 & 310 & 811 & 343 \\ 
+  BOD & 1088.00 & 334 & 182 & 226 & 168 \\ 
+  sunspots & 22992.00 & 22676 & 6482 & 22620 & 6742 \\ 
+  beaver2 & 4184.00 & 3423 & 751 & 3468 & 840 \\ 
+  anscombe & 2424.00 & 991 & 375 & 884 & 352 \\ 
+  esoph & 5624.00 & 3111 & 548 & 2240 & 665 \\ 
+  PlantGrowth & 1680.00 & 646 & 303 & 459 & 314 \\ 
+  infert & 15848.00 & 14328 & 1172 & 13197 & 1404 \\ 
+  BJsales & 1632.00 & 1316 & 496 & 1259 & 465 \\ 
+  stackloss & 1688.00 & 917 & 293 & 844 & 283 \\ 
+  crimtab & 7936.00 & 4641 & 713 & 1655 & 576 \\ 
+  LifeCycleSavings & 6048.00 & 3014 & 1420 & 2825 & 1407 \\ 
+  Harman74.cor & 9144.00 & 6056 & 2045 & 5861 & 2070 \\ 
+  nhtemp & 912.00 & 596 & 240 & 539 & 223 \\ 
+  faithful & 5136.00 & 4543 & 1339 & 4936 & 1776 \\ 
+  freeny & 5296.00 & 2465 & 1518 & 2271 & 1507 \\ 
+  discoveries & 1232.00 & 916 & 199 & 859 & 180 \\ 
+  state.x77 & 7168.00 & 4251 & 1754 & 4068 & 1756 \\ 
+  pressure & 1096.00 & 498 & 277 & 427 & 273 \\ 
+  fdeaths & 1008.00 & 692 & 291 & 635 & 272 \\ 
+  euro & 976.00 & 264 & 186 & 202 & 161 \\ 
+  LakeHuron & 1216.00 & 900 & 420 & 843 & 404 \\ 
+  mtcars & 6736.00 & 3798 & 1204 & 3633 & 1206 \\ 
+  precip & 4992.00 & 1793 & 813 & 1615 & 815 \\ 
+  state.area & 440.00 & 422 & 246 & 405 & 235 \\ 
+  attitude & 3024.00 & 1990 & 544 & 1920 & 561 \\ 
+  randu & 10496.00 & 9794 & 8859 & 10441 & 9558 \\ 
+  state.name & 3088.00 & 844 & 408 & 724 & 415 \\ 
+  airquality & 5496.00 & 4551 & 1241 & 2874 & 1294 \\ 
+  airmiles & 624.00 & 308 & 170 & 251 & 148 \\ 
+  quakes & 33112.00 & 32246 & 9898 & 29063 & 11595 \\ 
+  islands & 3496.00 & 1232 & 563 & 1098 & 561 \\ 
+  OrchardSprays & 3600.00 & 2164 & 445 & 1897 & 483 \\ 
+  WWWusage & 1232.00 & 916 & 274 & 859 & 251 \\ 
+   \hline
+\end{tabular}
+}
+\caption{Serialization sizes with R's built-in serialization and
+  RProtoBuf for 50 sample R datasets.}
+\label{tab:compression}
+\end{center}
+\end{table}
 
@@
 
-\subsection{Serialization comparison}
-
-TODO comparison of protobuf serialization sizes/times for various vectors.
-Compared to R's native serialization.
-Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.
-
 \section{Descriptor lookup}
@@
 
-\section{Other approaches}
-
-Saptarshi Guha wrote another package that deals with integration
-of protocol buffer messages with R, taking a different angle :
-serializing any R object as a message, based on a single catch-all
-\texttt{proto} file.  Saptarshi's package is available at
-\url{http://ml.stat.purdue.edu/rhipe/doc/html/ProtoBuffers.html}.
-
-Jeroen Ooms took a similar approach influenced by Saptarshi in his
-\pkg{RProtoBufUtils} package.  Unlike Saptarshi's package,
-RProtoBufUtils depends on RProtoBuf for underlying message operations.
-This package is available at
-\url{https://github.com/jeroenooms/RProtoBufUtils}.
+%\section{Other approaches}
 
@@
 
-\section{Basic usage example - tutorial.Person}
-
-\section{Application: distributed Data Collection with MapReduce}
-
-We could describe a common MapReduce pattern of having the MR written
-in another language output protocol buffers that are later pulled into
-R.  There is some text about this in section 2 of
-http://cran.r-project.org/web/packages/HistogramTools/vignettes/HistogramTools.pdf
+%\section{Basic usage example - tutorial.Person}
+
+\section{Application: Distributed Data Collection with MapReduce}
+
+TODO(mstokely): Make this better.
+
+Many large data sets in fields such as particle physics and
+information processing are stored in binned or histogram form in order
+to reduce the data storage requirements
+\citep{scott2009multivariate}. Protocol Buffers make a particularly
+good data transport format in distributed MapReduces environments
+where large numbers of computers process a large data set for analysis.
+
+There are two common patterns for generating histograms of large data
+sets with MapReduce.  In the first method, each mapper task can
+generate a histogram over a subset of the data that is has been
+assigned, and then the histograms of each mapper are sent to one or
+more reducer tasks to merge.
+
+In the second method, each mapper rounds a data point to a bucket
+width and outputs that bucket as a key and '1' as a value.  Reducers
+then sum up all of the values with the same key and output to a data store.
+
+In both methods, the mapper tasks must choose identical
+bucket boundaries even though they are analyzing disjoint parts of the
+input set that may cover different ranges, or we must implement
+multiple phases.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
+\end{center}
+\caption{Diagram of MapReduce Histogram Generation Pattern}
+\label{fig:mr-histogram-pattern1}
+\end{figure}
+
+Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
+described above for histogram generation of large data sets with
+MapReduce.
+
+This package is designed to be helpful if some of the Map or Reduce
+tasks are written in R, or if those components are written in other
+languages and only the resulting output histograms need to be
+manipulated in R.
 

### 2013-12-29 papers/rjournal/eddelbuettel-francois-stokely.Rnw
some edits for typos etc
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@627 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{by Dirk Eddelbuettel, Romain Fran\c{c}ois, and Murray Stokely}
+
+%% DE: I tend to have wider option(width=...) so this
+%%     guarantees better line breaks
+<<echo=FALSE,print=FALSE>>=
+options(width=65, prompt="R> ", digits=4)
+@
 
@@
 analysis pipeline may involve storing intermediate results in a
-file or sending them over the network.  Programming langauges such as
+file or sending them over the network.  Programming languages such as
 Java, Ruby, Python, and R include built-in serialization support, but
@@
 tabular datasets, lack of type-safety, inefficient text representation
-and parsing, and abiguities in the format involving special
-characters.  JSON is another widely supported format used mostly on
+and parsing, and ambiguities in the format involving special
+characters.  JSON is another widely-supported format used mostly on
 the web that removes many of these disadvantages, but it too suffers
@@
 
+%% TODO(de,ms)  What follows is oooooold and was lifted from the webpage
+%%              Rewrite?
 Protocol buffers are a language-neutral, platform-neutral, extensible
@@
 
+%% TODO(de) Can we make this not break the width of the page?
 \noindent
@@
 the \texttt{protobuf} C++ library.
-The \CRANpkg{Rcpp} \citep{eddelbuettel2011rcpp} package is used to
+The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp} is used to
 facilitate the integration of the R and C++ code for these objects.
@@
 <<echo=TRUE>>=
-datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
-datasets <- subset(datasets, valid.proto==TRUE)
+#datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
+#datasets <- subset(datasets, valid.proto==TRUE)
 m <- nrow(datasets)

### 2013-12-30 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Update the type table to note that uint32 and fixed32 are stored as doubles, not R integers since R doesn't have unsigned integers.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@633 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \hline
+uint32	  & \texttt{double} vector & \texttt{double} vector \\
+fixed32	  & \texttt{double} vector & \texttt{double} vector \\
+\hline
 int32	  & \texttt{integer} vector & \texttt{integer} vector \\
-uint32	  & \texttt{integer} vector & \texttt{integer} vector \\
 sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
-fixed32	  & \texttt{integer} vector & \texttt{integer} vector \\
 sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Remove Romain from authors list per his request, add an acknowledgements section which of course thanks him most of all.
Add a section on unsigned integers in the type coercion section, now
that they are handled better in RProtoBuf.

Add more text in the Descriptors subsection.

Add references to int64 (orphaned but still useful) and bit64 in the
64-bit integers subsection.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@686 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\author{by Dirk Eddelbuettel, Romain Fran\c{c}ois, and Murray Stokely}
+\author{by Dirk Eddelbuettel and Murray Stokely}
 
@@
 
-Message descriptors are represented in R with the \emph{Descriptor} S4
+Descriptors describe the type of a Message.  This includes what fields
+a message contains and what the types of those fields are.  Message
+descriptors are represented in R with the \emph{Descriptor} S4
 class. The class contains the slots \texttt{pointer} and
@@
 invoke pseudo-methods.
+
+When \CRANpkg{RProtoBuf} is first loaded it calls
+\texttt{readProtoFiles} to read in an example \texttt{.proto} file
+included with the package.  The \texttt{tutorial.Person} descriptor
+and any other descriptors defined in loaded \texttt{.proto} files are
+then available on the search path.
 
@@
 
+\subsection{Unsigned Integers}
+
+R lacks a native unsigned integer type.  Values between $2^{31}$ and
+$2^{32} - 1$ read from unsigned int protocol buffer fields must be
+stored as doubles in R.
+
+<<>>=
+as.integer(2^31-1)
+<<eval=F>>=
+as.integer(2^31 - 1) + as.integer(1)
+<<echo=FALSE,eval=TRUE,print=TRUE>>=
+try(as.integer(2^31 - 1) + as.integer(1))
+
+2^31
+class(2^31)
+@
+
 \subsection{64-bit integers}
@@
 will be returned if the \texttt{RProtoBuf.int64AsString} option is set
-to \texttt{TRUE}.
+to \texttt{TRUE}.  The character values are useful because they can
+accurately be used as unique identifiers and can easily be passed to R
+packages such as \CRANpkg{int64} \citep{int64} or \CRANpkg{bit64}
+\citep{bit64} which represent 64-bit integers in R.
 
@@
 
+\section{Acknowledgement}
+
+\CRANpkg{RProtoBuf} was originally written in 2009 by Romain
+Fran\c{c}ois and Dirk Eddelbuettel.  The authors would particularly
+like to thank Romain for his initial implementation and continued
+design discussions. Several features of this package are based
+on the design of the \CRANpkg{rJava} package by Simon Urbanek
+(dispatch on new, S4 class structures using external pointers). We'd
+like to thank Simon for his indirect involvment on
+\CRANpkg{RProtoBuf}.  The user defined table mechanism, implemented by
+Duncan Temple Lang for the purpose of the \pkg{RObjectTables}
+package allowed the dynamic symbol lookup. Many thanks to Duncan for
+this amazing feature.  Kenton Varda was generous with his time in
+reviewing code and explaining obscure protocol buffer semantics.  Karl
+Millar and Jeroen Ooms were helpful in reviewing code or offering
+suggestions.  Saptarshi Guha's contemporaneous work on \pkg{RHIPE} was a
+strong motivator.
+
 \bibliography{eddelbuettel-francois-stokely}
@@
 \email{edd@debian.org}
-
-\address{Author Two\\
-  Affiliation\\
-  Address\\
-  Country}
-\email{author2@work}
 

### 2014-01-03 papers/rjournal/JSSwrapper.tex
Add a JSSwrapper around the same content used for the RJournal article.  Edits should still be made to the same eddelbuettel-francois-stokely.Rnw file, and the Makefile still only builds the RJournal version by default, but now you can type pdflatex JSSwrapper instead and manually see what it would look like when formatted for JSS instead.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@687 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%
% Local helpers to make this more compatible with R Journal style.
%
\newcommand{\CRANpkg}[1]{\pkg{#1}}
\RequirePackage{fancyvrb}
\RequirePackage{alltt}
\DefineVerbatimEnvironment{example}{Verbatim}{}

%% almost as usual
\author{Dirk Eddelbuettel\\Debian and R Projects \And 
        Murray Stokely\\Google, Inc}
\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
Modern data collection and analysis pipelines often involve
a sophisticated mix of applications written in general purpose and
specialized programming languages.  Protocol Buffers are a popular
method of serializing structured data between applications---while remaining
independent of programming languages or operating system.  The
\CRANpkg{RProtoBuf} package provides a complete interface to this
library.
}
\Keywords{r, protocol buffers, serialization, cross-platform}
\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Murray Stokely\\
  Google, Inc.\\
  1600 Amphitheatre Parkway\\
  Mountain View, CA 94040\\
  USA\\
  E-mail: \email{mstokely@google.com}\\
  URL: \url{http://www.stokely.org/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\include{eddelbuettel-francois-stokely}
%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

%\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Add a JSSwrapper around the same content used for the RJournal article.  Edits should still be made to the same eddelbuettel-francois-stokely.Rnw file, and the Makefile still only builds the RJournal version by default, but now you can type pdflatex JSSwrapper instead and manually see what it would look like when formatted for JSS instead.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@687 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection{Importing Message Descriptors from \texttt{.proto} files}
+\subsection{Importing Message Descriptors from .proto files}
 
@@
 \texttt{NA}.  However, most other languages, including the protocol
-buffer schema, only accept \text{TRUE} or \text{FALSE}.  This means
+buffer schema, only accept \texttt{TRUE} or \texttt{FALSE}.  This means
 that we simply can not store R logical vectors that include all three
@@
 % Fri Dec 27 17:00:03 2013
-\begin{table}[ht]
+\begin{table}[h!]
 \begin{center}
@@
 \bibliography{eddelbuettel-francois-stokely}
-
-\address{Dirk Eddelbuettel\\
-  Debian and R Projects\\
-  711 Monroe Avenue, River Forest, IL 60305\\
-  USA}
-\email{edd@debian.org}
-
-\address{Murray Stokely\\
-  Google, Inc.\\
-  1600 Amphitheatre Parkway\\
-  Mountain View, CA 94043\\
-  USA}
-\email{mstokely@google.com}

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Improve introductory text.  Compare with the binary JSON formats like MessagePack and BSON.  Add the new high level figure about serializing/deserializing data from an interactive R session to other distributed systems.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@696 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 complexity through reusability, modularity, and fault
-isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  Different
-programming languages are often used for the different phases of data
+isolation \citep{Wegiel:2010:CTT:1932682.1869479}.  
+Data analysis patterns such as Split-Apply-Combine
+\citep{wickham2011split} explicitly break up large problems into
+manageable pieces.  These patterns are frequently employed with
+different programming languages used for the different phases of data
 analysis -- collection, cleaning, analysis, post-processing, and
@@
 analysis pipeline may involve storing intermediate results in a
-file or sending them over the network.  Programming languages such as
-Java, Ruby, Python, and R include built-in serialization support, but
-these formats are tied to the specific programming language in use.
+file or sending them over the network.
+
+Programming languages such as Java, Ruby, Python, and R include
+built-in serialization support, but these formats are tied to the
+specific programming language in use and thus lock the user into a
+single environment.
+%
+% do not facilitate
 % TODO(ms): and they often don't support versioning among other faults.
@@
 from being too slow to parse and also does not provide strong typing
-between integers and floating point.  Large numbers of JSON messages
-would also be required to duplicate the field names with each message.
-
-TODO(ms): Also work in reference to Split-Apply-Combine pattern for
-data analysis \citep{wickham2011split}, since that is a great pattern
-but it seems overly optimistic to expect all of those phases to always
-be done in the same language.
+between integers and floating point.  Because the schema information
+is not kept separately, multiple JSON messages of the same
+type needlessly duplicate the field names with each message.
+%
+%
+%
+A number of binary formats based on JSON have been proposed that
+reduce the parsing cost and improve the efficiency.  MessagePack
+\citep{msgpackR} and BSON \citep{rmongodb} both have R interfaces, but
+these formats lack a separate schema for the serialized data and thus
+still duplicate field names with each message sent over the network or
+stored in a file.  Such formats also lack support for versioning when
+data storage needs evolve over time.
+
+% TODO(mstokely): Take a more conversational tone here asking
+% questions and motivating protocol buffers?
 
@@
 
+\section{Protocol Buffers}
+% This content is good.  Maybe use and cite?
+% http://martin.kleppmann.com/2012/12/05/schema-evolution-in-avro-protocol-buffers-thrift.html
+
+Protocol Buffers are a widely used modern language-neutral,
+platform-neutral, extensible mechanism for sharing structured data.
+
+one of the more popular examples of the modern
+
 XXX Related work on IDLs (greatly expanded )
@@
 XXX Design tradeoffs: reflection vs proto compiler
-
-\section{Protocol Buffers}
 
@@
 project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
+
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=\textwidth]{protobuf-distributed-system-crop.pdf}
+\end{center}
+\caption{Example protobuf usage}
+\label{fig:protobuf-distributed-usecase}
+\end{figure}
 

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Further improve the introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@697 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Programming languages such as Java, Ruby, Python, and R include
-built-in serialization support, but these formats are tied to the
-specific programming language in use and thus lock the user into a
-single environment.
-%
-% do not facilitate
-% TODO(ms): and they often don't support versioning among other faults.
-CSV files can be read and written by many applications and so are
-often used for exporting tabular data.  However, CSV files have a
-number of disadvantages, such as a limitation of exporting only
+Given these requirements, how do we safely share intermediate results
+between different applications, possibly written in different
+languages, and possibly running on different computers?  Programming
+languages such as R, Java, Julia, and Python include built-in
+serialization support, but these formats are tied to the specific
+programming language in use and thus lock the user into a single
+environment.  CSV files can be read and written by many applications
+and so are often used for exporting tabular data.  However, CSV files
+have a number of disadvantages, such as a limitation of exporting only
 tabular datasets, lack of type-safety, inefficient text representation
@@
 between integers and floating point.  Because the schema information
-is not kept separately, multiple JSON messages of the same
-type needlessly duplicate the field names with each message.
+is not kept separately, multiple JSON messages of the same type
+needlessly duplicate the field names with each message.
 %
@@
 
+Once the data serialization needs of an application become complex
+enough, developers typically benefit from the use of an
+\emph{interface description language}, or \emph{IDL}.  IDLs like
+Google's Protocol Buffers, Apache Thrift, and Apache Avro provide a compact
+well-documented schema for cross-langauge data structures and
+efficient binary interchange formats.  The schema can be used to
+generate model classes for statically typed programming languages such
+as C++ and Java, or can be used with reflection for dynamically typed
+programming languages.  Since the schema is provided separately from
+the encoded data, the data can be efficiently encoded to minimize
+storage costs of the stored data when compared with simple
+``schema-less'' binary interchange formats.
+
 % TODO(mstokely): Take a more conversational tone here asking
@@
 
-Protocol Buffers are a widely used modern language-neutral,
-platform-neutral, extensible mechanism for sharing structured data.
+Protocol Buffers are a widely used modern language-neutral, platform-neutral, extensible mechanism for sharing structured data.
 
@@
 XXX Design tradeoffs: reflection vs proto compiler
-
-Once the data serialization needs get complex enough, application
-developers typically benefit from the use of an \emph{interface
-description language}, or \emph{IDL}.  IDLs like Google's Protocol
-Buffers and Apache Thrift provide a compact well-documented schema for
-cross-langauge data structures as well efficient binary interchange
-formats.  The schema can be used to generate model classes for
-statically typed programming languages such as C++ and Java, or can be
-used with reflection for dynamically typed programming languages.
-Since the schema is provided separately from the encoded data, the
-data can be efficiently encoded to minimize storage costs of the
-stored data when compared with simple ``schema-less'' binary
-interchange formats like BSON.
 

### 2014-01-03 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Improve section 2 on protocol buffers.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@698 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 languages, and possibly running on different computers?  Programming
-languages such as R, Java, Julia, and Python include built-in
+languages such as R, Julia, Java, and Python include built-in
 serialization support, but these formats are tied to the specific
@@
 \emph{interface description language}, or \emph{IDL}.  IDLs like
-Google's Protocol Buffers, Apache Thrift, and Apache Avro provide a compact
+Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro provide a compact
 well-documented schema for cross-langauge data structures and
@@
 
+% TODO(mstokely): If we go to JSS, include a larger paragraph here
+% referencing each numbered section.  I don't like these generally,
+% but its useful for this paper I think because we have a boring bit
+% in the middle (full class/method details) and interesting
+% applications at the end.
 This article describes the basics of Google's Protocol Buffers through
@@
 \section{Protocol Buffers}
+
+Introductory section which may include references in parentheses
+\citep{R}, or cite a reference such as \citet{R} in the text.
+
 % This content is good.  Maybe use and cite?
@@
 
-Protocol Buffers are a widely used modern language-neutral, platform-neutral, extensible mechanism for sharing structured data.
-
-one of the more popular examples of the modern
-
-XXX Related work on IDLs (greatly expanded )
-
-XXX Design tradeoffs: reflection vs proto compiler
-
-% TODO(ms) Also talk about versioning and why its useful.
-
-%BSON, msgpack, Thrift, and Protocol Buffers take this latter approach,
-%with the
-
-% There are references comparing these we should use here.
-
-TODO Also mention Thrift and msgpack and the references comparing some
-of these tradeoffs.
-
-Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
-
 %% TODO(de,ms)  What follows is oooooold and was lifted from the webpage
 %%              Rewrite?
-Protocol buffers are a language-neutral, platform-neutral, extensible
-way of serializing structured data for use in communications
-protocols, data storage, and more.
-
-Protocol Buffers offer key features such as an efficient data interchange
-format that is both language- and operating system-agnostic yet uses a
-lightweight and highly performant encoding, object serialization and
-de-serialization as well data and configuration management. Protocol
-buffers are also forward compatible: updates to the \texttt{proto}
-files do not break programs built against the previous specification.
-
-While benchmarks are not available, Google states on the project page that in
-comparison to XML, protocol buffers are at the same time \textsl{simpler},
-between three to ten times \textsl{smaller}, between twenty and one hundred
-times \textsl{faster}, as well as less ambiguous and easier to program.
-
-The protocol buffers code is released under an open-source (BSD) license. The
-protocol buffer project (\url{http://code.google.com/p/protobuf/})
-contains a C++ library and a set of runtime libraries and compilers for
-C++, Java and Python.
-
-With these languages, the workflow follows standard practice of so-called
-Interface Description Languages (IDL)
-(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
-  on IDL}).  This consists of compiling a protocol buffer description file
-(ending in \texttt{.proto}) into language specific classes that can be used
-to create, read, write and manipulate protocol buffer messages. In other
-words, given the 'proto' description file, code is automatically generated
-for the chosen target language(s). The project page contains a tutorial for
-each of these officially supported languages:
-\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}
-
-Besides the officially supported C++, Java and Python implementations, several projects have been
-created to support protocol buffers for many languages. The list of known
-languages to support protocol buffers is compiled as part of the
-project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
+Protocol Buffers are a modern language-neutral, platform-neutral,
+extensible mechanism for sharing and storing structured data.  They
+have been widely adopted in industry with applications as varied as Sony
+Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  While
+traditional IDLs were previously characterized by bloat and
+complexity, Protocol Buffers is based on a simple list and records
+model that is flexible and easy to use.  Some of the key features
+provided by Protocol Buffers for data analysis include:
+
+\begin{itemize}
+\item \emph{Portable}:  Allows users to send and receive data between
+  applications or different computers.
+\item \emph{Efficient}:  Data is serialized into a compact binary
+  representation for transmission or storage.
+\item \emph{Exentsible}:  New fields can be added to Protocol Buffer Schemas
+  in a forward-compatible way that do not break older applications.
+\item \emph{Stable}:  Protocol Buffers have been in wide use for over a
+  decade.
+\end{itemize}
+
+Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
+communication workflow with protocol buffers and an interactive R
+session.  Common use cases include populating a request RPC protocol
+buffer in R that is then serialized and sent over the network to a
+remote server.  The server would then deserialize the message, act on
+the request, and respond with a new protocol buffer over the network.
+
+%Protocol buffers are a language-neutral, platform-neutral, extensible
+%way of serializing structured data for use in communications
+%protocols, data storage, and more.
+
+%Protocol Buffers offer key features such as an efficient data interchange
+%format that is both language- and operating system-agnostic yet uses a
+%lightweight and highly performant encoding, object serialization and
+%de-serialization as well data and configuration management. Protocol
+%buffers are also forward compatible: updates to the \texttt{proto}
+%files do not break programs built against the previous specification.
+
+%While benchmarks are not available, Google states on the project page that in
+%comparison to XML, protocol buffers are at the same time \textsl{simpler},
+%between three to ten times \textsl{smaller}, between twenty and one hundred
+%times \textsl{faster}, as well as less ambiguous and easier to program.
+
+Many sources compare data serialization formats and show protocol
+buffers very favorably to the alternatives, such
+as \citep{Sumaray:2012:CDS:2184751.2184810}
+
+%The flexibility of the reflection-based API is particularly well
+%suited for interactive data analysis.
+
+% XXX Design tradeoffs: reflection vs proto compiler
+
+For added speed and efficiency, the C++, Java, and Python bindings to
+Protocol Buffers are used with a compiler that translates a protocol
+buffer schema description file (ending in \texttt{.proto}) into
+language-specific classes that can be used to create, read, write and
+manipulate protocol buffer messages.  The R interface, in contrast,
+uses a reflection-based API that is particularly well suited for
+interactive data analysis.  All messages in R have a single class
+structure, but different accessor methods are created at runtime based
+on the name fields of the specified message type.
+
+% In other words, given the 'proto'
+%description file, code is automatically generated for the chosen
+%target language(s). The project page contains a tutorial for each of
+%these officially supported languages:
+%\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}
+
+%The protocol buffers code is released under an open-source (BSD) license. The
+%protocol buffer project (\url{http://code.google.com/p/protobuf/})
+%contains a C++ library and a set of runtime libraries and compilers for
+%C++, Java and Python.
+
+%With these languages, the workflow follows standard practice of so-called
+%Interface Description Languages (IDL)
+%(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
+%  on IDL}).  This consists of compiling a protocol buffer description file
+%(ending in \texttt{.proto}) into language specific classes that can be used
+
+%Besides the officially supported C++, Java and Python implementations, several projects have been
+%created to support protocol buffers for many languages. The list of known
+%languages to support protocol buffers is compiled as part of the
+%project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
 
@@
 Message Descriptors are defined in \texttt{.proto} files and define a
-schema for a particular named class of messages.  This separation
-between schema and the message objects is in contrast to
-more verbose formats like JSON, and when combined with the efficient
-binary representation of any Message object explains a large part of
-the performance and storage-space advantage offered by Protocol
-Buffers. TODO(ms): we already said some of this above.  clean up.
+schema for a particular named class of messages.
 
@@
 populating its fields.
+
+% Commented out because we said this earlier.
+%This separation
+%between schema and the message objects is in contrast to
+%more verbose formats like JSON, and when combined with the efficient
+%binary representation of any Message object explains a large part of
+%the performance and storage-space advantage offered by Protocol
+%Buffers. TODO(ms): we already said some of this above.  clean up.
 
@@
 
-TODO(ms): random citations to work in:
-
-Many sources compare data serialization formats and show protocol
-buffers very favorably to the alternatives, such
-as \citep{Sumaray:2012:CDS:2184751.2184810}
+% RProtoBuf has been used.
 

### 2014-01-04 papers/rjournal/eddelbuettel-francois-stokely.Rnw
Submit dirk's rewrite of acknowledgments section.  Add todo about OpenGPU or Poly at the bottom.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@699 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 from RProtoBuf comes from its interoperability with other
-environments.
+environments, safe versioning,
 
@@
 
+TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
+Chu's Poly.
+
 \section{Summary}
@@
 
-\CRANpkg{RProtoBuf} was originally written in 2009 by Romain
-Fran\c{c}ois and Dirk Eddelbuettel.  The authors would particularly
-like to thank Romain for his initial implementation and continued
-design discussions. Several features of this package are based
-on the design of the \CRANpkg{rJava} package by Simon Urbanek
-(dispatch on new, S4 class structures using external pointers). We'd
-like to thank Simon for his indirect involvment on
-\CRANpkg{RProtoBuf}.  The user defined table mechanism, implemented by
-Duncan Temple Lang for the purpose of the \pkg{RObjectTables}
-package allowed the dynamic symbol lookup. Many thanks to Duncan for
-this amazing feature.  Kenton Varda was generous with his time in
-reviewing code and explaining obscure protocol buffer semantics.  Karl
-Millar and Jeroen Ooms were helpful in reviewing code or offering
-suggestions.  Saptarshi Guha's contemporaneous work on \pkg{RHIPE} was a
-strong motivator.
+The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010,
+with very significant contributions, both in code and design, made by
+Romain Fran\c{c}ois. His continued influence on design and code is
+appreciated. Several features of the package are influenced
+by on the design of the \CRANpkg{rJava} package by Simon Urbanek
+The user-defined table mechanism, implemented by Duncan Temple Lang for the
+purpose of the \pkg{RObjectTables} package allowed the dynamic symbol lookup.
+Kenton Varda was generous with his time in reviewing code and explaining
+obscure protocol buffer semantics.  Karl Millar and Jeroen Ooms were
+helpful in reviewing code or offering suggestions.  The contemporaneous
+work by Saptarshi Guha on \pkg{RHIPE} was a strong initial motivator.
 

### 2014-01-04 papers/rjournal/eddelbuettel-francois-stokely.Rnw
one incomplete round of comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@700 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
  independent of programming languages or operating system.  The
- \CRANpkg{RProtoBuf} package provides a complete interface to this
- library.
+ \CRANpkg{RProtoBuf} package provides a complete interface between this
+ library and the R environment for statistical computing.
  %TODO(ms) keep it less than 150 words.
@@
 file or sending them over the network.
+% DE: Nice!
 
@@
 between different applications, possibly written in different
-languages, and possibly running on different computers?  Programming
+languages, and possibly running on different computer system, possibly
+spanning different operating systems?  Programming
 languages such as R, Julia, Java, and Python include built-in
 serialization support, but these formats are tied to the specific
+% DE: need to define serialization?
 programming language in use and thus lock the user into a single
@@
 stored in a file.  Such formats also lack support for versioning when
-data storage needs evolve over time.
+data storage needs evolve over time, or when application logic and
+requirement changes dictate update to the message format.
+% DE: Need to talk about XML ?
 
@@
 efficient binary interchange formats.  The schema can be used to
-generate model classes for statically typed programming languages such
-as C++ and Java, or can be used with reflection for dynamically typed
+generate model classes for statically-typed programming languages such
+as C++ and Java, or can be used with reflection for dynamically-typed
 programming languages.  Since the schema is provided separately from
@@
 
-Introductory section which may include references in parentheses
+FIXME Introductory section which may include references in parentheses
 \citep{R}, or cite a reference such as \citet{R} in the text.
@@
 %%              Rewrite?
-Protocol Buffers are a modern language-neutral, platform-neutral,
-extensible mechanism for sharing and storing structured data.  They
-have been widely adopted in industry with applications as varied as Sony
-Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  While
-traditional IDLs were previously characterized by bloat and
-complexity, Protocol Buffers is based on a simple list and records
-model that is flexible and easy to use.  Some of the key features
-provided by Protocol Buffers for data analysis include:
+Protocol Buffers can be described as a modern, language-neutral, platform-neutral,
+extensible mechanism for sharing and storing structured data.  Since their
+introduction, Protocol Buffers have been widely adopted in industry with
+applications as varied as database-internal messaging (Drizzle), % DE: citation?
+Sony Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  While
+% TODO(DE): This either needs a citation, or remove the name drop
+traditional IDLs have at time been criticized for code bloat and
+complexity, Protocol Buffers are based on a simple list and records
+model that is compartively flexible and simple to use.
+
+Some of the key features provided by Protocol Buffers for data analysis
+include:
 
@@
 remote server.  The server would then deserialize the message, act on
-the request, and respond with a new protocol buffer over the network.
+the request, and respond with a new protocol buffer over the network. The key
+difference to, say, a request to an Rserve instance is that the remote server
+may not even know the R language.
 
@@
 buffers very favorably to the alternatives, such
-as \citep{Sumaray:2012:CDS:2184751.2184810}
+as \citet{Sumaray:2012:CDS:2184751.2184810}
 

### 2014-01-04 papers/rjournal/eddelbuettel-stokely.Rnw
renaming
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@701 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
table 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@702 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-04 papers/rjournal/eddelbuettel-francois-stokely.Rnw
@@
 \begin{table}
-\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .015\textwidth}|@{\hskip .015\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
-\hline
+\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .02\textwidth}@{\hskip .02\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
+\toprule
 Schema : \texttt{addressbook.proto} & Example R Session\\
-\hline
+\cmidrule{1-2}
 \begin{minipage}{.35\textwidth}
@@
 \vspace{2mm}
-\end{minipage} & \begin{minipage}{.45\textwidth}
+\end{minipage} & \begin{minipage}{.5\textwidth}
 <<echo=TRUE>>=
 library(RProtoBuf)
-person <- new(tutorial.Person, id=1, name="Dirk")
-person
-person$name
-person$name <- "Romain"
-cat(as.character(person))
-serialize(person, NULL)
+p <- new(tutorial.Person, id=1, name="Dirk")
+class(p)
+p$name
+p$name <- "Murray"
+cat(as.character(p))
+serialize(p, NULL)
 @
 \end{minipage} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 appreciated. Several features of the package are influenced
-by on the design of the \CRANpkg{rJava} package by Simon Urbanek
+by the design of the \CRANpkg{rJava} package by Simon Urbanek
 The user-defined table mechanism, implemented by Duncan Temple Lang for the
@@
 
-\bibliography{eddelbuettel-francois-stokely}
+\bibliography{eddelbuettel-stokely}

### 2014-01-04 papers/rjournal/eddelbuettel-stokely.Rnw
another incomplete round of edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@703 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Before we can create a new Protocol Buffer Message or parse a
-serialized stream of bytes as a Message, we must read in the message
+Before one can create a new Protocol Buffer Message or parse a
+serialized stream of bytes as a Message, one must first read in the message
 type specification from a \texttt{.proto} file.
@@
 <<>>=
-ls( "RProtoBuf:DescriptorPool" )
+ls("RProtoBuf:DescriptorPool")
 @
@@
 <<>>=
-p1 <- new( tutorial.Person )
-p <- new( tutorial.Person, name = "Romain", id = 1 )
+p1 <- new(tutorial.Person)
+p <- new(tutorial.Person, name = "Murray", id = 1)
 @
@@
 p$id
-p$email <- "francoisromain@free.fr"
+p$email <- "murray@stokely.org"
 @
@@
 <<>>=
-p[["name"]] <- "Romain Francois"
+p[["name"]] <- "Murray Stokely"
 p[[ 2 ]] <- 3
@@
 <<>>=
-writeLines( as.character( p ) )
+writeLines(as.character(p))
 @
@@
 <<>>=
-serialize( p, NULL )
+serialize(p, NULL)
 @
@@
 tf1 <- tempfile()
-serialize( p, tf1 )
-readBin( tf1, raw(0), 500 )
+serialize(p, tf1)
+readBin(tf1, raw(0), 500)
 @
@@
 tf2 <- tempfile()
-con <- file( tf2, open = "wb" )
-serialize( p, con )
-close( con )
-readBin( tf2, raw(0), 500 )
+con <- file(tf2, open = "wb")
+serialize(p, con)
+close(con)
+readBin(tf2, raw(0), 500)
 @
@@
 # serialize to a file
-p$serialize( tf1 )
+p$serialize(tf1)
 # serialize to a binary connection
-con <- file( tf2, open = "wb" )
-p$serialize( con )
-close( con )
+con <- file(tf2, open = "wb")
+p$serialize(con)
+close(con)
 @
@@
 <<>>=
-msg <- read( tutorial.Person, tf1 )
-writeLines( as.character( msg ) )
+msg <- read(tutorial.Person, tf1)
+writeLines(as.character(msg))
 @
@@
 <<>>=
-con <- file( tf2, open = "rb" )
-message <- read( tutorial.Person, con )
-close( con )
-writeLines( as.character( message ) )
+con <- file(tf2, open = "rb")
+message <- read(tutorial.Person, con)
+close(con)
+writeLines(as.character(message))
 @
@@
 # reading the raw vector payload of the message
-payload <- readBin( tf1, raw(0), 5000 )
-message <- read( tutorial.Person, payload )
+payload <- readBin(tf1, raw(0), 5000)
+message <- read(tutorial.Person, payload)
 @
@@
 # reading from a file
-message <- tutorial.Person$read( tf1 )
+message <- tutorial.Person$read(tf1)
 # reading from a binary connection
-con <- file( tf2, open = "rb" )
-message <- tutorial.Person$read( con )
-close( con )
+con <- file(tf2, open = "rb")
+message <- tutorial.Person$read(con)
+close(con)
 # read from the payload
-message <- tutorial.Person$read( payload )
+message <- tutorial.Person$read(payload)
 @
@@
 the \texttt{protobuf} C++ library.
-The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp} is used to
+The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to
 facilitate the integration of the R and C++ code for these objects.
@@
 implementation.  The RProtoBuf implementation in many ways motivated
-the development of Rcpp Modules \citep{eddelbuettel2010exposing},
+the development of Rcpp Modules \citep{eddelbuettel2013exposing},
 which provide a more concise way of wrapping C++ functions and classes
@@
 The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
-of the form \verb|method( object, arguments)| and the more traditional
+of the form \verb|method(object, arguments)| and the more traditional
 object oriented notation \verb|object$method(arguments)|.

### 2014-01-04 papers/rjournal/eddelbuettel-stokely.Rnw
increment minor version; small update re RProtoBufUtils
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@709 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 based on a single catch-all \texttt{proto} schema.  Jeroen Ooms took a
-similar approach influenced by Saptarshi in his \pkg{RProtoBufUtils}
-package.  Unlike Saptarshi's package, however, RProtoBufUtils depends
-on RProtoBuf for underlying message operations.  This package is
-available at \url{https://github.com/jeroenooms/RProtoBufUtils}.
-
-The \textbf{RProtoBufUtils} package by Jereoen Ooms provides a
+similar approach influenced by Saptarshi in the \pkg{RProtoBufUtils}
+package (which has now been integrated in \pkg{RProtoBuf}).  Unlike
+Saptarshi's package, however, RProtoBufUtils depends 
+on, and extends, RProtoBuf for underlying message operations.
+
+One key extension of \pkg{RProtoBufUtils} is the 
 \texttt{serialize\_pb} method to convert R objects into serialized
-protocol buffers in this format, and the \texttt{can\_serialize\_pb}
+protocol buffers in the catch-all schema. The \texttt{can\_serialize\_pb}
 method can be used to determine whether the given R object can safely
-be expressed in this way.  To show how how this method works, we
+be expressed in this way.  To illustrate how this method works, we
 attempt to convert all of the built-in datasets from R into this
@@
 <<echo=TRUE>>=
-library(RProtoBufUtils)
-
 datasets <- subset(as.data.frame(data()$results), Package=="datasets")

### 2014-01-07 papers/rjournal/eddelbuettel-stokely.Rnw
add section on opencpu
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@724 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
-% !TeX root = RJwrapper.tex
 % We don't want a left margin for Sinput or Soutput for our table 1.
@@
 
+\section{Application: Protocol Buffers for Data Interchange in Web Services}
+
+As the name implies, the primary application of protocol buffers is
+data interchange in the context of inter-system communications. 
+Network protocols such as HTTP describe procedures on client-server
+communication, i.e. how to iniate requests, authenticate, send messages, 
+etc. However, network 
+protocols generally do not regulate \emph{content} of messages: they allow
+transfer of any media type, such as web pages, files or video.
+When designing systems where various components require exchange of specific data
+structures, we need something on top of the protocol that prescribes 
+how these structures are to be respresented in messages (buffers) on the
+network. Protocol buffers solve exactly this problem by providing
+a cross platform method for serializing arbitrary structures into well defined
+messages, that can be exchanged using any protocol. The descriptors
+(\texttt{.proto} files) are used to formally define the interface of a
+remote API or network application. Libraries to parse and generate protobuf
+messages are available for many programming languages, making it 
+relatively straight forward to implement clients and servers.
+
+\subsection{Interacting with R through HTTPS and Protocol Buffers}
+
+One example of a system that supports protocol buffers to interact
+with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
+computation and reproducible research based on R and Latex. It exposes a 
+HTTP(S) API to access and manipulate R objects and allows for performing 
+remote R function calls. Clients do not need to understand 
+or generate any R code: HTTP requests are automatically mapped to 
+function calls, and arguments/return values can be posted/retrieved
+using several data interchange formats, such as protocol buffers.  
+OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
+from the \texttt{RProtoBuf} package to convert between R objects and protobuf
+messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
+earlier to parse and generate protobuf messages when interacting with OpenCPU.
+
+\subsection{HTTP GET: Retrieving an R object}
+
+The \texttt{HTTP GET} method is used to read a resource from OpenCPU. For example,
+to access the dataset \texttt{Animals} from the package \texttt{MASS}, a 
+client performs the following HTTP request:
+
+\begin{verbatim}
+  GET https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb
+\end{verbatim}
+The postfix \texttt{/pb} in the URL tells the server to send this
+object in the form of a protobuf message. Alternative formats include 
+\texttt{/json}, \texttt{/csv}, \texttt{/rds} and others. If the request
+is successful, OpenCPU returns the serialized object with HTTP status 
+code 200 and HTTP response header \texttt{Content-Type: application/x-protobuf}. 
+The latter is the conventional MIME type that formally notifies the client to
+interpret the response as a protobuf message.
+
+Because both HTTP and Protocol Buffers have libraries available for many 
+languages, clients can be implemented in just a few lines of code. Below
+example code for both R and Python that retrieve a dataset from R with 
+OpenCPU using a protobuf message. In R, we use the HTTP client from 
+the \texttt{httr} package \citep{httr}, and the protobuf
+parser from the \texttt{RProtoBuf} package. In this illustrative example we
+download a dataset which is part of the base R distribution, so we can actually
+verify that the object was transferred without loss of information.
+
+<<eval=FALSE>>=
+# Load packages
+library(RProtoBuf)
+library(httr)
+
+# Retrieve and parse message
+req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+output <- unserialize_pb(req$content)
+
+# Check that no information was lost
+identical(output, MASS::Animals)
+@
+This code suggests a method for exchanging objects between R servers, however this can 
+also be done without protocol buffers. The main advantage of using an inter-operable format 
+is that we can actually access R objects from within another
+programming language. For example, in a very similar fasion we can retrieve the same
+dataset in a Python client. To parse messages in Python, we first compile the 
+\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:
+
+\begin{verbatim}
+  protoc rexp.proto --python_out=.
+\end{verbatim}
+This generates python module called \texttt{rexp\_pb2.py}, containing both the 
+descriptor information as well as methods to read and manipulate the R object 
+message. In the example below we use the HTTP client from the \texttt{urllib2}
+module.
+
+\begin{verbatim}
+# Import modules
+import urllib2
+from rexp_pb2 import REXP
+
+# Retrieve message
+req = urllib2.Request('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+res = urllib2.urlopen(req)
+
+# Parse rexp.proto message
+msg = REXP()
+msg.ParseFromString(res.read())
+print(msg)
+\end{verbatim}
+The \texttt{msg} object contains all data from the Animals dataset. From here we
+can easily extract the desired fields for further use in Python.
+
+\subsection{HTTP POST: Calling an R function}
+
+The example above shows how the \texttt{HTTP GET} method retrieves a 
+resource from OpenCPU, for example an R object. The \texttt{HTTP POST} 
+method on the other hand is used for calling functions and running scripts, 
+which is the primary purpose of the framework. As before, the \texttt{/pb} 
+postfix requests to retrieve the output as a protobuf message, in this
+case the function return value. However, OpenCPU allows us to supply the
+arguments of the function call in the form of protobuf messages as well.
+This is a bit more work, because clients needs to both generate messages 
+containing R objects to post to the server, as well as retrieve and parse
+protobuf messages returned by the server. Using protocol buffers to post
+function arguments is not required, and for simple (scalar) arguments 
+the standard \texttt{appliation/www-url-encoded} format might be sufficient.
+However, with protocol buffers the client can perform function calls with
+more complex arguments such as R vectors or lists. The result is a complete
+RPC system to do arbitrary R function calls from within 
+any programming language.
+
+The following example R client code performs the remote function call 
+\texttt{stats::rnorm(n=42, mean=100)}. The function arguments (in this
+case \texttt{n} and \texttt{mean}) as well as the return value (a vector
+with 42 random numbers) are transferred using a protobuf message. RPC in
+OpenCPU works like the \texttt{do.call} function in R, hence all arguments
+are contained within a list.
+
+<<>>=
+#requires httr >= 0.2.99
+library(httr)
+library(RProtoBuf)
+
+args <- list(n=42, mean=100)
+payload <- serialize_pb(args, NULL)
+
+req <- POST (
+  url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
+  body = payload,
+  add_headers (
+    "Content-Type" = "application/x-protobuf"
+  )
+)
+
+#This is the output of stats::rnorm(n=42, mean=100)
+output <- unserialize_pb(req$content)
+print(output)
+@
+The OpenCPU server basically performs the following steps to process the above RPC request:
+
+<<eval=FALSE>>=
+fnargs <- unserialize_pb(inputmsg)
+val <- do.call(stats::rnorm, fnargs)
+outputmsg <- serialize_pb(val)
+@
+In reality the OpenCPU provides a lot of meta functionality such as handling
+of sessions, exceptions, security, and much more. OpenCPU also makes it possible to store
+output of a function call on the server, instead of directly retrieving it. Thereby 
+objects can be shared with other users or used as arguments in a subsequent
+function call. But in its essence, the HTTP API provides a simple way to perform remote 
+R function calls over HTTPS. The same request can be performed in Python as follows:
+
+\begin{verbatim}
+import urllib2;
+from rexp_pb2 import *;
+
+#create the post payload, i.e. list(n=42, mean=100)
+payload = REXP(
+  rclass = 5,
+    rexpValue = [
+      REXP(rclass = 2, realValue = [42]), 
+      REXP(rclass = 2, realValue = [100])
+    ],
+    attrName = [
+      "names"
+    ],
+    attrValue = [
+      REXP(rclass = 0, stringValue = [STRING(strval="n"), STRING(strval="mean")])
+    ]
+);
+
+#HTTP POST
+req = urllib2.Request(
+  "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb", 
+  data = payload.SerializeToString(), 
+  headers = {
+    'Content-type': 'application/x-protobuf'
+  }
+)
+res = urllib2.urlopen(req);
+
+#parse output pb
+msg = REXP();
+msg.ParseFromString(res.read());
+
+#the return value is a double vector in this case
+print(msg.realValue);
+\end{verbatim}
+
 \section{Summary}

### 2014-01-09 papers/rjournal/eddelbuettel-stokely.Rnw
Comment out an unneeded section now obviated by Jereoen's OpenCPU section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@725 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Application: Distributed Data Collection with MapReduce}
-
-TODO(mstokely): Make this better.
-
-Many large data sets in fields such as particle physics and
-information processing are stored in binned or histogram form in order
-to reduce the data storage requirements
-\citep{scott2009multivariate}. Protocol Buffers make a particularly
-good data transport format in distributed MapReduces environments
-where large numbers of computers process a large data set for analysis.
-
-There are two common patterns for generating histograms of large data
-sets with MapReduce.  In the first method, each mapper task can
-generate a histogram over a subset of the data that is has been
-assigned, and then the histograms of each mapper are sent to one or
-more reducer tasks to merge.
-
-In the second method, each mapper rounds a data point to a bucket
-width and outputs that bucket as a key and '1' as a value.  Reducers
-then sum up all of the values with the same key and output to a data store.
-
-In both methods, the mapper tasks must choose identical
-bucket boundaries even though they are analyzing disjoint parts of the
-input set that may cover different ranges, or we must implement
-multiple phases.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
-\end{center}
-\caption{Diagram of MapReduce Histogram Generation Pattern}
-\label{fig:mr-histogram-pattern1}
-\end{figure}
-
-Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
-described above for histogram generation of large data sets with
-MapReduce.
-
-This package is designed to be helpful if some of the Map or Reduce
-tasks are written in R, or if those components are written in other
-languages and only the resulting output histograms need to be
-manipulated in R.
-
-\section{Application: Sending/receiving Interaction With Servers}
-
-Unlike Apache Thrift, Protocol Buffers do not include a concrete RPC
-implementation.  However, serialized protocol buffers can trivially be
-sent over TCP or integrated with a proprietary RPC system.  Combined
-with an RPC system this means that one can interactively craft request
-messages, send the serialized message to a remote server, read back a
-response, and then parse the response protocol buffer interactively.
-
-TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
-Chu's Poly.
+\include{app-mapreduce}
+
+%\section{Application: Sending/receiving Interaction With Servers}
+%
+%Combined
+%with an RPC system this means that one can interactively craft request
+%messages, send the serialized message to a remote server, read back a
+%response, and then parse the response protocol buffer interactively.
+
+%TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
+%Chu's Poly.
 

### 2014-01-09 papers/rjournal/eddelbuettel-stokely.Rnw
Fix some typos / add misisng articles in Jereoen's new OpenCPU section. Move this example before the MapReduce one.
It uses the serialize_pb method but we haven't described that here.  A
short introduction of that concept is needed for the application.

I think some of the first paragraph here should be moved to the
introduction so this section can focus only on an application.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@726 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-07 papers/rjournal/eddelbuettel-stokely.Rnw
@@
 
-\section{Application: Distributed Data Collection with MapReduce}
-
-TODO(mstokely): Make this better.
-
-Many large data sets in fields such as particle physics and
-information processing are stored in binned or histogram form in order
-to reduce the data storage requirements
-\citep{scott2009multivariate}. Protocol Buffers make a particularly
-good data transport format in distributed MapReduces environments
-where large numbers of computers process a large data set for analysis.
-
-There are two common patterns for generating histograms of large data
-sets with MapReduce.  In the first method, each mapper task can
-generate a histogram over a subset of the data that is has been
-assigned, and then the histograms of each mapper are sent to one or
-more reducer tasks to merge.
-
-In the second method, each mapper rounds a data point to a bucket
-width and outputs that bucket as a key and '1' as a value.  Reducers
-then sum up all of the values with the same key and output to a data store.
-
-In both methods, the mapper tasks must choose identical
-bucket boundaries even though they are analyzing disjoint parts of the
-input set that may cover different ranges, or we must implement
-multiple phases.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
-\end{center}
-\caption{Diagram of MapReduce Histogram Generation Pattern}
-\label{fig:mr-histogram-pattern1}
-\end{figure}
-
-Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
-described above for histogram generation of large data sets with
-MapReduce.
-
-This package is designed to be helpful if some of the Map or Reduce
-tasks are written in R, or if those components are written in other
-languages and only the resulting output histograms need to be
-manipulated in R.
-
-\section{Application: Sending/receiving Interaction With Servers}
-
-Unlike Apache Thrift, Protocol Buffers do not include a concrete RPC
-implementation.  However, serialized protocol buffers can trivially be
-sent over TCP or integrated with a proprietary RPC system.  Combined
-with an RPC system this means that one can interactively craft request
-messages, send the serialized message to a remote server, read back a
-response, and then parse the response protocol buffer interactively.
-
-TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
-Chu's Poly.
-
-\section{Application: Protocol Buffers for Data Interchange in Web Services}
-
-As the name implies, the primary application of protocol buffers is
+\section{Application: Data Interchange in Web Services}
+
+% TODO(jeroen): I think maybe some of this should go earlier in the
+% paper, so this part can focus only on introducing the application,
+% Can you integrate some of this text earlier, maybe into the the
+% introduction?
+
+As described earlier, the primary application of protocol buffers is
 data interchange in the context of inter-system communications. 
-Network protocols such as HTTP describe procedures on client-server
-communication, i.e. how to iniate requests, authenticate, send messages, 
-etc. However, network 
+Network protocols such as HTTP provide mechanisms for client-server
+communication, i.e. how to initiate requests, authenticate, send messages, 
+etc.  However, many network 
 protocols generally do not regulate \emph{content} of messages: they allow
@@
 When designing systems where various components require exchange of specific data
-structures, we need something on top of the protocol that prescribes 
+structures, we need something on top of the network protocol that prescribes 
 how these structures are to be respresented in messages (buffers) on the
@@
 with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
-computation and reproducible research based on R and Latex. It exposes a 
+computation and reproducible research based on R and \LaTeX. It exposes a 
 HTTP(S) API to access and manipulate R objects and allows for performing 
@@
 languages, clients can be implemented in just a few lines of code. Below
-example code for both R and Python that retrieve a dataset from R with 
+is example code for both R and Python that retrieves a dataset from R with 
 OpenCPU using a protobuf message. In R, we use the HTTP client from 
-the \texttt{httr} package \citep{httr}, and the protobuf
-parser from the \texttt{RProtoBuf} package. In this illustrative example we
-download a dataset which is part of the base R distribution, so we can actually
+the \texttt{httr} package \citep{httr}.
+% superfluous?
+%, and the protobuf parser from the \texttt{RProtoBuf} package.
+In this example we
+download a dataset which is part of the base R distribution, so we can
 verify that the object was transferred without loss of information.
@@
 is that we can actually access R objects from within another
-programming language. For example, in a very similar fasion we can retrieve the same
+programming language. For example, in a very similar fashion we can retrieve the same
 dataset in a Python client. To parse messages in Python, we first compile the 
@@
 
+\section{Application: Distributed Data Collection with MapReduce}
+
+TODO(mstokely): Make this better.
+
+Many large data sets in fields such as particle physics and
+information processing are stored in binned or histogram form in order
+to reduce the data storage requirements
+\citep{scott2009multivariate}. Protocol Buffers make a particularly
+good data transport format in distributed MapReduces environments
+where large numbers of computers process a large data set for analysis.
+
+There are two common patterns for generating histograms of large data
+sets with MapReduce.  In the first method, each mapper task can
+generate a histogram over a subset of the data that is has been
+assigned, and then the histograms of each mapper are sent to one or
+more reducer tasks to merge.
+
+In the second method, each mapper rounds a data point to a bucket
+width and outputs that bucket as a key and '1' as a value.  Reducers
+then sum up all of the values with the same key and output to a data store.
+
+In both methods, the mapper tasks must choose identical
+bucket boundaries even though they are analyzing disjoint parts of the
+input set that may cover different ranges, or we must implement
+multiple phases.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
+\end{center}
+\caption{Diagram of MapReduce Histogram Generation Pattern}
+\label{fig:mr-histogram-pattern1}
+\end{figure}
+
+Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
+described above for histogram generation of large data sets with
+MapReduce.
+
+This package is designed to be helpful if some of the Map or Reduce
+tasks are written in R, or if those components are written in other
+languages and only the resulting output histograms need to be
+manipulated in R.
+
+%\section{Application: Sending/receiving Interaction With Servers}
+%
+%Combined
+%with an RPC system this means that one can interactively craft request
+%messages, send the serialized message to a remote server, read back a
+%response, and then parse the response protocol buffer interactively.
+
+%TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
+%Chu's Poly.
+
 \section{Summary}

### 2014-01-09 papers/jss/article.tex
Check in the new JSS article, and rewrite/improve my MapReduce example application section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@729 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-09 papers/rjournal/eddelbuettel-stokely.Rnw
@@
+\documentclass[article]{jss}
+\usepackage{booktabs}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%
+% Local helpers to make this more compatible with R Journal style.
+%
+\newcommand{\CRANpkg}[1]{\pkg{#1}}
+\RequirePackage{fancyvrb}
+\RequirePackage{alltt}
+\DefineVerbatimEnvironment{example}{Verbatim}{}
+
+%% almost as usual
+\author{Dirk Eddelbuettel\\Debian and R Projects \And 
+        Murray Stokely\\Google, Inc}
+\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}
+
+%% for pretty printing and a nice hypersummary also set:
+\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
+\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
+\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)
+
+%% an abstract and keywords
+\Abstract{
+Modern data collection and analysis pipelines often involve
+a sophisticated mix of applications written in general purpose and
+specialized programming languages.  Protocol Buffers are a popular
+method of serializing structured data between applications---while remaining
+independent of programming languages or operating system.  The
+\CRANpkg{RProtoBuf} package provides a complete interface to this
+library.
+}
+\Keywords{r, protocol buffers, serialization, cross-platform}
+\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
+%% at least one keyword must be supplied
+
+%% publication information
+%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{2012-06-04}
+%% \Acceptdate{2012-06-04}
+
+%% The address of (at least) one author should be given
+%% in the following format:
+\Address{
+  Dirk Eddelbuettel\\
+  \\
+  Murray Stokely\\
+  Google, Inc.\\
+  1600 Amphitheatre Parkway\\
+  Mountain View, CA 94040\\
+  USA\\
+  E-mail: \email{mstokely@google.com}\\
+  URL: \url{http://www.stokely.org/}
+}
+%% It is also possible to add a telephone and fax number
+%% before the e-mail in the following format:
+%% Telephone: +43/512/507-7103
+%% Fax: +43/512/507-2851
+
+%% for those who use Sweave please include the following line (with % symbols):
+%% need no \usepackage{Sweave.sty}
+
+%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+%% include your article here, just as usual
+%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
+
 % We don't want a left margin for Sinput or Soutput for our table 1.
@@
 %%     guarantees better line breaks
-<<echo=FALSE,print=FALSE>>=
-options(width=65, prompt="R> ", digits=4)
-@
 
@@
 \end{minipage} & \begin{minipage}{.5\textwidth}
-<<echo=TRUE>>=
-library(RProtoBuf)
-p <- new(tutorial.Person, id=1, name="Dirk")
-class(p)
-p$name
-p$name <- "Murray"
-cat(as.character(p))
-serialize(p, NULL)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library(RProtoBuf)
+R> p <- new(tutorial.Person, id=1, name="Dirk")
+R> class(p)
+\end{Sinput}
+\begin{Soutput}
+[1] "Message"
+attr(,"package")
+[1] "RProtoBuf"
+\end{Soutput}
+\begin{Sinput}
+R> p$name
+\end{Sinput}
+\begin{Soutput}
+[1] "Dirk"
+\end{Soutput}
+\begin{Sinput}
+R> p$name <- "Murray"
+R> cat(as.character(p))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray"
+id: 1
+\end{Soutput}
+\begin{Sinput}
+R> serialize(p, NULL)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 06 4d 75 72 72 61 79 10 01
+\end{Soutput}
+\end{Schunk}
 \end{minipage} \\
@@
 
-<<>>=
-ls("RProtoBuf:DescriptorPool")
-@
+\begin{Schunk}
+\begin{Sinput}
+R> ls("RProtoBuf:DescriptorPool")
+\end{Sinput}
+\begin{Soutput}
+ [1] "rexp.CMPLX"                  
+ [2] "rexp.REXP"                   
+ [3] "rexp.STRING"                 
+ [4] "rprotobuf.HelloWorldRequest" 
+ [5] "rprotobuf.HelloWorldResponse"
+ [6] "tutorial.AddressBook"        
+ [7] "tutorial.Person"             
+ [8] "tutorial.Test1"              
+ [9] "tutorial.Test2"              
+[10] "tutorial.Test3"              
+[11] "tutorial.Test4"              
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-p1 <- new(tutorial.Person)
-p <- new(tutorial.Person, name = "Murray", id = 1)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> p1 <- new(tutorial.Person)
+R> p <- new(tutorial.Person, name = "Murray", id = 1)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-p$name
-p$id
-p$email <- "murray@stokely.org"
-@
+\begin{Schunk}
+\begin{Sinput}
+R> p$name
+\end{Sinput}
+\begin{Soutput}
+[1] "Murray"
+\end{Soutput}
+\begin{Sinput}
+R> p$id
+\end{Sinput}
+\begin{Soutput}
+[1] 1
+\end{Soutput}
+\begin{Sinput}
+R> p$email <- "murray@stokely.org"
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-p[["name"]] <- "Murray Stokely"
-p[[ 2 ]] <- 3
-p[[ "email" ]]
-@
+\begin{Schunk}
+\begin{Sinput}
+R> p[["name"]] <- "Murray Stokely"
+R> p[[ 2 ]] <- 3
+R> p[[ "email" ]]
+\end{Sinput}
+\begin{Soutput}
+[1] "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-p
-@
+\begin{Schunk}
+\begin{Sinput}
+R> p
+\end{Sinput}
+\begin{Soutput}
+[1] "message of type 'tutorial.Person' with 3 fields set"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-writeLines(as.character(p))
-@
+\begin{Schunk}
+\begin{Sinput}
+R> writeLines(as.character(p))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray Stokely"
+id: 3
+email: "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-serialize(p, NULL)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> serialize(p, NULL)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
+[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-tf1 <- tempfile()
-serialize(p, tf1)
-readBin(tf1, raw(0), 500)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tf1 <- tempfile()
+R> serialize(p, tf1)
+R> readBin(tf1, raw(0), 500)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
+[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-tf2 <- tempfile()
-con <- file(tf2, open = "wb")
-serialize(p, con)
-close(con)
-readBin(tf2, raw(0), 500)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tf2 <- tempfile()
+R> con <- file(tf2, open = "wb")
+R> serialize(p, con)
+R> close(con)
+R> readBin(tf2, raw(0), 500)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12
+[21] 6d 75 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-# serialize to a file
-p$serialize(tf1)
-# serialize to a binary connection
-con <- file(tf2, open = "wb")
-p$serialize(con)
-close(con)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> # serialize to a file
+R> p$serialize(tf1)
+R> # serialize to a binary connection
+R> con <- file(tf2, open = "wb")
+R> p$serialize(con)
+R> close(con)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-msg <- read(tutorial.Person, tf1)
-writeLines(as.character(msg))
-@
+\begin{Schunk}
+\begin{Sinput}
+R> msg <- read(tutorial.Person, tf1)
+R> writeLines(as.character(msg))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray Stokely"
+id: 3
+email: "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-con <- file(tf2, open = "rb")
-message <- read(tutorial.Person, con)
-close(con)
-writeLines(as.character(message))
-@
+\begin{Schunk}
+\begin{Sinput}
+R> con <- file(tf2, open = "rb")
+R> message <- read(tutorial.Person, con)
+R> close(con)
+R> writeLines(as.character(message))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray Stokely"
+id: 3
+email: "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-# reading the raw vector payload of the message
-payload <- readBin(tf1, raw(0), 5000)
-message <- read(tutorial.Person, payload)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> # reading the raw vector payload of the message
+R> payload <- readBin(tf1, raw(0), 5000)
+R> message <- read(tutorial.Person, payload)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-# reading from a file
-message <- tutorial.Person$read(tf1)
-# reading from a binary connection
-con <- file(tf2, open = "rb")
-message <- tutorial.Person$read(con)
-close(con)
-# read from the payload
-message <- tutorial.Person$read(payload)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> # reading from a file
+R> message <- tutorial.Person$read(tf1)
+R> # reading from a binary connection
+R> con <- file(tf2, open = "rb")
+R> message <- tutorial.Person$read(con)
+R> close(con)
+R> # read from the payload
+R> message <- tutorial.Person$read(payload)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-# field descriptor
-tutorial.Person$email
-
-# enum descriptor
-tutorial.Person$PhoneType
-
-# nested type descriptor
-tutorial.Person$PhoneNumber
-# same as
-tutorial.Person.PhoneNumber
-@
+\begin{Schunk}
+\begin{Sinput}
+R> # field descriptor
+R> tutorial.Person$email
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for field 'email' of type 'tutorial.Person' "
+\end{Soutput}
+\begin{Sinput}
+R> # enum descriptor
+R> tutorial.Person$PhoneType
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for enum 'PhoneType' of type 'tutorial.Person' with 3 values"
+\end{Soutput}
+\begin{Sinput}
+R> # nested type descriptor
+R> tutorial.Person$PhoneNumber
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for type 'tutorial.Person.PhoneNumber' "
+\end{Soutput}
+\begin{Sinput}
+R> # same as
+R> tutorial.Person.PhoneNumber
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for type 'tutorial.Person.PhoneNumber' "
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-tutorial.Person$PhoneType
-tutorial.Person$PhoneType$WORK
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tutorial.Person$PhoneType
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for enum 'PhoneType' of type 'tutorial.Person' with 3 values"
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$WORK
+\end{Sinput}
+\begin{Soutput}
+[1] 2
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-f <- tutorial.Person$fileDescriptor()
-f
-f$Person
-@
+\begin{Schunk}
+\begin{Sinput}
+R> f <- tutorial.Person$fileDescriptor()
+R> f
+\end{Sinput}
+\begin{Soutput}
+[1] "file descriptor for package tutorial (/usr/local/google/R/RProtoBuf/proto/addressbook.proto)"
+\end{Soutput}
+\begin{Sinput}
+R> f$Person
+\end{Sinput}
+\begin{Soutput}
+[1] "descriptor for type 'tutorial.Person' "
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-tutorial.Person$PhoneType$value(1)
-tutorial.Person$PhoneType$value(name="HOME")
-tutorial.Person$PhoneType$value(number=1)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(1)
+\end{Sinput}
+\begin{Soutput}
+[1] "enum value descriptor tutorial.Person.MOBILE"
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(name="HOME")
+\end{Sinput}
+\begin{Soutput}
+[1] "enum value descriptor tutorial.Person.HOME"
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(number=1)
+\end{Sinput}
+\begin{Soutput}
+[1] "enum value descriptor tutorial.Person.HOME"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<echo=FALSE,print=FALSE>>=
-    if (!exists("protobuf_unittest.TestAllTypes",
-                "RProtoBuf:DescriptorPool")) {
-        unittest.proto.file <- system.file("unitTests", "data",
-                                           "unittest.proto",
-                                           package="RProtoBuf")
-        readProtoFiles(file=unittest.proto.file)
-    }
-@
-
-<<>>=
-a <- new(protobuf_unittest.TestAllTypes)
-a$optional_bool <- TRUE
-a$optional_bool <- FALSE
-<<eval=F>>=
-a$optional_bool <- NA
-<<echo=FALSE,eval=TRUE,print=TRUE>>=
-try(a$optional_bool <- NA,silent=TRUE)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> a <- new(protobuf_unittest.TestAllTypes)
+R> a$optional_bool <- TRUE
+R> a$optional_bool <- FALSE
+\end{Sinput}
+\end{Schunk}
+\begin{Schunk}
+\begin{Sinput}
+R> a$optional_bool <- NA
+\end{Sinput}
+\end{Schunk}
+\begin{Schunk}
+\begin{Soutput}
+[1] "Error in eval(expr, envir, enclos) : \n  NA boolean values can not be stored in bool protocol buffer fields\n"
+attr(,"class")
+[1] "try-error"
+attr(,"condition")
+<Rcpp::exception in eval(expr, envir, enclos): NA boolean values can not be stored in bool protocol buffer fields>
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-as.integer(2^31-1)
-<<eval=F>>=
-as.integer(2^31 - 1) + as.integer(1)
-<<echo=FALSE,eval=TRUE,print=TRUE>>=
-try(as.integer(2^31 - 1) + as.integer(1))
-
-2^31
-class(2^31)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> as.integer(2^31-1)
+\end{Sinput}
+\begin{Soutput}
+[1] 2147483647
+\end{Soutput}
+\end{Schunk}
+\begin{Schunk}
+\begin{Sinput}
+R> as.integer(2^31 - 1) + as.integer(1)
+\end{Sinput}
+\end{Schunk}
+\begin{Schunk}
+\begin{Soutput}
+[1] NA
+\end{Soutput}
+\begin{Soutput}
+[1] 2.147e+09
+\end{Soutput}
+\begin{Soutput}
+[1] "numeric"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-2^53 == (2^53 + 1)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> 2^53 == (2^53 + 1)
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<echo=FALSE,print=FALSE>>=
-if (!exists("protobuf_unittest.TestAllTypes",
-            "RProtoBuf:DescriptorPool")) {
-    unittest.proto.file <- system.file("unitTests", "data",
-                                       "unittest.proto",
-                                       package="RProtoBuf")
-    readProtoFiles(file=unittest.proto.file)
-}
-@
-
 If we try to set an int64 field in R to double values, we lose
@@
 
-<<>>=
-test <- new(protobuf_unittest.TestAllTypes)
-test$repeated_int64 <- c(2^53, 2^53+1)
-length(unique(test$repeated_int64))
-@
+\begin{Schunk}
+\begin{Sinput}
+R> test <- new(protobuf_unittest.TestAllTypes)
+R> test$repeated_int64 <- c(2^53, 2^53+1)
+R> length(unique(test$repeated_int64))
+\end{Sinput}
+\begin{Soutput}
+[1] 1
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<>>=
-test$repeated_int64 <- c("9007199254740992", "9007199254740993")
-@
+\begin{Schunk}
+\begin{Sinput}
+R> test$repeated_int64 <- c("9007199254740992", "9007199254740993")
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<>>=
-options("RProtoBuf.int64AsString" = FALSE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-options("RProtoBuf.int64AsString" = TRUE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-@
-
-<<echo=FALSE,print=FALSE>>=
-options("RProtoBuf.int64AsString" = FALSE)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> options("RProtoBuf.int64AsString" = FALSE)
+R> test$repeated_int64
+\end{Sinput}
+\begin{Soutput}
+[1] 9.007e+15 9.007e+15
+\end{Soutput}
+\begin{Sinput}
+R> length(unique(test$repeated_int64))
+\end{Sinput}
+\begin{Soutput}
+[1] 1
+\end{Soutput}
+\begin{Sinput}
+R> options("RProtoBuf.int64AsString" = TRUE)
+R> test$repeated_int64
+\end{Sinput}
+\begin{Soutput}
+[1] "9007199254740992" "9007199254740993"
+\end{Soutput}
+\begin{Sinput}
+R> length(unique(test$repeated_int64))
+\end{Sinput}
+\begin{Soutput}
+[1] 2
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<echo=TRUE>>=
-datasets <- subset(as.data.frame(data()$results), Package=="datasets")
-datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
-n <- nrow(datasets)
-@
-
-There are \Sexpr{n} standard data sets included in R.  We use the
+\begin{Schunk}
+\begin{Sinput}
+R> datasets <- subset(as.data.frame(data()$results), Package=="datasets")
+R> datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
+R> n <- nrow(datasets)
+\end{Sinput}
+\end{Schunk}
+
+There are 103 standard data sets included in R.  We use the
 \texttt{can\_serialize\_pb} method to determine how many of those can
@@
 
-<<echo=TRUE>>=
-#datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
-#datasets <- subset(datasets, valid.proto==TRUE)
-m <- nrow(datasets)
-@
-
-\Sexpr{m} data sets could be converted to Protocol Buffers
-(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
+\begin{Schunk}
+\begin{Sinput}
+R> #datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
+R> #datasets <- subset(datasets, valid.proto==TRUE)
+R> m <- nrow(datasets)
+\end{Sinput}
+\end{Schunk}
+
+103 data sets could be converted to Protocol Buffers
+(100\%).  The next section illustrates how
 many bytes were usued to store the data sets under four different
@@
 \label{sec:compression}
-
-<<echo=FALSE,print=FALSE>>=
-datasets$object.size <- unname(sapply(datasets$load.name, function(x) object.size(eval(as.name(x)))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))
-
-datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))
-
-datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))
-
-clean.df <- data.frame(dataset=datasets$load.name,
-                       object.size=datasets$object.size,
-                       "serialized"=datasets$R.serialize.size,
-                       "gzipped serialized"=datasets$R.serialize.size.gz,
-                       "RProtoBuf"=datasets$RProtoBuf.serialize.size,
-                       "gzipped RProtoBuf"=datasets$RProtoBuf.serialize.size.gz,
-                       check.names=FALSE)
-@
 
@@
 
-<<eval=FALSE>>=
-# Load packages
-library(RProtoBuf)
-library(httr)
-
-# Retrieve and parse message
-req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
-output <- unserialize_pb(req$content)
-
-# Check that no information was lost
-identical(output, MASS::Animals)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> # Load packages
+R> library(RProtoBuf)
+R> library(httr)
+R> # Retrieve and parse message
+R> req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+R> output <- unserialize_pb(req$content)
+R> # Check that no information was lost
+R> identical(output, MASS::Animals)
+\end{Sinput}
+\end{Schunk}
 This code suggests a method for exchanging objects between R servers, however this can 
@@
 
-<<>>=
-#requires httr >= 0.2.99
-library(httr)
-library(RProtoBuf)
-
-args <- list(n=42, mean=100)
-payload <- serialize_pb(args, NULL)
-
-req <- POST (
-  url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
-  body = payload,
-  add_headers (
-    "Content-Type" = "application/x-protobuf"
-  )
-)
-
-#This is the output of stats::rnorm(n=42, mean=100)
-output <- unserialize_pb(req$content)
-print(output)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> #requires httr >= 0.2.99
+R> library(httr)
+R> library(RProtoBuf)
+R> args <- list(n=42, mean=100)
+R> payload <- serialize_pb(args, NULL)
+R> req <- POST (
++   url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
++   body = payload,
++   add_headers (
++     "Content-Type" = "application/x-protobuf"
++   )
++ )
+R> #This is the output of stats::rnorm(n=42, mean=100)
+R> output <- unserialize_pb(req$content)
+R> print(output)
+\end{Sinput}
+\begin{Soutput}
+ [1] 100.02 100.75  99.91  98.92  97.19  98.07 102.34  98.48
+ [9] 100.31 100.58 100.65 100.64  98.02  97.91 100.57 100.41
+[17] 100.11 103.32 100.88 100.48 101.06 100.52 101.58 101.23
+[25] 100.73 100.40  98.49 101.50  99.32  99.12  99.57  98.80
+[33]  98.60  99.83  99.62 100.43 100.24 100.30  99.13  99.56
+[41]  99.78 101.01
+\end{Soutput}
+\end{Schunk}
 The OpenCPU server basically performs the following steps to process the above RPC request:
 
-<<eval=FALSE>>=
-fnargs <- unserialize_pb(inputmsg)
-val <- do.call(stats::rnorm, fnargs)
-outputmsg <- serialize_pb(val)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> fnargs <- unserialize_pb(inputmsg)
+R> val <- do.call(stats::rnorm, fnargs)
+R> outputmsg <- serialize_pb(val)
+\end{Sinput}
+\end{Schunk}
 In reality the OpenCPU provides a lot of meta functionality such as handling
@@
 
-TODO(mstokely): Make this better.
-
-Many large data sets in fields such as particle physics and
-information processing are stored in binned or histogram form in order
-to reduce the data storage requirements
-\citep{scott2009multivariate}. Protocol Buffers make a particularly
-good data transport format in distributed MapReduces environments
-where large numbers of computers process a large data set for analysis.
-
+The MapReduce programming model \citep{dean2008mapreduce} has emerged
+in the last decade as a popular framework for processing large data
+sets in parallel on large compute clusters.  Protocol Buffers
+provide a convenient mechanism to send structured data between tasks
+in a MapReduce cluster.  In particular, the large data sets in fields
+such as particle physics and information processing are frequently
+stored in binned or histogram form in order to reduce the data storage
+requirements for later data analysis \citep{scott2009multivariate}.
+
+In such environments, analysts may be interested in very rare
+phenomenon or be dealing with highly skewed data sets or inflexible
+raw data storage systems from which unbiased sampling is not feasible.
 There are two common patterns for generating histograms of large data
-sets with MapReduce.  In the first method, each mapper task can
-generate a histogram over a subset of the data that is has been
-assigned, and then the histograms of each mapper are sent to one or
-more reducer tasks to merge.
-
-In the second method, each mapper rounds a data point to a bucket
-width and outputs that bucket as a key and '1' as a value.  Reducers
-then sum up all of the values with the same key and output to a data store.
-
-In both methods, the mapper tasks must choose identical
-bucket boundaries even though they are analyzing disjoint parts of the
-input set that may cover different ranges, or we must implement
-multiple phases.
+sets in a single pass with MapReduce.  In the first method, each
+mapper task generates a histogram over a subset of the data that it
+has been assigned, serializes this histogram and sends it to one or
+more reducer tasks which merge the intermediate histograms from the
+mappers.
+
+In the second method, illustrated in
+Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
+point to a bucket width and outputs that bucket as a key and '1' as a
+value.  Reducers then sum up all of the values with the same key and
+output to a data store.
+
+In both methods, the mapper tasks must choose identical bucket
+boundaries in advance if we are to construct the histogram in a single
+pass, even though they are analyzing disjoint parts of the input set
+that may cover different ranges.  The \pkg{HistogramTools} package
+\citep{histogramtools} enhances \pkg{RProtoBuf} by providing a concise
+schema for R histogram objects.  The histogram message type is
+designed to be helpful if some of the Map or Reduce tasks are written
+in R, or if those components are written in other languages and only
+the resulting output histograms need to be manipulated in R.
 
@@
 \end{figure}
-
-Figure~\ref{fig:mr-histogram-pattern1} illustrates the second method
-described above for histogram generation of large data sets with
-MapReduce.
-
-This package is designed to be helpful if some of the Map or Reduce
-tasks are written in R, or if those components are written in other
-languages and only the resulting output histograms need to be
-manipulated in R.
 
@@
 \bibliography{eddelbuettel-stokely}
+
+%\section[About Java]{About \proglang{Java}}
+%% Note: If there is markup in \(sub)section, then it has to be escape as above.
+
+\end{document}

### 2014-01-09 papers/jss/article.Rnw
Oops check in the Rnw rather than the generated tex file.  Also, add section \labels and add a roadmap section at the end of the introduction that highlights the main sections.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@730 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-09 papers/rjournal/eddelbuettel-stokely.Rnw
@@
+\documentclass[article]{jss}
+\usepackage{booktabs}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%
+% Local helpers to make this more compatible with R Journal style.
+%
+\newcommand{\CRANpkg}[1]{\pkg{#1}}
+\RequirePackage{fancyvrb}
+\RequirePackage{alltt}
+\DefineVerbatimEnvironment{example}{Verbatim}{}
+
+%% almost as usual
+\author{Dirk Eddelbuettel\\Debian and R Projects \And 
+        Murray Stokely\\Google, Inc}
+\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}
+
+%% for pretty printing and a nice hypersummary also set:
+\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
+\Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
+\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)
+
+%% an abstract and keywords
+\Abstract{
+Modern data collection and analysis pipelines often involve
+a sophisticated mix of applications written in general purpose and
+specialized programming languages.  Protocol Buffers are a popular
+method of serializing structured data between applications---while remaining
+independent of programming languages or operating system.  The
+\CRANpkg{RProtoBuf} package provides a complete interface between this
+library and the R environment for statistical computing.
+%TODO(ms) keep it less than 150 words.
+}
+\Keywords{r, protocol buffers, serialization, cross-platform}
+\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
+%% at least one keyword must be supplied
+
+%% publication information
+%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
+%% \Volume{50}
+%% \Issue{9}
+%% \Month{June}
+%% \Year{2012}
+%% \Submitdate{2012-06-04}
+%% \Acceptdate{2012-06-04}
+
+%% The address of (at least) one author should be given
+%% in the following format:
+\Address{
+  Dirk Eddelbuettel\\
+  \\
+  Murray Stokely\\
+  Google, Inc.\\
+  1600 Amphitheatre Parkway\\
+  Mountain View, CA 94040\\
+  USA\\
+  E-mail: \email{mstokely@google.com}\\
+  URL: \url{http://www.stokely.org/}
+}
+%% It is also possible to add a telephone and fax number
+%% before the e-mail in the following format:
+%% Telephone: +43/512/507-7103
+%% Fax: +43/512/507-2851
+
+%% for those who use Sweave please include the following line (with % symbols):
+%% need no \usepackage{Sweave.sty}
+
+%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+
+%% include your article here, just as usual
+%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
+
 % We don't want a left margin for Sinput or Soutput for our table 1.
@@
 \maketitle
-
-\abstract{Modern data collection and analysis pipelines often involve
- a sophisticated mix of applications written in general purpose and
- specialized programming languages.  Protocol Buffers are a popular
- method of serializing structured data between applications---while remaining
- independent of programming languages or operating system.  The
- \CRANpkg{RProtoBuf} package provides a complete interface between this
- library and the R environment for statistical computing.
- %TODO(ms) keep it less than 150 words.
-}
 
@@
 % applications at the end.
-This article describes the basics of Google's Protocol Buffers through
-an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
-basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
-several common use cases for protocol buffers in data analysis.
+
+Section~\ref{sec:protobuf} provides a general overview of Protocol
+Buffers.  Section~\ref{sec:rprotobuf-basic} describes the interactive
+R interface provided by \CRANpkg{RProtoBuf} and introduces the two
+main abstractions: \emph{Messages} and \emph{Descriptors}.
+Section~\ref{sec:rprotobuf-classes} describes the implementation
+details of the main S4 classes making up this package.
+Section~\ref{sec:types} describes the challenges of type coercion
+between R and other languages.  Section~\ref{sec:evaluation}
+introduces a general R language schema for serializing arbitrary R
+objects and evaluates it against R's built-in serialization.
+Sections~\label{sec:opencpu} and \label{sec:mapreduce} provide
+real-world use cases of \CRANpkg{RProtoBuf} in web service and
+MapReduce environments, respectively.
+
+%This article describes the basics of Google's Protocol Buffers through
+%an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
+%basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
+%several common use cases for protocol buffers in data analysis.
 
 \section{Protocol Buffers}
-
-FIXME Introductory section which may include references in parentheses
-\citep{R}, or cite a reference such as \citet{R} in the text.
+\label{sec:protobuf}
+
+%FIXME Introductory section which may include references in parentheses
+%\citep{R}, or cite a reference such as \citet{R} in the text.
 
@@
 \section{Basic Usage: Messages and Descriptors}
+\label{sec:rprotobuf-basic}
 
@@
 \section{Under the hood: S4 Classes, Methods, and Pseudo Methods}
+\label{sec:rprotobuf-classes}
 
@@
 \section{Type Coercion}
+\label{sec:types}
 
@@
 \section{Evaluation: data.frame to Protocol Buffer Serialization}
+\label{sec:evaluation}
 
@@
 \section{Application: Data Interchange in Web Services}
+\label{sec:opencpu}
 
@@
 \section{Application: Distributed Data Collection with MapReduce}
-
-TODO(mstokely): Make this better.
-
-Many large data sets in fields such as particle physics and
-information processing are stored in binned or histogram form in order
-to reduce the data storage requirements
-\citep{scott2009multivariate}. Protocol Buffers make a particularly
-good data transport format in distributed MapReduces environments
-where large numbers of computers process a large data set for analysis.
-
+\label{sec:mapreduce}
+
+The MapReduce programming model \citep{dean2008mapreduce} has emerged
+in the last decade as a popular framework for processing large data
+sets in parallel on large compute clusters.  Protocol Buffers
+provide a convenient mechanism to send structured data between tasks
+in a MapReduce cluster.  In particular, the large data sets in fields
+such as particle physics and information processing are frequently
+stored in binned or histogram form in order to reduce the data storage
+requirements for later data analysis \citep{scott2009multivariate}.
+
+In such environments, analysts may be interested in very rare
+phenomenon or be dealing with highly skewed data sets or inflexible
+raw data storage systems from which unbiased sampling is not feasible.
 There are two common patterns for generating histograms of large data
-sets with MapReduce.  In the first method, each mapper task can
-generate a histogram over a subset of the data that is has been
-assigned, and then the histograms of each mapper are sent to one or
-more reducer tasks to merge.
-
-In the second method, each mapper rounds a data point to a bucket
-width and outputs that bucket as a key and '1' as a value.  Reducers
-then sum up all of the values with the same key and output to a data store.
-
-In both methods, the mapper tasks must choose identical
-bucket boundaries even though they are analyzing disjoint parts of the
-input set that may cover different ranges, or we must implement
-multiple phases.
+sets in a single pass with MapReduce.  In the first method, each
+mapper task generates a histogram over a subset of the data that it
+has been assigned, serializes this histogram and sends it to one or
+more reducer tasks which merge the intermediate histograms from the
+mappers.
+
+In the second method, illustrated in
+Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
+point to a bucket width and outputs that bucket as a key and '1' as a
+value.  Reducers then sum up all of the values with the same key and
+output to a data store.
+
+In both methods, the mapper tasks must choose identical bucket
+boundaries in advance if we are to construct the histogram in a single
+pass, even though they are analyzing disjoint parts of the input set
+that may cover different ranges.  The \pkg{HistogramTools} package
+\citep{histogramtools} enhances \pkg{RProtoBuf} by providing a concise
+schema for R histogram objects.  The histogram message type is
+designed to be helpful if some of the Map or Reduce tasks are written
+in R, or if those components are written in other languages and only
+the resulting output histograms need to be manipulated in R.
 

### 2014-01-09 papers/jss/article.Rnw
small edits paper does NOT 'compile' without network connectivity which is BAD BAD BAD BAD
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@734 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 needlessly duplicate the field names with each message.
+Lastly, XML is a well-established and widely-supported protocol with the ability to define
+just about any arbitrarily complex schema. However, it pays for this
+complexity with comparatively large and verbose messages, and added
+complexities at the parsing side. 
 %
@@
 reduce the parsing cost and improve the efficiency.  MessagePack
-\citep{msgpackR} and BSON \citep{rmongodb} both have R interfaces, but
+and BSON both have R interfaces \citep{msgpackR,rmongodb}, but
+% DE Why do we cite these packages, but not the numerous JSON packages?
 these formats lack a separate schema for the serialized data and thus
@@
 requirement changes dictate update to the message format.
-% DE: Need to talk about XML ?
+% DE: Need to talk about XML -- added a few lines at previous paragraph
 
@@
 a$optional_bool <- FALSE
-<<eval=F>>=
+<<eval=FALSE>>=
 a$optional_bool <- NA
@@
 as.integer(2^31-1)
-<<eval=F>>=
+<<eval=FALSE>>=
 as.integer(2^31 - 1) + as.integer(1)
@@
 @
+
 This code suggests a method for exchanging objects between R servers, however this can 
@@
 
-<<>>=
+<<eval=FALSE>>=
 #requires httr >= 0.2.99

### 2014-01-11 papers/jss/article.Rnw
a few fixes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@742 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The \verb|[[| operator can also be used to query and set fields
-of a mesages, supplying either their name or their tag number :
+of a messages, supplying either their name or their tag number :
 
@@
 Table~\ref{Descriptor-methods-table} provides a complete list of the
-slots and avalailable methods for Descriptors.
+slots and available methods for Descriptors.
 
@@
 the most appropriate type in different environments.
-Table~\ref{table-get-types} details the correspondance between the
+Table~\ref{table-get-types} details the correspondence between the
 field type and the type of data that is retrieved by \verb|$| and \verb|[[|
@@
 \end{small}
-\caption{\label{table-get-types}Correspondance between field type and
+\caption{\label{table-get-types}Correspondence between field type and
   R type retrieved by the extractors. \footnotesize{1. R lacks native
@@
 (\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
-many bytes were usued to store the data sets under four different
+many bytes were used to store the data sets under four different
 situations (1) normal R serialization, (2) R serialization followed by
@@
 structures, we need something on top of the network protocol that prescribes 
-how these structures are to be respresented in messages (buffers) on the
+how these structures are to be represented in messages (buffers) on the
 network. Protocol buffers solve exactly this problem by providing
@@
 function arguments is not required, and for simple (scalar) arguments 
-the standard \texttt{appliation/www-url-encoded} format might be sufficient.
+the standard \texttt{application/www-url-encoded} format might be sufficient.
 However, with protocol buffers the client can perform function calls with

### 2014-01-11 papers/jss/article.Rnw
another one
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@743 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
fix
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@746 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   representation for transmission or storage.
-\item \emph{Exentsible}:  New fields can be added to Protocol Buffer Schemas
+\item \emph{Extensible}:  New fields can be added to Protocol Buffer Schemas
   in a forward-compatible way that do not break older applications.
@@
 function arguments is not required, and for simple (scalar) arguments 
-the standard \texttt{application/www-url-encoded} format might be sufficient.
+the standard \texttt{application/x-www-form-urlencoded} format might be sufficient.
 However, with protocol buffers the client can perform function calls with

### 2014-01-11 papers/jss/article.Rnw
rename bib file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@749 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %% almost as usual
-\author{Dirk Eddelbuettel\\Debian and R Projects \And 
-        Murray Stokely\\Google, Inc}
+\author{Dirk Eddelbuettel\\Debian Project \And 
+        Murray Stokely\\Google, Inc \And
+        Jeroen Ooms\\UCLA}
 \title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}
@@
 \Address{
-  Dirk Eddelbuettel\\
+  Dirk Eddelbuettel \\
+  Debian Project \\
+  River Forest, IL, USA\\
+  E-mail: \email{edd@debian.org}\\
+  URL: \url{http://dirk.eddelbuettel.com}\\
   \\
@@
   E-mail: \email{mstokely@google.com}\\
-  URL: \url{http://www.stokely.org/}
+  URL: \url{http://www.stokely.org/}\\
+  \\
+  Jeroen Ooms\\
+  UCLA Department of Statistics\\
+  University of California\\
+  E-mail: \email{jeroen.ooms@stat.ucla.edu}\\
+  URL: \url{http://jeroenooms.github.io}
 }
@@
 \centering
-\begin{tabular}{|l|c|c|l|}
-\hline
-\textbf{Class} & \textbf{Slots} & \textbf{Methods} & \textbf{Dynamic Dispatch}\\
-\hline
-\hline
+\begin{tabular}{lccl}
+\toprule
+\textbf{Class}      & 
+     \textbf{Slots} & 
+     \textbf{Methods} & 
+     \textbf{Dynamic Dispatch}\\
+\cmidrule{1-4}
 Message             & 2 & 20 & yes (field names)\\
-\hline
 Descriptor          & 2 & 16 & yes (field names, enum types, nested types)\\
-\hline
 FieldDescriptor     & 4 & 18 & no\\
-\hline
 EnumDescriptor      & 4 & 11 & yes (enum constant names)\\
-\hline
-FileDescriptor & 3 & 6 & yes (message/field definitions)\\
-\hline
-EnumValueDescriptor & 3 & 6 & no\\
-\hline
+FileDescriptor      & 3 & \phantom{1}6 & yes (message/field definitions)\\
+EnumValueDescriptor & 3 & \phantom{1}6 & no\\
+\bottomrule
 \end{tabular}
+\caption{\label{Message-methods-table}Overview of Class, Slot, Method and
+  Dispatch Relationships}
 \end{table}
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
 \textbf{Slot} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{pointer} & External pointer to the \texttt{Message} object of the C++ proto library. Documentation for the
@@
 \url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message} \\
-\hline
 \texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
 has its \texttt{type} slot set to \texttt{tutorial.Person} \\[.3cm]
-\hline
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{has} & Indicates if a message has a given field.   \\
@@
 \texttt{size} & The number of elements in a message field\\
-\texttt{bytesize} & The number of bytes the message would take once serialized\\
-\hline
+\texttt{bytesize} & The number of bytes the message would take once serialized\\[3mm]
+%
 \texttt{swap} & swap elements of a repeated field of a message\\
@@
 \texttt{getExtension} & get the value of an extension of a message\\
-\texttt{add} & add elements to a repeated field \\
-\hline
+\texttt{add} & add elements to a repeated field \\[3mm]
+%
 \texttt{str} & the R structure of the message\\
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
 \textbf{Slot} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
@@
 \url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
-\hline
 \texttt{type} & Fully qualified path of the message type. \\[.3cm]
-\hline
+%
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{new} & Creates a prototype of a message described by this descriptor.\\
@@
 text connection.\\
-\hline
 \texttt{name} & Retrieve the name of the message type associated with
@@
 \texttt{asMessage} & return DescriptorProto message. \\
-\hline
 \texttt{fileDescriptor} & Retrieve the file descriptor of this
@@
 type in this descriptor.\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
 \textbf{Slot} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{pointer} & External pointer to the \texttt{FieldDescriptor} C++ variable \\
-\hline
 \texttt{name} & Simple name of the field \\
-\hline
 \texttt{full\_name} & Fully qualified name of the field \\
-\hline
 \texttt{type} & Name of the message type where the field is declared \\[.3cm]
-\hline
+%
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{as.character} & Character representation of a descriptor\\
-\texttt{toString} & Character
-representation of a descriptor (same as \texttt{as.character}) \\
+\texttt{toString} & Character representation of a descriptor (same as \texttt{as.character}) \\
 \texttt{asMessage} & Return FieldDescriptorProto message. \\
@@
 \texttt{enum\_type} & Return the enum type if this is an enum type field.\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
 \textbf{Slot} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{pointer} & External pointer to the \texttt{EnumDescriptor} C++ variable \\
-\hline
 \texttt{name} & Simple name of the enum \\
-\hline
 \texttt{full\_name} & Fully qualified name of the enum \\
-\hline
 \texttt{type} & Name of the message type where the enum is declared \\[.3cm]
-\hline
+%
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{as.list} & return a named
@@
 \texttt{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
-\textbf{slot} & \textbf{description} \\
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
+\textbf{Slot} & \textbf{Description} \\
+\cmidrule(r){2-2}
 \texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the C++ proto library. Documentation for the
@@
 \url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
-\hline
 \texttt{filename} & fully qualified pathname of the \texttt{.proto} file.\\
-\hline
 \texttt{package} & package name defined in this \texttt{.proto} file.\\[.3cm]
-\hline
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{name} & Return the filename for this FileDescriptorProto.\\
@@
 \texttt{as.list} & return named list of descriptors defined in this file descriptor.\\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{small}
-\begin{tabular}{l|p{10cm}}
-\hline
-\textbf{slot} & \textbf{description} \\
-\hline
+\begin{tabular}{lp{10cm}}
+\toprule
+\textbf{Slot} & \textbf{Description} \\
+\cmidrule(r){2-2}
 \texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} C++ variable \\
-\hline
 \texttt{name} & simple name of the enum value \\
-\hline
 \texttt{full\_name} & fully qualified name of the enum value \\[.3cm]
-\hline
+%
 \textbf{Method} & \textbf{Description} \\
-\hline
+\cmidrule(r){2-2}
 \texttt{number} & return the number of this EnumValueDescriptor. \\
@@
 \texttt{asMessage} & return EnumValueDescriptorProto message. \\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{small}
-\begin{tabular}{|c|p{5cm}p{5cm}|}
-\hline
-field type & R type (non repeated) & R type (repeated) \\
-\hline
-\hline
+\begin{tabular}{lp{5cm}p{5cm}}
+\toprule
+Field type & R type (non repeated) & R type (repeated) \\
+\cmidrule(r){2-3}
 double	& \texttt{double} vector & \texttt{double} vector \\
-float	& \texttt{double} vector & \texttt{double} vector \\
-\hline
+float	& \texttt{double} vector & \texttt{double} vector \\[3mm]
 uint32	  & \texttt{double} vector & \texttt{double} vector \\
-fixed32	  & \texttt{double} vector & \texttt{double} vector \\
-\hline
+fixed32	  & \texttt{double} vector & \texttt{double} vector \\[3mm]
 int32	  & \texttt{integer} vector & \texttt{integer} vector \\
 sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
-sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\
-\hline
+sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\[3mm]
 int64	  & \texttt{integer} or \texttt{character}
@@
 fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-\hline
-bool	& \texttt{logical} vector & \texttt{logical} vector \\
-\hline
+sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\\hline
+bool	& \texttt{logical} vector & \texttt{logical} vector \\[3mm]
 string	& \texttt{character} vector & \texttt{character} vector \\
-bytes	& \texttt{character} vector & \texttt{character} vector \\
-\hline
-enum & \texttt{integer} vector & \texttt{integer} vector \\
-\hline
+bytes	& \texttt{character} vector & \texttt{character} vector \\[3mm]
+enum & \texttt{integer} vector & \texttt{integer} vector \\[3mm]
 message & \texttt{S4} object of class \texttt{Message} & \texttt{list} of \texttt{S4} objects of class \texttt{Message} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 \begin{center}
+  \small
 \scalebox{0.9}{
-\begin{tabular}{l|r|r|r|r|r}
-  \hline
-Data Set & object.size & \multicolumn{2}{c|}{R Serialization} &
-\multicolumn{2}{c}{RProtoBuf Serialization} \\
- & & Default & gzipped & Default & gzipped \\
-  \hline
-uspop & 584.00 & 268 & 172 & 211 & 148 \\
-  Titanic & 1960.00 & 633 & 257 & 481 & 249 \\
-  volcano & 42656.00 & 42517 & 5226 & 42476 & 4232 \\
-  euro.cross & 2728.00 & 1319 & 910 & 1207 & 891 \\
-  attenu & 14568.00 & 8234 & 2165 & 7771 & 2336 \\
-  ToothGrowth & 2568.00 & 1486 & 349 & 1239 & 391 \\
-  lynx & 1344.00 & 1028 & 429 & 971 & 404 \\
-  nottem & 2352.00 & 2036 & 627 & 1979 & 641 \\
-  sleep & 2752.00 & 746 & 282 & 483 & 260 \\
-  co2 & 4176.00 & 3860 & 1473 & 3803 & 1453 \\
-  austres & 1144.00 & 828 & 439 & 771 & 410 \\
-  ability.cov & 1944.00 & 716 & 357 & 589 & 341 \\
-  EuStockMarkets & 60664.00 & 59785 & 21232 & 59674 & 19882 \\
-  treering & 64272.00 & 63956 & 17647 & 63900 & 17758 \\
-  freeny.x & 1944.00 & 1445 & 1311 & 1372 & 1289 \\
-  Puromycin & 2088.00 & 813 & 306 & 620 & 320 \\
-  warpbreaks & 2768.00 & 1231 & 310 & 811 & 343 \\
-  BOD & 1088.00 & 334 & 182 & 226 & 168 \\
-  sunspots & 22992.00 & 22676 & 6482 & 22620 & 6742 \\
-  beaver2 & 4184.00 & 3423 & 751 & 3468 & 840 \\
-  anscombe & 2424.00 & 991 & 375 & 884 & 352 \\
-  esoph & 5624.00 & 3111 & 548 & 2240 & 665 \\
-  PlantGrowth & 1680.00 & 646 & 303 & 459 & 314 \\
-  infert & 15848.00 & 14328 & 1172 & 13197 & 1404 \\
-  BJsales & 1632.00 & 1316 & 496 & 1259 & 465 \\
-  stackloss & 1688.00 & 917 & 293 & 844 & 283 \\
-  crimtab & 7936.00 & 4641 & 713 & 1655 & 576 \\
-  LifeCycleSavings & 6048.00 & 3014 & 1420 & 2825 & 1407 \\
-  Harman74.cor & 9144.00 & 6056 & 2045 & 5861 & 2070 \\
-  nhtemp & 912.00 & 596 & 240 & 539 & 223 \\
-  faithful & 5136.00 & 4543 & 1339 & 4936 & 1776 \\
-  freeny & 5296.00 & 2465 & 1518 & 2271 & 1507 \\
-  discoveries & 1232.00 & 916 & 199 & 859 & 180 \\
-  state.x77 & 7168.00 & 4251 & 1754 & 4068 & 1756 \\
-  pressure & 1096.00 & 498 & 277 & 427 & 273 \\
-  fdeaths & 1008.00 & 692 & 291 & 635 & 272 \\
-  euro & 976.00 & 264 & 186 & 202 & 161 \\
-  LakeHuron & 1216.00 & 900 & 420 & 843 & 404 \\
-  mtcars & 6736.00 & 3798 & 1204 & 3633 & 1206 \\
-  precip & 4992.00 & 1793 & 813 & 1615 & 815 \\
-  state.area & 440.00 & 422 & 246 & 405 & 235 \\
-  attitude & 3024.00 & 1990 & 544 & 1920 & 561 \\
-  randu & 10496.00 & 9794 & 8859 & 10441 & 9558 \\
-  state.name & 3088.00 & 844 & 408 & 724 & 415 \\
-  airquality & 5496.00 & 4551 & 1241 & 2874 & 1294 \\
-  airmiles & 624.00 & 308 & 170 & 251 & 148 \\
-  quakes & 33112.00 & 32246 & 9898 & 29063 & 11595 \\
-  islands & 3496.00 & 1232 & 563 & 1098 & 561 \\
-  OrchardSprays & 3600.00 & 2164 & 445 & 1897 & 483 \\
-  WWWusage & 1232.00 & 916 & 274 & 859 & 251 \\
-   \hline
+\begin{tabular}{lrrrrr}
+  \toprule
+  Data Set & object.size & \multicolumn{2}{c}{R Serialization} &
+  \multicolumn{2}{c}{RProtoBuf Serial.} \\
+  & & default & gzipped & default & gzipped \\
+  \cmidrule(r){2-6}
+  uspop & 584 & 268 & 172 & 211 & 148 \\
+  Titanic & 1960 & 633 & 257 & 481 & 249 \\
+  volcano & 42656 & 42517 & 5226 & 42476 & 4232 \\
+  euro.cross & 2728 & 1319 & 910 & 1207 & 891 \\
+  attenu & 14568 & 8234 & 2165 & 7771 & 2336 \\
+  ToothGrowth & 2568 & 1486 & 349 & 1239 & 391 \\
+  lynx & 1344 & 1028 & 429 & 971 & 404 \\
+  nottem & 2352 & 2036 & 627 & 1979 & 641 \\
+  sleep & 2752 & 746 & 282 & 483 & 260 \\
+  co2 & 4176 & 3860 & 1473 & 3803 & 1453 \\
+  austres & 1144 & 828 & 439 & 771 & 410 \\
+  ability.cov & 1944 & 716 & 357 & 589 & 341 \\
+  EuStockMarkets & 60664 & 59785 & 21232 & 59674 & 19882 \\
+  treering & 64272 & 63956 & 17647 & 63900 & 17758 \\
+  freeny.x & 1944 & 1445 & 1311 & 1372 & 1289 \\
+  Puromycin & 2088 & 813 & 306 & 620 & 320 \\
+  warpbreaks & 2768 & 1231 & 310 & 811 & 343 \\
+  BOD & 1088 & 334 & 182 & 226 & 168 \\
+  sunspots & 22992 & 22676 & 6482 & 22620 & 6742 \\
+  beaver2 & 4184 & 3423 & 751 & 3468 & 840 \\
+  anscombe & 2424 & 991 & 375 & 884 & 352 \\
+  esoph & 5624 & 3111 & 548 & 2240 & 665 \\
+  PlantGrowth & 1680 & 646 & 303 & 459 & 314 \\
+  infert & 15848 & 14328 & 1172 & 13197 & 1404 \\
+  BJsales & 1632 & 1316 & 496 & 1259 & 465 \\
+  stackloss & 1688 & 917 & 293 & 844 & 283 \\
+  crimtab & 7936 & 4641 & 713 & 1655 & 576 \\
+  LifeCycleSavings & 6048 & 3014 & 1420 & 2825 & 1407 \\
+  Harman74.cor & 9144 & 6056 & 2045 & 5861 & 2070 \\
+  nhtemp & 912 & 596 & 240 & 539 & 223 \\
+  faithful & 5136 & 4543 & 1339 & 4936 & 1776 \\
+  freeny & 5296 & 2465 & 1518 & 2271 & 1507 \\
+  discoveries & 1232 & 916 & 199 & 859 & 180 \\
+  state.x77 & 7168 & 4251 & 1754 & 4068 & 1756 \\
+  pressure & 1096 & 498 & 277 & 427 & 273 \\
+  fdeaths & 1008 & 692 & 291 & 635 & 272 \\
+  euro & 976 & 264 & 186 & 202 & 161 \\
+  LakeHuron & 1216 & 900 & 420 & 843 & 404 \\
+  mtcars & 6736 & 3798 & 1204 & 3633 & 1206 \\
+  precip & 4992 & 1793 & 813 & 1615 & 815 \\
+  state.area & 440 & 422 & 246 & 405 & 235 \\
+  attitude & 3024 & 1990 & 544 & 1920 & 561 \\
+  randu & 10496 & 9794 & 8859 & 10441 & 9558 \\
+  state.name & 3088 & 844 & 408 & 724 & 415 \\
+  airquality & 5496 & 4551 & 1241 & 2874 & 1294 \\
+  airmiles & 624 & 308 & 170 & 251 & 148 \\
+  quakes & 33112 & 32246 & 9898 & 29063 & 11595 \\
+  islands & 3496 & 1232 & 563 & 1098 & 561 \\
+  OrchardSprays & 3600 & 2164 & 445 & 1897 & 483 \\
+  WWWusage & 1232 & 916 & 274 & 859 & 251 \\
+  \bottomrule
 \end{tabular}
 }
-\caption{Serialization sizes with R's built-in serialization and
-  RProtoBuf for 50 sample R datasets.}
+\caption{Serialization sizes for default serialization in R and
+  RProtoBuf for 50 R datasets.}
 \label{tab:compression}
@@
 
-\bibliography{eddelbuettel-stokely}
+\bibliography{article}
 

### 2014-01-11 papers/jss/article.Rnw
bunch of edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@750 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Dirk Eddelbuettel, Murray Stokely} %% comma-separated
+\Plainauthor{Dirk Eddelbuettel, Murray Stokely, Jeroen Ooms} %% comma-separated
 \Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
@@
 different programming languages used for the different phases of data
-analysis -- collection, cleaning, analysis, post-processing, and
+analysis -- collection, cleaning, modeling, analysis, post-processing, and
 presentation in order to take advantage of the unique combination of
 performance, speed of development, and library support offered by
-different environments.  Each stage of the data
+different environments and languages.  Each stage of such a data
 analysis pipeline may involve storing intermediate results in a
 file or sending them over the network.
-% DE: Nice!
 
@@
 programming language in use and thus lock the user into a single
-environment.  CSV files can be read and written by many applications
-and so are often used for exporting tabular data.  However, CSV files
-have a number of disadvantages, such as a limitation of exporting only
-tabular datasets, lack of type-safety, inefficient text representation
-and parsing, and ambiguities in the format involving special
-characters.  JSON is another widely-supported format used mostly on
-the web that removes many of these disadvantages, but it too suffers
-from being too slow to parse and also does not provide strong typing
-between integers and floating point.  Because the schema information
-is not kept separately, multiple JSON messages of the same type
-needlessly duplicate the field names with each message.
-Lastly, XML is a well-established and widely-supported protocol with the ability to define
-just about any arbitrarily complex schema. However, it pays for this
-complexity with comparatively large and verbose messages, and added
-complexities at the parsing side. 
-%
-%
-%
+environment.
+
+\emph{Comma-separated values} (CSV) files can be read and written by many
+applications and so are often used for exporting tabular data.  However, CSV
+files have a number of disadvantages, such as a limitation of exporting only
+tabular datasets, lack of type-safety, inefficient text representation and
+parsing, possibly limited precision and ambiguities in the format involving
+special characters.  \emph{JavaScript Object Notation} (JSON) is another
+widely-supported format used mostly on the web that removes many of these
+disadvantages, but it too suffers from being too slow to parse and also does
+not provide strong typing between integers and floating point.  Because the
+schema information is not kept separately, multiple JSON messages of the same
+type needlessly duplicate the field names with each message.  Lastly,
+\emph{Extensible Markup Language} (XML) is a well-established and widely-supported
+protocol with the ability to define just about any arbitrarily complex
+schema. However, it pays for this complexity with comparatively large and
+verbose messages, and added complexities at the parsing side (which are
+somewhat metigated by the availability of mature libraries and
+parsers).
+
 A number of binary formats based on JSON have been proposed that
 reduce the parsing cost and improve the efficiency.  MessagePack
-and BSON both have R interfaces \citep{msgpackR,rmongodb}, but
+and BSON both have R interfaces, but % \citep{msgpackR,rmongodb}, but
 % DE Why do we cite these packages, but not the numerous JSON packages?
@@
 requirement changes dictate update to the message format.
-% DE: Need to talk about XML -- added a few lines at previous paragraph
 
@@
 
-Section~\ref{sec:protobuf} provides a general overview of Protocol
-Buffers.  Section~\ref{sec:rprotobuf-basic} describes the interactive
-R interface provided by \CRANpkg{RProtoBuf} and introduces the two
-main abstractions: \emph{Messages} and \emph{Descriptors}.
-Section~\ref{sec:rprotobuf-classes} describes the implementation
-details of the main S4 classes making up this package.
-Section~\ref{sec:types} describes the challenges of type coercion
-between R and other languages.  Section~\ref{sec:evaluation}
-introduces a general R language schema for serializing arbitrary R
-objects and evaluates it against R's built-in serialization.
-Sections~\label{sec:opencpu} and \label{sec:mapreduce} provide
-real-world use cases of \CRANpkg{RProtoBuf} in web service and
-MapReduce environments, respectively.
+The rest of the paper is organized as follows. Section~\ref{sec:protobuf}
+provides a general overview of Protocol Buffers.
+Section~\ref{sec:rprotobuf-basic} describes the interactive R interface
+provided by \CRANpkg{RProtoBuf} and introduces the two main abstractions:
+\emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
+describes the implementation details of the main S4 classes making up this
+package.  Section~\ref{sec:types} describes the challenges of type coercion
+between R and other languages.  Section~\ref{sec:evaluation} introduces a
+general R language schema for serializing arbitrary R objects and evaluates
+it against R's built-in serialization.  Sections~\ref{sec:opencpu}
+and \ref{sec:mapreduce} provide real-world use cases of \CRANpkg{RProtoBuf}
+in web service and MapReduce environments, respectively, before
+Section~\ref{sec:summary} concludes.
 
@@
 introduction, Protocol Buffers have been widely adopted in industry with
-applications as varied as database-internal messaging (Drizzle), % DE: citation?
-Sony Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  While
+applications as varied as %database-internal messaging (Drizzle), % DE: citation?
+Sony Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  
 % TODO(DE): This either needs a citation, or remove the name drop
-traditional IDLs have at time been criticized for code bloat and
+While traditional IDLs have at time been criticized for code bloat and
 complexity, Protocol Buffers are based on a simple list and records
@@
 \begin{itemize}
-\item \emph{Portable}:  Allows users to send and receive data between
-  applications or different computers.
+\item \emph{Portable}:  Enable users to send and receive data between
+  applications as well as different computers or operating systems.
 \item \emph{Efficient}:  Data is serialized into a compact binary
@@
 \item \emph{Extensible}:  New fields can be added to Protocol Buffer Schemas
-  in a forward-compatible way that do not break older applications.
+  in a forward-compatible way that does not break older applications.
 \item \emph{Stable}:  Protocol Buffers have been in wide use for over a
@@
 Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
-communication workflow with protocol buffers and an interactive R
-session.  Common use cases include populating a request RPC protocol
-buffer in R that is then serialized and sent over the network to a
-remote server.  The server would then deserialize the message, act on
-the request, and respond with a new protocol buffer over the network. The key
+communication workflow with Protocol Buffers and an interactive R session.
+Common use cases include populating a request remote-procedure call (RPC)
+Protocol Buffer in R that is then serialized and sent over the network to a
+remote server.  The server would then deserialize the message, act on the
+request, and respond with a new Protocol Buffer over the network. The key
 difference to, say, a request to an Rserve instance is that the remote server
@@
 
-Many sources compare data serialization formats and show protocol
-buffers very favorably to the alternatives, such
-as \citet{Sumaray:2012:CDS:2184751.2184810}
+Many sources compare data serialization formats and show Protocol
+Buffers very favorably to the alternatives; see
+\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
 
@@
 For added speed and efficiency, the C++, Java, and Python bindings to
-Protocol Buffers are used with a compiler that translates a protocol
-buffer schema description file (ending in \texttt{.proto}) into
+Protocol Buffers are used with a compiler that translates a Protocol
+Buffer schema description file (ending in \texttt{.proto}) into
 language-specific classes that can be used to create, read, write and
-manipulate protocol buffer messages.  The R interface, in contrast,
-uses a reflection-based API that is particularly well suited for
+manipulate Protocol Buffer messages.  The R interface, in contrast,
+uses a reflection-based API that is particularly well-suited for
 interactive data analysis.  All messages in R have a single class
@@
 
-The two fundamental building blocks of Protocol Buffers are Messages
-and Descriptors.  Messages provide a common abstract encapsulation of
+The two fundamental building blocks of Protocol Buffers are \emph{Messages}
+and \emph{Descriptors}.  Messages provide a common abstract encapsulation of
 structured data fields of the type specified in a Message Descriptor.
@@
 \begin{table}
-\begin{tabular}{@{\hskip .01\textwidth}p{.40\textwidth}@{\hskip .02\textwidth}@{\hskip .02\textwidth}p{0.55\textwidth}@{\hskip .01\textwidth}}
+\begin{tabular}{p{.40\textwidth}p{0.55\textwidth}}
 \toprule
@@
 \cmidrule{1-2}
-\begin{minipage}{.35\textwidth}
+\begin{minipage}{.40\textwidth}
 \vspace{2mm}
@@
 \vspace{2mm}
-\end{minipage} & \begin{minipage}{.5\textwidth}
+\end{minipage} & \begin{minipage}{.55\textwidth}
 <<echo=TRUE>>=
 library(RProtoBuf)
-p <- new(tutorial.Person, id=1, name="Dirk")
+p <- new(tutorial.Person,id=1,name="Dirk")
 class(p)
@@
 The \texttt{.proto} file syntax for defining the structure of protocol
-buffer data is described comprehensively on Google Code:
-\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.
+buffer data is described comprehensively on Google Code\footnote{See 
+\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
 
@@
 and the name must be given entirely.
-
 The \verb|[[| operator can also be used to query and set fields
@@
 
-Protocol buffers include a 64-bit integer type, but R lacks native
+Protocol Buffers include a 64-bit integer type, but R lacks native
 64-bit integer support.  A workaround is available and described in
@@
 
-Protocol buffer messages and descriptors implement \texttt{show}
+Protocol Buffer messages and descriptors implement \texttt{show}
 methods that provide basic information about the message :
@@
 
-However, the main focus of protocol buffer messages is
+However, the main focus of Protocol Buffer messages is
 efficiency. Therefore, messages are transported as a sequence
 of bytes. The \texttt{serialize} method is implemented for
-protocol buffer messages to serialize a message into a sequence of
+Protocol Buffer messages to serialize a message into a sequence of
 bytes that represents the message.
@@
 
-\texttt{read} can also be used as a pseudo method of the descriptor
+\texttt{read} can also be used as a pseudo-method of the descriptor
 object :
@@
 Each R object stores an external pointer to an object managed by
-the \texttt{protobuf} C++ library.
-The \CRANpkg{Rcpp} package \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to
+the \texttt{protobuf} C++ library which implements the core Protocol Buffer
+functionality.  The \CRANpkg{Rcpp} package
+\citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
 facilitate the integration of the R and C++ code for these objects.
@@
 
-The \texttt{RProtoBuf} package combines the \emph{R typical} dispatch
-of the form \verb|method(object, arguments)| and the more traditional
-object oriented notation \verb|object$method(arguments)|.
+The \texttt{RProtoBuf} package combines a dispatch mechanism 
+of the form \verb|method(object, arguments)| (common to R) and the more
+traditional object oriented notation \verb|object$method(arguments)|.
 Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
 (defined in the \texttt{utils} package) for all classes to enable tab
-completion.  Completion possibilities include pseudo method names for all
+completion.  Completion possibilities include pseudo-method names for all
 classes, plus dynamic dispatch on names or types specific to a given object.
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{Message} object of the C++ proto library. Documentation for the
-\texttt{Message} class is available from the protocol buffer project page:
-\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message} \\
+\texttt{pointer} & External pointer to the \texttt{Message} object of the C++ protobuf library. Documentation for the
+\texttt{Message} class is available from the Protocol Buffer project page. \\
+%(\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message}) \\
 \texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
@@
 \texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
-\texttt{Descriptor} class is available from the protocol buffer project page:
-\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
+\texttt{Descriptor} class is available from the Protocol Buffer project page.\\
+%\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
 \texttt{type} & Fully qualified path of the message type. \\[.3cm]
@@
 like Protocol Buffers is that it provides a highly portable basic type
-system that different language and hardware implementations can map to
+system. This permits different language and hardware implementations to map to
 the most appropriate type in different environments.
+
 Table~\ref{table-get-types} details the correspondence between the
@@
 int64	  & \texttt{integer} or \texttt{character}
-vector \footnotemark & \texttt{integer} or \texttt{character} vector \\
+vector    & \texttt{integer} or \texttt{character} vector \\
 uint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
@@
 fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\\hline
+sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\[3mm]
 bool	& \texttt{logical} vector & \texttt{logical} vector \\[3mm]
@@
 \caption{\label{table-get-types}Correspondence between field type and
-  R type retrieved by the extractors. \footnotesize{1. R lacks native
+  R type retrieved by the extractors. Note that R lacks native
   64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
   available to return large integers as characters to avoid losing
-  precision.  This option is described in Section~\ref{sec:int64}}.}
+  precision.  This option is described in Section~\ref{sec:int64}.}
 \end{table}
@@
 R booleans can accept three values: \texttt{TRUE}, \texttt{FALSE}, and
-\texttt{NA}.  However, most other languages, including the protocol
-buffer schema, only accept \texttt{TRUE} or \texttt{FALSE}.  This means
+\texttt{NA}.  However, most other languages, including the Protocol
+Buffer schema, only accept \texttt{TRUE} or \texttt{FALSE}.  This means
 that we simply can not store R logical vectors that include all three
@@
 R lacks a native unsigned integer type.  Values between $2^{31}$ and
-$2^{32} - 1$ read from unsigned int protocol buffer fields must be
+$2^{32} - 1$ read from unsigned into Protocol Buffer fields must be
 stored as doubles in R.
@@
 
-Saptarshi Guha wrote the RHIPE package \citep{rhipe} which includes
-protocol buffer integration with R.  However, this implementation
-takes a different approach: any R object is serialized into a message
-based on a single catch-all \texttt{proto} schema.  Jeroen Ooms took a
-similar approach influenced by Saptarshi in the \pkg{RProtoBufUtils}
-package (which has now been integrated in \pkg{RProtoBuf}).  Unlike
-Saptarshi's package, however, RProtoBufUtils depends 
-on, and extends, RProtoBuf for underlying message operations.
+The \pkg{RHIPE} package \citep{rhipe} also includes a Protocol integration with R.
+However, its implementation takes a different approach: any R object is
+serialized into a message based on a single catch-all \texttt{proto} schema.
+A similar approach was taken by \pkg{RProtoBufUtils} package (which has now been integrated in
+\pkg{RProtoBuf}). Unlike \pkg{RHIPE}, however, \pkg{RProtoBufUtils}
+depended upon on, and extended, \pkg{RProtoBuf} for underlying message operations.
+%DE Shall this go away now that we sucket RPBUtils into RBP?
 
@@
 \texttt{serialize\_pb} method to convert R objects into serialized
-protocol buffers in the catch-all schema. The \texttt{can\_serialize\_pb}
-method can be used to determine whether the given R object can safely
+Protocol Buffers in the catch-all schema. The \texttt{can\_serialize\_pb}
+method can then be used to determine whether the given R object can safely
 be expressed in this way.  To illustrate how this method works, we
 attempt to convert all of the built-in datasets from R into this
-serialized protocol buffer representation.
+serialized Protocol Buffer representation.
 
@@
 \texttt{can\_serialize\_pb} method to determine how many of those can
-be safely converted to a serialized protocol buffer representation.
+be safely converted to a serialized Protocol Buffer representation.
 
@@
 situations (1) normal R serialization, (2) R serialization followed by
-gzip, (3) normal protocol buffer serialization, (4) protocol buffer
+gzip, (3) normal Protocol Buffer serialization, (4) Protocol Buffer
 serialization followed by gzip.
@@
 \section{Summary}
+\label{sec:summary}
 

### 2014-01-11 papers/jss/article.Rnw
more edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@751 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-One key extension of \pkg{RProtoBufUtils} is the 
+One key extension which \pkg{RProtoBufUtils} brought to \pkg{RProtoBuf} is the 
 \texttt{serialize\_pb} method to convert R objects into serialized
@@
 <<echo=TRUE>>=
-#datasets$valid.proto <- sapply(datasets$load.name, function(x) can_serialize_pb(eval(as.name(x))))
-#datasets <- subset(datasets, valid.proto==TRUE)
+datasets$valid.proto <- sapply(datasets$load.name,
+                        function(x) can_serialize_pb(eval(as.name(x))))
+datasets <- subset(datasets, valid.proto==TRUE)
 m <- nrow(datasets)
@@
 many bytes were used to store the data sets under four different
-situations (1) normal R serialization, (2) R serialization followed by
-gzip, (3) normal Protocol Buffer serialization, (4) Protocol Buffer
-serialization followed by gzip.
+situations:
+\begin{itemize}
+\item normal R serialization,
+\item R serialization followed by gzip,
+\item normal Protocol Buffer serialization, and
+\item Protocol Buffer serialization followed by gzip.
+\end{itemize}
 
@@
 returned by object.size() compared to the serialized sizes.
-The summary compression sizes are listed below, and a full table for a
-sample of 50 datasets is included on the next page.  Sizes are comparable
-but protocol buffers provide simple getters and setters in multiple
-languages instead of requiring other programs to parse the R
-serialization format \citep{serialization}.  One takeaway from this
-table is that RProtoBuf does not in general provide any significant
-space-savings over R's normal serialization mechanism.  The benefit
-from RProtoBuf comes from its interoperability with other
-environments, safe versioning,
+%The summary compression sizes are listed below, and a full table for a
+%sample of 50 datasets is included on the next page.  
+Sizes are comparable but Protocol Buffers provide simple getters and setters
+in multiple languages instead of requiring other programs to parse the R
+serialization format.% \citep{serialization}.
+One takeaway from this table is that RProtoBuf does not in general provide
+any significant saving in file size compared to the normal serialization
+mechanism in R which is seen as equally compact.  The benefit from RProtoBuf
+comes from its interoperability with other environments, as well as its safe
+versioning,
 
 TODO comparison of protobuf serialization sizes/times for various vectors.
-Compared to R's native serialization.
-Discussion of the RHIPE approach of serializing any/all R objects, vs more specific protocol buffers for specific R objects.
+Compared to R's native serialization.  Discussion of the RHIPE approach of
+serializing any/all R objects, vs more specific Protocol Buffers for specific
+R objects.
 
@@
 analysis, but it is not designed for certain classes of operations one
-might like to do with protocol buffers.  For example, taking a list of
-10,000 protocol buffers, extracting a named field from each one, and
+might like to do with Protocol Buffers.  For example, taking a list of
+10,000 Protocol Buffers, extracting a named field from each one, and
 computing a aggregate statistics on those values would be extremely
@@
 
-As described earlier, the primary application of protocol buffers is
-data interchange in the context of inter-system communications. 
-Network protocols such as HTTP provide mechanisms for client-server
-communication, i.e. how to initiate requests, authenticate, send messages, 
-etc.  However, many network 
-protocols generally do not regulate \emph{content} of messages: they allow
-transfer of any media type, such as web pages, files or video.
-When designing systems where various components require exchange of specific data
-structures, we need something on top of the network protocol that prescribes 
-how these structures are to be represented in messages (buffers) on the
-network. Protocol buffers solve exactly this problem by providing
-a cross platform method for serializing arbitrary structures into well defined
-messages, that can be exchanged using any protocol. The descriptors
-(\texttt{.proto} files) are used to formally define the interface of a
-remote API or network application. Libraries to parse and generate protobuf
-messages are available for many programming languages, making it 
-relatively straight forward to implement clients and servers.
+As described earlier, the primary application of Protocol Buffers is data
+interchange in the context of inter-system communications.  Network protocols
+such as HTTP provide mechanisms for client-server communication, i.e. how to
+initiate requests, authenticate, send messages, etc.  However, many network
+protocols generally do not regulate the \emph{content} of messages: they
+allow transfer of any media type, such as web pages, static files or
+multimedia content.  When designing systems where various components require
+exchange of specific data structures, we need something on top of the network
+protocol that prescribes how these structures are to be represented in
+messages (buffers) on the network. Protocol Buffers solve exactly this
+problem by providing a cross-platform method for serializing arbitrary
+structures into well defined messages, which can then be exchanged using any
+protocol. The descriptors (\texttt{.proto} files) are used to formally define
+the interface of a remote API or network application. Libraries to parse and
+generate protobuf messages are available for many programming languages,
+making it relatively straightforward to implement clients and servers.
 
@@
 
-One example of a system that supports protocol buffers to interact
+One example of a system that supports Protocol Buffers to interact
 with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
@@
 # Retrieve and parse message
-req <- GET ('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+req <- GET('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
 output <- unserialize_pb(req$content)
@@
 This code suggests a method for exchanging objects between R servers, however this can 
-also be done without protocol buffers. The main advantage of using an inter-operable format 
+also be done without Protocol Buffers. The main advantage of using an inter-operable format 
 is that we can actually access R objects from within another
@@
 \end{verbatim}
-This generates python module called \texttt{rexp\_pb2.py}, containing both the 
+This generates Python module called \texttt{rexp\_pb2.py}, containing both the 
 descriptor information as well as methods to read and manipulate the R object 
@@
 containing R objects to post to the server, as well as retrieve and parse
-protobuf messages returned by the server. Using protocol buffers to post
+protobuf messages returned by the server. Using Protocol Buffers to post
 function arguments is not required, and for simple (scalar) arguments 
@@
 @
-In reality the OpenCPU provides a lot of meta functionality such as handling
-of sessions, exceptions, security, and much more. OpenCPU also makes it possible to store
+
+OpenCPU also provides a lot of meta-functionality such as handling
+of sessions, exceptions, security, and more. OpenCPU also makes it possible to store
 output of a function call on the server, instead of directly retrieving it. Thereby 
@@
 
-The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010,
-with very significant contributions, both in code and design, made by
-Romain Fran\c{c}ois. His continued influence on design and code is
-appreciated. Several features of the package are influenced
+The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
+Very significant contributions, both in code and design, were made by
+Romain Fran\c{c}ois whose continued influence on design and code is
+greatly appreciated. Several features of the package are influenced
 by the design of the \CRANpkg{rJava} package by Simon Urbanek
 The user-defined table mechanism, implemented by Duncan Temple Lang for the
-purpose of the \pkg{RObjectTables} package allowed the dynamic symbol lookup.
+purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
 Kenton Varda was generous with his time in reviewing code and explaining
-obscure protocol buffer semantics.  Karl Millar and Jeroen Ooms were
-helpful in reviewing code or offering suggestions.  The contemporaneous
-work by Saptarshi Guha on \pkg{RHIPE} was a strong initial motivator.
+obscure protocol buffer semantics.  Karl Millar was very
+helpful in reviewing code and offering suggestions.  
+%The contemporaneous work by Saptarshi Guha on \pkg{RHIPE} was a strong
+%initial motivator.
 

### 2014-01-12 papers/jss/article.Rnw
tiny changes for that my IDE needs
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@754 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
move a strayed sentence
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@755 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
rewrite paragraph
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@757 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
+\SweaveOpts{concordance=FALSE}
 
@@
 ``schema-less'' binary interchange formats.
+Many sources compare data serialization formats and show Protocol
+Buffers very favorably to the alternatives; see
+\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
 
@@
 %times \textsl{faster}, as well as less ambiguous and easier to program.
-
-Many sources compare data serialization formats and show Protocol
-Buffers very favorably to the alternatives; see
-\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
 
@@
 
-Before one can create a new Protocol Buffer Message or parse a
-serialized stream of bytes as a Message, one must first read in the message
-type specification from a \texttt{.proto} file.
-
-New \texttt{.proto} files are imported with the \code{readProtoFiles}
-function, which can import a single file, all files in a directory, or
-all \texttt{.proto} files provided by another R package.
-
+To create or parse a Protocol Buffer Message, one must first read in 
+the message type specification from a \texttt{.proto} file. The 
+\texttt{.proto} files are imported using the \code{readProtoFiles}
+function, which can either import a single file, all files in a directory,
+or every \texttt{.proto} file provided by a particular R package.
 The \texttt{.proto} file syntax for defining the structure of protocol
@@
 
-Once the proto files are imported, all message descriptors are
-are available in the R search path in the \texttt{RProtoBuf:DescriptorPool}
-special environment. The underlying mechanism used here is
+After importing proto files, the corresponding message descriptors are
+available from the \texttt{RProtoBuf:DescriptorPool} environment in 
+the R search path. The underlying mechanism used here is
 described in more detail in Section~\ref{sec-lookup}.
+
+%JO: can we just move the section 7 to here? It's only one paragraph%
 

### 2014-01-12 papers/jss/article.Rnw
words
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@758 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-The same method can also be used to serialize messages to files :
+The same method can be used to serialize messages to files :
 
@@
 
-\texttt{serialize} can also be used in a more traditional
-object oriented fashion using the dollar operator :
+\texttt{serialize} can also be called in a more traditional
+object oriented fashion using the dollar operator:
 

### 2014-01-12 papers/jss/article.Rnw
tiny changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@759 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-The \texttt{RProtoBuf} package combines a dispatch mechanism 
+The \texttt{RProtoBuf} package combines a functional dispatch mechanism 
 of the form \verb|method(object, arguments)| (common to R) and the more
@@
 \texttt{NA}s in protocol buffer boolean fields, and users must instead
-choose another type (such as integers) capable of storing three
+choose another type (such as enum or integer) capable of storing three
 distinct values.

### 2014-01-12 papers/jss/article.Rnw
some rephrasing on int64
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@760 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-R does not have native 64-bit integer support.  Instead, R treats
-large integers as doubles which have limited precision.  For example,
-it loses the ability to distinguish some distinct integers:
+R also does not support the native 64-bit integer type. Numeric vectors
+with values $\geq 2^{31}$ can only be stored as doubles, which have
+limited precision. Thereby R loses the ability to distinguish some
+distinct integers:
 
@@
 
-Protocol Buffers are frequently used to pass data between different
-systems, however, and most other modern systems do have support for
-64-bit integers.  To work around this, RProtoBuf allows users to get
-and set 64-bit integer types by treating them as characters.
+However, most modern languages do have support for 64-bit integers, 
+which becomes problematic when \pkg{RProtoBuf} is used to exchange data 
+with a system that requires this integer type. To work around this, 
+RProtoBuf allows users to get and set 64-bit integer values by specifying 
+them as character strings.
 
@@
 
-However, we can specify the values as character strings so that the
-C++ library on which RProtoBuf is based can store a true 64-bit
-integer representation of the data.
+But when the values are specified as character strings, RProtoBuf
+will automatically coerse them into a true 64-bit integer types 
+before storing them in the Protocol Buffer message:
 

### 2014-01-12 papers/jss/article.Rnw
first pass at data frame section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@761 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<echo=TRUE>>=
-datasets <- subset(as.data.frame(data()$results), Package=="datasets")
-datasets$load.name <- sub("\\s+.*$", "", datasets$Item)
+datasets <- as.data.frame(data(package="datasets")$results)
+datasets$name <- sub("\\s+.*$", "", datasets$Item)
 n <- nrow(datasets)
@@
 
-There are \Sexpr{n} standard data sets included in R.  We use the
-\texttt{can\_serialize\_pb} method to determine how many of those can
-be safely converted to a serialized Protocol Buffer representation.
+There are \Sexpr{n} standard data sets included in the base-r \pkg{datasets}
+package. These datasets include data frames, matices, timeseries, tables lists,
+and some more exotic data classes. The \texttt{can\_serialize\_pb} method can be 
+used to determine which of those can fully be converted to the \textt{rexp.proto}
+Protocol Buffer representation:
 
 <<echo=TRUE>>=
-datasets$valid.proto <- sapply(datasets$load.name,
-                        function(x) can_serialize_pb(eval(as.name(x))))
+datasets$valid.proto <- sapply(datasets$name, function(x) can_serialize_pb(get(x)))
 datasets <- subset(datasets, valid.proto==TRUE)
@@
 <<echo=FALSE,print=FALSE>>=
-datasets$object.size <- unname(sapply(datasets$load.name, function(x) object.size(eval(as.name(x)))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))
-
-datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$load.name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))
-
-datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$load.name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))
-
-clean.df <- data.frame(dataset=datasets$load.name,
+datasets$object.size <- unname(sapply(datasets$name, function(x) object.size(eval(as.name(x)))))
+
+datasets$R.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize(eval(as.name(x)), NULL))))
+
+datasets$R.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize(eval(as.name(x)), NULL))))
+
+datasets$R.serialize.size.gz <- unname(sapply(datasets$name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))
+
+datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))
+
+datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))
+
+clean.df <- data.frame(dataset=datasets$name,
                        object.size=datasets$object.size,

### 2014-01-12 papers/jss/article.Rnw
mostly spelling
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@762 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\author{by Dirk Eddelbuettel and Murray Stokely}
+\author{by Dirk Eddelbuettel, Murray Stokely and Jeroen Ooms}
 
@@
 verbose messages, and added complexities at the parsing side (which are
-somewhat metigated by the availability of mature libraries and
+somewhat mitigated by the availability of mature libraries and
 parsers).
@@
 Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro provide a compact
-well-documented schema for cross-langauge data structures and
+well-documented schema for cross-language data structures and
 efficient binary interchange formats.  The schema can be used to
@@
 But when the values are specified as character strings, RProtoBuf
-will automatically coerse them into a true 64-bit integer types 
+will automatically coerce them into a true 64-bit integer types 
 before storing them in the Protocol Buffer message:
@@
 There are \Sexpr{n} standard data sets included in the base-r \pkg{datasets}
-package. These datasets include data frames, matices, timeseries, tables lists,
+package. These datasets include data frames, matrices, time series, tables lists,
 and some more exotic data classes. The \texttt{can\_serialize\_pb} method can be 

### 2014-01-13 papers/jss/article.Rnw
sync
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@763 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\Sexpr{m} data sets could be converted to Protocol Buffers
-(\Sexpr{format(100*m/n,digits=1)}\%).  The next section illustrates how
-many bytes were used to store the data sets under four different
-situations:
+\Sexpr{m} data sets can be converted to Protocol Buffers
+(\Sexpr{format(100*m/n,digits=1)}\%) without loss of information. Upon closer
+inspection, all other datasets are objects of class \texttt{nfnGroupedData}.
+This class represents a special type of data frame that has some additional 
+attributes used by the \pkg{nlme} package, among which a \emph{formula} object.
+Because formulas are R \emph{language} objects, they have little meaning to
+other systems, and are not supported by the \texttt{rexp.proto} descriptor.
+When \texttt{serialize_pb} is used on objects of this class (or other objects
+containing unsupported data types), it will serialize all other values and 
+attributes of the object, but skip over the unsupported types with a warning.
+
+\subsection{Compression Performance}
+\label{sec:compression}
+
+This section compares how many bytes are used to store data sets
+using four different methods:
+
 \begin{itemize}
@@
 \end{itemize}
-
-\subsection{Compression Performance}
-\label{sec:compression}
 

### 2014-01-13 papers/jss/article.Rnw
underscore FTW
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@764 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 other systems, and are not supported by the \texttt{rexp.proto} descriptor.
-When \texttt{serialize_pb} is used on objects of this class (or other objects
+When \texttt{serialize\_pb} is used on objects of this class (or other objects
 containing unsupported data types), it will serialize all other values and 

### 2014-01-13 papers/jss/article.Rnw
second pass at data frames
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@765 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{booktabs}
+\usepackage[toc,page]{appendix}
 
@@
 
-\section{Evaluation: data.frame to Protocol Buffer Serialization}
+\section{Converting R Data Structures into Protocol Buffers}
 \label{sec:evaluation}
 
-The \pkg{RHIPE} package \citep{rhipe} also includes a Protocol integration with R.
-However, its implementation takes a different approach: any R object is
-serialized into a message based on a single catch-all \texttt{proto} schema.
-A similar approach was taken by \pkg{RProtoBufUtils} package (which has now been integrated in
-\pkg{RProtoBuf}). Unlike \pkg{RHIPE}, however, \pkg{RProtoBufUtils}
-depended upon on, and extended, \pkg{RProtoBuf} for underlying message operations.
-%DE Shall this go away now that we sucket RPBUtils into RBP?
-
-One key extension which \pkg{RProtoBufUtils} brought to \pkg{RProtoBuf} is the 
-\texttt{serialize\_pb} method to convert R objects into serialized
-Protocol Buffers in the catch-all schema. The \texttt{can\_serialize\_pb}
-method can then be used to determine whether the given R object can safely
-be expressed in this way.  To illustrate how this method works, we
-attempt to convert all of the built-in datasets from R into this
-serialized Protocol Buffer representation.
+The previous sections discussed functionality in the \pkg{RProtoBuf} package
+for creating, manipulating, parsing and serializing Protocol Buffer messages.
+In addition to these low-level methods, the package also has some high level
+functionality for automatically converting R data structures into protocol
+buffers and vice versa. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
+functions convert arbitrary R objects into a universal Protocol Buffer structure:
+
+<<>>=
+msg <- serialize_pb(iris, NULL)
+identical(iris, unserialize_pb(msg))
+@
+
+In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
+schema that \pkg{RHIPE} uses for exchanging R data with Hadoop \citep{rhipe}. This 
+schema, which we will refer to as \texttt{rexp.proto} is printed in appendix 
+\ref{rex.proto}. Even though the \texttt{RHIPE} implementation is written in Java and
+\texttt{RProtoBuf} is writting in R and \texttt{C++}, the Protocol Buffer messages
+are naturally compatible between the two systems because they use the same schema. 
+This shows the power of using a schema based cross-platform format such as Protocol 
+Buffers: interoperability is archieved without tight coordination or collaboration.
+
+\subsection{Evaluation: Converting R Data Sets}
+
+To illustrate how this method works, we attempt to convert all of the built-in 
+datasets from R into this serialized Protocol Buffer representation.
 
@@
 and some more exotic data classes. The \texttt{can\_serialize\_pb} method can be 
-used to determine which of those can fully be converted to the \textt{rexp.proto}
+used to determine which of those can fully be converted to the \texttt{rexp.proto}
 Protocol Buffer representation:
@@
 
+\newpage
+\begin{appendices}
+
+\section{The rexp.proto schema descriptor}
+\label{rexp.proto}
+
+Below a print of the \texttt{rexp.proto} schema (originally designed by \cite{rhipe})
+that is included with the \pkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
+\texttt{unserialize\_pb}.
+
+\begin{verbatim}
+package rexp;
+
+message REXP {
+  enum RClass {
+    STRING = 0;
+    RAW = 1;
+    REAL = 2;
+    COMPLEX = 3;
+    INTEGER = 4;
+    LIST = 5;
+    LOGICAL = 6;
+    NULLTYPE = 7;
+  }
+  enum RBOOLEAN {
+    F=0;
+    T=1;
+    NA=2;
+  }
+
+  required RClass rclass = 1 ; 
+  repeated double realValue = 2 [packed=true];
+  repeated sint32 intValue = 3 [packed=true];
+  repeated RBOOLEAN booleanValue = 4;
+  repeated STRING stringValue = 5;
+  optional bytes rawValue = 6;
+  repeated CMPLX complexValue = 7;
+  repeated REXP rexpValue = 8;
+  repeated string attrName = 11;
+  repeated REXP attrValue = 12;
+}
+message STRING {
+  optional string strval = 1;
+  optional bool isNA = 2 [default=false];
+}
+message CMPLX {
+  optional double real = 1 [default=0];
+  required double imag = 2;
+}
+\end{verbatim}
+\end{appendices}
+
 \bibliography{article}

### 2014-01-13 papers/jss/article.Rnw
pass 3
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@766 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 buffers and vice versa. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
-functions convert arbitrary R objects into a universal Protocol Buffer structure:
+functions serialize arbitrary R objects into a universal Protocol Buffer 
+message:
 
@@
 In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
-schema that \pkg{RHIPE} uses for exchanging R data with Hadoop \citep{rhipe}. This 
+schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
 schema, which we will refer to as \texttt{rexp.proto} is printed in appendix 
-\ref{rex.proto}. Even though the \texttt{RHIPE} implementation is written in Java and
-\texttt{RProtoBuf} is writting in R and \texttt{C++}, the Protocol Buffer messages
-are naturally compatible between the two systems because they use the same schema. 
-This shows the power of using a schema based cross-platform format such as Protocol 
-Buffers: interoperability is archieved without tight coordination or collaboration.
+\ref{rexp.proto}. The Protocol Buffer messages generated by \pkg{RProtoBuf} and
+\pkg{RHIPE} are naturally compatible between the two systems because they use the 
+same schema. This shows the power of using a schema based cross-platform format such
+as Protocol Buffers: interoperability is archieved without effort or close coordination.
+
+The \texttt{rexp.proto} schema supports all main R storage types holding \emph{data}.
+These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
+\texttt{character}, \texttt{double}, \texttt{integer} and \texttt{complex}. In addition,
+every type can contain a named set of attributes, as is the case in R. The \texttt{rexp.proto}
+schema does not support some of the special R specific storage types, such as \texttt{function},
+\texttt{language} or \texttt{environment}. Such objects have no native equivalent 
+type in Protocol Buffers, and have little meaning outside the context of R.
+When serializing R objects using \texttt{serialize\_pb}, values or attributes of
+unsupported types are skipped with a warning. If the user really wishes to serialize these 
+objects, they need to be converted into a supported type. For example, the  can use 
+\texttt{deparse} to convert functions or language objects into strings, or \texttt{as.list}
+for environments.
 
@@
 package. These datasets include data frames, matrices, time series, tables lists,
-and some more exotic data classes. The \texttt{can\_serialize\_pb} method can be 
+and some more exotic data classes. The \texttt{can\_serialize\_pb} method is 
 used to determine which of those can fully be converted to the \texttt{rexp.proto}
-Protocol Buffer representation:
+Protocol Buffer representation. This method simply checks if any of the values or
+attributes in an object is of an unsupported type:
 
 <<echo=TRUE>>=
-datasets$valid.proto <- sapply(datasets$name, function(x) can_serialize_pb(get(x)))
-datasets <- subset(datasets, valid.proto==TRUE)
-m <- nrow(datasets)
+m <- sum(sapply(datasets$name, function(x) can_serialize_pb(get(x))))
 @
@@
 other systems, and are not supported by the \texttt{rexp.proto} descriptor.
-When \texttt{serialize\_pb} is used on objects of this class (or other objects
-containing unsupported data types), it will serialize all other values and 
-attributes of the object, but skip over the unsupported types with a warning.
+When \texttt{serialize\_pb} is used on objects of this class, it will serialize
+the data frame and all attributes, except for the formula.
+
+<<>>=
+attr(CO2, "formula")
+msg <- serialize_pb(CO2, NULL)
+object <- unserialize_pb(msg)
+identical(CO2, object)
+identical(class(CO2), class(object))
+identical(dim(CO2), dim(object))
+attr(object, "formula")
+@
 
@@
 versioning,
-
-TODO comparison of protobuf serialization sizes/times for various vectors.
-Compared to R's native serialization.  Discussion of the RHIPE approach of
-serializing any/all R objects, vs more specific Protocol Buffers for specific
-R objects.
 
@@
 
+%JO: is this section really relevant? Maybe just a citation will do instead?
+
 The \texttt{RProtoBuf} package uses the user defined tables framework
@@
 such as HTTP provide mechanisms for client-server communication, i.e. how to
-initiate requests, authenticate, send messages, etc.  However, many network
+initiate requests, authenticate, send messages, etc.  However, network
 protocols generally do not regulate the \emph{content} of messages: they
@@
 OpenCPU using a protobuf message. In R, we use the HTTP client from 
-the \texttt{httr} package \citep{httr}.
-% superfluous?
-%, and the protobuf parser from the \texttt{RProtoBuf} package.
-In this example we
+the \texttt{httr} package \citep{httr}. In this example we
 download a dataset which is part of the base R distribution, so we can
@@
 
-This code suggests a method for exchanging objects between R servers, however this can 
-also be done without Protocol Buffers. The main advantage of using an inter-operable format 
+This code suggests a method for exchanging objects between R servers, however this might as 
+well be done without Protocol Buffers. The main advantage of using an inter-operable format 
 is that we can actually access R objects from within another
@@
 function call. But in its essence, the HTTP API provides a simple way to perform remote 
-R function calls over HTTPS. The same request can be performed in Python as follows:
+R function calls over HTTPS. The same request can be performed in Python as demonstrated
+below. The code is a bit verbose because to show how the REXP message is created from 
+scratch. In practice would probably write a function or small module construct a Protocol
+Buffer message representing an R list from a Python dictionary object.
 
@@
 
-The MapReduce programming model \citep{dean2008mapreduce} has emerged
-in the last decade as a popular framework for processing large data
+Over the past years, the MapReduce programming model \citep{dean2008mapreduce}
+has emerged as a poweful design pattern for processing large data
 sets in parallel on large compute clusters.  Protocol Buffers

### 2014-01-13 papers/jss/article.Rnw
Remove descriptor lookup section and incorporate a few sentences and the citation from that section into an earlier section where we first talk about the descriptorpool environment, per Jeroen's comment.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@771 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 available from the \texttt{RProtoBuf:DescriptorPool} environment in 
-the R search path. The underlying mechanism used here is
-described in more detail in Section~\ref{sec-lookup}.
+the R search path.  This environment is implemented with the user
+defined tables framework from the \pkg{RObjectTables} package
+available from the OmegaHat project \citep{RObjectTables}.  Instead of
+being associated with a static hash table, this environment
+dynamically queries the in-memory database of loaded descriptors
+during normal variable lookup.
 
 %JO: can we just move the section 7 to here? It's only one paragraph%
+%MS: I replaced section 7 with 2 sentences above.
 
@@
 @
+
+% The old section 7 in entirety:
+%The \texttt{RProtoBuf} package uses the user defined tables framework
+%that is defined as part of the \texttt{RObjectTables} package available
+%from the OmegaHat project \citep{RObjectTables}.
+%
+%The feature allows \texttt{RProtoBuf} to install the
+%special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
+%The environment is special in that, instead of being associated with a
+%static hash table, it is dynamically queried by R as part of R's usual
+%variable lookup. In other words, it means that when the R interpreter
+%looks for a binding to a symbol (foo) in its search path,
+%it asks to our package if it knows the binding "foo", this is then
+%implemented by the \texttt{RProtoBuf} package by calling an internal
+%method of the \texttt{protobuf} C++ library.
 
@@
 
-\section{Descriptor lookup}
-\label{sec-lookup}
-
-%JO: is this section really relevant? Maybe just a citation will do instead?
-
-The \texttt{RProtoBuf} package uses the user defined tables framework
-that is defined as part of the \texttt{RObjectTables} package available
-from the OmegaHat project \citep{RObjectTables}.
-
-The feature allows \texttt{RProtoBuf} to install the
-special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
-The environment is special in that, instead of being associated with a
-static hash table, it is dynamically queried by R as part of R's usual
-variable lookup. In other words, it means that when the R interpreter
-looks for a binding to a symbol (foo) in its search path,
-it asks to our package if it knows the binding "foo", this is then
-implemented by the \texttt{RProtoBuf} package by calling an internal
-method of the \texttt{protobuf} C++ library.
-
 %\section{Other approaches}

### 2014-01-14 papers/jss/article.Rnw
Edits to section 1 suggested by Karl.
1) Remove the Hadley Split-Apply-Combine reference for now as it is
confusing and more narrowly R-only than the
multi-platform-data-analysis pipeline pattern otherwise being
discussed in the first paragraph.  We might be able to add it back
with suitable distinctions added.

2) Note that our goal is not just to safely transfer the data, but to
safely _and efficiently_ do so.

3) Rewrite the second to last paragraph a bit, and add a note
specifically at the end that this paper describes an R implementation
of protocol buffers.



git-svn-id: svn://svn.r-for

@@
 isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
-Data analysis patterns such as Split-Apply-Combine
-\citep{wickham2011split} explicitly break up large problems into
-manageable pieces.  These patterns are frequently employed with
+% This is really a different pattern not connected well here.
+%Data analysis patterns such as Split-Apply-Combine
+%\citep{wickham2011split} explicitly break up large problems into manageable pieces.  
+These pipelines are frequently built with
 different programming languages used for the different phases of data
@@
 
-Given these requirements, how do we safely share intermediate results
+Given these requirements, how do we safely and efficiently share intermediate results
 between different applications, possibly written in different
@@
 well-documented schema for cross-language data structures and
-efficient binary interchange formats.  The schema can be used to
-generate model classes for statically-typed programming languages such
-as C++ and Java, or can be used with reflection for dynamically-typed
-programming languages.  Since the schema is provided separately from
-the encoded data, the data can be efficiently encoded to minimize
-storage costs of the stored data when compared with simple
-``schema-less'' binary interchange formats.
-Many sources compare data serialization formats and show Protocol
-Buffers very favorably to the alternatives; see
+efficient binary interchange formats.
+Since the schema is provided separately from the encoded data, the data can be
+efficiently encoded to minimize storage costs of the stored data when compared with simple
+``schema-less'' binary interchange formats. Many sources compare data serialization formats
+and show Protocol Buffers compare very favorably to the alternatives; see
 \citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
+The schema can be used to generate classes for statically-typed programming languages
+such as C++ and Java, or can be used with reflection for dynamically-typed programming
+languages
+
+%  The schema can be used to
+%generate model classes for statically-typed programming languages such
+%as C++ and Java, or can be used with reflection for dynamically-typed
+%programming languages.  Since the schema is provided separately from
+%the encoded data, the data can be efficiently encoded to minimize
+%storage costs of the stored data when compared with simple
+%``schema-less'' binary interchange formats.
+%Many sources compare data serialization formats and show Protocol
+%Buffers very
 
@@
 
+This paper describes an R interface to protocol buffers.
 The rest of the paper is organized as follows. Section~\ref{sec:protobuf}

### 2014-01-14 papers/jss/article.Rnw
Correct a typo: s/at time/at times/
Implement some suggestions from Karl:
* Improve the sentence comparing to Rserve
* Add more context about the structure of the protocol buffer schema.
  Section 2 is very high level / abstract, then section 3 dives into the
  R interface without ever describing field types, tag numbers, etc.
  Added three sentences or so to the beginning of section 3 before
  diving into the R examples.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@773 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % TODO(DE): This either needs a citation, or remove the name drop
-While traditional IDLs have at time been criticized for code bloat and
+% MS: These are mostly from blog posts, I can't find a good reference
+% that has a long list, and the name and year citation style seems
+% less conducive to long lists of marginal citations like blog posts
+% compared to say concise CS/math style citations [3,4,5,6]. Thoughts?
+While traditional IDLs have at times been criticized for code bloat and
 complexity, Protocol Buffers are based on a simple list and records
@@
 remote server.  The server would then deserialize the message, act on the
-request, and respond with a new Protocol Buffer over the network. The key
-difference to, say, a request to an Rserve instance is that the remote server
-may not even know the R language.
+request, and respond with a new Protocol Buffer over the network. 
+The key difference to, say, a request to an Rserve instance is that
+the remote server may be implemented in any language, with no
+dependence on R.
 
@@
 uses a reflection-based API that is particularly well-suited for
-interactive data analysis.  All messages in R have a single class
+interactive data analysis.  
+All messages in R have a single class
 structure, but different accessor methods are created at runtime based
-on the name fields of the specified message type.
+on the named fields of the specified message type.
 
@@
 column shows an example of creating a new message of this type and
-populating its fields.
+populating its fields.  A \texttt{.proto} file may contain one or more
+message types, and each message type has one or more fields.  A field
+is specified with a unique number, a name, a value type, and a field
+rule specifying whether the field is optional, required, or repeated.
+The supported value types are numbers, enumerations, booleans,
+strings, raw bytes, or other nested message types.
+The \texttt{.proto} file syntax for defining the structure of protocol
+buffer data is described comprehensively on Google Code\footnote{See 
+\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
 
@@
 or every \texttt{.proto} file provided by a particular R package.
-The \texttt{.proto} file syntax for defining the structure of protocol
-buffer data is described comprehensively on Google Code\footnote{See 
-\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
 

### 2014-01-14 papers/jss/article.Rnw
Section 5 (types): Add a sentence noting that three types in particular need special attention, to transition to the three subsections on booleans, uint32s, and int64s.
Section 6 (Converting R data structures to Protocol Buffers):
Rewrite first two sentences to make the distinction between the
previous sections describing methods for working with messages and a
specific pre-defined schema, and the function in this section
describing a universal schema that can be used on arbitrary R data
structures.  This seems like the more accurate distinction than
calling it low-level vs high-level as was d

@@
 field type and the type of data that is retrieved by \verb|$| and \verb|[[|
-extractors.
+extractors.  Three types in particular need further attention due to
+specific differences in the R language.
 
@@
 The previous sections discussed functionality in the \pkg{RProtoBuf} package
-for creating, manipulating, parsing and serializing Protocol Buffer messages.
-In addition to these low-level methods, the package also has some high level
-functionality for automatically converting R data structures into protocol
-buffers and vice versa. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
+for creating, manipulating, parsing and serializing Protocol Buffer
+messages of a specific pre-defined schema.  The package also provides
+methods for converting arbitrary R data structures into protocol
+buffers and vice versa with a universal R object schema. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
 functions serialize arbitrary R objects into a universal Protocol Buffer 
@@
 schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
-schema, which we will refer to as \texttt{rexp.proto} is printed in appendix 
+schema, which we will refer to as \texttt{rexp.proto}, is printed in appendix 
 \ref{rexp.proto}. The Protocol Buffer messages generated by \pkg{RProtoBuf} and

### 2014-01-14 papers/jss/article.Rnw
Move the MapReduce / Histogram example before the OpenCPU one and greatly improve it.  Add a very simple python example of using the protoc and a few lines of python to store arrays of bin counts and breaks as a Histogram protocol buffer.
Then, use HistogramTools to read in this histogram into R, convert it
to a native R histogram object, and plot it.

Add a reference to blocker's theoretical work on preprocessing, and a
self-citation at the end of this example section to show a real
application of the design pattern described here.

TODO: All three steps - the python code, the R code, and the

@@
 
-%\section{Basic usage example - tutorial.Person}
+\section{Application: Distributed Data Collection with MapReduce}
+\label{sec:mapreduce}
+
+Many large data sets in fields such as particle physics and information
+processing are stored in binned or histogram form in order to reduce
+the data storage requirements \citep{scott2009multivariate}.  In the
+last decade, the MapReduce programming model \citep{dean2008mapreduce}
+has emerged as a popular design pattern that enables the processing of
+very large data sets on large compute clusters.
+
+Many types of data analysis over large data sets may involve very rare
+phenomenon or be dealing with highly skewed data sets or inflexible
+raw data storage systems from which unbiased sampling is not feasible.
+In such situations, MapReduce and binning may be combined as a
+pre-processing step for a wide range of statistical and scientific
+analyses \citep{blocker2013}.
+
+There are two common patterns for generating histograms of large data
+sets in a single pass with MapReduce.  In the first method, each
+mapper task generates a histogram over a subset of the data that it
+has been assigned, serializes this histogram and sends it to one or
+more reducer tasks which merge the intermediate histograms from the
+mappers.
+
+In the second method, illustrated in
+Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
+point to a bucket width and outputs that bucket as a key and '1' as a
+value.  Reducers then sum up all of the values with the same key and
+output to a data store.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
+\end{center}
+\caption{Diagram of MapReduce Histogram Generation Pattern}
+\label{fig:mr-histogram-pattern1}
+\end{figure}
+
+In both methods, the mapper tasks must choose identical bucket
+boundaries in advance if we are to construct the histogram in a single
+pass, even though they are analyzing disjoint parts of the input set
+that may cover different ranges.  All distributed tasks involved in
+the pre-processing as well as any downstream data analysis tasks must
+share a schema of the histogram representation to coordinate
+effectively.
+
+The \pkg{HistogramTools} package \citep{histogramtools} enhances
+\pkg{RProtoBuf} by providing a concise schema for R histogram objects:
+
+\begin{example}
+package HistogramTools;
+
+message HistogramState {
+  repeated double breaks = 1;
+  repeated int32 counts = 2;
+  optional string name = 3;
+}
+\end{example}
+
+This HistogramState message type is designed to be helpful if some of
+the Map or Reduce tasks are written in R, or if those components are
+written in other languages and only the resulting output histograms
+need to be manipulated in R.  For example, to create HistogramState
+messages in Python for later consumption by R, we first compile the 
+\texttt{histogram.proto} descriptor into a python module using the
+\texttt{protoc} compiler:
+
+\begin{verbatim}
+  protoc histogram.proto --python_out=.
+\end{verbatim}
+This generates Python module called \texttt{histogram\_pb2.py}, containing both the 
+descriptor information as well as methods to read and manipulate the R object 
+message.
+
+\begin{verbatim}
+# Import modules
+from histogram_pb2 import HistogramState;
+
+# Create empty Histogram message
+hist = HistogramState()
+
+# Add breakpoints and binned data set.
+hist.counts.extend([2, 6, 2, 4, 6])
+hist.breaks.extend(range(6))
+hist.name="Example Histogram Created in Python"
+
+# Output the histogram
+outfile = open("/tmp/hist.pb", "wb")
+outfile.write(hist.SerializeToString())
+outfile.close()
+\end{verbatim}
+
+We can then read in the histogram into R and plot it with :
+
+\begin{verbatim}
+library(RProtoBuf)
+library(HistogramTools)
+
+# Read the Histogram schema
+readProtoFiles(package="HistogramTools")
+
+# Read the serialized histogram file.
+hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
+hist
+[1] "message of type 'HistogramTools.HistogramState' with 3 fields set"
+
+# Convert to native R histogram object and plot
+plot(as.histogram(hist))
+\end{verbatim}
+
+<<echo=FALSE,fig=TRUE>>=
+require(RProtoBuf)
+require(HistogramTools)
+readProtoFiles(package="HistogramTools")
+hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
+plot(as.histogram(hist))
+@
+
+One of the authors has used this design pattern for several large
+scale studies of distributed filesystems \citep{janus}.
 
@@
 \end{verbatim}
-
-\section{Application: Distributed Data Collection with MapReduce}
-\label{sec:mapreduce}
-
-Over the past years, the MapReduce programming model \citep{dean2008mapreduce}
-has emerged as a poweful design pattern for processing large data
-sets in parallel on large compute clusters.  Protocol Buffers
-provide a convenient mechanism to send structured data between tasks
-in a MapReduce cluster.  In particular, the large data sets in fields
-such as particle physics and information processing are frequently
-stored in binned or histogram form in order to reduce the data storage
-requirements for later data analysis \citep{scott2009multivariate}.
-
-In such environments, analysts may be interested in very rare
-phenomenon or be dealing with highly skewed data sets or inflexible
-raw data storage systems from which unbiased sampling is not feasible.
-There are two common patterns for generating histograms of large data
-sets in a single pass with MapReduce.  In the first method, each
-mapper task generates a histogram over a subset of the data that it
-has been assigned, serializes this histogram and sends it to one or
-more reducer tasks which merge the intermediate histograms from the
-mappers.
-
-In the second method, illustrated in
-Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
-point to a bucket width and outputs that bucket as a key and '1' as a
-value.  Reducers then sum up all of the values with the same key and
-output to a data store.
-
-In both methods, the mapper tasks must choose identical bucket
-boundaries in advance if we are to construct the histogram in a single
-pass, even though they are analyzing disjoint parts of the input set
-that may cover different ranges.  The \pkg{HistogramTools} package
-\citep{histogramtools} enhances \pkg{RProtoBuf} by providing a concise
-schema for R histogram objects.  The histogram message type is
-designed to be helpful if some of the Map or Reduce tasks are written
-in R, or if those components are written in other languages and only
-the resulting output histograms need to be manipulated in R.
-
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
-\end{center}
-\caption{Diagram of MapReduce Histogram Generation Pattern}
-\label{fig:mr-histogram-pattern1}
-\end{figure}
 

### 2014-01-14 papers/jss/article.Rnw
Dirk likes my tongue in cheek subtitle for the intro so change the section title to "Introduction: Friends don't let friends use CSV".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@784 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Introduction}
+\section{Introduction: Friends don't let friends use CSV}
 
@@
 readProtoFiles(package="HistogramTools")
-hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
+hist <- HistogramTools.HistogramState$read("hist.pb")
 plot(as.histogram(hist))

### 2014-01-14 papers/jss/article.Rnw
intermediate commit cause i’m getting lunch
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@786 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Modern data collection and analysis pipelines are increasingly being
-built using collections of components to better manage software
-complexity through reusability, modularity, and fault
-isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
+Modern data collection and analysis pipelines involve collections
+of components to enhance conrol of complex systems through
+reusability, modularity, and fault isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
 % This is really a different pattern not connected well here.
@@
 %\citep{wickham2011split} explicitly break up large problems into manageable pieces.  
-These pipelines are frequently built with
-different programming languages used for the different phases of data
-analysis -- collection, cleaning, modeling, analysis, post-processing, and
+These pipelines are frequently built using different programming 
+languages for various phases of data analysis -- collection, 
+cleaning, modeling, analysis, post-processing, and
 presentation in order to take advantage of the unique combination of
 performance, speed of development, and library support offered by
-different environments and languages.  Each stage of such a data
-analysis pipeline may involve storing intermediate results in a
-file or sending them over the network.
-
-Given these requirements, how do we safely and efficiently share intermediate results
-between different applications, possibly written in different
-languages, and possibly running on different computer system, possibly
-spanning different operating systems?  Programming
-languages such as R, Julia, Java, and Python include built-in
-serialization support, but these formats are tied to the specific
-% DE: need to define serialization?
-programming language in use and thus lock the user into a single
+each environment or language. Every stage of such a data
+analysis pipeline may produce intermediate results that need to be
+stored or sent over the network for further processing. 
+% JO Perhaps also mention that serialization is needed for distributed
+% systems to make systems scale up?
+
+Such systems require reliable and efficient exchange of intermediate
+results between the individual components, using formats that are
+independent of platform, language, operating system or architecture.
+Most technical computing languages such as R, Julia, Java, and Python
+include built-in support for serialization, but the default formats 
+are usually language specific and thereby lock the user into a single
 environment.
 
-\emph{Comma-separated values} (CSV) files can be read and written by many
-applications and so are often used for exporting tabular data.  However, CSV
-files have a number of disadvantages, such as a limitation of exporting only
-tabular datasets, lack of type-safety, inefficient text representation and
-parsing, possibly limited precision and ambiguities in the format involving
-special characters.  \emph{JavaScript Object Notation} (JSON) is another
-widely-supported format used mostly on the web that removes many of these
-disadvantages, but it too suffers from being too slow to parse and also does
-not provide strong typing between integers and floating point.  Because the
-schema information is not kept separately, multiple JSON messages of the same
-type needlessly duplicate the field names with each message.  Lastly,
-\emph{Extensible Markup Language} (XML) is a well-established and widely-supported
-protocol with the ability to define just about any arbitrarily complex
-schema. However, it pays for this complexity with comparatively large and
-verbose messages, and added complexities at the parsing side (which are
-somewhat mitigated by the availability of mature libraries and
-parsers).
-
-A number of binary formats based on JSON have been proposed that
-reduce the parsing cost and improve the efficiency.  MessagePack
-and BSON both have R interfaces, but % \citep{msgpackR,rmongodb}, but
-% DE Why do we cite these packages, but not the numerous JSON packages?
-these formats lack a separate schema for the serialized data and thus
-still duplicate field names with each message sent over the network or
-stored in a file.  Such formats also lack support for versioning when
-data storage needs evolve over time, or when application logic and
-requirement changes dictate update to the message format.
-
+Traditionally, scientists and statisticians often use character seperated
+text formats such as \texttt{CSV} \citep{shafranovich2005common} to 
+export and import data. However, anyone who has ever used this will have
+noticed that this method has many limitations: it is restricted to tabular 
+datasets, lacks type-safety, and has limited precision for numeric values.
+Moreover, ambiguities in the format itself frequently cause problems. 
+For example the default characters used as seperator and decimal point 
+are different in various parts of the world.
+\emph{Extensible Markup Language} (\texttt{XML}) is another text-based 
+well-established and widely-supported format with the ability to define
+just about any arbitrarily complex schema \citep{nolan2013xml}. However,
+it pays for this complexity with comparatively large and verbose messages,
+and added complexities at the parsing side (which are somewhat mitigated
+by the availability of mature libraries and parsers).
+A more modern, widely used format is \emph{JavaScript Object Notation}
+(\texttt{JSON}), which is derived from the object literals of
+\proglang{JavaScript}. This format is text-based as well and used mostly
+on the web. Several R packages implement functions to parse and generate
+\texttt{JSON} data from R objects. A number of \texttt{JSON} dialects has
+been proposed, such as \texttt{BSON} and \texttt{MessagePack} which both
+add binary support. However, these derivatives are not compatible with 
+existing JSON software, and have not been widely adopted.
+
+\subsection{Why Protocol Buffers}
+
+- This paper introduces another format: protocol buffers
+- unique combination of features that make it very suitable for numerical computing:
+- binary, schema, versioned, mature, high quality cross language implementations
+- we argue that (complex) statistical applications will benefit from using this format
+
+%we should probably explain what a schema is%
 Once the data serialization needs of an application become complex

### 2014-01-15 papers/jss/article.Rnw
rewrite of section 1
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@788 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Modern data collection and analysis pipelines involve collections
-of components to enhance conrol of complex systems through
-reusability, modularity, and fault isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
+of decoupled components in order to manage and control complexity 
+through reusability, modularity, and fault isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
 % This is really a different pattern not connected well here.
@@
 text formats such as \texttt{CSV} \citep{shafranovich2005common} to 
-export and import data. However, anyone who has ever used this will have
+export and import data. However, anyone who has ever used \texttt{CSV} will have
 noticed that this method has many limitations: it is restricted to tabular 
-datasets, lacks type-safety, and has limited precision for numeric values.
+data, lacks type-safety, and has limited precision for numeric values.
 Moreover, ambiguities in the format itself frequently cause problems. 
-For example the default characters used as seperator and decimal point 
-are different in various parts of the world.
-\emph{Extensible Markup Language} (\texttt{XML}) is another text-based 
+For example, conventions on which characters used as seperator and decimal
+point vary by country.
+\emph{Extensible Markup Language} (\texttt{XML}) is another
 well-established and widely-supported format with the ability to define
@@
 it pays for this complexity with comparatively large and verbose messages,
-and added complexities at the parsing side (which are somewhat mitigated
-by the availability of mature libraries and parsers).
+and added complexitiy at the parsing side (which are somewhat mitigated
+by the availability of mature libraries and parsers). Because \texttt{XML}
+is text based and has no native notion of numeric types or arrays, it 
+usually not a very practical format to store numeric datasets as they appear
+in statistical applications.
 A more modern, widely used format is \emph{JavaScript Object Notation}
 (\texttt{JSON}), which is derived from the object literals of
-\proglang{JavaScript}. This format is text-based as well and used mostly
-on the web. Several R packages implement functions to parse and generate
-\texttt{JSON} data from R objects. A number of \texttt{JSON} dialects has
-been proposed, such as \texttt{BSON} and \texttt{MessagePack} which both
-add binary support. However, these derivatives are not compatible with 
-existing JSON software, and have not been widely adopted.
+\proglang{JavaScript}, and used mostly on the web. \texttt{JSON} natively
+supports arrays and distinguishes 4 primitive types: numbers, strings, 
+booleans and null. However, because it is a text-based format, numbers are
+stored as human-readable decimal notation which is somewhat inefficient and
+leads to loss of type (double vs integer) and precision. Several R packages
+implement functions to parse and generate \texttt{JSON} data from R objects. 
+A number of \texttt{JSON} dialects has been proposed, such as \texttt{BSON} and 
+\texttt{MessagePack} which both add binary support. However, these derivatives
+are not compatible with existing JSON software, and have not been widely adopted.
 
@@
 
-- This paper introduces another format: protocol buffers
-- unique combination of features that make it very suitable for numerical computing:
-- binary, schema, versioned, mature, high quality cross language implementations
-- we argue that (complex) statistical applications will benefit from using this format
-
-%we should probably explain what a schema is%
-Once the data serialization needs of an application become complex
-enough, developers typically benefit from the use of an
-\emph{interface description language}, or \emph{IDL}.  IDLs like
-Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro provide a compact
-well-documented schema for cross-language data structures and
-efficient binary interchange formats.
-Since the schema is provided separately from the encoded data, the data can be
-efficiently encoded to minimize storage costs of the stored data when compared with simple
-``schema-less'' binary interchange formats. Many sources compare data serialization formats
-and show Protocol Buffers compare very favorably to the alternatives; see
-\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
-The schema can be used to generate classes for statically-typed programming languages
-such as C++ and Java, or can be used with reflection for dynamically-typed programming
-languages
+In 2008, Google released an open source version of Protocol Buffers: the data 
+interchange format that was designed and used for their internal infrastructure.
+Google officially provides high quality parsing libraries for \texttt{Java}, 
+\texttt{C++} and \texttt{Python}, and community developed open source implementations
+are available for many other languages. 
+Protocol Buffers take a quite different approach from many other popular formats.
+They offer a unique combination of features, performance, and maturity that seems
+particulary well suited for data-driven applications and numerical computing.
+Protocol Buffers are a binary format that natively supports all common primitive types
+found in modern programming languages. The advantage of this is that numeric values
+are serialized exactly the same way as they are stored in memory. Therefore there is
+no loss of precision, no overhead, and parsing messages is very efficient: the system can 
+simply copy bytes to memory without any further processing. 
+But the most powerful feature of protocol buffers is that it decouples the content
+from the structure using a schema, very similar to a database. This further increases
+performance by eliminating redundancy, while at the same time providing foundations
+for defining an \emph{Interface Description Language}, or \emph{IDL}.
+Many sources compare data serialization formats and show Protocol Buffers compare 
+very favorably to the alternatives; see \citet{Sumaray:2012:CDS:2184751.2184810} 
+for one such comparison.
 
@@
 
+% JO: I'm not sure where to put this paragraph. I think it is too technical
+% for the introduction section. Maybe start this section with some explanation
+% of what a schema is and then continue with showing how PB implement this?
+Once the data serialization needs of an application become complex
+enough, developers typically benefit from the use of an
+\emph{interface description language}, or \emph{IDL}.  IDLs like
+Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro provide a compact
+well-documented schema for cross-language data structures and
+efficient binary interchange formats.
+Since the schema is provided separately from the encoded data, the data can be
+efficiently encoded to minimize storage costs of the stored data when compared with simple
+``schema-less'' binary interchange formats. 
+The schema can be used to generate classes for statically-typed programming languages
+such as C++ and Java, or can be used with reflection for dynamically-typed programming
+languages.
+
 %FIXME Introductory section which may include references in parentheses
@@
 % compared to say concise CS/math style citations [3,4,5,6]. Thoughts?
+
 While traditional IDLs have at times been criticized for code bloat and

### 2014-01-15 papers/jss/article.Rnw
some tweaks to intro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@789 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Introduction: Friends don't let friends use CSV}
+\section{Introduction} % TODO(DE) More sober: Friends don't let friends use CSV}
 
@@
 
-Traditionally, scientists and statisticians often use character seperated
-text formats such as \texttt{CSV} \citep{shafranovich2005common} to 
-export and import data. However, anyone who has ever used \texttt{CSV} will have
-noticed that this method has many limitations: it is restricted to tabular 
-data, lacks type-safety, and has limited precision for numeric values.
-Moreover, ambiguities in the format itself frequently cause problems. 
-For example, conventions on which characters used as seperator and decimal
-point vary by country.
-\emph{Extensible Markup Language} (\texttt{XML}) is another
-well-established and widely-supported format with the ability to define
-just about any arbitrarily complex schema \citep{nolan2013xml}. However,
-it pays for this complexity with comparatively large and verbose messages,
-and added complexitiy at the parsing side (which are somewhat mitigated
-by the availability of mature libraries and parsers). Because \texttt{XML}
-is text based and has no native notion of numeric types or arrays, it 
-usually not a very practical format to store numeric datasets as they appear
-in statistical applications.
-A more modern, widely used format is \emph{JavaScript Object Notation}
-(\texttt{JSON}), which is derived from the object literals of
-\proglang{JavaScript}, and used mostly on the web. \texttt{JSON} natively
+%\paragraph*{Friends don't let friends use CSV!}
+Data analysts and researchers often use character seperated text formats such
+as \texttt{CSV} \citep{shafranovich2005common} to export and import
+data. However, anyone who has ever used \texttt{CSV} files will have noticed
+that this method has many limitations: it is restricted to tabular data,
+lacks type-safety, and has limited precision for numeric values.  Moreover,
+ambiguities in the format itself frequently cause problems.  For example,
+conventions on which characters is used as seperator or decimal point vary by
+country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
+well-established and widely-supported format with the ability to define just
+about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
+for this complexity with comparatively large and verbose messages, and added
+complexitiy at the parsing side (which are somewhat mitigated by the
+availability of mature libraries and parsers). Because \texttt{XML} is text
+based and has no native notion of numeric types or arrays, it usually not a
+very practical format to store numeric datasets as they appear in statistical
+applications.  A more modern, widely used format is \emph{JavaScript Object
+  Notation} (\texttt{JSON}), which is derived from the object literals of
+\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
 supports arrays and distinguishes 4 primitive types: numbers, strings,
-booleans and null. However, because it is a text-based format, numbers are
+booleans and null. However, as it too is a text-based format, numbers are
 stored as human-readable decimal notation which is somewhat inefficient and
-leads to loss of type (double vs integer) and precision. Several R packages
+leads to loss of type (double versus integer) and precision. Several R packages
 implement functions to parse and generate \texttt{JSON} data from R objects.
-A number of \texttt{JSON} dialects has been proposed, such as \texttt{BSON} and 
-\texttt{MessagePack} which both add binary support. However, these derivatives
-are not compatible with existing JSON software, and have not been widely adopted.
-
-\subsection{Why Protocol Buffers}
-
-In 2008, Google released an open source version of Protocol Buffers: the data 
+A number of \texttt{JSON} variants has been proposed, such as \texttt{BSON}
+and \texttt{MessagePack} which both add binary support. However, these
+derivatives are not compatible with existing JSON software, and have not seen
+wide adoption.
+
+%\paragraph*{Enter Protocol Buffers:}
+In 2008, and following several years of internal use, Google released an open
+source version of Protocol Buffers. It provides data  
 interchange format that was designed and used for their internal infrastructure.
-Google officially provides high quality parsing libraries for \texttt{Java}, 
-\texttt{C++} and \texttt{Python}, and community developed open source implementations
+Google officially provides high-quality parsing libraries for \texttt{Java}, 
+\texttt{C++} and \texttt{Python}, and community-developed open source implementations
 are available for many other languages. 
@@
 Protocol Buffers are a binary format that natively supports all common primitive types
-found in modern programming languages. The advantage of this is that numeric values
-are serialized exactly the same way as they are stored in memory. Therefore there is
+found in modern programming languages. A key advantage is that numeric values
+are serialized exactly the same way as they are stored in memory. There is
 no loss of precision, no overhead, and parsing messages is very efficient: the system can 
 simply copy bytes to memory without any further processing. 
-But the most powerful feature of protocol buffers is that it decouples the content
+But the most powerful feature of Protocol Buffers is that it decouples the content
 from the structure using a schema, very similar to a database. This further increases
@@
 for one such comparison.
+% TODO(DE): Mention "future proof" forward compatibility of schemata
 
@@
 
-This paper describes an R interface to protocol buffers.
-The rest of the paper is organized as follows. Section~\ref{sec:protobuf}
+This paper describes an R interface to Protocol Buffer, 
+and is organized as follows. Section~\ref{sec:protobuf}
 provides a general overview of Protocol Buffers.

### 2014-01-15 papers/jss/article.Rnw
Move the first sentence of the introduction a little back towards the direction of my original version.  Specifically, "manage and control complexity" is redundant, so move back to my "better manage software complexity".  Also add back the "increasingly", as I mentioned in email wanting to make this point specifically -- the utility of this multi-platform pattern becomes more evident with scale and larger code bases, which happens increasingly often now as computer systems get larger.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@790 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Modern data collection and analysis pipelines involve collections
-of decoupled components in order to manage and control complexity 
+Modern data collection and analysis pipelines increasingly involve collections
+of decoupled components in order to better manage software complexity 
 through reusability, modularity, and fault isolation \citep{Wegiel:2010:CTT:1932682.1869479}.

### 2014-01-15 papers/jss/article.Rnw
Add back the article 'the' in the second sentence of the introduction, and revert back to my use of 'different' rather than various.
I believe this flows better but don't have a strong preference if
someone can explain why they like the change.

Adding the article 'the' makes it more concrete/definitive.  Certainly
others could come up with a different break down of the phases of data
analysis, but that doesn't make our breakdown wrong, and thus I see no
reason to avoid the article 'the' here, if that was indeed the reason
for the change.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/

@@
 These pipelines are frequently built using different programming 
-languages for various phases of data analysis -- collection, 
+languages for the different phases of data analysis -- collection,
 cleaning, modeling, analysis, post-processing, and

### 2014-01-15 papers/jss/article.Rnw
Each is more correct than every as it is used in this sentence, so revert back to my previous usage of 'each' here, but keep some of the other nice improvements that were made to this sentence.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@792 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-15 papers/jss/article.Rnw
@@
 performance, speed of development, and library support offered by
-each environment or language. Every stage of such a data
+different environments and languages.  Each stage of such a data
 analysis pipeline may produce intermediate results that need to be
-stored or sent over the network for further processing. 
+stored in a file or sent over the network for further processing. 
 % JO Perhaps also mention that serialization is needed for distributed

### 2014-01-15 papers/jss/article.Rnw
Revert back to my motivating question to start the second paragraph of the introduction "Given these requirements, how do we X, Y, Z?"
This to me is much more engaging than the text that replaced it, and
motivating questions like this are not uncommon in JSS introductions.

The next sentence, "Most technical computing languages such as X, Y,
Z, and W" does not seem as correct as the simpler "Programming
languages X, Y, Z, and W" that I used previously.  X, Y, Z, and W do
not necessarily equate with "most technical computing languages", and
"technical computing languages" is more cumbersome tha

@@
 % systems to make systems scale up?
-
-Such systems require reliable and efficient exchange of intermediate
-results between the individual components, using formats that are
-independent of platform, language, operating system or architecture.
-Most technical computing languages such as R, Julia, Java, and Python
-include built-in support for serialization, but the default formats 
+% MS: yes perhaps somewhere near here we could define serialization
+% and describe this.
+
+Given these requirements, how do we safely and efficiently share intermediate results
+between different applications, possibly written in different
+languages, and possibly running on different computer systems?
+% Reverted to my original above, because the replacement below puts me
+% to sleep:
+%Such systems require reliable and efficient exchange of intermediate
+%results between the individual components, using formats that are
+%independent of platform, language, operating system or architecture.
+Programming
+languages such as R, Julia, Java, and Python include built-in
+support for serialization, but the default formats 
 are usually language specific and thereby lock the user into a single

### 2014-01-21 papers/jss/article.Rnw
Switch order of mapreduce/web services sections in the last paragraph of the intro since we've moved the order of them in the text.
For a figure in the mapreduce example section, make the histogram
more vertically compact to not waste so much space.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@799 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 general R language schema for serializing arbitrary R objects and evaluates
-it against R's built-in serialization.  Sections~\ref{sec:opencpu}
-and \ref{sec:mapreduce} provide real-world use cases of \CRANpkg{RProtoBuf}
-in web service and MapReduce environments, respectively, before
+it against R's built-in serialization.  Sections~\ref{sec:mapreduce}
+and \ref{sec:opencpu} provide real-world use cases of \CRANpkg{RProtoBuf}
+in MapReduce and web service environments, respectively, before
 Section~\ref{sec:summary} concludes.
@@
 
-<<echo=FALSE,fig=TRUE>>=
+<<echo=FALSE,fig=TRUE,width=8,height=4>>=
 require(RProtoBuf)

### 2014-01-21 papers/jss/article.Rnw
Revert Jereoen's two sentences about MessagePack and BSON back to my original three sentences, and add back the citations to the relevant R packages.
Jereon's version is very dismissive of these formats as not being
widely used or compatible with the almighty JSON.

My version notes that these address real deficiencies of JSON for the
application domain we are talking about in this paper, and points out
the shortcomings they still have compared to protocol buffers.

This introduction is a natural flow of alternatives each slightly
better than the last one discussed: starting with CSV, then XML

@@
 %TODO(de) 'protocol buffers' or 'Protocol Buffers' ?
+% MS: Lets standardize on 'Protocol Buffers'?
 
 \section{Introduction} % TODO(DE) More sober: Friends don't let friends use CSV}
-
+% NOTE(MS): I really do think we can use add back:
+% \section{Introduction: Friends Don't Let Friends Use CSV}
+% I didn't use proper Title Caps the first time around but really I
+% think it makes the paper more readable to have a tl;dr intro title
+% that is fun and engaging since this paper is still on the dry/boring
+% side.
 Modern data collection and analysis pipelines increasingly involve collections
@@
 very practical format to store numeric datasets as they appear in statistical
-applications.  A more modern, widely used format is \emph{JavaScript Object
+applications.
+%
+A more modern, widely used format is \emph{JavaScript Object
   Notation} (\texttt{JSON}), which is derived from the object literals of
@@
 booleans and null. However, as it too is a text-based format, numbers are
-stored as human-readable decimal notation which is somewhat inefficient and
+stored as human-readable decimal notation which is inefficient and
 leads to loss of type (double versus integer) and precision. Several R packages
 implement functions to parse and generate \texttt{JSON} data from R objects.
-A number of \texttt{JSON} variants has been proposed, such as \texttt{BSON}
-and \texttt{MessagePack} which both add binary support. However, these
-derivatives are not compatible with existing JSON software, and have not seen
-wide adoption.
+
+A number of binary formats based on \texttt{JSON} have been proposed
+that reduce the parsing cost and improve efficiency.  \pkg{MessagePack}
+\citep{msgpackR} and \pkg{BSON} \citep{rmongodb} both have R
+interfaces, but these formats lack a separate schema for the seralized
+data and thus still duplicate field names with each message sent over
+the network or stored in a file.  Such formats also lack support for
+versioning when data storage needs evolve over time, or when
+application logic and requirement changes dictate updates to the
+message format.
+
+%and \texttt{MessagePack} which both add binary support. However, these
+%derivatives are not compatible with existing JSON software, and have not seen
+%wide adoption.
 

### 2014-01-21 papers/jss/article.Rnw
Add citations for rjson and RJSONIO in the JSON section to make this comparable to the other formats described in the introduction which include citations to R packages.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@801 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Add a short sentence to define serialization early in the intro, addressing a todo that multiple people had mentioned.  Reference the C++ FAQ for lack of a better reference for now.
Revert the second to last paragraph of the introduction to my earlier
version that was instead moved to section 2.  Remove one of the more
techni

@@
 %TODO(ms) keep it less than 150 words.
+% Maybe add Jeroen's sentence:
+% They offer a unique combination of features, performance, and maturity that seems
+% particulary well suited for data-driven applications and numerical
+% computing.
 }
@@
 % systems to make systems scale up?
-% MS: yes perhaps somewhere near here we could define serialization
-% and describe this.
 
@@
 languages, and possibly running on different computer systems?
+In computer programming, \emph{serialization} is the process of
+translating data structures, variables, and session state into a
+format that can be stored or transmitted and then reconstructed in the
+original form later \citep{clinec++}.
 % Reverted to my original above, because the replacement below puts me
@@
 leads to loss of type (double versus integer) and precision. Several R packages
-implement functions to parse and generate \texttt{JSON} data from R objects.
+implement functions to parse and generate \texttt{JSON} data from R
+objects \citep{rjson,RJSONIO}.
 
@@
 
-%and \texttt{MessagePack} which both add binary support. However, these
-%derivatives are not compatible with existing JSON software, and have not seen
-%wide adoption.
+Once the data serialization needs of an application become complex
+enough, developers typically benefit from the use of an
+\emph{interface description language}, or \emph{IDL}.  IDLs like
+Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro
+provide a compact well-documented schema for cross-language data
+structures and efficient binary interchange formats.  Since the schema
+is provided separately from the encoded data, the data can be
+efficiently encoded to minimize storage costs when
+compared with simple ``schema-less'' binary interchange formats.
+Many sources compare data serialization formats
+and show Protocol Buffers compare very favorably to the alternatives; see
+\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
+
+% Too technical, move to section 2.
+% The schema can be used to generate model classes for statically-typed programming languages
+%such as C++ and Java, or can be used with reflection for dynamically-typed programming
+%languages.
+
+% TODO(mstokely): Will need to define reflection if we use it here.
+% Maybe in the next section since its not as key as 'serialization'
+% which we already defined.
 
 %\paragraph*{Enter Protocol Buffers:}
-In 2008, and following several years of internal use, Google released an open
-source version of Protocol Buffers. It provides data  
-interchange format that was designed and used for their internal infrastructure.
-Google officially provides high-quality parsing libraries for \texttt{Java}, 
-\texttt{C++} and \texttt{Python}, and community-developed open source implementations
-are available for many other languages. 
-Protocol Buffers take a quite different approach from many other popular formats.
-They offer a unique combination of features, performance, and maturity that seems
-particulary well suited for data-driven applications and numerical computing.
-Protocol Buffers are a binary format that natively supports all common primitive types
-found in modern programming languages. A key advantage is that numeric values
-are serialized exactly the same way as they are stored in memory. There is
-no loss of precision, no overhead, and parsing messages is very efficient: the system can 
-simply copy bytes to memory without any further processing. 
-But the most powerful feature of Protocol Buffers is that it decouples the content
-from the structure using a schema, very similar to a database. This further increases
-performance by eliminating redundancy, while at the same time providing foundations
-for defining an \emph{Interface Description Language}, or \emph{IDL}.
-Many sources compare data serialization formats and show Protocol Buffers compare 
-very favorably to the alternatives; see \citet{Sumaray:2012:CDS:2184751.2184810} 
-for one such comparison.
+
+% In 2008, and following several years of internal use, Google released an open
+% source version of Protocol Buffers. It provides data  
+% interchange format that was designed and used for their internal infrastructure.
+% Google officially provides high-quality parsing libraries for \texttt{Java}, 
+% \texttt{C++} and \texttt{Python}, and community-developed open source implementations
+% are available for many other languages. 
+% Protocol Buffers take a quite different approach from many other popular formats.
+
+% TODO(mstokely): Good sentence from Jeroen, add it here or sec 2.
+% They offer a unique combination of features, performance, and maturity that seems
+% particulary well suited for data-driven applications and numerical
+% computing.
+
 % TODO(DE): Mention "future proof" forward compatibility of schemata
-
-%  The schema can be used to
-%generate model classes for statically-typed programming languages such
-%as C++ and Java, or can be used with reflection for dynamically-typed
-%programming languages.  Since the schema is provided separately from
-%the encoded data, the data can be efficiently encoded to minimize
-%storage costs of the stored data when compared with simple
-%``schema-less'' binary interchange formats.
-%Many sources compare data serialization formats and show Protocol
-%Buffers very
 
@@
 
-% TODO(mstokely): If we go to JSS, include a larger paragraph here
-% referencing each numbered section.  I don't like these generally,
-% but its useful for this paper I think because we have a boring bit
-% in the middle (full class/method details) and interesting
-% applications at the end.
-
-This paper describes an R interface to Protocol Buffer, 
+% NOTE(mstokely): I don't like these roadmap paragraphs in general,
+% but it seems ueful here because we have a boring bit in the middle
+% (full class/method details) and interesting applications at the end.
+
+This paper describes an R interface to Protocol Buffers,
 and is organized as follows. Section~\ref{sec:protobuf}

### 2014-01-21 papers/jss/article.Rnw
Mostly work on section 2 to address some flaws identified by Jeroen, Karl, and others.
Move up the basic description of the protocol buffer schema from
section 3 to section 2, including the example of how protocol buffers
are manipulated with this package.

Revert a regression -- fix the reference to BSON and MessagePack by
putting the citations next to the text about the R interfaces rather
than the formats themselves (re-apply fix from Dirk).

Add an explicit transition from section 2 to section 3 as the last sentence of 2.

Define payload at the beginning of section 3 just once, so we don't

@@
 that reduce the parsing cost and improve efficiency.  \pkg{MessagePack}
-\citep{msgpackR} and \pkg{BSON} \citep{rmongodb} both have R
-interfaces, but these formats lack a separate schema for the seralized
+and \pkg{BSON} both have R
+interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
 data and thus still duplicate field names with each message sent over
@@
 % of what a schema is and then continue with showing how PB implement this?
-Once the data serialization needs of an application become complex
-enough, developers typically benefit from the use of an
-\emph{interface description language}, or \emph{IDL}.  IDLs like
-Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro provide a compact
-well-documented schema for cross-language data structures and
-efficient binary interchange formats.
-Since the schema is provided separately from the encoded data, the data can be
-efficiently encoded to minimize storage costs of the stored data when compared with simple
-``schema-less'' binary interchange formats. 
-The schema can be used to generate classes for statically-typed programming languages
-such as C++ and Java, or can be used with reflection for dynamically-typed programming
-languages.
+% MS: Yes I agree, tried to address below.
 
@@
 %%              Rewrite?
-Protocol Buffers can be described as a modern, language-neutral, platform-neutral,
-extensible mechanism for sharing and storing structured data.  Since their
-introduction, Protocol Buffers have been widely adopted in industry with
-applications as varied as %database-internal messaging (Drizzle), % DE: citation?
-Sony Playstations, Twitter, Google Search, Hadoop, and Open Street Map.  
-% TODO(DE): This either needs a citation, or remove the name drop
-% MS: These are mostly from blog posts, I can't find a good reference
-% that has a long list, and the name and year citation style seems
-% less conducive to long lists of marginal citations like blog posts
-% compared to say concise CS/math style citations [3,4,5,6]. Thoughts?
-
-While traditional IDLs have at times been criticized for code bloat and
-complexity, Protocol Buffers are based on a simple list and records
-model that is compartively flexible and simple to use.
-
-Some of the key features provided by Protocol Buffers for data analysis
-include:
+Protocol Buffers are a modern, language-neutral, platform-neutral,
+extensible mechanism for sharing and storing structured data.  Some of
+the key features provided by Protocol Buffers for data analysis include:
 
@@
 
+% Lets place this at the top of the page or the bottom, or on a float
+% page, but not just here in the middle of the page.
+\begin{figure}[tbp]
+\begin{center}
+\includegraphics[width=\textwidth]{protobuf-distributed-system-crop.pdf}
+\end{center}
+\caption{Example protobuf usage}
+\label{fig:protobuf-distributed-usecase}
+\end{figure}
+
 Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
@@
 
-%Protocol buffers are a language-neutral, platform-neutral, extensible
-%way of serializing structured data for use in communications
-%protocols, data storage, and more.
-
-%Protocol Buffers offer key features such as an efficient data interchange
-%format that is both language- and operating system-agnostic yet uses a
-%lightweight and highly performant encoding, object serialization and
-%de-serialization as well data and configuration management. Protocol
-%buffers are also forward compatible: updates to the \texttt{proto}
-%files do not break programs built against the previous specification.
-
-%While benchmarks are not available, Google states on the project page that in
-%comparison to XML, protocol buffers are at the same time \textsl{simpler},
-%between three to ten times \textsl{smaller}, between twenty and one hundred
-%times \textsl{faster}, as well as less ambiguous and easier to program.
-
-%The flexibility of the reflection-based API is particularly well
-%suited for interactive data analysis.
-
-% XXX Design tradeoffs: reflection vs proto compiler
-
-For added speed and efficiency, the C++, Java, and Python bindings to
-Protocol Buffers are used with a compiler that translates a Protocol
-Buffer schema description file (ending in \texttt{.proto}) into
-language-specific classes that can be used to create, read, write and
-manipulate Protocol Buffer messages.  The R interface, in contrast,
-uses a reflection-based API that is particularly well-suited for
-interactive data analysis.  
-All messages in R have a single class
-structure, but different accessor methods are created at runtime based
-on the named fields of the specified message type.
-
-% In other words, given the 'proto'
-%description file, code is automatically generated for the chosen
-%target language(s). The project page contains a tutorial for each of
-%these officially supported languages:
-%\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}
-
-%The protocol buffers code is released under an open-source (BSD) license. The
-%protocol buffer project (\url{http://code.google.com/p/protobuf/})
-%contains a C++ library and a set of runtime libraries and compilers for
-%C++, Java and Python.
-
-%With these languages, the workflow follows standard practice of so-called
-%Interface Description Languages (IDL)
-%(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
-%  on IDL}).  This consists of compiling a protocol buffer description file
-%(ending in \texttt{.proto}) into language specific classes that can be used
-
-%Besides the officially supported C++, Java and Python implementations, several projects have been
-%created to support protocol buffers for many languages. The list of known
-%languages to support protocol buffers is compiled as part of the
-%project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
-
-\begin{figure}[t]
-\begin{center}
-\includegraphics[width=\textwidth]{protobuf-distributed-system-crop.pdf}
-\end{center}
-\caption{Example protobuf usage}
-\label{fig:protobuf-distributed-usecase}
-\end{figure}
-
-\section{Basic Usage: Messages and Descriptors}
-\label{sec:rprotobuf-basic}
-
-This section describes how to use the R API to create and manipulate
-protocol buffer messages in R, and how to read and write the
-binary \emph{payload} of the messages to files and arbitrary binary
-R connections.
-
-The two fundamental building blocks of Protocol Buffers are \emph{Messages}
-and \emph{Descriptors}.  Messages provide a common abstract encapsulation of
-structured data fields of the type specified in a Message Descriptor.
-Message Descriptors are defined in \texttt{.proto} files and define a
-schema for a particular named class of messages.
-
+While traditional IDLs have at times been criticized for code bloat and
+complexity, Protocol Buffers are based on a simple list and records
+model that is flexible and simple to use.  The schema for structured
+protocol buffer data is defined in \texttt{.proto} files which may
+contain one or more message types.  Each message type has one or more
+fields.  A field is specified with a unique number, a name, a value
+type, and a field rule specifying whether the field is optional,
+required, or repeated.  The supported value types are numbers,
+enumerations, booleans, strings, raw bytes, or other nested message
+types.  The \texttt{.proto} file syntax for defining the structure of protocol
+buffer data is described comprehensively on Google Code\footnote{See 
+\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
 Table~\ref{tab:proto} shows an example \texttt{.proto} file which
@@
 column shows an example of creating a new message of this type and
-populating its fields.  A \texttt{.proto} file may contain one or more
-message types, and each message type has one or more fields.  A field
-is specified with a unique number, a name, a value type, and a field
-rule specifying whether the field is optional, required, or repeated.
-The supported value types are numbers, enumerations, booleans,
-strings, raw bytes, or other nested message types.
-The \texttt{.proto} file syntax for defining the structure of protocol
-buffer data is described comprehensively on Google Code\footnote{See 
-\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
-
-% Commented out because we said this earlier.
-%This separation
-%between schema and the message objects is in contrast to
-%more verbose formats like JSON, and when combined with the efficient
-%binary representation of any Message object explains a large part of
-%the performance and storage-space advantage offered by Protocol
-%Buffers. TODO(ms): we already said some of this above.  clean up.
-
-% lifted from protobuf page:
-%With Protocol Buffers you define how you want your data to be
-%structured once, and then you can read or write structured data to and
-%from a variety of data streams using a variety of different
-%languages.  The definition
+populating its fields.
 
@@
 
+% The schema can be used to generate model classes for statically-typed programming languages
+%such as C++ and Java, or can be used with reflection for dynamically-typed programming
+%languages.
+
+% TODO(mstokely): Maybe find a place to add this?  
+% Since their
+% introduction, Protocol Buffers have been widely adopted in industry with
+% applications as varied as %database-internal messaging (Drizzle), % DE: citation?
+% Sony Playstations, Twitter, Google Search, Hadoop, and Open Street
+% Map.
+
+% TODO(DE): This either needs a citation, or remove the name drop
+% MS: These are mostly from blog posts, I can't find a good reference
+% that has a long list, and the name and year citation style seems
+% less conducive to long lists of marginal citations like blog posts
+% compared to say concise CS/math style citations [3,4,5,6]. Thoughts?
+
+% The schema can be used to generate classes for statically-typed programming languages
+% such as C++ and Java, or can be used with reflection for dynamically-typed programming
+% languages.
+
+%Protocol buffers are a language-neutral, platform-neutral, extensible
+%way of serializing structured data for use in communications
+%protocols, data storage, and more.
+
+%Protocol Buffers offer key features such as an efficient data interchange
+%format that is both language- and operating system-agnostic yet uses a
+%lightweight and highly performant encoding, object serialization and
+%de-serialization as well data and configuration management. Protocol
+%buffers are also forward compatible: updates to the \texttt{proto}
+%files do not break programs built against the previous specification.
+
+For added speed and efficiency, the C++, Java, and Python bindings to
+Protocol Buffers are used with a compiler that translates a Protocol
+Buffer schema description file (ending in \texttt{.proto}) into
+language-specific classes that can be used to create, read, write and
+manipulate Protocol Buffer messages.  The R interface, in contrast,
+uses a reflection-based API that is particularly well-suited for
+interactive data analysis.  
+All messages in R have a single class
+structure, but different accessor methods are created at runtime based
+on the named fields of the specified message type, as described in the
+next section.
+
+% In other words, given the 'proto'
+%description file, code is automatically generated for the chosen
+%target language(s). The project page contains a tutorial for each of
+%these officially supported languages:
+%\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}
+
+%The protocol buffers code is released under an open-source (BSD) license. The
+%protocol buffer project (\url{http://code.google.com/p/protobuf/})
+%contains a C++ library and a set of runtime libraries and compilers for
+%C++, Java and Python.
+
+%With these languages, the workflow follows standard practice of so-called
+%Interface Description Languages (IDL)
+%(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
+%  on IDL}).  This consists of compiling a protocol buffer description file
+%(ending in \texttt{.proto}) into language specific classes that can be used
+
+%Besides the officially supported C++, Java and Python implementations, several projects have been
+%created to support protocol buffers for many languages. The list of known
+%languages to support protocol buffers is compiled as part of the
+%project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
+
+\section{Basic Usage: Messages and Descriptors}
+\label{sec:rprotobuf-basic}
+
+This section describes how to use the R API to create and manipulate
+protocol buffer messages in R, and how to read and write the
+binary representation of the message (often called the \emph{payload}) to files and arbitrary binary
+R connections.
+The two fundamental building blocks of Protocol Buffers are \emph{Messages}
+and \emph{Descriptors}.  Messages provide a common abstract encapsulation of
+structured data fields of the type specified in a Message Descriptor.
+Message Descriptors are defined in \texttt{.proto} files and define a
+schema for a particular named class of messages.
+
+% Commented out because we said this earlier.
+%This separation
+%between schema and the message objects is in contrast to
+%more verbose formats like JSON, and when combined with the efficient
+%binary representation of any Message object explains a large part of
+%the performance and storage-space advantage offered by Protocol
+%Buffers. TODO(ms): we already said some of this above.  clean up.
+
+% lifted from protobuf page:
+%With Protocol Buffers you define how you want your data to be
+%structured once, and then you can read or write structured data to and
+%from a variety of data streams using a variety of different
+%languages.  The definition
+
 %This section may contain a figure such as Figure~\ref{figure:rlogo}.
@@
 
-The binary representation of the message (often called the payload)
+The binary representation of the message
 does not contain information that can be used to dynamically
@@
 for creating, manipulating, parsing and serializing Protocol Buffer
-messages of a specific pre-defined schema.  The package also provides
-methods for converting arbitrary R data structures into protocol
+messages of a pre-defined schema.  This is useful when there are
+pre-existing systems with defined schemas or significant software
+components written in other languages that need to be accessed from
+within R.
+
+The package also provides methods for converting arbitrary R data structures into protocol
 buffers and vice versa with a universal R object schema. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}

### 2014-01-21 papers/jss/article.Rnw
spellcheck.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@806 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %\paragraph*{Friends don't let friends use CSV!}
-Data analysts and researchers often use character seperated text formats such
+Data analysts and researchers often use character separated text formats such
 as \texttt{CSV} \citep{shafranovich2005common} to export and import
@@
 ambiguities in the format itself frequently cause problems.  For example,
-conventions on which characters is used as seperator or decimal point vary by
+conventions on which characters is used as separator or decimal point vary by
 country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
@@
 for this complexity with comparatively large and verbose messages, and added
-complexitiy at the parsing side (which are somewhat mitigated by the
+complexity at the parsing side (which are somewhat mitigated by the
 availability of mature libraries and parsers). Because \texttt{XML} is text
@@
 same schema. This shows the power of using a schema based cross-platform format such
-as Protocol Buffers: interoperability is archieved without effort or close coordination.
+as Protocol Buffers: interoperability is achieved without effort or close coordination.
 

### 2014-01-21 papers/jss/article.Rnw
Add back a reference to Tierney's serialization doc.
Move my incomplete TODO performance summary subsection to the final
conclusion/summary section, where I will likely delete most of it but
may use a bit of those ideas in wrapping everything up in a
conclusion.  Ditto for the commented out other approaches section.

Section 6 previously ended on a pessimistic note and a sentence
fragment (the table shows protobuf doesn't do much better in
compression size than normal R serialization).

End on a more complete note that mentions that RProtoBuf is most
benefetial when multiple languages are invo

@@
 % MS: Yes I agree, tried to address below.
-
-%FIXME Introductory section which may include references in parentheses
-%\citep{R}, or cite a reference such as \citet{R} in the text.
 
@@
 for creating, manipulating, parsing and serializing Protocol Buffer
-messages of a pre-defined schema.  This is useful when there are
+messages of a defined schema.  This is useful when there are
 pre-existing systems with defined schemas or significant software
@@
 \begin{itemize}
-\item normal R serialization,
+\item normal R serialization \citep{serialization},
 \item R serialization followed by gzip,
@@
 in multiple languages instead of requiring other programs to parse the R
-serialization format.% \citep{serialization}.
-One takeaway from this table is that RProtoBuf does not in general provide
+serialization format. % \citep{serialization}.
+One takeaway from this table is that the universal R object schema
+included in RProtoBuf does not in general provide
 any significant saving in file size compared to the normal serialization
-mechanism in R which is seen as equally compact.  The benefit from RProtoBuf
-comes from its interoperability with other environments, as well as its safe
-versioning,
+mechanism in R.
+% redundant: which is seen as equally compact.
+The benefits of RProtoBuf accrue more naturally in applications where
+multiple programming languages are involved, or when a more concise
+application-specific schema has been defined.  The example in the next
+section provides both of these conditions.
 
@@
 
-\subsection{Performance considerations}
-
-TODO RProtoBuf is quite flexible and easy to use for interactive
-analysis, but it is not designed for certain classes of operations one
-might like to do with Protocol Buffers.  For example, taking a list of
-10,000 Protocol Buffers, extracting a named field from each one, and
-computing a aggregate statistics on those values would be extremely
-slow with RProtoBuf, and while this is a useful class of operations,
-it is outside of the scope of RProtoBuf.  We should be very clear
-about this to clarify the goals and strengths of RProtoBuf and its
-reflection and object mapping.
-
-%\section{Other approaches}
-
-% Phillip Yelland wrote another implementation, currently proprietary,
-% that has significant speed advantages when querying fields from a
-% large number of protocol buffers, but is less user friendly for the
-% basic cases documented here.
-
 \section{Application: Distributed Data Collection with MapReduce}
@@
 \label{sec:summary}
+
+TODO RProtoBuf is quite flexible and easy to use for interactive
+analysis, but it is not designed for certain classes of operations one
+might like to do with Protocol Buffers.  For example, taking a list of
+10,000 Protocol Buffers, extracting a named field from each one, and
+computing a aggregate statistics on those values would be extremely
+slow with RProtoBuf, and while this is a useful class of operations,
+it is outside of the scope of RProtoBuf.  We should be very clear
+about this to clarify the goals and strengths of RProtoBuf and its
+reflection and object mapping.
+
+%\section{Other approaches}
+
+% Phillip Yelland wrote another implementation, currently proprietary,
+% that has significant speed advantages when querying fields from a
+% large number of protocol buffers, but is less user friendly for the
+% basic cases documented here.
 

### 2014-01-21 papers/jss/article.Rnw
Comment out the OpenCPU section about generating R objects from within Python.  This code could be added to OpenCPU docs somewhere, but it is confusing in an article about RProtoBuf because it doesn't at all look language neutral as we describe many times here.
It was a good addition before we had any python code in the Histogram
application section, but now we have a more concise and clear example
of python <-> R protocol buffer usage, so we don't need this any more.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@809 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-OpenCPU also provides a lot of meta-functionality such as handling
-of sessions, exceptions, security, and more. OpenCPU also makes it possible to store
-output of a function call on the server, instead of directly retrieving it. Thereby 
-objects can be shared with other users or used as arguments in a subsequent
-function call. But in its essence, the HTTP API provides a simple way to perform remote 
-R function calls over HTTPS. The same request can be performed in Python as demonstrated
-below. The code is a bit verbose because to show how the REXP message is created from 
-scratch. In practice would probably write a function or small module construct a Protocol
-Buffer message representing an R list from a Python dictionary object.
-
-\begin{verbatim}
-import urllib2;
-from rexp_pb2 import *;
-
-#create the post payload, i.e. list(n=42, mean=100)
-payload = REXP(
-  rclass = 5,
-    rexpValue = [
-      REXP(rclass = 2, realValue = [42]), 
-      REXP(rclass = 2, realValue = [100])
-    ],
-    attrName = [
-      "names"
-    ],
-    attrValue = [
-      REXP(rclass = 0, stringValue = [STRING(strval="n"), STRING(strval="mean")])
-    ]
-);
-
-#HTTP POST
-req = urllib2.Request(
-  "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb", 
-  data = payload.SerializeToString(), 
-  headers = {
-    'Content-type': 'application/x-protobuf'
-  }
-)
-res = urllib2.urlopen(req);
-
-#parse output pb
-msg = REXP();
-msg.ParseFromString(res.read());
-
-#the return value is a double vector in this case
-print(msg.realValue);
-\end{verbatim}
-
-%\section{Application: Sending/receiving Interaction With Servers}
+% OpenCPU also provides a lot of meta-functionality such as handling
+% of sessions, exceptions, security, and more. OpenCPU also makes it possible to store
+% output of a function call on the server, instead of directly retrieving it. Thereby 
+% objects can be shared with other users or used as arguments in a subsequent
+% function call.
+%  But in its essence, the HTTP API provides a simple way to perform remote 
+% R function calls over HTTPS. The same request can be performed in Python as demonstrated
+% below. The code is a bit verbose because to show how the REXP message is created from 
+% scratch. In practice would probably write a function or small module construct a Protocol
+% Buffer message representing an R list from a Python dictionary object. 
 %
-%Combined
-%with an RPC system this means that one can interactively craft request
-%messages, send the serialized message to a remote server, read back a
-%response, and then parse the response protocol buffer interactively.
-
-%TODO(mstokely): Talk about Jeroen Ooms OpenCPU, or talk about Andy
-%Chu's Poly.
+% \begin{verbatim}
+% import urllib2;
+% from rexp_pb2 import *;
+%
+% #create the post payload, i.e. list(n=42, mean=100)
+% payload = REXP(
+%  rclass = 5,
+%    rexpValue = [
+%      REXP(rclass = 2, realValue = [42]), 
+%      REXP(rclass = 2, realValue = [100])
+%    ],
+%    attrName = [
+%      "names"
+%    ],
+%    attrValue = [
+%      REXP(rclass = 0, stringValue = [STRING(strval="n"), STRING(strval="mean")])
+%    ]
+%);
+%
+%#HTTP POST
+%req = urllib2.Request(
+%  "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb", 
+%  data = payload.SerializeToString(), 
+%  headers = {
+%    'Content-type': 'application/x-protobuf'
+%  }
+%)
+%res = urllib2.urlopen(req);
+%        
+%#parse output pb
+%msg = REXP();
+%msg.ParseFromString(res.read());
+%
+%#the return value is a double vector in this case
+%print(msg.realValue);
+%\end{verbatim}
 

### 2014-01-21 papers/jss/article.Rnw
Add an initial conclusions and commentary section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@810 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Summary}
+\section{Conclusion and Commentary}
 \label{sec:summary}
-
-TODO RProtoBuf is quite flexible and easy to use for interactive
-analysis, but it is not designed for certain classes of operations one
-might like to do with Protocol Buffers.  For example, taking a list of
-10,000 Protocol Buffers, extracting a named field from each one, and
-computing a aggregate statistics on those values would be extremely
-slow with RProtoBuf, and while this is a useful class of operations,
-it is outside of the scope of RProtoBuf.  We should be very clear
-about this to clarify the goals and strengths of RProtoBuf and its
-reflection and object mapping.
-
-%\section{Other approaches}
-
-% Phillip Yelland wrote another implementation, currently proprietary,
-% that has significant speed advantages when querying fields from a
-% large number of protocol buffers, but is less user friendly for the
-% basic cases documented here.
-
-% RProtoBuf has been used.
-
-%Its pretty useful.  Murray to see if he can get approval to talk a
-%tiny bit about how much its used at Google.
-
-%This file is only a basic article template.
-For full details of \emph{The R Journal} style and information on how to prepare your article for submission, see the \href{http://journal.r-project.org/latex/RJauthorguide.pdf}{Instructions for Authors}.
+% TODO(mstokely): Get cibona approval for these two sentences before
+% publishing.
+Schema-less text formats such as CSV and JSON will continue to be
+widely used in many contexts, but we hope that the availability of
+\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
+pipelines to embrace schemas such as Protocol Buffers for type-safe
+and performant data serialization between applications.
+
+\pkg{RProtoBuf} has been heavily used inside Google for the past three
+years by statisticians and software engineers.  At the time of this
+writing there are more than XXX 30-day active users of RProtoBuf using
+it to read data from and otherwise interact with other distributed
+systems written in C++, Java, Python, and other languages.
+
+\paragraph*{Other Approaches}
+
+\pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
+but it is not designed for efficient high-speed manipulation of large
+numbers of protocol buffers once they have been read into R.  For
+example, taking a list of 100,000 Protocol Buffers, extracting a named
+field from each one, and computing an aggregate statistic on those
+values would be relatively slow with RProtoBuf.  Instead for such a
+use case, the current design of RProtoBuf relies on other database
+systems to provide query and aggregation semantics before the
+resulting protocol buffers are read into R.  Such queries could be
+supported in a future version of \pkg{RProtoBuf} by supporting a
+vector of messages type such that \emph{slicing} operations over a
+given field across a large number of messages could be done
+efficiently in C++.
 

### 2014-01-22 papers/jss/article.Rnw
Reword the beginning of section 4 a bit, and move the \verb|method(..)| lines into an enumerated list so that we don't overflow into the margin.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@811 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 information about descriptors and messages.  Using the S4 system
-allows the \texttt{RProtoBuf} package to dispatch methods that are not
+allows the package to dispatch methods that are not
 generic in the S3 sense, such as \texttt{new} and
 \texttt{serialize}.
-
-Each R object stores an external pointer to an object managed by
-the \texttt{protobuf} C++ library which implements the core Protocol Buffer
-functionality.  The \CRANpkg{Rcpp} package
+Table~\ref{class-summary-table} lists the six
+primary Message and Descriptor classes in RProtoBuf.  Each R object
+contains an external pointer to an object managed by the
+\texttt{protobuf} C++ library, and the R objects make calls into more
+than 100 C++ functions that provide the
+glue code between the R language classes and the underlying C++
+classes.
+
+The \CRANpkg{Rcpp} package
 \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
-facilitate the integration of the R and C++ code for these objects.
+facilitate this integration of the R and C++ code for these objects.
+Each method is wrapped individually which alllows us to add user
+friendly custom error handling, type coercion, and performance
+improvements at the cost of a more verbose implementation.
+The RProtoBuf package in many ways motivated
+the development of Rcpp Modules \citep{eddelbuettel2013exposing},
+which provide a more concise way of wrapping C++ functions and classes
+in a single entity.
 
@@
 
-There are over 100 C++ functions that provide the glue code between
-the member functions of the 6 primary Message and Descriptor classes
-in the protobuf library.  Wrapping each method individually allows us
-to add user friendly custom error handling, type coercion, and
-performance improvements at the cost of a more verbose
-implementation.  The RProtoBuf implementation in many ways motivated
-the development of Rcpp Modules \citep{eddelbuettel2013exposing},
-which provide a more concise way of wrapping C++ functions and classes
-in a single entity.
-
-The \texttt{RProtoBuf} package combines a functional dispatch mechanism 
-of the form \verb|method(object, arguments)| (common to R) and the more
-traditional object oriented notation \verb|object$method(arguments)|.
+The \texttt{RProtoBuf} package supports two forms for calling
+functions with these S4 classes:
+\begin{itemize}
+\item The functional dispatch mechanism of the the form
+  \verb|method(object, arguments)| (common to R), and
+\item The traditional object oriented notation
+  \verb|object$method(arguments)|.
+\end{itemize}
+
 Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
@@
 \end{tabular}
-\caption{\label{Message-methods-table}Overview of Class, Slot, Method and
+\caption{\label{class-summary-table}Overview of Class, Slot, Method and
   Dispatch Relationships}

### 2014-01-22 papers/jss/article.Rnw
Add \shortcites{janus} so that the Google paper with ~8 authors just shows up as authorOne, et al, per the jss.pdf style guide.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@812 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Label the appendix simply as "Appendix: The rexp.proto schema descriptor" and refer to it in line as "the appendix" instead of adding a section "Appendices" with only a single appendix "A" in it.
Suggested by Dirk.

Also, add missing \proglang{} in the intro.


git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@813 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \DefineVerbatimEnvironment{example}{Verbatim}{}
-
+\shortcites{janus}
 %% almost as usual
@@
 Programming
-languages such as R, Julia, Java, and Python include built-in
+languages such as \proglang{R}, \proglang{Julia}, \proglang{Java}, and \proglang{Python} include built-in
 support for serialization, but the default formats 
@@
 schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
-schema, which we will refer to as \texttt{rexp.proto}, is printed in appendix 
-\ref{rexp.proto}. The Protocol Buffer messages generated by \pkg{RProtoBuf} and
+schema, which we will refer to as \texttt{rexp.proto}, is printed in
+%appendix \ref{rexp.proto}.
+the appendix.
+The Protocol Buffer messages generated by \pkg{RProtoBuf} and
 \pkg{RHIPE} are naturally compatible between the two systems because they use the 
@@
 
-\section{Acknowledgement}
+\section*{Acknowledgments}
 
@@
 \newpage
-\begin{appendices}
-
-\section{The rexp.proto schema descriptor}
+\appendix
+\setcounter{secnumdepth}{0}
+%\begin{appendices}
+
+\section*{Appendix: The rexp.proto schema descriptor}
 \label{rexp.proto}
@@
 \end{verbatim}
-\end{appendices}
-
+% \end{appendices}
+\newpage
 \bibliography{article}

### 2014-01-22 papers/jss/article.Rnw
copy/paste typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@814 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 This generates Python module called \texttt{histogram\_pb2.py}, containing both the 
-descriptor information as well as methods to read and manipulate the R object 
-message.
+descriptor information as well as methods to read and manipulate the histogram 
+message data.
 

### 2014-01-22 papers/jss/article.Rnw
Add reference to jsonlite
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@815 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Update the other approaches section of conclusion with some suggestions by Phillip Yelland.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@816 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \DefineVerbatimEnvironment{example}{Verbatim}{}
-\shortcites{janus}
+\shortcites{janus,dremel}
 %% almost as usual
@@
 implement functions to parse and generate \texttt{JSON} data from R
-objects \citep{rjson,RJSONIO}.
+objects \citep{rjson,RJSONIO,jsonlite}.
 
@@
 
-\section{Conclusion and Commentary}
+\section{Concluding remarks}
 \label{sec:summary}
@@
 systems written in C++, Java, Python, and other languages.
-
-\paragraph*{Other Approaches}
+\\
+
+\emph{Other Approaches}
+\\
 
@@
 field from each one, and computing an aggregate statistic on those
-values would be relatively slow with RProtoBuf.  Instead for such a
-use case, the current design of RProtoBuf relies on other database
-systems to provide query and aggregation semantics before the
-resulting protocol buffers are read into R.  Such queries could be
-supported in a future version of \pkg{RProtoBuf} by supporting a
-vector of messages type such that \emph{slicing} operations over a
-given field across a large number of messages could be done
-efficiently in C++.
+values would be relatively slow with RProtoBuf.  Mechanisms to address
+such use cases are under investigation for possible incorporation into
+future releases of RProtoBuf, but currently, the package relies on
+other database systems to provide query and aggregation semantics
+before the resulting protocol buffers are read into R.  Inside Google,
+the Dremel query system \citep{dremel} is often employed in this role
+in conjunction with \pkg{RProtoBuf}.
+
+% Such queries could be
+%supported in a future version of \pkg{RProtoBuf} by supporting a
+%vector of messages type such that \emph{slicing} operations over a
+%given field across a large number of messages could be done
+%efficiently in C++.
 

### 2014-01-22 papers/jss/article.Rnw
update conclusion
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@817 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % Maybe add Jeroen's sentence:
+% JO: added this sentence to the conclustion, but could use it in abstract as well.
 % They offer a unique combination of features, performance, and maturity that seems
@@
 % TODO(mstokely): Get cibona approval for these two sentences before
-% publishing.
-Schema-less text formats such as CSV and JSON will continue to be
-widely used in many contexts, but we hope that the availability of
-\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
-pipelines to embrace schemas such as Protocol Buffers for type-safe
-and performant data serialization between applications.
-
-\pkg{RProtoBuf} has been heavily used inside Google for the past three
-years by statisticians and software engineers.  At the time of this
-writing there are more than XXX 30-day active users of RProtoBuf using
-it to read data from and otherwise interact with other distributed
-systems written in C++, Java, Python, and other languages.
+% publishing
+Over the past decade, many formats have become available for interoperable
+data exchange, each with their unique features, strengths and weaknesses. 
+Text based formats such as CSV and JSON are easy to use and will likely 
+remain popular among statisticians for many years to come. However, in the 
+context of increasingly complex stacks and applications involving 
+distributed computing and mixed language analysis pipelines, choosing a more 
+sophisticated data interchange format will bring many benefits. 
+Protocol Buffers offer a unique combination of features, performance,
+and maturity that seems particulary well suited for data-driven 
+applications and numerical computing.
+
+The \pkg{RProtoBuf} package implements functionality to generate, 
+parse and manipulate Protocol Buffer messages in R. We hope that 
+this package will make Protocol Buffers more accessible to the R 
+community, and contributes towards better integration of R with
+other software. \pkg{RProtoBuf} has been heavily used inside Google
+for the past three years by statisticians and software engineers.
+At the time of this writing there are more than XXX 30-day active
+users of RProtoBuf using it to read data from and otherwise interact
+with other distributed systems written in C++, Java, Python, and 
+other languages.
 \\
@@
 \emph{Other Approaches}
-\\
+
+== JO: I don't really like this section here, it gives the entire paper a bit of a 
+sour aftertaste. Perhaps we can mention performance caveats in the technical
+sections? I think it's nicer to leave it at the above paragraphs.==
 

### 2014-01-22 papers/jss/article.Rnw
quick round over Summary
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@818 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-01-22 papers/jss/article.Rnw
@@
 
-\section{Concluding remarks}
+\section{Summary}  % DE Simpler title
 \label{sec:summary}
 % TODO(mstokely): Get cibona approval for these two sentences before
-% publishing.
-Schema-less text formats such as CSV and JSON will continue to be
-widely used in many contexts, but we hope that the availability of
-\pkg{RProtoBuf} makes it easy for many mixed-language data analysis
-pipelines to embrace schemas such as Protocol Buffers for type-safe
-and performant data serialization between applications.
-
-\pkg{RProtoBuf} has been heavily used inside Google for the past three
-years by statisticians and software engineers.  At the time of this
-writing there are more than XXX 30-day active users of RProtoBuf using
-it to read data from and otherwise interact with other distributed
-systems written in C++, Java, Python, and other languages.
-\\
-
-\emph{Other Approaches}
-\\
-
-\pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
-but it is not designed for efficient high-speed manipulation of large
-numbers of protocol buffers once they have been read into R.  For
-example, taking a list of 100,000 Protocol Buffers, extracting a named
-field from each one, and computing an aggregate statistic on those
-values would be relatively slow with RProtoBuf.  Mechanisms to address
-such use cases are under investigation for possible incorporation into
-future releases of RProtoBuf, but currently, the package relies on
-other database systems to provide query and aggregation semantics
-before the resulting protocol buffers are read into R.  Inside Google,
-the Dremel query system \citep{dremel} is often employed in this role
-in conjunction with \pkg{RProtoBuf}.
+% publishing
+Over the past decade, many formats for interoperable
+data exchange have become available, each with their unique features,
+strengths and weaknesses.  
+Text based formats such as CSV and JSON are easy to use, and will likely 
+remain popular among statisticians for many years to come. However, in the 
+context of increasingly complex analysis stacks and applications involving 
+distributed computing as well as mixed language analysis pipelines, choosing a more 
+sophisticated data interchange format may reap considerable benefits. 
+The Protocol Buffers protocol and librart offers a unique combination of features, performance,
+maturity, and forward-compatibility that seems particulary well suited for data-driven 
+applications and numerical computing.
+
+%% DE Re-ordering so that we end on RProtoBuf
+The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
+extends the R system with the ability to create, read and write Protocol
+Buffer message. \pkg{RProtoBuf} has been used extensively inside Google 
+for the past three years by statisticians, analysts and software engineers.
+At the time of this writing there are more than XXX 30-day active
+users of \pkg{RProtoBuf} using it to read data from and otherwise interact
+with other distributed systems written in C++, Java, Python, and 
+other languages.
+
+As the \pkg{RProtoBuf} package provides users with the ability to generate,
+parse and manipulate Protocol Buffer messages in R, it is our hope that this
+package will make Protocol Buffers more accessible to the R community, and
+thereby makes a small contribution towards better integration between R and
+other software systems and applications.
+
+%\emph{Other Approaches}
+%
+%== JO: I don't really like this section here, it gives the entire paper a bit of a 
+%sour aftertaste. Perhaps we can mention performance caveats in the technical
+%sections? I think it's nicer to leave it at the above paragraphs.==
+%
+% DE: Agreed -- commenting out
+
+%% \pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
+%% but it is not designed for efficient high-speed manipulation of large
+%% numbers of protocol buffers once they have been read into R.  For
+%% example, taking a list of 100,000 Protocol Buffers, extracting a named
+%% field from each one, and computing an aggregate statistic on those
+%% values would be relatively slow with RProtoBuf.  Mechanisms to address
+%% such use cases are under investigation for possible incorporation into
+%% future releases of RProtoBuf, but currently, the package relies on
+%% other database systems to provide query and aggregation semantics
+%% before the resulting protocol buffers are read into R.  Inside Google,
+%% the Dremel query system \citep{dremel} is often employed in this role
+%% in conjunction with \pkg{RProtoBuf}.
 
@@
 Romain Fran\c{c}ois whose continued influence on design and code is
-greatly appreciated. Several features of the package are influenced
-by the design of the \CRANpkg{rJava} package by Simon Urbanek
+greatly appreciated. Several features of the package reflect
+the design of the \CRANpkg{rJava} package by Simon Urbanek
 The user-defined table mechanism, implemented by Duncan Temple Lang for the

### 2014-01-22 papers/jss/article.Rnw
Fix a typo and then replace XXX with 300 for now.  This summary looks pretty good to me, thanks Jeroen and Dirk for greatly improving it.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@819 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 sophisticated data interchange format may reap considerable benefits. 
-The Protocol Buffers protocol and librart offers a unique combination of features, performance,
+The Protocol Buffers protocol and library offers a unique combination of features, performance,
 maturity, and forward-compatibility that seems particulary well suited for data-driven 
@@
 for the past three years by statisticians, analysts and software engineers.
-At the time of this writing there are more than XXX 30-day active
+At the time of this writing there are more than 300 30-day active
 users of \pkg{RProtoBuf} using it to read data from and otherwise interact

### 2014-01-22 papers/jss/article.Rnw
Remove the wooden 30-day qualifier about active users.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@820 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 for the past three years by statisticians, analysts and software engineers.
-At the time of this writing there are more than 300 30-day active
+At the time of this writing there are more than 300 active
 users of \pkg{RProtoBuf} using it to read data from and otherwise interact

### 2014-01-22 papers/jss/article.Rnw
Add a missing article in the mapreduce/histogram application section and try using \begin{Code} instead of \begin{verbatim} in that section (seems to be typeset the same, I may try more changes to improve the typsetting of this example).
Steve Scott noticed that one sentence in the conclusion had the word
'other' in it 3 times -- remove one superfluous word as he suggested.

Split another run-on sentence in the conclusion into two sentences as
suggested by Steve as well.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@821 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{verbatim}
-This generates Python module called \texttt{histogram\_pb2.py}, containing both the 
+This generates a Python module called \texttt{histogram\_pb2.py}, containing both the 
 descriptor information as well as methods to read and manipulate the histogram 
@@
 
-\begin{verbatim}
+\begin{Code}
 # Import modules
@@
 outfile.close()
-\end{verbatim}
+\end{Code}
 
@@
 users of \pkg{RProtoBuf} using it to read data from and otherwise interact
-with other distributed systems written in C++, Java, Python, and 
+with distributed systems written in C++, Java, Python, and 
 other languages.
 
-As the \pkg{RProtoBuf} package provides users with the ability to generate,
-parse and manipulate Protocol Buffer messages in R, it is our hope that this
+The \pkg{RProtoBuf} package provides users with the ability to generate,
+parse and manipulate Protocol Buffer messages in R.  It is our hope that this
 package will make Protocol Buffers more accessible to the R community, and

### 2014-01-22 papers/jss/article.Rnw
John Wilkes feels we should end the session in table1 with class(p) instead of beginning with it.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@822 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 p <- new(tutorial.Person,id=1,name="Dirk")
-class(p)
 p$name
@@
 serialize(p, NULL)
+class(p)
 @

### 2014-01-22 papers/jss/article.Rnw
Use \proglang and \pkg in Keywords list for proper typesetting, add Rcpp keyword.
Improve the table float placement significantly in section 4 by
letting latex do its own thing with [tbp] instead of trying to place
them [h] which never works.  With the default [tbp] it gets most of
the floats placed inside section 4 so section 5 is much more readable
now without being interspersed with floats from the previous section.

Improve the histogram / mr example slightly.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@823 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 }
-\Keywords{r, protocol buffers, serialization, cross-platform}
-\Plainkeywords{r, protocol buffers, serialization, cross-platform} %% without formatting
+\Keywords{\proglang{R}, \pkg{Rcpp}, protocol buffers, serialization, cross-platform}
+\Plainkeywords{r, Rcpp, protocol buffers, serialization, cross-platform} %% without formatting
 %% at least one keyword must be supplied
@@
 
-The \CRANpkg{Rcpp} package
-\citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
-facilitate this integration of the R and C++ code for these objects.
-Each method is wrapped individually which alllows us to add user
-friendly custom error handling, type coercion, and performance
-improvements at the cost of a more verbose implementation.
-The RProtoBuf package in many ways motivated
-the development of Rcpp Modules \citep{eddelbuettel2013exposing},
-which provide a more concise way of wrapping C++ functions and classes
-in a single entity.
-
-% Message, Descriptor, FieldDescriptor, EnumDescriptor,
-% FileDescriptor, EnumValueDescriptor
-%
-% grep RPB_FUNC * | grep -v define|wc -l
-% 84
-% grep RPB_ * | grep -v RPB_FUNCTION | grep METHOD|wc -l
-% 33
-
-The \texttt{RProtoBuf} package supports two forms for calling
-functions with these S4 classes:
-\begin{itemize}
-\item The functional dispatch mechanism of the the form
-  \verb|method(object, arguments)| (common to R), and
-\item The traditional object oriented notation
-  \verb|object$method(arguments)|.
-\end{itemize}
-
-Additionally, \texttt{RProtoBuf} implements the \texttt{.DollarNames} S3 generic function
-(defined in the \texttt{utils} package) for all classes to enable tab
-completion.  Completion possibilities include pseudo-method names for all
-classes, plus dynamic dispatch on names or types specific to a given object.
-
-% TODO(ms): Add column check box for doing dynamic dispatch based on type.
-\begin{table}[h]
+% MS: I think this looks better at the bottom of the page.
+% so it appears after the new section starts where it is referenced.
+\begin{table}[bp]
 \centering
@@
 \end{table}
+
+The \CRANpkg{Rcpp} package
+\citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
+facilitate this integration of the R and C++ code for these objects.
+Each method is wrapped individually which alllows us to add user
+friendly custom error handling, type coercion, and performance
+improvements at the cost of a more verbose implementation.
+The RProtoBuf package in many ways motivated
+the development of Rcpp Modules \citep{eddelbuettel2013exposing},
+which provide a more concise way of wrapping C++ functions and classes
+in a single entity.
+
+% Message, Descriptor, FieldDescriptor, EnumDescriptor,
+% FileDescriptor, EnumValueDescriptor
+%
+% grep RPB_FUNC * | grep -v define|wc -l
+% 84
+% grep RPB_ * | grep -v RPB_FUNCTION | grep METHOD|wc -l
+% 33
+
+The \texttt{RProtoBuf} package supports two forms for calling
+functions with these S4 classes:
+\begin{itemize}
+\item The functional dispatch mechanism of the the form
+  \verb|method(object, arguments)| (common to R), and
+\item The traditional object oriented notation
+  \verb|object$method(arguments)|.
+\end{itemize}
+
+Additionally, \pkg{RProtoBuf} supports tab completion for all
+classes.  Completion possibilities include pseudo-method names for all
+classes, plus dynamic dispatch on names or types specific to a given
+object.  This functionality is implemented with the
+\texttt{.DollarNames} S3 generic function defined in the \pkg{utils}
+package.
 
@@
 <<>>=
-# field descriptor
-tutorial.Person$email
-
-# enum descriptor
-tutorial.Person$PhoneType
-
-# nested type descriptor
-tutorial.Person$PhoneNumber
+tutorial.Person$email # field descriptor
+
+tutorial.Person$PhoneType # enum descriptor
+
+tutorial.Person$PhoneNumber # nested type descriptor
 # same as
@@
 
-\begin{table}[h]
+\begin{table}[tbp]
 \centering
@@
 The class \emph{FieldDescriptor} represents field
-descriptor in R. This is a wrapper S4 class around the
+descriptors in R. This is a wrapper S4 class around the
 \texttt{google::protobuf::FieldDescriptor} C++ class.
@@
 
-The class \emph{EnumDescriptor} is an R wrapper
-class around the C++ class \texttt{google::protobuf::EnumDescriptor}.
+The class \emph{EnumDescriptor} represents enum descriptors in R.
+This is a wrapper S4 class around the
+\texttt{google::protobuf::EnumDescriptor} C++ class.
 Table~\ref{enumdescriptor-methods-table} describes the methods
@@
 
-The class \emph{FileDescriptor} is an R wrapper
-class around the C++ class \texttt{google::protobuf::FileDescriptor}.
+The class \emph{FileDescriptor} represents file descriptors in R.
+This is a wrapper S4 class around the
+\texttt{google::protobuf::FileDescriptor} C++ class.
 Table~\ref{filedescriptor-methods-table} describes the methods
@@
 
-The class \emph{EnumValueDescriptor} is an R wrapper
-class around the C++ class \texttt{google::protobuf::EnumValueDescriptor}.
+The class \emph{EnumValueDescriptor} represents enumeration value
+descriptors in R.  This is a wrapper S4 class around the
+\texttt{google::protobuf::EnumValueDescriptor} C++ class.
 Table~\ref{EnumValueDescriptor-methods-table} describes the methods
@@
 Many types of data analysis over large data sets may involve very rare
-phenomenon or be dealing with highly skewed data sets or inflexible
+phenomenon or deal with highly skewed data sets or inflexible
 raw data storage systems from which unbiased sampling is not feasible.
@@
 descriptor information as well as methods to read and manipulate the histogram 
-message data.
+message data.  The following simple Python script uses this generated
+module to create a histogram and write out the protocol buffer
+representation to a file:
 
 \begin{Code}
-# Import modules
 from histogram_pb2 import HistogramState;
@@
 
-We can then read in the histogram into R and plot it with :
-
-\begin{verbatim}
+The protocol buffer can then be read into R and converted to a native
+R histogram object for plotting:
+
+\begin{Code}
 library(RProtoBuf)
@@
 plot(as.histogram(hist))
-\end{verbatim}
+\end{Code}
 

### 2014-01-23 papers/jss/article.Rnw
Improve the typesetting for the section 5 type coercion examples.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@824 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<>>=
-a <- new(protobuf_unittest.TestAllTypes)
-a$optional_bool <- TRUE
-a$optional_bool <- FALSE
-<<eval=FALSE>>=
-a$optional_bool <- NA
-<<echo=FALSE,eval=TRUE,print=TRUE>>=
-try(a$optional_bool <- NA,silent=TRUE)
-@
+% We want a cleaner error message here.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> a <- new(protobuf_unittest.TestAllTypes)
+R> a$optional_bool <- TRUE
+R> a$optional_bool <- FALSE
+R> a$optional_bool <- NA
+\end{CodeInput}
+\begin{CodeOutput}
+Error: NA boolean values can not be stored in bool protocol buffer fields
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 as.integer(2^31-1)
-<<eval=FALSE>>=
 as.integer(2^31 - 1) + as.integer(1)
-<<echo=FALSE,eval=TRUE,print=TRUE>>=
-try(as.integer(2^31 - 1) + as.integer(1))
-
 2^31
@@
 them as character strings.
-
-<<echo=FALSE,print=FALSE>>=
-if (!exists("protobuf_unittest.TestAllTypes",
-            "RProtoBuf:DescriptorPool")) {
-    unittest.proto.file <- system.file("unitTests", "data",
-                                       "unittest.proto",
-                                       package="RProtoBuf")
-    readProtoFiles(file=unittest.proto.file)
-}
-@
 

### 2014-01-23 papers/jss/article.Rnw
Liberal sprinkling of \proglang{} anywhere we mention R, C++, Python, or Java.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@825 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \CRANpkg{RProtoBuf} package provides a complete interface between this
-library and the R environment for statistical computing.
+library and the \proglang{R} environment for statistical computing.
 %TODO(ms) keep it less than 150 words.
@@
 \Keywords{\proglang{R}, \pkg{Rcpp}, protocol buffers, serialization, cross-platform}
-\Plainkeywords{r, Rcpp, protocol buffers, serialization, cross-platform} %% without formatting
+\Plainkeywords{R, Rcpp, protocol buffers, serialization, cross-platform} %% without formatting
 %% at least one keyword must be supplied
@@
 stored as human-readable decimal notation which is inefficient and
-leads to loss of type (double versus integer) and precision. Several R packages
-implement functions to parse and generate \texttt{JSON} data from R
+leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
+implement functions to parse and generate \texttt{JSON} data from \proglang{R}
 objects \citep{rjson,RJSONIO,jsonlite}.
@@
 
-This paper describes an R interface to Protocol Buffers,
+This paper describes an \proglang{R} interface to Protocol Buffers,
 and is organized as follows. Section~\ref{sec:protobuf}
 provides a general overview of Protocol Buffers.
-Section~\ref{sec:rprotobuf-basic} describes the interactive R interface
+Section~\ref{sec:rprotobuf-basic} describes the interactive \proglang{R} interface
 provided by \CRANpkg{RProtoBuf} and introduces the two main abstractions:
@@
 package.  Section~\ref{sec:types} describes the challenges of type coercion
-between R and other languages.  Section~\ref{sec:evaluation} introduces a
-general R language schema for serializing arbitrary R objects and evaluates
+between \proglang{R} and other languages.  Section~\ref{sec:evaluation} introduces a
+general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and evaluates
 it against R's built-in serialization.  Sections~\ref{sec:mapreduce}
@@
 %This article describes the basics of Google's Protocol Buffers through
-%an easy to use R package, \CRANpkg{RProtoBuf}.  After describing the
+%an easy to use \proglang{R} package, \CRANpkg{RProtoBuf}.  After describing the
 %basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
@@
 Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
-communication workflow with Protocol Buffers and an interactive R session.
+communication workflow with Protocol Buffers and an interactive \proglang{R} session.
 Common use cases include populating a request remote-procedure call (RPC)
-Protocol Buffer in R that is then serialized and sent over the network to a
+Protocol Buffer in \proglang{R} that is then serialized and sent over the network to a
 remote server.  The server would then deserialize the message, act on the
@@
 Table~\ref{tab:proto} shows an example \texttt{.proto} file which
-defines the \texttt{tutorial.Person} type.  The R code in the right
+defines the \texttt{tutorial.Person} type.  The \proglang{R} code in the right
 column shows an example of creating a new message of this type and
@@
 \toprule
-Schema : \texttt{addressbook.proto} & Example R Session\\
+Schema : \texttt{addressbook.proto} & Example \proglang{R} Session\\
 \cmidrule{1-2}
@@
 \caption{The schema representation from a \texttt{.proto} file for the
-  \texttt{tutorial.Person} class (left) and simple R code for creating
+  \texttt{tutorial.Person} class (left) and simple \proglang{R} code for creating
   an object of this class and accessing its fields (right).}
@@
 
-For added speed and efficiency, the C++, Java, and Python bindings to
+For added speed and efficiency, the \proglang{C++}, \proglang{Java},
+and \proglang{Python} bindings to
 Protocol Buffers are used with a compiler that translates a Protocol
@@
 language-specific classes that can be used to create, read, write and
-manipulate Protocol Buffer messages.  The R interface, in contrast,
+manipulate Protocol Buffer messages.  The \proglang{R} interface, in contrast,
 uses a reflection-based API that is particularly well-suited for
 interactive data analysis.  
-All messages in R have a single class
+All messages in \proglang{R} have a single class
 structure, but different accessor methods are created at runtime based
@@
 
-This section describes how to use the R API to create and manipulate
-protocol buffer messages in R, and how to read and write the
+This section describes how to use the \proglang{R} API to create and manipulate
+protocol buffer messages in \proglang{R}, and how to read and write the
 binary representation of the message (often called the \emph{payload}) to files and arbitrary binary
@@
 
-\subsection{Importing Message Descriptors from .proto files}
+\subsection[Importing Message Descriptors from .proto files]{Importing Message Descriptors from \texttt{.proto} files}
 
@@
 function, which can either import a single file, all files in a directory,
-or every \texttt{.proto} file provided by a particular R package.
+or every \texttt{.proto} file provided by a particular \proglang{R} package.
 
@@
 available from the \texttt{RProtoBuf:DescriptorPool} environment in 
-the R search path.  This environment is implemented with the user
+the \proglang{R} search path.  This environment is implemented with the user
 defined tables framework from the \pkg{RObjectTables} package
@@
 %The feature allows \texttt{RProtoBuf} to install the
-%special environment \emph{RProtoBuf:DescriptorPool} in the R search path.
+%special environment \emph{RProtoBuf:DescriptorPool} in the \proglang{R} search path.
 %The environment is special in that, instead of being associated with a
-%static hash table, it is dynamically queried by R as part of R's usual
-%variable lookup. In other words, it means that when the R interpreter
+%static hash table, it is dynamically queried by \proglang{R} as part of R's usual
+%variable lookup. In other words, it means that when the \proglang{R} interpreter
 %looks for a binding to a symbol (foo) in its search path,
@@
 %initial step of statically compiling the proto file into C++ code that is
-%then accessed by R code is necessary. Instead, \texttt{proto} files are
+%then accessed by \proglang{R} code is necessary. Instead, \texttt{proto} files are
 %parsed and processed \textsl{at runtime} by the protobuf C++ library---which
@@
 
-However, as opposed to R lists, no partial matching is performed
+However, as opposed to \proglang{R} lists, no partial matching is performed
 and the name must be given entirely.
@@
 
-Protocol Buffers include a 64-bit integer type, but R lacks native
+Protocol Buffers include a 64-bit integer type, but \proglang{R} lacks native
 64-bit integer support.  A workaround is available and described in
@@
 bytes that represents the message.
-%(raw vector in R speech) that represents the message.
+%(raw vector in \proglang{R} speech) that represents the message.
 
@@
 The \texttt{input} argument of \texttt{read} can also be a binary
-readable R connection, such as a binary file connection:
+readable \proglang{R} connection, such as a binary file connection:
 
@@
 Table~\ref{class-summary-table} lists the six
-primary Message and Descriptor classes in RProtoBuf.  Each R object
+primary Message and Descriptor classes in RProtoBuf.  Each \proglang{R} object
 contains an external pointer to an object managed by the
-\texttt{protobuf} C++ library, and the R objects make calls into more
-than 100 C++ functions that provide the
-glue code between the R language classes and the underlying C++
+\texttt{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
+than 100 \proglang{C++} functions that provide the
+glue code between the \proglang{R} language classes and the underlying \proglang{C++}
 classes.
@@
 \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
-facilitate this integration of the R and C++ code for these objects.
+facilitate this integration of the \proglang{R} and \proglang{C++} code for these objects.
 Each method is wrapped individually which alllows us to add user
@@
 improvements at the cost of a more verbose implementation.
-The RProtoBuf package in many ways motivated
-the development of Rcpp Modules \citep{eddelbuettel2013exposing},
-which provide a more concise way of wrapping C++ functions and classes
+The \pkg{RProtoBuf} package in many ways motivated
+the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
+which provide a more concise way of wrapping \proglang{C++} functions and classes
 in a single entity.
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{Message} object of the C++ protobuf library. Documentation for the
+\texttt{pointer} & External pointer to the \texttt{Message} object of the \proglang{C++} protobuf library. Documentation for the
 \texttt{Message} class is available from the Protocol Buffer project page. \\
@@
 %
-\texttt{str} & the R structure of the message\\
+\texttt{str} & the \proglang{R} structure of the message\\
 \texttt{as.character} & character representation of a message\\
 \texttt{toString} & character representation of a message (same as \texttt{as.character}) \\
-\texttt{as.list} & converts message to a named R list\\
+\texttt{as.list} & converts message to a named \proglang{R} list\\
 \texttt{update} & updates several fields of a message at once\\
@@
 a message contains and what the types of those fields are.  Message
-descriptors are represented in R with the \emph{Descriptor} S4
+descriptors are represented in \proglang{R} with the \emph{Descriptor} S4
 class. The class contains the slots \texttt{pointer} and
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the C++ proto library. Documentation for the
+\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the \proglang{C++} proto library. Documentation for the
 \texttt{Descriptor} class is available from the Protocol Buffer project page.\\
@@
 descriptors in R. This is a wrapper S4 class around the
-\texttt{google::protobuf::FieldDescriptor} C++ class.
+\texttt{google::protobuf::FieldDescriptor} \proglang{C++} class.
 Table~\ref{fielddescriptor-methods-table} describes the methods
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{FieldDescriptor} C++ variable \\
+\texttt{pointer} & External pointer to the \texttt{FieldDescriptor} \proglang{C++} variable \\
 \texttt{name} & Simple name of the field \\
@@
 \texttt{type} & Gets the type of the field.\\
-\texttt{cpp\_type} & Gets the C++ type of the field.\\
+\texttt{cpp\_type} & Gets the \proglang{C++} type of the field.\\
 \texttt{label} & Gets the label of a field (optional, required, or repeated).\\
@@
 This is a wrapper S4 class around the
-\texttt{google::protobuf::EnumDescriptor} C++ class.
+\texttt{google::protobuf::EnumDescriptor} \proglang{C++} class.
 Table~\ref{enumdescriptor-methods-table} describes the methods
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{EnumDescriptor} C++ variable \\
+\texttt{pointer} & External pointer to the \texttt{EnumDescriptor} \proglang{C++} variable \\
 \texttt{name} & Simple name of the enum \\
@@
 This is a wrapper S4 class around the
-\texttt{google::protobuf::FileDescriptor} C++ class.
+\texttt{google::protobuf::FileDescriptor} \proglang{C++} class.
 Table~\ref{filedescriptor-methods-table} describes the methods
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the C++ proto library. Documentation for the
+\texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the \proglang{C++} proto library. Documentation for the
 \texttt{FileDescriptor} class is available from the protocol buffer project page:
@@
 descriptors in R.  This is a wrapper S4 class around the
-\texttt{google::protobuf::EnumValueDescriptor} C++ class.
+\texttt{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
 Table~\ref{EnumValueDescriptor-methods-table} describes the methods
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} C++ variable \\
+\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} \proglang{C++} variable \\
 \texttt{name} & simple name of the enum value \\
@@
 extractors.  Three types in particular need further attention due to
-specific differences in the R language.
+specific differences in the \proglang{R} language.
 
@@
 \toprule
-Field type & R type (non repeated) & R type (repeated) \\
+Field type & \proglang{R} type (non repeated) & \proglang{R} type (repeated) \\
 \cmidrule(r){2-3}
@@
 \caption{\label{table-get-types}Correspondence between field type and
-  R type retrieved by the extractors. Note that R lacks native
+  \proglang{R} type retrieved by the extractors. Note that \proglang{R} lacks native
   64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
@@
 Buffer schema, only accept \texttt{TRUE} or \texttt{FALSE}.  This means
-that we simply can not store R logical vectors that include all three
+that we simply can not store \proglang{R} logical vectors that include all three
 possible values as booleans.  The library will refuse to store
@@
 with values $\geq 2^{31}$ can only be stored as doubles, which have
-limited precision. Thereby R loses the ability to distinguish some
+limited precision. Thereby \proglang{R} loses the ability to distinguish some
 distinct integers:
@@
 
-If we try to set an int64 field in R to double values, we lose
+If we try to set an int64 field in \proglang{R} to double values, we lose
 precision:
@@
 
-\section{Converting R Data Structures into Protocol Buffers}
+\section[Converting R Data Structures into Protocol Buffers]{Converting \proglang{R} Data Structures into Protocol Buffers}
 \label{sec:evaluation}
@@
 
-The package also provides methods for converting arbitrary R data structures into protocol
-buffers and vice versa with a universal R object schema. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
-functions serialize arbitrary R objects into a universal Protocol Buffer 
+The package also provides methods for converting arbitrary \proglang{R} data structures into protocol
+buffers and vice versa with a universal \proglang{R} object schema. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
+functions serialize arbitrary \proglang{R} objects into a universal Protocol Buffer 
 message:
@@
 In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
-schema used by \pkg{RHIPE} for exchanging R data with Hadoop \citep{rhipe}. This 
+schema used by \pkg{RHIPE} for exchanging \proglang{R} data with Hadoop \citep{rhipe}. This 
 schema, which we will refer to as \texttt{rexp.proto}, is printed in
@@
 
-The \texttt{rexp.proto} schema supports all main R storage types holding \emph{data}.
+The \texttt{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
 These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
@@
 every type can contain a named set of attributes, as is the case in R. The \texttt{rexp.proto}
-schema does not support some of the special R specific storage types, such as \texttt{function},
+schema does not support some of the special \proglang{R} specific storage types, such as \texttt{function},
 \texttt{language} or \texttt{environment}. Such objects have no native equivalent 
 type in Protocol Buffers, and have little meaning outside the context of R.
-When serializing R objects using \texttt{serialize\_pb}, values or attributes of
+When serializing \proglang{R} objects using \texttt{serialize\_pb}, values or attributes of
 unsupported types are skipped with a warning. If the user really wishes to serialize these 
@@
 
-\subsection{Evaluation: Converting R Data Sets}
+\subsection[Evaluation: Converting R Data Sets]{Evaluation: Converting \proglang{R} Data Sets}
 
 To illustrate how this method works, we attempt to convert all of the built-in 
-datasets from R into this serialized Protocol Buffer representation.
+datasets from \proglang{R} into this serialized Protocol Buffer representation.
 
@@
 attributes used by the \pkg{nlme} package, among which a \emph{formula} object.
-Because formulas are R \emph{language} objects, they have little meaning to
+Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
 other systems, and are not supported by the \texttt{rexp.proto} descriptor.
@@
 \begin{itemize}
-\item normal R serialization \citep{serialization},
-\item R serialization followed by gzip,
+\item normal \proglang{R} serialization \citep{serialization},
+\item \proglang{R} serialization followed by gzip,
 \item normal Protocol Buffer serialization, and
@@
 
-Table~\ref{tab:compression} shows the sizes of 50 sample R datasets as
+Table~\ref{tab:compression} shows the sizes of 50 sample \proglang{R} datasets as
 returned by object.size() compared to the serialized sizes.
@@
 serialization format. % \citep{serialization}.
-One takeaway from this table is that the universal R object schema
+One takeaway from this table is that the universal \proglang{R} object schema
 included in RProtoBuf does not in general provide
@@
 %
-% latex table generated in R 3.0.2 by xtable 1.7-0 package
+% latex table generated in \proglang{R} 3.0.2 by xtable 1.7-0 package
 % Fri Dec 27 17:00:03 2013
@@
 }
-\caption{Serialization sizes for default serialization in R and
-  RProtoBuf for 50 R datasets.}
+\caption{Serialization sizes for default serialization in \proglang{R} and
+  RProtoBuf for 50 \proglang{R} datasets.}
 \label{tab:compression}
@@
 The \pkg{HistogramTools} package \citep{histogramtools} enhances
-\pkg{RProtoBuf} by providing a concise schema for R histogram objects:
+\pkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
 
@@
 
-The protocol buffer can then be read into R and converted to a native
+The protocol buffer can then be read into \proglang{R} and converted to a native
 R histogram object for plotting:
@@
 
-# Convert to native R histogram object and plot
+# Convert to native \proglang{R} histogram object and plot
 plot(as.histogram(hist))
@@
 
-\subsection{Interacting with R through HTTPS and Protocol Buffers}
+\subsection[Interacting with R through HTTPS and Protocol Buffers]{Interacting with \proglang{R} through HTTPS and Protocol Buffers}
 
 One example of a system that supports Protocol Buffers to interact
-with R is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
-computation and reproducible research based on R and \LaTeX. It exposes a 
-HTTP(S) API to access and manipulate R objects and allows for performing 
-remote R function calls. Clients do not need to understand 
-or generate any R code: HTTP requests are automatically mapped to 
+with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
+computation and reproducible research based on \proglang{R} and \LaTeX. It exposes a 
+HTTP(S) API to access and manipulate \proglang{R} objects and allows for performing 
+remote \proglang{R} function calls. Clients do not need to understand 
+or generate any \proglang{R} code: HTTP requests are automatically mapped to 
 function calls, and arguments/return values can be posted/retrieved
@@
 OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
-from the \texttt{RProtoBuf} package to convert between R objects and protobuf
+from the \texttt{RProtoBuf} package to convert between \proglang{R} objects and protobuf
 messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
@@
 
-\subsection{HTTP GET: Retrieving an R object}
+\subsection[HTTP GET: Retrieving an R object]{HTTP GET: Retrieving an \proglang{R} object}
 
@@
 languages, clients can be implemented in just a few lines of code. Below
-is example code for both R and Python that retrieves a dataset from R with 
+is example code for both \proglang{R} and Python that retrieves a dataset from \proglang{R} with 
 OpenCPU using a protobuf message. In R, we use the HTTP client from 
 the \texttt{httr} package \citep{httr}. In this example we
-download a dataset which is part of the base R distribution, so we can
+download a dataset which is part of the base \proglang{R} distribution, so we can
 verify that the object was transferred without loss of information.
@@
 
-This code suggests a method for exchanging objects between R servers, however this might as 
+This code suggests a method for exchanging objects between \proglang{R} servers, however this might as 
 well be done without Protocol Buffers. The main advantage of using an inter-operable format 
-is that we can actually access R objects from within another
+is that we can actually access \proglang{R} objects from within another
 programming language. For example, in a very similar fashion we can retrieve the same
@@
 This generates Python module called \texttt{rexp\_pb2.py}, containing both the 
-descriptor information as well as methods to read and manipulate the R object 
+descriptor information as well as methods to read and manipulate the \proglang{R} object 
 message. In the example below we use the HTTP client from the \texttt{urllib2}
@@
 
-\subsection{HTTP POST: Calling an R function}
+\subsection[HTTP POST: Calling an R function]{HTTP POST: Calling an \proglang{R} function}
 
 The example above shows how the \texttt{HTTP GET} method retrieves a 
-resource from OpenCPU, for example an R object. The \texttt{HTTP POST} 
+resource from OpenCPU, for example an \proglang{R} object. The \texttt{HTTP POST} 
 method on the other hand is used for calling functions and running scripts, 
@@
 This is a bit more work, because clients needs to both generate messages 
-containing R objects to post to the server, as well as retrieve and parse
+containing \proglang{R} objects to post to the server, as well as retrieve and parse
 protobuf messages returned by the server. Using Protocol Buffers to post
@@
 However, with protocol buffers the client can perform function calls with
-more complex arguments such as R vectors or lists. The result is a complete
-RPC system to do arbitrary R function calls from within 
+more complex arguments such as \proglang{R} vectors or lists. The result is a complete
+RPC system to do arbitrary \proglang{R} function calls from within 
 any programming language.
 
-The following example R client code performs the remote function call 
+The following example \proglang{R} client code performs the remote function call 
 \texttt{stats::rnorm(n=42, mean=100)}. The function arguments (in this
@@
 %  But in its essence, the HTTP API provides a simple way to perform remote 
-% R function calls over HTTPS. The same request can be performed in Python as demonstrated
+% \proglang{R} function calls over HTTPS. The same request can be performed in Python as demonstrated
 % below. The code is a bit verbose because to show how the REXP message is created from 
 % scratch. In practice would probably write a function or small module construct a Protocol
-% Buffer message representing an R list from a Python dictionary object. 
+% Buffer message representing an \proglang{R} list from a Python dictionary object. 
 %
@@
 The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
-extends the R system with the ability to create, read and write Protocol
+extends the \proglang{R} system with the ability to create, read and write Protocol
 Buffer message. \pkg{RProtoBuf} has been used extensively inside Google 
@@
 users of \pkg{RProtoBuf} using it to read data from and otherwise interact
-with distributed systems written in C++, Java, Python, and 
+with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
 other languages.
@@
 parse and manipulate Protocol Buffer messages in R.  It is our hope that this
-package will make Protocol Buffers more accessible to the R community, and
-thereby makes a small contribution towards better integration between R and
+package will make Protocol Buffers more accessible to the \proglang{R} community, and
+thereby makes a small contribution towards better integration between \proglang{R} and
 other software systems and applications.
@@
 %given field across a large number of messages could be done
-%efficiently in C++.
+%efficiently in \proglang{C++}.
 

### 2014-01-23 papers/jss/article.Rnw
mix of minor fixes from the train ride
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@826 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   1600 Amphitheatre Parkway\\
-  Mountain View, CA 94040\\
-  USA\\
+  Mountain View, CA, USA\\
   E-mail: \email{mstokely@google.com}\\
@@
   University of California\\
+  Los Angeles, CA, USA\\
   E-mail: \email{jeroen.ooms@stat.ucla.edu}\\
@@
 and is organized as follows. Section~\ref{sec:protobuf}
-provides a general overview of Protocol Buffers.
+provides a general high-level overview of Protocol Buffers as well as a basic
+motivation for their use.
 Section~\ref{sec:rprotobuf-basic} describes the interactive \proglang{R} interface
-provided by \CRANpkg{RProtoBuf} and introduces the two main abstractions:
+provided by the \CRANpkg{RProtoBuf} package, and introduces the two main abstractions:
 \emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
-describes the implementation details of the main S4 classes making up this
+details the implementation details of the main S4 classes and methods
+contained in this
 package.  Section~\ref{sec:types} describes the challenges of type coercion
@@
 general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and evaluates
-it against R's built-in serialization.  Sections~\ref{sec:mapreduce}
+it against the serialization capbilities built directly into R.  Sections~\ref{sec:mapreduce}
 and \ref{sec:opencpu} provide real-world use cases of \CRANpkg{RProtoBuf}
@@
 This class represents a special type of data frame that has some additional 
-attributes used by the \pkg{nlme} package, among which a \emph{formula} object.
+attributes (such as a \emph{formula} object) used by the \pkg{nlme} package.
 Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
@@
 One takeaway from this table is that the universal \proglang{R} object schema
-included in RProtoBuf does not in general provide
+included in \pkg{RProtoBuf} does not in general provide
 any significant saving in file size compared to the normal serialization
@@
 % redundant: which is seen as equally compact.
-The benefits of RProtoBuf accrue more naturally in applications where
+The benefits of \pkg{RProtoBuf} accrue more naturally in applications where
 multiple programming languages are involved, or when a more concise
 application-specific schema has been defined.  The example in the next
-section provides both of these conditions.
+section satisfies both of these conditions.
 
@@
 
+\begin{center}
 <<echo=FALSE,fig=TRUE,width=8,height=4>>=
@@
 @
+\end{center}
 
@@
 interchange in the context of inter-system communications.  Network protocols
-such as HTTP provide mechanisms for client-server communication, i.e. how to
+such as HTTP provide mechanisms for client-server communication, i.e.~how to
 initiate requests, authenticate, send messages, etc.  However, network
@@
 
-%\section[About Java]{About \proglang{Java}}
-%% Note: If there is markup in \(sub)section, then it has to be escape as above.
-
 \end{document}

### 2014-01-23 papers/jss/article.Rnw
Remove duplicate \title line, add a few missing \proglangs, and emphasize dynamic dispatch on first use since its a rather technical term which we only partially define.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@827 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \DefineVerbatimEnvironment{example}{Verbatim}{}
+% Articles with many authors we should shorten to FirstAuthor, et al.
 \shortcites{janus,dremel}
-%% almost as usual
 \author{Dirk Eddelbuettel\\Debian Project \And 
@@
         Jeroen Ooms\\UCLA}
-\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in R}
+\title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}}
 
@@
 \Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in R} %% a short title (if necessary)
+\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
 
@@
 
-\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\author{by Dirk Eddelbuettel, Murray Stokely and Jeroen Ooms}
+%\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
+%\author{by Dirk Eddelbuettel, Murray Stokely and Jeroen Ooms}
 
@@
 classes.  Completion possibilities include pseudo-method names for all
-classes, plus dynamic dispatch on names or types specific to a given
+classes, plus \emph{dynamic dispatch} on names or types specific to a given
 object.  This functionality is implemented with the

### 2014-01-23 papers/jss/article.Rnw
Add more \proglangs, we now have \proglang{R} at least 119 times in this document, which might be a bit much.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@828 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 that reduce the parsing cost and improve efficiency.  \pkg{MessagePack}
-and \pkg{BSON} both have R
+and \pkg{BSON} both have \proglang{R}
 interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
@@
 general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and evaluates
-it against the serialization capbilities built directly into R.  Sections~\ref{sec:mapreduce}
+it against the serialization capbilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
 and \ref{sec:opencpu} provide real-world use cases of \CRANpkg{RProtoBuf}
@@
 request, and respond with a new Protocol Buffer over the network. 
-The key difference to, say, a request to an Rserve instance is that
+The key difference to, say, a request to an \pkg{Rserve} instance is that
 the remote server may be implemented in any language, with no
-dependence on R.
+dependence on \proglang{R}.
 
@@
 binary representation of the message (often called the \emph{payload}) to files and arbitrary binary
-R connections.
+\proglang{R} connections.
 The two fundamental building blocks of Protocol Buffers are \emph{Messages}
@@
 %languages.  The definition
-
-%This section may contain a figure such as Figure~\ref{figure:rlogo}.
-%
-%\begin{figure}[htbp]
-%  \centering
-%  \includegraphics{Rlogo}
-%  \caption{The logo of R.}
-%  \label{figure:rlogo}
-%\end{figure}
 
@@
 Once the message is created, its fields can be queried
-and modified using the dollar operator of R, making protocol
+and modified using the dollar operator of \proglang{R}, making protocol
 buffer messages seem like lists.
@@
 Table~\ref{class-summary-table} lists the six
-primary Message and Descriptor classes in RProtoBuf.  Each \proglang{R} object
+primary Message and Descriptor classes in \CRANpkg{RProtoBuf}.  Each \proglang{R} object
 contains an external pointer to an object managed by the
@@
 \item The functional dispatch mechanism of the the form
-  \verb|method(object, arguments)| (common to R), and
+  \verb|method(object, arguments)| (common to \proglang{R}), and
 \item The traditional object oriented notation
@@
 The class \emph{FieldDescriptor} represents field
-descriptors in R. This is a wrapper S4 class around the
+descriptors in \proglang{R}. This is a wrapper S4 class around the
 \texttt{google::protobuf::FieldDescriptor} \proglang{C++} class.
@@
 
-The class \emph{EnumDescriptor} represents enum descriptors in R.
+The class \emph{EnumDescriptor} represents enum descriptors in \proglang{R}.
 This is a wrapper S4 class around the
@@
 
-The class \emph{FileDescriptor} represents file descriptors in R.
+The class \emph{FileDescriptor} represents file descriptors in \proglang{R}.
 This is a wrapper S4 class around the
@@
 The class \emph{EnumValueDescriptor} represents enumeration value
-descriptors in R.  This is a wrapper S4 class around the
+descriptors in \proglang{R}.  This is a wrapper S4 class around the
 \texttt{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
@@
 
-R booleans can accept three values: \texttt{TRUE}, \texttt{FALSE}, and
+\proglang{R} booleans can accept three values: \texttt{TRUE}, \texttt{FALSE}, and
 \texttt{NA}.  However, most other languages, including the Protocol
@@
 
-R lacks a native unsigned integer type.  Values between $2^{31}$ and
+\proglang{R} lacks a native unsigned integer type.  Values between $2^{31}$ and
 $2^{32} - 1$ read from unsigned into Protocol Buffer fields must be
-stored as doubles in R.
+stored as doubles in \proglang{R}.
 
@@
 
-R also does not support the native 64-bit integer type. Numeric vectors
+\proglang{R} also does not support the native 64-bit integer type. Numeric vectors
 with values $\geq 2^{31}$ can only be stored as doubles, which have
@@
 However, most modern languages do have support for 64-bit integers, 
-which becomes problematic when \pkg{RProtoBuf} is used to exchange data 
+which becomes problematic when \CRANpkg{RProtoBuf} is used to exchange data 
 with a system that requires this integer type. To work around this, 
-RProtoBuf allows users to get and set 64-bit integer values by specifying 
+\CRANpkg{RProtoBuf} allows users to get and set 64-bit integer values by specifying 
 them as character strings.
@@
 
-But when the values are specified as character strings, RProtoBuf
+But when the values are specified as character strings, \CRANpkg{RProtoBuf}
 will automatically coerce them into a true 64-bit integer types 
@@
 
-When reading the value back into R, numeric types are returned by
+When reading the value back into \proglang{R}, numeric types are returned by
 default, but when the full precision is required a character value
@@
 to \texttt{TRUE}.  The character values are useful because they can
-accurately be used as unique identifiers and can easily be passed to R
+accurately be used as unique identifiers and can easily be passed to \proglang{R}
 packages such as \CRANpkg{int64} \citep{int64} or \CRANpkg{bit64}
-\citep{bit64} which represent 64-bit integers in R.
+\citep{bit64} which represent 64-bit integers in \proglang{R}.
 
@@
 components written in other languages that need to be accessed from
-within R.
+within \proglang{R}.
 
@@
 \texttt{character}, \texttt{double}, \texttt{integer} and \texttt{complex}. In addition,
-every type can contain a named set of attributes, as is the case in R. The \texttt{rexp.proto}
+every type can contain a named set of attributes, as is the case in \proglang{R}. The \texttt{rexp.proto}
 schema does not support some of the special \proglang{R} specific storage types, such as \texttt{function},
 \texttt{language} or \texttt{environment}. Such objects have no native equivalent 
-type in Protocol Buffers, and have little meaning outside the context of R.
+type in Protocol Buffers, and have little meaning outside the context of \proglang{R}.
 When serializing \proglang{R} objects using \texttt{serialize\_pb}, values or attributes of
@@
 Sizes are comparable but Protocol Buffers provide simple getters and setters
-in multiple languages instead of requiring other programs to parse the R
+in multiple languages instead of requiring other programs to parse the \proglang{R}
 serialization format. % \citep{serialization}.
@@
 any significant saving in file size compared to the normal serialization
-mechanism in R.
+mechanism in \proglang{R}.
 % redundant: which is seen as equally compact.
@@
   \toprule
-  Data Set & object.size & \multicolumn{2}{c}{R Serialization} &
+  Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
   \multicolumn{2}{c}{RProtoBuf Serial.} \\
@@
 This HistogramState message type is designed to be helpful if some of
-the Map or Reduce tasks are written in R, or if those components are
+the Map or Reduce tasks are written in \proglang{R}, or if those components are
 written in other languages and only the resulting output histograms
-need to be manipulated in R.  For example, to create HistogramState
-messages in Python for later consumption by R, we first compile the 
+need to be manipulated in \proglang{R}.  For example, to create HistogramState
+messages in Python for later consumption by \proglang{R}, we first compile the 
 \texttt{histogram.proto} descriptor into a python module using the
@@
 The protocol buffer can then be read into \proglang{R} and converted to a native
-R histogram object for plotting:
+\proglang{R} histogram object for plotting:
 
@@
 is example code for both \proglang{R} and Python that retrieves a dataset from \proglang{R} with 
-OpenCPU using a protobuf message. In R, we use the HTTP client from 
+OpenCPU using a protobuf message. In \proglang{R}, we use the HTTP client from 
 the \texttt{httr} package \citep{httr}. In this example we
@@
 with 42 random numbers) are transferred using a protobuf message. RPC in
-OpenCPU works like the \texttt{do.call} function in R, hence all arguments
+OpenCPU works like the \texttt{do.call} function in \proglang{R}, hence all arguments
 are contained within a list.
@@
 The \pkg{RProtoBuf} package provides users with the ability to generate,
-parse and manipulate Protocol Buffer messages in R.  It is our hope that this
+parse and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
 package will make Protocol Buffers more accessible to the \proglang{R} community, and

### 2014-01-23 papers/jss/article.Rnw
Add second reference where the mapreduce histogram pattern is used since we note we are aware of multiple applications.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@829 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Standardize on using the oxford/serial comma with lists of 3 or more. This seemed the far more prevalent convention in the file but there were 5 instances where it was not used.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@830 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % Articles with many authors we should shorten to FirstAuthor, et al.
-\shortcites{janus,dremel}
+\shortcites{sciencecloud,janus,dremel}
 \author{Dirk Eddelbuettel\\Debian Project \And 
@@
 supports arrays and distinguishes 4 primitive types: numbers, strings,
-booleans and null. However, as it too is a text-based format, numbers are
+booleans, and null. However, as it too is a text-based format, numbers are
 stored as human-readable decimal notation which is inefficient and
@@
 Buffer schema description file (ending in \texttt{.proto}) into
-language-specific classes that can be used to create, read, write and
+language-specific classes that can be used to create, read, write, and
 manipulate Protocol Buffer messages.  The \proglang{R} interface, in contrast,
@@
 These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
-\texttt{character}, \texttt{double}, \texttt{integer} and \texttt{complex}. In addition,
+\texttt{character}, \texttt{double}, \texttt{integer}, and \texttt{complex}. In addition,
 every type can contain a named set of attributes, as is the case in \proglang{R}. The \texttt{rexp.proto}
@@
 One of the authors has used this design pattern for several large
-scale studies of distributed filesystems \citep{janus}.
+scale studies of distributed storage systems \citep{sciencecloud,janus}.
 
@@
 The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
-extends the \proglang{R} system with the ability to create, read and write Protocol
-Buffer message. \pkg{RProtoBuf} has been used extensively inside Google 
-for the past three years by statisticians, analysts and software engineers.
+extends the \proglang{R} system with the ability to create, read, and write Protocol
+Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
+for the past three years by statisticians, analysts, and software engineers.
 At the time of this writing there are more than 300 active
@@
 The \pkg{RProtoBuf} package provides users with the ability to generate,
-parse and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
+parse, and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
 package will make Protocol Buffers more accessible to the \proglang{R} community, and

### 2014-01-23 papers/jss/article.Rnw
Minor grammar improvements suggested by my office mate Arif Merchant. Mostly involving that/which.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@831 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 complexity, Protocol Buffers are based on a simple list and records
-model that is flexible and simple to use.  The schema for structured
-protocol buffer data is defined in \texttt{.proto} files which may
+model that is flexible and easy to use.  The schema for structured
+protocol buffer data is defined in \texttt{.proto} files, which may
 contain one or more message types.  Each message type has one or more
@@
 \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
-Table~\ref{tab:proto} shows an example \texttt{.proto} file which
+Table~\ref{tab:proto} shows an example \texttt{.proto} file that
 defines the \texttt{tutorial.Person} type.  The \proglang{R} code in the right
@@
 
-However, as opposed to \proglang{R} lists, no partial matching is performed
+As opposed to \proglang{R} lists, no partial matching is performed
 and the name must be given entirely.

### 2014-01-23 papers/jss/article.Rnw
Add two minor text improvements suggested by Tim Hesterberg to sections 5 and 6.
The first sentence of each of the last two paragraphs of the summary
was saying the same thing -- merge the second duplicate sentence into
the first, which then leaves the third paragraph of the summary as
just a single sentence.  Add it to the end of the second sentence and
we now have I think a slightly crisper summary in just two paragraphs.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@832 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 extractors.  Three types in particular need further attention due to
-specific differences in the \proglang{R} language.
+specific differences in the \proglang{R} language: booleans, unsigned
+integers, and 64-bit integers.
 
@@
 \Sexpr{m} data sets can be converted to Protocol Buffers
-(\Sexpr{format(100*m/n,digits=1)}\%) without loss of information. Upon closer
+without loss of information (\Sexpr{format(100*m/n,digits=1)}\%). Upon closer
 inspection, all other datasets are objects of class \texttt{nfnGroupedData}.
@@
 \label{sec:summary}
-% TODO(mstokely): Get cibona approval for these two sentences before
-% publishing
 Over the past decade, many formats for interoperable
@@
 The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
-extends the \proglang{R} system with the ability to create, read, and write Protocol
+extends the \proglang{R} system with the ability to create, read,
+write, parse, and manipulate Protocol
 Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
@@
 with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
-other languages.
-
-The \pkg{RProtoBuf} package provides users with the ability to generate,
-parse, and manipulate Protocol Buffer messages in \proglang{R}.  It is our hope that this
+other languages.  It is our hope that this
 package will make Protocol Buffers more accessible to the \proglang{R} community, and

### 2014-01-23 papers/jss/article.Rnw
large scale -> large-scale protocol buffers -> Protocol Buffers
And remove some old commented out sections now.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@833 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[toc,page]{appendix}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+% Spelling Standardization:
+% Protocol Buffers, not protocol buffers
 
@@
 
-%\title{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-%\author{by Dirk Eddelbuettel, Murray Stokely and Jeroen Ooms}
-
 %% DE: I tend to have wider option(width=...) so this
@@
 \maketitle
-
-%TODO(de) 'protocol buffers' or 'Protocol Buffers' ?
-% MS: Lets standardize on 'Protocol Buffers'?
 
@@
 
-%JO: can we just move the section 7 to here? It's only one paragraph%
-%MS: I replaced section 7 with 2 sentences above.
-
 <<>>=
@@
 @
-
-% The old section 7 in entirety:
-%The \texttt{RProtoBuf} package uses the user defined tables framework
-%that is defined as part of the \texttt{RObjectTables} package available
-%from the OmegaHat project \citep{RObjectTables}.
-%
-%The feature allows \texttt{RProtoBuf} to install the
-%special environment \emph{RProtoBuf:DescriptorPool} in the \proglang{R} search path.
-%The environment is special in that, instead of being associated with a
-%static hash table, it is dynamically queried by \proglang{R} as part of R's usual
-%variable lookup. In other words, it means that when the \proglang{R} interpreter
-%looks for a binding to a symbol (foo) in its search path,
-%it asks to our package if it knows the binding "foo", this is then
-%implemented by the \texttt{RProtoBuf} package by calling an internal
-%method of the \texttt{protobuf} C++ library.
 
@@
 
-One of the authors has used this design pattern for several large
-scale studies of distributed storage systems \citep{sciencecloud,janus}.
+One of the authors has used this design pattern for several
+large-scale studies of distributed storage systems
+\citep{sciencecloud,janus}.
 
@@
 function calls, and arguments/return values can be posted/retrieved
-using several data interchange formats, such as protocol buffers.  
+using several data interchange formats, such as Protocol Buffers.  
 OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
@@
 the standard \texttt{application/x-www-form-urlencoded} format might be sufficient.
-However, with protocol buffers the client can perform function calls with
+However, with Protocol Buffers the client can perform function calls with
 more complex arguments such as \proglang{R} vectors or lists. The result is a complete
@@
 @
-
-% OpenCPU also provides a lot of meta-functionality such as handling
-% of sessions, exceptions, security, and more. OpenCPU also makes it possible to store
-% output of a function call on the server, instead of directly retrieving it. Thereby 
-% objects can be shared with other users or used as arguments in a subsequent
-% function call.
-%  But in its essence, the HTTP API provides a simple way to perform remote 
-% \proglang{R} function calls over HTTPS. The same request can be performed in Python as demonstrated
-% below. The code is a bit verbose because to show how the REXP message is created from 
-% scratch. In practice would probably write a function or small module construct a Protocol
-% Buffer message representing an \proglang{R} list from a Python dictionary object. 
-%
-% \begin{verbatim}
-% import urllib2;
-% from rexp_pb2 import *;
-%
-% #create the post payload, i.e. list(n=42, mean=100)
-% payload = REXP(
-%  rclass = 5,
-%    rexpValue = [
-%      REXP(rclass = 2, realValue = [42]), 
-%      REXP(rclass = 2, realValue = [100])
-%    ],
-%    attrName = [
-%      "names"
-%    ],
-%    attrValue = [
-%      REXP(rclass = 0, stringValue = [STRING(strval="n"), STRING(strval="mean")])
-%    ]
-%);
-%
-%#HTTP POST
-%req = urllib2.Request(
-%  "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb", 
-%  data = payload.SerializeToString(), 
-%  headers = {
-%    'Content-type': 'application/x-protobuf'
-%  }
-%)
-%res = urllib2.urlopen(req);
-%        
-%#parse output pb
-%msg = REXP();
-%msg.ParseFromString(res.read());
-%
-%#the return value is a double vector in this case
-%print(msg.realValue);
-%\end{verbatim}
 

### 2014-01-23 papers/jss/article.Rnw
Go back to dissing MessagePack and BSON since Jeroen pointed out some limitations in the messagepack R package.  Remove citations to both, and call them "not widely supported".
I think this wording is a fair compromise in that it still makes the
point I wanted to make that they address a shortcoming of text based
JSON for certain classes of applications by providing more efficient
binary encoding, but they make Jeroen's point that they still suck due
to lack of adoption and weak implementations for R packages so
shouldn't really be held up on the same level as JSON.

Add a paragraph break befo

@@
 % Protocol Buffers, not protocol buffers
+% large-scale, not large scale
 
@@
 support for serialization, but the default formats 
-are usually language specific and thereby lock the user into a single
+are usually language-specific and thereby lock the user into a single
 environment.
@@
 %
+
 A more modern, widely used format is \emph{JavaScript Object
@@
 objects \citep{rjson,RJSONIO,jsonlite}.
-
 A number of binary formats based on \texttt{JSON} have been proposed
-that reduce the parsing cost and improve efficiency.  \pkg{MessagePack}
-and \pkg{BSON} both have \proglang{R}
-interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
-data and thus still duplicate field names with each message sent over
-the network or stored in a file.  Such formats also lack support for
-versioning when data storage needs evolve over time, or when
-application logic and requirement changes dictate updates to the
-message format.
+that reduce the parsing cost and improve efficiency, but these formats
+are not widely supported.  Furthermore, such formats lack a separate
+schema for the serialized data and thus still duplicate field names
+with each message sent over the network or stored in a file.
+% and still must send duplicate field names
+% with each message since there is no separate schema.
+% \pkg{MessagePack}
+% and \pkg{BSON} both have \proglang{R}
+% interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
+% data and thus still duplicate field names with each message sent over
+% the network or stored in a file.  Such formats also lack support for
+% versioning when data storage needs evolve over time, or when
+% application logic and requirement changes dictate updates to the
+%message format.
 

### 2014-01-23 papers/jss/article.Rnw
Note at the end of section 2 that the reflection based API is slightly slower for some operations, but that its much more convenient for interactive data analysis.
I want to tone down the implication that protocol buffers are the
best/fastest thing out there, and I always thought the 'particularly
well-suited' was a tad vague here.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@835 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 manipulate Protocol Buffer messages.  The \proglang{R} interface, in contrast,
-uses a reflection-based API that is particularly well-suited for
-interactive data analysis.  
+uses a reflection-based API that makes some operations slightly
+slower but which is much more convenient for interactive data analysis.
+%particularly well-suited for
+%interactive data analysis.  
 All messages in \proglang{R} have a single class
@@
 p[[ 2 ]] <- 3
-p[[ "email" ]]
+p[["email"]]
 @

### 2014-01-23 papers/jss/article.Rnw
Improve the text about serializing messages.  Don't start the section with "However".
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@836 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-However, the main focus of Protocol Buffer messages is
-efficiency. Therefore, messages are transported as a sequence
-of bytes. The \texttt{serialize} method is implemented for
+One of the primary benefits of Protocol Buffers is the efficient
+binary wire-format representation.  
+The \texttt{serialize} method is implemented for
 Protocol Buffer messages to serialize a message into a sequence of
-bytes that represents the message.
-%(raw vector in \proglang{R} speech) that represents the message.
+bytes (raw vector) that represents the message.
+The raw bytes can then be parsed back into the original message safely
+as long as the message type is known and its descriptor is available.
 

### 2014-01-23 papers/jss/article.Rnw
Standardize on no space before ':' in the text.  Use tbp instead of h for the Message table float placement, and correct a typo in the type coercion section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@837 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The \verb|[[| operator can also be used to query and set fields
-of a messages, supplying either their name or their tag number :
+of a messages, supplying either their name or their tag number:
 
@@
 Protocol Buffer messages and descriptors implement \texttt{show}
-methods that provide basic information about the message :
+methods that provide basic information about the message:
 
@@
 
-The same method can be used to serialize messages to files :
+The same method can be used to serialize messages to files:
 
@@
 infer the message type, so we have to provide this information
-to the \texttt{read} function in the form of a descriptor :
+to the \texttt{read} function in the form of a descriptor:
 
@@
 
-Finally, the payload of the message can be used :
+Finally, the payload of the message can be used:
 
@@
 \texttt{read} can also be used as a pseudo-method of the descriptor
-object :
+object:
 
@@
 facilitate this integration of the \proglang{R} and \proglang{C++} code for these objects.
-Each method is wrapped individually which alllows us to add user
+Each method is wrapped individually which allows us to add user
 friendly custom error handling, type coercion, and performance
@@
 \proglang{R} lacks a native unsigned integer type.  Values between $2^{31}$ and
-$2^{32} - 1$ read from unsigned into Protocol Buffer fields must be
+$2^{32} - 1$ read from unsigned integer Protocol Buffer fields must be
 stored as doubles in \proglang{R}.

### 2014-01-23 papers/jss/article.Rnw
Section 6:
Add the oxford comma somewhere I missed it earlier.
Use \code{} instead of \texttt{} for a function name.


git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@838 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 The previous sections discussed functionality in the \pkg{RProtoBuf} package
-for creating, manipulating, parsing and serializing Protocol Buffer
+for creating, manipulating, parsing, and serializing Protocol Buffer
 messages of a defined schema.  This is useful when there are
@@
 within \proglang{R}.
-
 The package also provides methods for converting arbitrary \proglang{R} data structures into protocol
-buffers and vice versa with a universal \proglang{R} object schema. The \texttt{serialize\_pb} and \texttt{unserialize\_pb}
+buffers and vice versa with a universal \proglang{R} object schema. The \code{serialize\_pb} and \code{unserialize\_pb}
 functions serialize arbitrary \proglang{R} objects into a universal Protocol Buffer 

### 2014-01-23 papers/jss/article.Rnw
Add new, better ack for saptarshi.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@839 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Switch around a sentence to get a better flow
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@840 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-A more modern, widely used format is \emph{JavaScript Object
-  Notation} (\texttt{JSON}), which is derived from the object literals of
-\proglang{JavaScript}, and used increasingly on the world wide web. \texttt{JSON} natively
-supports arrays and distinguishes 4 primitive types: numbers, strings,
+A more modern, widely used format is \emph{JavaScript ObjectNotation} 
+(\texttt{JSON}), which is derived from the object literals of
+\proglang{JavaScript}, and used increasingly on the world wide web. 
+Several \proglang{R} packages implement functions to parse and generate
+\texttt{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
+\texttt{JSON} natively supports arrays and 4 primitive types: numbers, strings,
 booleans, and null. However, as it too is a text-based format, numbers are
 stored as human-readable decimal notation which is inefficient and
-leads to loss of type (double versus integer) and precision. Several \proglang{R} packages
-implement functions to parse and generate \texttt{JSON} data from \proglang{R}
-objects \citep{rjson,RJSONIO,jsonlite}.
+leads to loss of type (double versus integer) and precision. 
 A number of binary formats based on \texttt{JSON} have been proposed
@@
 obscure protocol buffer semantics.  Karl Millar was very
-helpful in reviewing code and offering suggestions.  
-%The contemporaneous work by Saptarshi Guha on \pkg{RHIPE} was a strong
-%initial motivator.
+helpful in reviewing code and offering suggestions.  Saptarshi Guha's
+work on RHIPE and implementation of a universal message type for \proglang{R}
+language objects allowed us to add the \code{serialize_pb} and \code{unserialize\_pb}
+methods for turning arbitrary R objects into protocol buffers without
+a specialized pre-defined schema.
 

### 2014-01-23 papers/jss/article.Rnw
another pass at the summary
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@841 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 strengths and weaknesses.  
-Text based formats such as CSV and JSON are easy to use, and will likely 
+Text based formats such as \texttt{CSV} and \texttt{JSON} are easy to use, and will likely 
 remain popular among statisticians for many years to come. However, in the 
@@
 sophisticated data interchange format may reap considerable benefits. 
-The Protocol Buffers protocol and library offers a unique combination of features, performance,
-maturity, and forward-compatibility that seems particulary well suited for data-driven 
+%Protocol Buffers is itself not a protocol.
+%Forward-compatibility is one of the features. No need to re-iterate those 
+The Protocol Buffers standard and library offer a unique combination of features, 
+performance, and maturity, that seems particulary well suited for data-driven 
 applications and numerical computing.
@@
 %% DE Re-ordering so that we end on RProtoBuf
-The \pkg{RProtoBuf} package builds on the Protocol Buffers library, and
-extends the \proglang{R} system with the ability to create, read,
+The \pkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
+and extends the \proglang{R} system with the ability to create, read,
 write, parse, and manipulate Protocol
@@
 for the past three years by statisticians, analysts, and software engineers.
-At the time of this writing there are more than 300 active
+At the time of this writing there are over 300 active
 users of \pkg{RProtoBuf} using it to read data from and otherwise interact
 with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
-other languages.  It is our hope that this
-package will make Protocol Buffers more accessible to the \proglang{R} community, and
-thereby makes a small contribution towards better integration between \proglang{R} and
-other software systems and applications.
+other languages. We hope that making Protocol Buffers available to the
+\proglang{R} community will contribute towards better software integration
+and allow for building even more advanced applications and analysis pipelines 
+with \proglang{R}.
 

### 2014-01-23 papers/jss/article.Rnw
a missing dot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@842 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 greatly appreciated. Several features of the package reflect
-the design of the \CRANpkg{rJava} package by Simon Urbanek
+the design of the \CRANpkg{rJava} package by Simon Urbanek.
 The user-defined table mechanism, implemented by Duncan Temple Lang for the

### 2014-01-23 papers/jss/article.Rnw
Add a few improvements to the abstract suggested by Arif Merchant to make it more self contained.  After the first sentence explains that pipelines of applications need to communicate with each other, briefly describe the problem with common solutions such as CSV files in a new sentence, before transitioning to the third sentence to talk about protocol buffers.  Then after this, uncomment Jeroen's sentence from elsewhere in the paper which I had moved here which describes how they offer a unique combination of ...
After this change, section 2 starts at the top of page 3 instead of
the last lin

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Spelling Standardization:
@@
 % large-scale, not large scale
+% Oxford comma: foo, bar, and baz.
 
@@
 a sophisticated mix of applications written in general purpose and
-specialized programming languages.  Protocol Buffers are a popular
+specialized programming languages.  
+There are many formats used to import and export data between
+different programs or systems, such as CSV, but they are usually
+verbose, inefficient, not type-safe, or tied to a specific programming language.
+Protocol Buffers are a popular
 method of serializing structured data between applications---while remaining
-independent of programming languages or operating system.  The
-\CRANpkg{RProtoBuf} package provides a complete interface between this
-library and the \proglang{R} environment for statistical computing.
+independent of programming languages or operating system.
+They offer a unique combination of features, performance, and maturity that seems
+particulary well suited for data-driven applications and numerical
+computing.
+The
+\CRANpkg{RProtoBuf} package provides a complete interface to Protocol
+Buffers from the
+\proglang{R} environment for statistical computing.
 %TODO(ms) keep it less than 150 words.

### 2014-01-23 papers/jss/article.Rnw
Add one more sentence to end the abstract by describing what the paper is about:
"This paper outlines the general class of data serialization
requirements for statistical computing, describes the implementation
of the \CRANpkg{RProtoBuf} package, and illustrates its use with
examples applications in large-scale data collection pipelines and web
services."

We still have the nice page break between pages 2 and 3 where section
2 begins at the very top of page 3.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@844 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
a few micro-edits in Section

@@
 method of serializing structured data between applications---while remaining
-independent of programming languages or operating system.
+independent of programming languages or operating systems.
 They offer a unique combination of features, performance, and maturity that seems
@@
 \proglang{R} environment for statistical computing.
-%TODO(ms) keep it less than 150 words.
-% Maybe add Jeroen's sentence:
-% JO: added this sentence to the conclustion, but could use it in abstract as well.
-% They offer a unique combination of features, performance, and maturity that seems
-% particulary well suited for data-driven applications and numerical
-% computing.
+This paper outlines the general class of data serialization
+requirements for statistical computing, describes the implementation
+of the \CRANpkg{RProtoBuf} package, and illustrates its use with
+example applications in large-scale data collection pipelines and web
+services.
+%TODO(ms) keep it less than 150 words. -- I think this may be 154,
+%depending how emacs is counting.
 }
@@
 cleaning, modeling, analysis, post-processing, and
-presentation in order to take advantage of the unique combination of
+presentation -- in order to take advantage of the unique combination of
 performance, speed of development, and library support offered by
@@
 analysis pipeline may produce intermediate results that need to be
-stored in a file or sent over the network for further processing. 
+stored in a file, or sent over the network for further processing. 
 % JO Perhaps also mention that serialization is needed for distributed
@@
 %\paragraph*{Friends don't let friends use CSV!}
-Data analysts and researchers often use character separated text formats such
+Data analysts and researchers often use character-separated text formats such
 as \texttt{CSV} \citep{shafranovich2005common} to export and import
@@
 complexity at the parsing side (which are somewhat mitigated by the
-availability of mature libraries and parsers). Because \texttt{XML} is text
-based and has no native notion of numeric types or arrays, it usually not a
+availability of mature libraries and parsers). Because \texttt{XML} is 
+text-based and has no native notion of numeric types or arrays, it usually not a
 very practical format to store numeric datasets as they appear in statistical
@@
 
-A more modern, widely used format is \emph{JavaScript ObjectNotation} 
+A more modern format is \emph{JavaScript ObjectNotation} 
 (\texttt{JSON}), which is derived from the object literals of
-\proglang{JavaScript}, and used increasingly on the world wide web. 
+\proglang{JavaScript}, and already widely-used on the world wide web. 
 Several \proglang{R} packages implement functions to parse and generate
 \texttt{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
-\texttt{JSON} natively supports arrays and 4 primitive types: numbers, strings,
+\texttt{JSON} natively supports arrays and four primitive types: numbers, strings,
 booleans, and null. However, as it too is a text-based format, numbers are
@@
 structures and efficient binary interchange formats.  Since the schema
-is provided separately from the encoded data, the data can be
+is provided separately from the data, the data can be
 efficiently encoded to minimize storage costs when
@@
 \emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
-details the implementation details of the main S4 classes and methods
-contained in this
-package.  Section~\ref{sec:types} describes the challenges of type coercion
+details the implementation details of the main S4 classes and methods.  
+Section~\ref{sec:types} describes the challenges of type coercion
 between \proglang{R} and other languages.  Section~\ref{sec:evaluation} introduces a

### 2014-01-23 papers/jss/article.Rnw
concisification
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@846 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 specialized programming languages.  
-There are many formats used to import and export data between
-different programs or systems, such as CSV, but they are usually
+Many formats commonly used to import and export data between
+different programs or systems, such as \texttt{CSV} or \texttt{JSON}, are
 verbose, inefficient, not type-safe, or tied to a specific programming language.

### 2014-01-23 papers/jss/article.Rnw
Consistently use \CRANpkg for RProtoBuf and also use \code more.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@847 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
+\Shorttitle{\CRANpkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
 
@@
 After importing proto files, the corresponding message descriptors are
-available from the \texttt{RProtoBuf:DescriptorPool} environment in 
+available from the \code{RProtoBuf:DescriptorPool} environment in 
 the \proglang{R} search path.  This environment is implemented with the user
@@
 
-The \texttt{RProtoBuf} package defines the \texttt{read} and
-\texttt{readASCII} functions to read messages from files, raw vectors,
-or arbitrary connections.  \texttt{read} expects to read the message
-payload from binary files or connections and \texttt{readASCII} parses
+The \CRANpkg{RProtoBuf} package defines the \code{read} and
+\code{readASCII} functions to read messages from files, raw vectors,
+or arbitrary connections.  \code{read} expects to read the message
+payload from binary files or connections and \code{readASCII} parses
 the human-readable ASCII output that is created with
@@
 infer the message type, so we have to provide this information
-to the \texttt{read} function in the form of a descriptor:
+to the \code{read} function in the form of a descriptor:
 
@@
 improvements at the cost of a more verbose implementation.
-The \pkg{RProtoBuf} package in many ways motivated
-the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
+The \CRANpkg{RProtoBuf} package in many ways motivated
+the development of \CRANpkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
 which provide a more concise way of wrapping \proglang{C++} functions and classes
@@
 
-The \texttt{RProtoBuf} package supports two forms for calling
+The \CRANpkg{RProtoBuf} package supports two forms for calling
 functions with these S4 classes:
@@
 
-Additionally, \pkg{RProtoBuf} supports tab completion for all
+Additionally, \CRANpkg{RProtoBuf} supports tab completion for all
 classes.  Completion possibilities include pseudo-method names for all
@@
   \proglang{R} type retrieved by the extractors. Note that \proglang{R} lacks native
-  64-bit integers, so the \texttt{RProtoBuf.int64AsString} option is
+  64-bit integers, so the \code{RProtoBuf.int64AsString} option is
   available to return large integers as characters to avoid losing
@@
 default, but when the full precision is required a character value
-will be returned if the \texttt{RProtoBuf.int64AsString} option is set
+will be returned if the \code{RProtoBuf.int64AsString} option is set
 to \texttt{TRUE}.  The character values are useful because they can
@@
 
-The previous sections discussed functionality in the \pkg{RProtoBuf} package
+The previous sections discussed functionality in the \CRANpkg{RProtoBuf} package
 for creating, manipulating, parsing, and serializing Protocol Buffer
@@
 
-In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
+In order to accomplish this, \CRANpkg{RProtoBuf} uses the same catch-all \texttt{proto}
 schema used by \pkg{RHIPE} for exchanging \proglang{R} data with Hadoop \citep{rhipe}. This 
@@
 the appendix.
-The Protocol Buffer messages generated by \pkg{RProtoBuf} and
+The Protocol Buffer messages generated by \CRANpkg{RProtoBuf} and
 \pkg{RHIPE} are naturally compatible between the two systems because they use the 
@@
 One takeaway from this table is that the universal \proglang{R} object schema
-included in \pkg{RProtoBuf} does not in general provide
+included in \CRANpkg{RProtoBuf} does not in general provide
 any significant saving in file size compared to the normal serialization
@@
 % redundant: which is seen as equally compact.
-The benefits of \pkg{RProtoBuf} accrue more naturally in applications where
+The benefits of \CRANpkg{RProtoBuf} accrue more naturally in applications where
 multiple programming languages are involved, or when a more concise
@@
 \caption{Serialization sizes for default serialization in \proglang{R} and
-  RProtoBuf for 50 \proglang{R} datasets.}
+  \CRANpkg{RProtoBuf} for 50 \proglang{R} datasets.}
 \label{tab:compression}
@@
 
-The \pkg{HistogramTools} package \citep{histogramtools} enhances
-\pkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
+The \CRANpkg{HistogramTools} package \citep{histogramtools} enhances
+\CRANpkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
 
@@
 
-# Convert to native \proglang{R} histogram object and plot
+# Convert to native R histogram object and plot
 plot(as.histogram(hist))
@@
 OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
-from the \texttt{RProtoBuf} package to convert between \proglang{R} objects and protobuf
+from the \CRANpkg{RProtoBuf} package to convert between \proglang{R} objects and protobuf
 messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
@@
 %% DE Re-ordering so that we end on RProtoBuf
-The \pkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
+The \CRANpkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
 and extends the \proglang{R} system with the ability to create, read,
 write, parse, and manipulate Protocol
-Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
+Buffer messages. \CRANpkg{RProtoBuf} has been used extensively inside Google 
 for the past three years by statisticians, analysts, and software engineers.
 At the time of this writing there are over 300 active
-users of \pkg{RProtoBuf} using it to read data from and otherwise interact
+users of \CRANpkg{RProtoBuf} using it to read data from and otherwise interact
 with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
@@
 work on RHIPE and implementation of a universal message type for \proglang{R}
-language objects allowed us to add the \code{serialize_pb} and \code{unserialize\_pb}
+language objects allowed us to add the \code{serialize_pb} and \code{unserialize_pb}
 methods for turning arbitrary R objects into protocol buffers without
@@
 Below a print of the \texttt{rexp.proto} schema (originally designed by \cite{rhipe})
-that is included with the \pkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
+that is included with the \CRANpkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
 \texttt{unserialize\_pb}.

### 2014-01-23 papers/jss/article.Rnw
Add a total summary row giving the relative size reduction for different serialization formats, and reference the finding more specifically in the text.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@848 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %sample of 50 datasets is included on the next page.  
-Sizes are comparable but Protocol Buffers provide simple getters and setters
-in multiple languages instead of requiring other programs to parse the \proglang{R}
-serialization format. % \citep{serialization}.
+Note that protocol buffer serialization results in slightly
+smaller byte streams compared to native \proglang{R} serialization in most cases,
+but this difference disappears if the results are compressed with gzip.
+%Sizes are comparable but Protocol Buffers provide simple getters and setters
+%in multiple languages instead of requiring other programs to parse the \proglang{R}
+%serialization format. % \citep{serialization}.
 One takeaway from this table is that the universal \proglang{R} object schema
@@
   \bottomrule
+%  Total & 391176 & 327537 & 99161 & 313456 & 100308 \\
+  Relative Size & 100\% & 83.7\% & 25.3\% & 80.1\% & 25.6\%\\
+  \bottomrule
 \end{tabular}

### 2014-01-24 papers/jss/article.Rnw
more hyphen for adverbial use, more Prot. Buf capitalization
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@849 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 }
-\Keywords{\proglang{R}, \pkg{Rcpp}, protocol buffers, serialization, cross-platform}
-\Plainkeywords{R, Rcpp, protocol buffers, serialization, cross-platform} %% without formatting
+\Keywords{\proglang{R}, \pkg{Rcpp}, Protocol Buffers, serialization, cross-platform}
+\Plainkeywords{R, Rcpp, Protocol Buffers, serialization, cross-platform} %% without formatting
 %% at least one keyword must be supplied
@@
 model that is flexible and easy to use.  The schema for structured
-protocol buffer data is defined in \texttt{.proto} files, which may
+Protocol Buffer data is defined in \texttt{.proto} files, which may
 contain one or more message types.  Each message type has one or more
@@
 enumerations, booleans, strings, raw bytes, or other nested message
-types.  The \texttt{.proto} file syntax for defining the structure of protocol
-buffer data is described comprehensively on Google Code\footnote{See 
+types.  The \texttt{.proto} file syntax for defining the structure of Protocol
+Buffer data is described comprehensively on Google Code\footnote{See 
 \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
@@
 This section describes how to use the \proglang{R} API to create and manipulate
-protocol buffer messages in \proglang{R}, and how to read and write the
+Protocol Buffer messages in \proglang{R}, and how to read and write the
 binary representation of the message (often called the \emph{payload}) to files and arbitrary binary
@@
 available from the \code{RProtoBuf:DescriptorPool} environment in 
-the \proglang{R} search path.  This environment is implemented with the user
-defined tables framework from the \pkg{RObjectTables} package
+the \proglang{R} search path.  This environment is implemented with the 
+user-defined tables framework from the \pkg{RObjectTables} package
 available from the OmegaHat project \citep{RObjectTables}.  Instead of
@@
 Once the message is created, its fields can be queried
-and modified using the dollar operator of \proglang{R}, making protocol
-buffer messages seem like lists.
+and modified using the dollar operator of \proglang{R}, making Protocol
+Buffer messages seem like lists.
 
@@
 facilitate this integration of the \proglang{R} and \proglang{C++} code for these objects.
-Each method is wrapped individually which allows us to add user
-friendly custom error handling, type coercion, and performance
+Each method is wrapped individually which allows us to add 
+user-friendly custom error handling, type coercion, and performance
 improvements at the cost of a more verbose implementation.
@@
 \texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the \proglang{C++} proto library. Documentation for the
-\texttt{FileDescriptor} class is available from the protocol buffer project page:
+\texttt{FileDescriptor} class is available from the Protocol Buffer project page:
 \url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
@@
 possible values as booleans.  The library will refuse to store
-\texttt{NA}s in protocol buffer boolean fields, and users must instead
+\texttt{NA}s in Protocol Buffer boolean fields, and users must instead
 choose another type (such as enum or integer) capable of storing three
@@
 \begin{CodeOutput}
-Error: NA boolean values can not be stored in bool protocol buffer fields
+Error: NA boolean values can not be stored in bool Protocol Buffer fields
 \end{CodeOutput}
@@
 within \proglang{R}.
-The package also provides methods for converting arbitrary \proglang{R} data structures into protocol
-buffers and vice versa with a universal \proglang{R} object schema. The \code{serialize\_pb} and \code{unserialize\_pb}
+The package also provides methods for converting arbitrary \proglang{R} data structures into Protocol
+Buffers and vice versa with a universal \proglang{R} object schema. The \code{serialize\_pb} and \code{unserialize\_pb}
 functions serialize arbitrary \proglang{R} objects into a universal Protocol Buffer 
@@
 %sample of 50 datasets is included on the next page.  
-Note that protocol buffer serialization results in slightly
+Note that Protocol Buffer serialization results in slightly
 smaller byte streams compared to native \proglang{R} serialization in most cases,
@@
 message data.  The following simple Python script uses this generated
-module to create a histogram and write out the protocol buffer
+module to create a histogram and write out the Protocol Buffer
 representation to a file:
@@
 
-The protocol buffer can then be read into \proglang{R} and converted to a native
+The Protocol Buffer can then be read into \proglang{R} and converted to a native
 \proglang{R} histogram object for plotting:
@@
 Kenton Varda was generous with his time in reviewing code and explaining
-obscure protocol buffer semantics.  Karl Millar was very
+obscure Protocol Buffer semantics.  Karl Millar was very
 helpful in reviewing code and offering suggestions.  Saptarshi Guha's
@@
 language objects allowed us to add the \code{serialize_pb} and \code{unserialize_pb}
-methods for turning arbitrary R objects into protocol buffers without
+methods for turning arbitrary R objects into Protocol Buffers without
 a specialized pre-defined schema.

### 2014-01-26 papers/jss/article.Rnw
make the directory structure a little more organized
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@850 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
section and subsection in sentence style
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@851 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{document}
-\SweaveOpts{concordance=FALSE}
+\SweaveOpts{concordance=FALSE,prefix.string=figures/fig}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{protobuf-distributed-system-crop.pdf}
+\includegraphics[width=\textwidth]{figures/protobuf-distributed-system-crop.pdf}
 \end{center}
@@
 
-\section{Basic Usage: Messages and Descriptors}
+\section{Basic Usage: Messages and descriptors}
 \label{sec:rprotobuf-basic}
@@
 
-\subsection[Importing Message Descriptors from .proto files]{Importing Message Descriptors from \texttt{.proto} files}
+\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \texttt{.proto} files}
 
@@
 
-\section{Under the hood: S4 Classes, Methods, and Pseudo Methods}
+\section{Under the hood: S4 classes, methods, and pseudo methods}
 \label{sec:rprotobuf-classes}
@@
 
-\subsection{Field Descriptors}
+\subsection{Field descriptors}
 \label{subsec-field-descriptor}
@@
 
-\subsection{Enum Descriptors}
+\subsection{Enum descriptors}
 \label{subsec-enum-descriptor}
@@
 
-\subsection{File Descriptors}
+\subsection{File descriptors}
 \label{subsec-file-descriptor}
@@
 
-\subsection{Enum Value Descriptors}
+\subsection{Enum value descriptors}
 \label{subsec-enumvalue-descriptor}
@@
 
-\section{Type Coercion}
+\section{Type coercion}
 \label{sec:types}
@@
 
-\subsection{Unsigned Integers}
+\subsection{Unsigned integers}
 
@@
 
-\section[Converting R Data Structures into Protocol Buffers]{Converting \proglang{R} Data Structures into Protocol Buffers}
+\section[Converting R data structures into Protocol Buffers]{Converting \proglang{R} data structures into Protocol Buffers}
 \label{sec:evaluation}
@@
 
-\subsection[Evaluation: Converting R Data Sets]{Evaluation: Converting \proglang{R} Data Sets}
+\subsection[Evaluation: Converting R data sets]{Evaluation: Converting \proglang{R} data sets}
 
@@
 
-\subsection{Compression Performance}
+\subsection{Compression performance}
 \label{sec:compression}
@@
 
-\section{Application: Distributed Data Collection with MapReduce}
+\section{Application: Distributed data collection with MapReduce}
 \label{sec:mapreduce}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{histogram-mapreduce-diag1.pdf}
+\includegraphics[width=\textwidth]{figures/histogram-mapreduce-diag1.pdf}
 \end{center}
@@
 
-\section{Application: Data Interchange in Web Services}
+\section{Application: Data Interchange in web Services}
 \label{sec:opencpu}

### 2014-01-26 papers/jss/article.Rnw
sentence style in captions
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@852 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{tabular}
-\caption{\label{class-summary-table}Overview of Class, Slot, Method and
-  Dispatch Relationships}
+\caption{\label{class-summary-table}Overview of class, slot, method and
+  dispatch relationships}
 \end{table}
@@
 \end{center}
-\caption{Diagram of MapReduce Histogram Generation Pattern}
+\caption{Diagram of MapReduce histogram generation pattern}
 \label{fig:mr-histogram-pattern1}

### 2014-01-26 papers/jss/article.Rnw
i.e. and e.g. should have comma (see Style Guide, Misc)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@853 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
MASSIVE killing of comments which were becoming ballast
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@854 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\section{Introduction} % TODO(DE) More sober: Friends don't let friends use CSV}
-% NOTE(MS): I really do think we can use add back:
-% \section{Introduction: Friends Don't Let Friends Use CSV}
-% I didn't use proper Title Caps the first time around but really I
-% think it makes the paper more readable to have a tl;dr intro title
-% that is fun and engaging since this paper is still on the dry/boring
-% side.
+\section{Introduction}
+
 Modern data collection and analysis pipelines increasingly involve collections
@@
 through reusability, modularity, and fault isolation \citep{Wegiel:2010:CTT:1932682.1869479}.
-% This is really a different pattern not connected well here.
-%Data analysis patterns such as Split-Apply-Combine
-%\citep{wickham2011split} explicitly break up large problems into manageable pieces.  
 These pipelines are frequently built using different programming 
@@
 stored in a file, or sent over the network for further processing.
-% JO Perhaps also mention that serialization is needed for distributed
-% systems to make systems scale up?
 
@@
 original form later \citep{clinec++}.
-% Reverted to my original above, because the replacement below puts me
-% to sleep:
-%Such systems require reliable and efficient exchange of intermediate
-%results between the individual components, using formats that are
-%independent of platform, language, operating system or architecture.
 Programming
@@
 
-%\paragraph*{Friends don't let friends use CSV!}
 Data analysts and researchers often use character-separated text formats such
@@
 applications.
-%
 
@@
 with each message sent over the network or stored in a file.
-% and still must send duplicate field names
-% with each message since there is no separate schema.
-% \pkg{MessagePack}
-% and \pkg{BSON} both have \proglang{R}
-% interfaces \citep{msgpackR,rmongodb}, but these formats lack a separate schema for the serialized
-% data and thus still duplicate field names with each message sent over
-% the network or stored in a file.  Such formats also lack support for
-% versioning when data storage needs evolve over time, or when
-% application logic and requirement changes dictate updates to the
-%message format.
 
@@
 \citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
-
-% Too technical, move to section 2.
-% The schema can be used to generate model classes for statically-typed programming languages
-%such as C++ and Java, or can be used with reflection for dynamically-typed programming
-%languages.
-
-% TODO(mstokely): Will need to define reflection if we use it here.
-% Maybe in the next section since its not as key as 'serialization'
-% which we already defined.
-
-%\paragraph*{Enter Protocol Buffers:}
-
-% In 2008, and following several years of internal use, Google released an open
-% source version of Protocol Buffers. It provides data  
-% interchange format that was designed and used for their internal infrastructure.
-% Google officially provides high-quality parsing libraries for \texttt{Java}, 
-% \texttt{C++} and \texttt{Python}, and community-developed open source implementations
-% are available for many other languages. 
-% Protocol Buffers take a quite different approach from many other popular formats.
-
-% TODO(mstokely): Good sentence from Jeroen, add it here or sec 2.
-% They offer a unique combination of features, performance, and maturity that seems
-% particulary well suited for data-driven applications and numerical
-% computing.
-
-% TODO(DE): Mention "future proof" forward compatibility of schemata
-
-% TODO(mstokely): Take a more conversational tone here asking
-% questions and motivating protocol buffers?
-
-% NOTE(mstokely): I don't like these roadmap paragraphs in general,
-% but it seems ueful here because we have a boring bit in the middle
-% (full class/method details) and interesting applications at the end.
 
@@
 
-%This article describes the basics of Google's Protocol Buffers through
-%an easy to use \proglang{R} package, \CRANpkg{RProtoBuf}.  After describing the
-%basics of protocol buffers and \CRANpkg{RProtoBuf}, we illustrate
-%several common use cases for protocol buffers in data analysis.
-
 \section{Protocol Buffers}
@@
 
-% JO: I'm not sure where to put this paragraph. I think it is too technical
-% for the introduction section. Maybe start this section with some explanation
-% of what a schema is and then continue with showing how PB implement this?
-% MS: Yes I agree, tried to address below.
-
-% This content is good.  Maybe use and cite?
-% http://martin.kleppmann.com/2012/12/05/schema-evolution-in-avro-protocol-buffers-thrift.html
-
-%% TODO(de,ms)  What follows is oooooold and was lifted from the webpage
-%%              Rewrite?
 Protocol Buffers are a modern, language-neutral, platform-neutral,
@@
 
-% Lets place this at the top of the page or the bottom, or on a float
-% page, but not just here in the middle of the page.
 \begin{figure}[tbp]
@@
 
-%% TODO(de) Can we make this not break the width of the page?
 \noindent
@@
 
-% The schema can be used to generate model classes for statically-typed programming languages
-%such as C++ and Java, or can be used with reflection for dynamically-typed programming
-%languages.
-
-% TODO(mstokely): Maybe find a place to add this?  
-% Since their
-% introduction, Protocol Buffers have been widely adopted in industry with
-% applications as varied as %database-internal messaging (Drizzle), % DE: citation?
-% Sony Playstations, Twitter, Google Search, Hadoop, and Open Street
-% Map.
-
-% TODO(DE): This either needs a citation, or remove the name drop
-% MS: These are mostly from blog posts, I can't find a good reference
-% that has a long list, and the name and year citation style seems
-% less conducive to long lists of marginal citations like blog posts
-% compared to say concise CS/math style citations [3,4,5,6]. Thoughts?
-
-% The schema can be used to generate classes for statically-typed programming languages
-% such as C++ and Java, or can be used with reflection for dynamically-typed programming
-% languages.
-
-%Protocol buffers are a language-neutral, platform-neutral, extensible
-%way of serializing structured data for use in communications
-%protocols, data storage, and more.
-
-%Protocol Buffers offer key features such as an efficient data interchange
-%format that is both language- and operating system-agnostic yet uses a
-%lightweight and highly performant encoding, object serialization and
-%de-serialization as well data and configuration management. Protocol
-%buffers are also forward compatible: updates to the \texttt{proto}
-%files do not break programs built against the previous specification.
-
 For added speed and efficiency, the \proglang{C++}, \proglang{Java},
@@
 slower but which is much more convenient for interactive data analysis.
-%particularly well-suited for
-%interactive data analysis.  
 All messages in \proglang{R} have a single class
@@
 next section.
-
-% In other words, given the 'proto'
-%description file, code is automatically generated for the chosen
-%target language(s). The project page contains a tutorial for each of
-%these officially supported languages:
-%\url{http://code.google.com/apis/protocolbuffers/docs/tutorials.html}
-
-%The protocol buffers code is released under an open-source (BSD) license. The
-%protocol buffer project (\url{http://code.google.com/p/protobuf/})
-%contains a C++ library and a set of runtime libraries and compilers for
-%C++, Java and Python.
-
-%With these languages, the workflow follows standard practice of so-called
-%Interface Description Languages (IDL)
-%(c.f. \href{http://en.wikipedia.org/wiki/Interface_description_language}{Wikipedia
-%  on IDL}).  This consists of compiling a protocol buffer description file
-%(ending in \texttt{.proto}) into language specific classes that can be used
-
-%Besides the officially supported C++, Java and Python implementations, several projects have been
-%created to support protocol buffers for many languages. The list of known
-%languages to support protocol buffers is compiled as part of the
-%project page: \url{http://code.google.com/p/protobuf/wiki/ThirdPartyAddOns}
 
@@
 
-% Commented out because we said this earlier.
-%This separation
-%between schema and the message objects is in contrast to
-%more verbose formats like JSON, and when combined with the efficient
-%binary representation of any Message object explains a large part of
-%the performance and storage-space advantage offered by Protocol
-%Buffers. TODO(ms): we already said some of this above.  clean up.
-
-% lifted from protobuf page:
-%With Protocol Buffers you define how you want your data to be
-%structured once, and then you can read or write structured data to and
-%from a variety of data streams using a variety of different
-%languages.  The definition
-
 \subsection[Importing message descriptors from .proto files]{Importing message descriptors from \texttt{.proto} files}
-
-%The three basic abstractions of \CRANpkg{RProtoBuf} are Messages,
-%which encapsulate a data structure, Descriptors, which define the
-%schema used by one or more messages, and DescriptorPools, which
-%provide access to descriptors.
 
@@
 
-%\subsection{Importing proto files}
-%In contrast to the other languages (Java, C++, Python) that are officially
-%supported by Google, the implementation used by the \texttt{RProtoBuf}
-%package does not rely on the \texttt{protoc} compiler (with the exception of
-%the two functions discussed in the previous section). This means that no
-%initial step of statically compiling the proto file into C++ code that is
-%then accessed by \proglang{R} code is necessary. Instead, \texttt{proto} files are
-%parsed and processed \textsl{at runtime} by the protobuf C++ library---which
-%is much more appropriate for a dynamic language.
-
 \subsection{Creating a message}
@@
 Section~\ref{sec:int64} for working with large integer values.
-
-% TODO(mstokely): Document extensions here.
-% There are none in addressbook.proto though.
 
@@
 
-% MS: I think this looks better at the bottom of the page.
-% so it appears after the new section starts where it is referenced.
 \begin{table}[bp]
@@
 in a single entity.
-
-% Message, Descriptor, FieldDescriptor, EnumDescriptor,
-% FileDescriptor, EnumValueDescriptor
-%
-% grep RPB_FUNC * | grep -v define|wc -l
-% 84
-% grep RPB_ * | grep -v RPB_FUNCTION | grep METHOD|wc -l
-% 33
 
@@
 
-% TODO(ms): Useful distinction to make -- FieldDescriptor does not do
-% separate '$' dispatch like Messages, Descriptors, and
-% EnumDescriptors do.  Should it?
-
 \subsection{Enum descriptors}
@@
 
-% N.B. see table.Rnw for how this table is created.
-%
 % latex table generated in \proglang{R} 3.0.2 by xtable 1.7-0 package
@@
 
-% TODO(jeroen): I think maybe some of this should go earlier in the
-% paper, so this part can focus only on introducing the application,
-% Can you integrate some of this text earlier, maybe into the the
-% introduction?
-
 As described earlier, the primary application of Protocol Buffers is data
 interchange in the context of inter-system communications.  Network protocols
-such as HTTP provide mechanisms for client-server communication, i.e.~how to
+such as HTTP provide mechanisms for client-server communication, i.e., how to
 initiate requests, authenticate, send messages, etc.  However, network
@@
 
-\section{Summary}  % DE Simpler title
+\section{Summary}  
 \label{sec:summary}
@@
 
-%% DE Re-ordering so that we end on RProtoBuf
 The \CRANpkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
@@
 with \proglang{R}.
-
-%\emph{Other Approaches}
-%
-%== JO: I don't really like this section here, it gives the entire paper a bit of a 
-%sour aftertaste. Perhaps we can mention performance caveats in the technical
-%sections? I think it's nicer to leave it at the above paragraphs.==
-%
-% DE: Agreed -- commenting out
-
-%% \pkg{RProtoBuf} is quite flexible and easy to use for interactive use,
-%% but it is not designed for efficient high-speed manipulation of large
-%% numbers of protocol buffers once they have been read into R.  For
-%% example, taking a list of 100,000 Protocol Buffers, extracting a named
-%% field from each one, and computing an aggregate statistic on those
-%% values would be relatively slow with RProtoBuf.  Mechanisms to address
-%% such use cases are under investigation for possible incorporation into
-%% future releases of RProtoBuf, but currently, the package relies on
-%% other database systems to provide query and aggregation semantics
-%% before the resulting protocol buffers are read into R.  Inside Google,
-%% the Dremel query system \citep{dremel} is often employed in this role
-%% in conjunction with \pkg{RProtoBuf}.
-
-% Such queries could be
-%supported in a future version of \pkg{RProtoBuf} by supporting a
-%vector of messages type such that \emph{slicing} operations over a
-%given field across a large number of messages could be done
-%efficiently in \proglang{C++}.
 

### 2014-01-26 papers/jss/article.Rnw
one spell checks, lots of 'data set' instead of dataset, one 'work flow'
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@855 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 These pipelines are frequently built using different programming 
-languages for the different phases of data analysis -- collection,
+languages for the different phases of data analysis --- collection,
 cleaning, modeling, analysis, post-processing, and
-presentation -- in order to take advantage of the unique combination of
+presentation --- in order to take advantage of the unique combination of
 performance, speed of development, and library support offered by
@@
 text-based and has no native notion of numeric types or arrays, it usually not a
-very practical format to store numeric datasets as they appear in statistical
+very practical format to store numeric data sets as they appear in statistical
 applications.
@@
 general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and evaluates
-it against the serialization capbilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
+it against the serialization capabilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
 and \ref{sec:opencpu} provide real-world use cases of \CRANpkg{RProtoBuf}
@@
   representation for transmission or storage.
-\item \emph{Extensible}:  New fields can be added to Protocol Buffer Schemas
+\item \emph{Extensible}:  New fields can be added to Protocol Buffer schemas
   in a forward-compatible way that does not break older applications.
@@
 Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
-communication workflow with Protocol Buffers and an interactive \proglang{R} session.
+communication work flow with Protocol Buffers and an interactive \proglang{R} session.
 Common use cases include populating a request remote-procedure call (RPC)
@@
 To illustrate how this method works, we attempt to convert all of the built-in 
-datasets from \proglang{R} into this serialized Protocol Buffer representation.
+data sets from \proglang{R} into this serialized Protocol Buffer representation.
 
@@
 
-There are \Sexpr{n} standard data sets included in the base-r \pkg{datasets}
-package. These datasets include data frames, matrices, time series, tables lists,
+There are \Sexpr{n} standard data sets included in the \pkg{datasets}
+package included with \proglang{R}. These data sets include data frames, matrices, time series, tables lists,
 and some more exotic data classes. The \texttt{can\_serialize\_pb} method is 
@@
 without loss of information (\Sexpr{format(100*m/n,digits=1)}\%). Upon closer
-inspection, all other datasets are objects of class \texttt{nfnGroupedData}.
+inspection, all other data sets are objects of class \texttt{nfnGroupedData}.
 This class represents a special type of data frame that has some additional 
@@
 
-Table~\ref{tab:compression} shows the sizes of 50 sample \proglang{R} datasets as
+Table~\ref{tab:compression} shows the sizes of 50 sample \proglang{R} data sets as
 returned by object.size() compared to the serialized sizes.
 %The summary compression sizes are listed below, and a full table for a
-%sample of 50 datasets is included on the next page.  
+%sample of 50 data sets is included on the next page.  
 Note that Protocol Buffer serialization results in slightly
@@
 \caption{Serialization sizes for default serialization in \proglang{R} and
-  \CRANpkg{RProtoBuf} for 50 \proglang{R} datasets.}
+  \CRANpkg{RProtoBuf} for 50 \proglang{R} data sets.}
 \label{tab:compression}
@@
 The \texttt{HTTP GET} method is used to read a resource from OpenCPU. For example,
-to access the dataset \texttt{Animals} from the package \texttt{MASS}, a 
+to access the data set \texttt{Animals} from the package \texttt{MASS}, a 
 client performs the following HTTP request:
@@
 languages, clients can be implemented in just a few lines of code. Below
-is example code for both \proglang{R} and Python that retrieves a dataset from \proglang{R} with 
+is example code for both \proglang{R} and Python that retrieves a data set from \proglang{R} with 
 OpenCPU using a protobuf message. In \proglang{R}, we use the HTTP client from 
 the \texttt{httr} package \citep{httr}. In this example we
-download a dataset which is part of the base \proglang{R} distribution, so we can
+download a data set which is part of the base \proglang{R} distribution, so we can
 verify that the object was transferred without loss of information.
@@
 programming language. For example, in a very similar fashion we can retrieve the same
-dataset in a Python client. To parse messages in Python, we first compile the 
+data set in a Python client. To parse messages in Python, we first compile the 
 \texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:
@@
 \end{verbatim}
-The \texttt{msg} object contains all data from the Animals dataset. From here we
+The \texttt{msg} object contains all data from the Animals data set. From here we
 can easily extract the desired fields for further use in Python.
@@
 The Protocol Buffers standard and library offer a unique combination of features, 
-performance, and maturity, that seems particulary well suited for data-driven 
+performance, and maturity, that seems particularly well suited for data-driven 
 applications and numerical computing.

### 2014-01-27 papers/jss/article.Rnw
add nlme citation as we have to cite software anyway
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@856 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 % Articles with many authors we should shorten to FirstAuthor, et al.
-\shortcites{sciencecloud,janus,dremel}
+\shortcites{sciencecloud,janus,dremel,nlme}
 \author{Dirk Eddelbuettel\\Debian Project \And 
@@
 This class represents a special type of data frame that has some additional 
-attributes (such as a \emph{formula} object) used by the \pkg{nlme} package.
+attributes (such as a \emph{formula} object) used by the \pkg{nlme} package \citep{nlme}.
 Because formulas are \proglang{R} \emph{language} objects, they have little meaning to

### 2014-01-27 papers/jss/article.Rnw
another hyphen between nouns
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@857 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{RHIPE} are naturally compatible between the two systems because they use the 
-same schema. This shows the power of using a schema based cross-platform format such
+same schema. This shows the power of using a schema-based cross-platform format such
 as Protocol Buffers: interoperability is achieved without effort or close coordination.

### 2014-01-27 papers/jss/article.Rnw
preferred options() specified at jstatsoft.org
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@858 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Replace the second use of 'compare' in a sentence with another more specific and less duplicative word ('perform')
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@859 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 <<echo=FALSE,print=FALSE>>=
-options(width=65, prompt="R> ", digits=4)
+## cf http://www.jstatsoft.org/style#q12
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE, digits=4)
 @
@@
 Many sources compare data serialization formats
-and show Protocol Buffers compare very favorably to the alternatives; see
+and show Protocol Buffers perform favorably to the alternatives; see
 \citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.

### 2014-01-27 papers/jss/article.Rnw
Move figure1 float to [bp] from [tbp] so it is placed at the bottom of the page.  This looks better so that section 2 with its title header start at the top of the page, instead of the previous layout of starting with a not-yet-referenced figure and then a section title header beneath it.
I think this looks a bit better, but feel free to revert this change
if you disagree.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@860 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Add a short example to the messages section again illustrating how these objects are created, and modify a table float

@@
 
-\begin{figure}[tbp]
+\begin{figure}[bp]
 \begin{center}
@@
 contain one or more message types.  Each message type has one or more
-fields.  A field is specified with a unique number, a name, a value
+fields.  A field is specified with a unique number (called a \emph{tag number}), a name, a value
 type, and a field rule specifying whether the field is optional,
@@
 Table~\ref{tab:proto} shows an example \texttt{.proto} file that
-defines the \texttt{tutorial.Person} type.  The \proglang{R} code in the right
+defines the \texttt{tutorial.Person} type\footnote{The compound name
+  \texttt{tutorial.Person} in R is derived from the name of the
+  message and the name of the package defined at the top of the
+  \texttt{.proto} file in which it is defined.}.  The \proglang{R} code in the right
 column shows an example of creating a new message of this type and
@@
 is available in Table~\ref{Message-methods-table}.
+
+<<>>=
+new(tutorial.Person)
+@
 
@@
 a message contains and what the types of those fields are.  Message
-descriptors are represented in \proglang{R} with the \emph{Descriptor} S4
+descriptors are represented in \proglang{R} by the \emph{Descriptor} S4
 class. The class contains the slots \texttt{pointer} and
@@
 
-\begin{table}[tbp]
+\begin{table}[p]
 \centering
@@
 pseudo-methods.
+
+The \texttt{EnumDescriptor} contains information about what values this type
+defines, while the \texttt{EnumValueDescriptor} describes an
+individual enum constant of a particular type.
 

### 2014-01-27 papers/jss/article.Rnw
Switch the order of the FileDescriptor and EnumValueDescriptor subsections so that EnumDescriptors and EnumValueDescriptors are described in adjacent sections.
Also manually move table definition earlier and adjust float placement
options so that all the tables appear in section 4.

Section 5 now begins at top of a page, and includes no floats from
section 4, which is desirable.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@863 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   \texttt{tutorial.Person} in R is derived from the name of the
-  message and the name of the package defined at the top of the
-  \texttt{.proto} file in which it is defined.}.  The \proglang{R} code in the right
+  message (\emph{Person}) and the name of the package defined at the top of the
+  \texttt{.proto} file in which it is defined (\emph{tutorial}).}.  The \proglang{R} code in the right
 column shows an example of creating a new message of this type and
@@
 EnumDescriptor      & 4 & 11 & yes (enum constant names)\\
+EnumValueDescriptor & 3 & \phantom{1}6 & no\\
 FileDescriptor      & 3 & \phantom{1}6 & yes (message/field definitions)\\
-EnumValueDescriptor & 3 & \phantom{1}6 & no\\
 \bottomrule
@@
 
-\begin{table}[p]
+\begin{table}[tbp]
 \centering
@@
 
+\subsection{Enum value descriptors}
+\label{subsec-enumvalue-descriptor}
+
+The class \emph{EnumValueDescriptor} represents enumeration value
+descriptors in \proglang{R}.  This is a wrapper S4 class around the
+\texttt{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
+Table~\ref{EnumValueDescriptor-methods-table} describes the methods
+defined for the \texttt{EnumValueDescriptor} class.
+
+The \verb|$| operator can be used to invoke pseudo-methods.
+
+<<>>=
+tutorial.Person$PhoneType$value(1)
+tutorial.Person$PhoneType$value(name="HOME")
+tutorial.Person$PhoneType$value(number=1)
+@
+
+\begin{table}[tbp]
+\centering
+\begin{small}
+\begin{tabular}{lp{10cm}}
+\toprule
+\textbf{Slot} & \textbf{Description} \\
+\cmidrule(r){2-2}
+\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} \proglang{C++} variable \\
+\texttt{name} & simple name of the enum value \\
+\texttt{full\_name} & fully qualified name of the enum value \\[.3cm]
+%
+\textbf{Method} & \textbf{Description} \\
+\cmidrule(r){2-2}
+\texttt{number} & return the number of this EnumValueDescriptor. \\
+\texttt{name} & Return the name of the enum value descriptor.\\
+\texttt{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
+\texttt{as.character} & character representation of a descriptor. \\
+\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
+\texttt{asMessage} & return EnumValueDescriptorProto message. \\
+\bottomrule
+\end{tabular}
+\end{small}
+\caption{\label{EnumValueDescriptor-methods-table}Description of slots
+  and methods for the \texttt{EnumValueDescriptor} S4 class}
+\end{table}
+
 \subsection{File descriptors}
 \label{subsec-file-descriptor}
-
-The class \emph{FileDescriptor} represents file descriptors in \proglang{R}.
-This is a wrapper S4 class around the
-\texttt{google::protobuf::FileDescriptor} \proglang{C++} class.
-Table~\ref{filedescriptor-methods-table} describes the methods
-defined for the \texttt{FileDescriptor} class.
-
-The \verb|$| operator can be used to retrieve named fields defined in
-the FileDescriptor, or to invoke pseudo-methods.
-
-<<>>=
-f <- tutorial.Person$fileDescriptor()
-f
-f$Person
-@
 
@@
 
-\subsection{Enum value descriptors}
-\label{subsec-enumvalue-descriptor}
-
-The class \emph{EnumValueDescriptor} represents enumeration value
-descriptors in \proglang{R}.  This is a wrapper S4 class around the
-\texttt{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
-Table~\ref{EnumValueDescriptor-methods-table} describes the methods
-defined for the \texttt{EnumValueDescriptor} class.
-
-The \verb|$| operator can be used to invoke pseudo-methods.
-
-<<>>=
-tutorial.Person$PhoneType$value(1)
-tutorial.Person$PhoneType$value(name="HOME")
-tutorial.Person$PhoneType$value(number=1)
-@
-
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-\textbf{Slot} & \textbf{Description} \\
-\cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} \proglang{C++} variable \\
-\texttt{name} & simple name of the enum value \\
-\texttt{full\_name} & fully qualified name of the enum value \\[.3cm]
-%
-\textbf{Method} & \textbf{Description} \\
-\cmidrule(r){2-2}
-\texttt{number} & return the number of this EnumValueDescriptor. \\
-\texttt{name} & Return the name of the enum value descriptor.\\
-\texttt{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
-\texttt{as.character} & character representation of a descriptor. \\
-\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
-\texttt{asMessage} & return EnumValueDescriptorProto message. \\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{EnumValueDescriptor-methods-table}Description of slots
-  and methods for the \texttt{EnumValueDescriptor} S4 class}
-\end{table}
+The class \emph{FileDescriptor} represents file descriptors in \proglang{R}.
+This is a wrapper S4 class around the
+\texttt{google::protobuf::FileDescriptor} \proglang{C++} class.
+Table~\ref{filedescriptor-methods-table} describes the methods
+defined for the \texttt{FileDescriptor} class.
+
+The \verb|$| operator can be used to retrieve named fields defined in
+the FileDescriptor, or to invoke pseudo-methods.
+
+<<>>=
+f <- tutorial.Person$fileDescriptor()
+f
+f$Person
+@
 

### 2014-01-27 papers/jss/article.Rnw
Add two words suggested by Phillip for clarity.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@864 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 After importing proto files, the corresponding message descriptors are
-available from the \code{RProtoBuf:DescriptorPool} environment in 
+available by name from the \code{RProtoBuf:DescriptorPool} environment in 
 the \proglang{R} search path.  This environment is implemented with the 

### 2014-01-27 papers/jss/article.Rnw
One more comment made by Phillip that I had missed.
Be more specific about how readProtoFiles is automatically called to
read in addressbook.proto on package load, and add a footnote pointing
out that this is why we didn't have to explicitly call readprotofiles
in the example in table1.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@865 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 When \CRANpkg{RProtoBuf} is first loaded it calls
-\texttt{readProtoFiles} to read in an example \texttt{.proto} file
+\texttt{readProtoFiles} to read in the example \texttt{addressbook.proto} file
 included with the package.  The \texttt{tutorial.Person} descriptor
-and any other descriptors defined in loaded \texttt{.proto} files are
-then available on the search path.
+and all other descriptors defined in the loaded \texttt{.proto} files are
+then available on the search path\footnote{This explains why the example in
+Table~\ref{tab:proto} lacked an explicit call to
+\texttt{readProtoFiles}.}.
 

### 2014-01-29 papers/jss/article.Rnw
typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@868 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 They offer a unique combination of features, performance, and maturity that seems
-particulary well suited for data-driven applications and numerical
+particularly well suited for data-driven applications and numerical
 computing.

### 2014-03-16 papers/jss/article.Rnw
first pass concerning JSS comments
- cite all software: added Rserve
- "replication more accessible": huh?
- sentence style for \section, \subsection
- sentence style for \caption
- sentence style for column captions in tables
- quote argument to data() and library() 
- remove comments from quoted code
- adjust linebreak of first code example table
- reference to 'Springer-Verlag'
- title style for references


git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@874 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \end{center}
-\caption{Example protobuf usage}
+\caption{Example ussage of Protocol Buffers.}
 \label{fig:protobuf-distributed-usecase}
@@
 request, and respond with a new Protocol Buffer over the network. 
-The key difference to, say, a request to an \pkg{Rserve} instance is that
+The key difference to, say, a request to an \pkg{Rserve} 
+\citep{Urbanek:2003:Rserve,CRAN:Rserve} instance is that
 the remote server may be implemented in any language, with no
@@
 \begin{table}
-\begin{tabular}{p{.40\textwidth}p{0.55\textwidth}}
+\begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
 \toprule
-Schema : \texttt{addressbook.proto} & Example \proglang{R} Session\\
+Schema : \texttt{addressbook.proto} & Example \proglang{R} session\\
 \cmidrule{1-2}
@@
   enum PhoneType {
-   MOBILE = 0; HOME = 1;
+    MOBILE = 0; 
+    HOME = 1;
     WORK = 2;
@@
 <<echo=TRUE>>=
-library(RProtoBuf)
-p <- new(tutorial.Person,id=1,name="Dirk")
+library("RProtoBuf")
+p <- new(tutorial.Person, id=1,
+         name="Dirk")
 p$name
@@
 
-\section{Basic Usage: Messages and descriptors}
+\section{Basic usage: Messages and descriptors}
 \label{sec:rprotobuf-basic}
@@
 \texttt{serialize} can also be called in a more traditional
-object oriented fashion using the dollar operator:
-
-<<>>=
-# serialize to a file
+object oriented fashion using the dollar operator.
+
+<<>>=
 p$serialize(tf1)
-# serialize to a binary connection
 con <- file(tf2, open = "wb")
@@
 @
+
+Here, we first serialize to a file \code{tf1} before we serialize to a binary
+connection to file \code{tf2}.
 
@@
 
-Finally, the payload of the message can be used:
-
-<<>>=
-# reading the raw vector payload of the message
+Finally, the raw vector payload of the message can be used:
+
+<<>>=
 payload <- readBin(tf1, raw(0), 5000)
@@
 <<>>=
-# reading from a file
 message <- tutorial.Person$read(tf1)
-# reading from a binary connection
 con <- file(tf2, open = "rb")
@@
 close(con)
-# read from the payload
 message <- tutorial.Person$read(payload)
 @
+
+Here we read first from a file, then from a binary connection and lastly from
+a message payload.
 
@@
 \toprule
-\textbf{Class}      & 
-     \textbf{Slots} & 
-     \textbf{Methods} & 
-     \textbf{Dynamic Dispatch}\\
-\cmidrule{1-4}
+Class               & Slots & Methods & Dynamic dispatch\\
+\cmidrule{2-4}
 Message             & 2 & 20 & yes (field names)\\
@@
 \caption{\label{class-summary-table}Overview of class, slot, method and
-  dispatch relationships}
+  dispatch relationships.}
 \end{table}
@@
 \texttt{.DollarNames} S3 generic function defined in the \pkg{utils}
-package.
+package that is included with \pronglang{R} \citep{r}.
 
@@
 \toprule
-\textbf{Slot} & \textbf{Description} \\
+Slot & Description \\
 \cmidrule(r){2-2}
@@
 has its \texttt{type} slot set to \texttt{tutorial.Person} \\[.3cm]
-\textbf{Method} & \textbf{Description} \\
+
+Method & Description \\
 \cmidrule(r){2-2}
@@
 \end{small}
-\caption{\label{Message-methods-table}Description of slots and methods for the \texttt{Message} S4 class}
+\caption{\label{Message-methods-table}Description of slots and methods for the \texttt{Message} S4 class.}
 \end{table}
@@
 <<>>=
-tutorial.Person$email # field descriptor
-
-tutorial.Person$PhoneType # enum descriptor
-
-tutorial.Person$PhoneNumber # nested type descriptor
-# same as
+tutorial.Person$email
+
+tutorial.Person$PhoneType
+
+tutorial.Person$PhoneNumber
+
 tutorial.Person.PhoneNumber
@@
 %
-\textbf{Method} & \textbf{Description} \\
+
+Method & Description \\
 \cmidrule(r){2-2}
@@
 \end{small}
-\caption{\label{Descriptor-methods-table}Description of slots and methods for the \texttt{Descriptor} S4 class}
+\caption{\label{Descriptor-methods-table}Description of slots and methods for the \texttt{Descriptor} S4 class.}
 \end{table}
@@
 \caption{\label{fielddescriptor-methods-table}Description of slots and
-  methods for the \texttt{FieldDescriptor} S4 class}
+  methods for the \texttt{FieldDescriptor} S4 class.}
 \end{table}
@@
 \end{small}
-\caption{\label{enumdescriptor-methods-table}Description of slots and methods for the \texttt{EnumDescriptor} S4 class}
+\caption{\label{enumdescriptor-methods-table}Description of slots and methods
+  for the \texttt{EnumDescriptor} S4 class.} 
 \end{table}
@@
 \caption{\label{EnumValueDescriptor-methods-table}Description of slots
-  and methods for the \texttt{EnumValueDescriptor} S4 class}
+  and methods for the \texttt{EnumValueDescriptor} S4 class.}
 \end{table}
@@
 \texttt{package} & package name defined in this \texttt{.proto} file.\\[.3cm]
-\textbf{Method} & \textbf{Description} \\
+
+Method & Description \\
 \cmidrule(r){2-2}
@@
 \end{small}
-\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \texttt{FileDescriptor} S4 class}
+\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \texttt{FileDescriptor} S4 class.}
 \end{table}
@@
 \end{center}
-\caption{Diagram of MapReduce histogram generation pattern}
+\caption{Diagram of MapReduce histogram generation pattern.}
 \label{fig:mr-histogram-pattern1}
@@
 message data.  The following simple Python script uses this generated
-module to create a histogram and write out the Protocol Buffer
+module to create a histogram (to which breakpoints and binned data are
+added), and writes out the Protocol Buffer
 representation to a file:
@@
 
-# Create empty Histogram message
 hist = HistogramState()
 
-# Add breakpoints and binned data set.
 hist.counts.extend([2, 6, 2, 4, 6])
@@
 
-# Output the histogram
 outfile = open("/tmp/hist.pb", "wb")
@@
 The Protocol Buffer can then be read into \proglang{R} and converted to a native
-\proglang{R} histogram object for plotting:
+\proglang{R} histogram object for plotting. Here, the schema is read first,
+then the (serialized) histogram is read into the variable \code{hist} which
+is then converted a histogram object which is display as a plot.
 
 \begin{Code}
-library(RProtoBuf)
-library(HistogramTools)
-
-# Read the Histogram schema
+library("RProtoBuf")
+library("HistogramTools")
+
 readProtoFiles(package="HistogramTools")
 
-# Read the serialized histogram file.
 hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
@@
 
-# Convert to native R histogram object and plot
 plot(as.histogram(hist))
@@
 
-\section{Application: Data Interchange in web Services}
+\section{Application: Data interchange in web services}
 \label{sec:opencpu}
@@
 <<eval=FALSE>>=
-# Load packages
-library(RProtoBuf)
-library(httr)
-
-# Retrieve and parse message
+library("RProtoBuf")
+library("httr")
+
 req <- GET('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
@@
 
-# Check that no information was lost
 identical(output, MASS::Animals)
@@
 \begin{verbatim}
-# Import modules
 import urllib2
@@
 
-# Retrieve message
 req = urllib2.Request('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
@@
 
-# Parse rexp.proto message
 msg = REXP()
@@
 <<eval=FALSE>>=
-#requires httr >= 0.2.99
-library(httr)
-library(RProtoBuf)
+library("httr")       #requires httr >= 0.2.99
+library("RProtoBuf")
 
@@
 
-#This is the output of stats::rnorm(n=42, mean=100)
 output <- unserialize_pb(req$content)
@@
 \section*{Appendix: The rexp.proto schema descriptor}
+
 \label{rexp.proto}

### 2014-03-23 papers/jss/article.Rnw
one char typo
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@875 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Fix typo in Dirks last change.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@876 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
two new citation (as, I think, suggested by the note) minor twiddling with floats; slightly narrower figures
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@877 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 %
-\newcommand{\CRANpkg}[1]{\pkg{#1}}
 \RequirePackage{fancyvrb}
@@
 \Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\Shorttitle{\CRANpkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
+\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
 
@@
   computing.
-The
-\CRANpkg{RProtoBuf} package provides a complete interface to Protocol
+  The \pkg{RProtoBuf} package provides a complete interface to Protocol
   Buffers from the
@@
   requirements for statistical computing, describes the implementation
-of the \CRANpkg{RProtoBuf} package, and illustrates its use with
+  of the \pkg{RProtoBuf} package, and illustrates its use with
   example applications in large-scale data collection pipelines and web
   services.
-%TODO(ms) keep it less than 150 words. -- I think this may be 154,
-%depending how emacs is counting.
+  %% TODO(ms) keep it less than 150 words. -- I think this may be 154,
+  %% depending how emacs is counting.
 }
@@
 \emph{interface description language}, or \emph{IDL}.  IDLs like
-Protocol Buffers \citep{protobuf}, Apache Thrift, and Apache Avro
+Protocol Buffers \citep{protobuf}, Apache Thrift \citep{Apache:Thrift}, and Apache Avro \citep{Apache:Avro}
 provide a compact well-documented schema for cross-language data
@@
 compared with simple ``schema-less'' binary interchange formats.
-Many sources compare data serialization formats
-and show Protocol Buffers perform favorably to the alternatives; see
-\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
+%Many sources compare data serialization formats
+%and show Protocol Buffers perform favorably to the alternatives; see
+%\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
+Protocol Buffers performs well in the comparison of such formats by
+\citet{Sumaray:2012:CDS:2184751.2184810}.
 
@@
 Section~\ref{sec:rprotobuf-basic} describes the interactive \proglang{R} interface
-provided by the \CRANpkg{RProtoBuf} package, and introduces the two main abstractions:
+provided by the \pkg{RProtoBuf} package, and introduces the two main abstractions:
 \emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
@@
 it against the serialization capabilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
-and \ref{sec:opencpu} provide real-world use cases of \CRANpkg{RProtoBuf}
+and \ref{sec:opencpu} provide real-world use cases of \pkg{RProtoBuf}
 in MapReduce and web service environments, respectively, before
@@
 
-\begin{figure}[bp]
+%\begin{figure}[bp]
+\begin{figure}[h!]
 \begin{center}
-\includegraphics[width=\textwidth]{figures/protobuf-distributed-system-crop.pdf}
+\includegraphics[width=0.9\textwidth]{figures/protobuf-distributed-system-crop.pdf}
 \end{center}
-\caption{Example ussage of Protocol Buffers.}
+\caption{Example usage of Protocol Buffers.}
 \label{fig:protobuf-distributed-usecase}
@@
 \citep{Urbanek:2003:Rserve,CRAN:Rserve} instance is that
-the remote server may be implemented in any language, with no
-dependence on \proglang{R}.
+the remote server may be implemented in any language.
+%, with no dependence on \proglang{R}.
 
@@
 
-The \CRANpkg{RProtoBuf} package defines the \code{read} and
+The \pkg{RProtoBuf} package defines the \code{read} and
 \code{readASCII} functions to read messages from files, raw vectors,
@@
 
-The \CRANpkg{RProtoBuf} package uses the S4 system to store
+The \pkg{RProtoBuf} package uses the S4 system to store
 information about descriptors and messages.  Using the S4 system
@@
 Table~\ref{class-summary-table} lists the six
-primary Message and Descriptor classes in \CRANpkg{RProtoBuf}.  Each \proglang{R} object
+primary Message and Descriptor classes in \pkg{RProtoBuf}.  Each \proglang{R} object
 contains an external pointer to an object managed by the
@@
 
-The \CRANpkg{Rcpp} package
+The \pkg{Rcpp} package
 \citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
@@
 improvements at the cost of a more verbose implementation.
-The \CRANpkg{RProtoBuf} package in many ways motivated
-the development of \CRANpkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
+The \pkg{RProtoBuf} package in many ways motivated
+the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
 which provide a more concise way of wrapping \proglang{C++} functions and classes
@@
 
-The \CRANpkg{RProtoBuf} package supports two forms for calling
+The \pkg{RProtoBuf} package supports two forms for calling
 functions with these S4 classes:
@@
 
-Additionally, \CRANpkg{RProtoBuf} supports tab completion for all
+Additionally, \pkg{RProtoBuf} supports tab completion for all
 classes.  Completion possibilities include pseudo-method names for all
@@
 \texttt{.DollarNames} S3 generic function defined in the \pkg{utils}
-package that is included with \pronglang{R} \citep{r}.
+package that is included with \proglang{R} \citep{r}.
 
@@
 The \texttt{Message} S4 class represents Protocol Buffer Messages and
-is the core abstraction of \CRANpkg{RProtoBuf}. Each \texttt{Message}
+is the core abstraction of \pkg{RProtoBuf}. Each \texttt{Message}
 contains a pointer to a \texttt{Descriptor} which defines the schema
@@
 
-When \CRANpkg{RProtoBuf} is first loaded it calls
+When \pkg{RProtoBuf} is first loaded it calls
 \texttt{readProtoFiles} to read in the example \texttt{addressbook.proto} file
@@
 However, most modern languages do have support for 64-bit integers, 
-which becomes problematic when \CRANpkg{RProtoBuf} is used to exchange data 
+which becomes problematic when \pkg{RProtoBuf} is used to exchange data 
 with a system that requires this integer type. To work around this, 
-\CRANpkg{RProtoBuf} allows users to get and set 64-bit integer values by specifying 
+\pkg{RProtoBuf} allows users to get and set 64-bit integer values by specifying 
 them as character strings.
@@
 
-But when the values are specified as character strings, \CRANpkg{RProtoBuf}
+But when the values are specified as character strings, \pkg{RProtoBuf}
 will automatically coerce them into a true 64-bit integer types 
@@
 accurately be used as unique identifiers and can easily be passed to \proglang{R}
-packages such as \CRANpkg{int64} \citep{int64} or \CRANpkg{bit64}
+packages such as \pkg{int64} \citep{int64} or \pkg{bit64}
 \citep{bit64} which represent 64-bit integers in \proglang{R}.
@@
 
-The previous sections discussed functionality in the \CRANpkg{RProtoBuf} package
+The previous sections discussed functionality in the \pkg{RProtoBuf} package
 for creating, manipulating, parsing, and serializing Protocol Buffer
@@
 
-In order to accomplish this, \CRANpkg{RProtoBuf} uses the same catch-all \texttt{proto}
+In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
 schema used by \pkg{RHIPE} for exchanging \proglang{R} data with Hadoop \citep{rhipe}. This 
@@
 the appendix.
-The Protocol Buffer messages generated by \CRANpkg{RProtoBuf} and
+The Protocol Buffer messages generated by \pkg{RProtoBuf} and
 \pkg{RHIPE} are naturally compatible between the two systems because they use the 
@@
 One takeaway from this table is that the universal \proglang{R} object schema
-included in \CRANpkg{RProtoBuf} does not in general provide
+included in \pkg{RProtoBuf} does not in general provide
 any significant saving in file size compared to the normal serialization
@@
 % redundant: which is seen as equally compact.
-The benefits of \CRANpkg{RProtoBuf} accrue more naturally in applications where
+The benefits of \pkg{RProtoBuf} accrue more naturally in applications where
 multiple programming languages are involved, or when a more concise
@@
 \caption{Serialization sizes for default serialization in \proglang{R} and
-  \CRANpkg{RProtoBuf} for 50 \proglang{R} data sets.}
+  \pkg{RProtoBuf} for 50 \proglang{R} data sets.}
 \label{tab:compression}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/histogram-mapreduce-diag1.pdf}
+\includegraphics[width=0.9\textwidth]{figures/histogram-mapreduce-diag1.pdf}
 \end{center}
@@
 
-The \CRANpkg{HistogramTools} package \citep{histogramtools} enhances
-\CRANpkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
+The \pkg{HistogramTools} package \citep{histogramtools} enhances
+\pkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
 
@@
 OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
-from the \CRANpkg{RProtoBuf} package to convert between \proglang{R} objects and protobuf
+from the \pkg{RProtoBuf} package to convert between \proglang{R} objects and protobuf
 messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
@@
 <<eval=FALSE>>=
-library("httr")       #requires httr >= 0.2.99
+library("httr")       
 library("RProtoBuf")
@@
 
-The \CRANpkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
+The \pkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
 and extends the \proglang{R} system with the ability to create, read,
 write, parse, and manipulate Protocol
-Buffer messages. \CRANpkg{RProtoBuf} has been used extensively inside Google 
+Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
 for the past three years by statisticians, analysts, and software engineers.
 At the time of this writing there are over 300 active
-users of \CRANpkg{RProtoBuf} using it to read data from and otherwise interact
+users of \pkg{RProtoBuf} using it to read data from and otherwise interact
 with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
@@
 
-The first versions of \CRANpkg{RProtoBuf} were written during 2009-2010.
+The first versions of \pkg{RProtoBuf} were written during 2009 - 2010.
 Very significant contributions, both in code and design, were made by
@@
 greatly appreciated. Several features of the package reflect
-the design of the \CRANpkg{rJava} package by Simon Urbanek.
+the design of the \pkg{rJava} package by Simon Urbanek.
 The user-defined table mechanism, implemented by Duncan Temple Lang for the
@@
 Below a print of the \texttt{rexp.proto} schema (originally designed by \cite{rhipe})
-that is included with the \CRANpkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
+that is included with the \pkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
 \texttt{unserialize\_pb}.

### 2014-03-23 papers/jss/article.Rnw
also 'R CMD Stangle' to create single R file with code -- is that what they ask for?
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@878 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 ## cf http://www.jstatsoft.org/style#q12
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE, digits=4)
+options(prompt = "R> ", 
+        continue = "+  ", 
+        width = 70, 
+        useFancyQuotes = FALSE, 
+        digits = 4)
 @

### 2014-03-23 papers/jss/article.Rnw
s/texttt/code/ as the later is a JSS-recommended macro
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@879 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   Many formats commonly used to import and export data between
-  different programs or systems, such as \texttt{CSV} or \texttt{JSON}, are
+  different programs or systems, such as \code{CSV} or \code{JSON}, are
   verbose, inefficient, not type-safe, or tied to a specific programming language.
@@
 Data analysts and researchers often use character-separated text formats such
-as \texttt{CSV} \citep{shafranovich2005common} to export and import
-data. However, anyone who has ever used \texttt{CSV} files will have noticed
+as \code{CSV} \citep{shafranovich2005common} to export and import
+data. However, anyone who has ever used \code{CSV} files will have noticed
 that this method has many limitations: it is restricted to tabular data,
@@
 conventions on which characters is used as separator or decimal point vary by
-country.  \emph{Extensible Markup Language} (\texttt{XML}) is another
+country.  \emph{Extensible Markup Language} (\code{XML}) is another
 well-established and widely-supported format with the ability to define just
@@
 complexity at the parsing side (which are somewhat mitigated by the
-availability of mature libraries and parsers). Because \texttt{XML} is 
+availability of mature libraries and parsers). Because \code{XML} is 
 text-based and has no native notion of numeric types or arrays, it usually not a
@@
 A more modern format is \emph{JavaScript ObjectNotation} 
-(\texttt{JSON}), which is derived from the object literals of
+(\code{JSON}), which is derived from the object literals of
 \proglang{JavaScript}, and already widely-used on the world wide web. 
 Several \proglang{R} packages implement functions to parse and generate
-\texttt{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
-\texttt{JSON} natively supports arrays and four primitive types: numbers, strings,
+\code{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
+\code{JSON} natively supports arrays and four primitive types: numbers, strings,
 booleans, and null. However, as it too is a text-based format, numbers are
@@
 leads to loss of type (double versus integer) and precision. 
-A number of binary formats based on \texttt{JSON} have been proposed
+A number of binary formats based on \code{JSON} have been proposed
 that reduce the parsing cost and improve efficiency, but these formats
@@
 model that is flexible and easy to use.  The schema for structured
-Protocol Buffer data is defined in \texttt{.proto} files, which may
+Protocol Buffer data is defined in \code{.proto} files, which may
 contain one or more message types.  Each message type has one or more
@@
 enumerations, booleans, strings, raw bytes, or other nested message
-types.  The \texttt{.proto} file syntax for defining the structure of Protocol
+types.  The \code{.proto} file syntax for defining the structure of Protocol
 Buffer data is described comprehensively on Google Code\footnote{See 
 \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
-Table~\ref{tab:proto} shows an example \texttt{.proto} file that
-defines the \texttt{tutorial.Person} type\footnote{The compound name
-  \texttt{tutorial.Person} in R is derived from the name of the
+Table~\ref{tab:proto} shows an example \code{.proto} file that
+defines the \code{tutorial.Person} type\footnote{The compound name
+  \code{tutorial.Person} in R is derived from the name of the
   message (\emph{Person}) and the name of the package defined at the top of the
-  \texttt{.proto} file in which it is defined (\emph{tutorial}).}.  The \proglang{R} code in the right
+  \code{.proto} file in which it is defined (\emph{tutorial}).}.  The \proglang{R} code in the right
 column shows an example of creating a new message of this type and
@@
 \toprule
-Schema : \texttt{addressbook.proto} & Example \proglang{R} session\\
+Schema : \code{addressbook.proto} & Example \proglang{R} session\\
 \cmidrule{1-2}
@@
 \end{tabular}
-\caption{The schema representation from a \texttt{.proto} file for the
-  \texttt{tutorial.Person} class (left) and simple \proglang{R} code for creating
+\caption{The schema representation from a \code{.proto} file for the
+  \code{tutorial.Person} class (left) and simple \proglang{R} code for creating
   an object of this class and accessing its fields (right).}
@@
 Protocol Buffers are used with a compiler that translates a Protocol
-Buffer schema description file (ending in \texttt{.proto}) into
+Buffer schema description file (ending in \code{.proto}) into
 language-specific classes that can be used to create, read, write, and
@@
 structured data fields of the type specified in a Message Descriptor.
-Message Descriptors are defined in \texttt{.proto} files and define a
+Message Descriptors are defined in \code{.proto} files and define a
 schema for a particular named class of messages.
 
-\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \texttt{.proto} files}
+\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \code{.proto} files}
 
 To create or parse a Protocol Buffer Message, one must first read in 
-the message type specification from a \texttt{.proto} file. The 
-\texttt{.proto} files are imported using the \code{readProtoFiles}
+the message type specification from a \code{.proto} file. The 
+\code{.proto} files are imported using the \code{readProtoFiles}
 function, which can either import a single file, all files in a directory,
-or every \texttt{.proto} file provided by a particular \proglang{R} package.
+or every \code{.proto} file provided by a particular \proglang{R} package.
 
@@
 
-New messages are created with the \texttt{new} function which accepts
+New messages are created with the \code{new} function which accepts
 a Message Descriptor and optionally a list of ``name = value'' pairs
@@
 %discussed in detail in another part of this document, but for the
-%purpose of this section, descriptors are just used with the \texttt{new}
+%purpose of this section, descriptors are just used with the \code{new}
 %function to create messages.
@@
 
-Protocol Buffer messages and descriptors implement \texttt{show}
+Protocol Buffer messages and descriptors implement \code{show}
 methods that provide basic information about the message:
@@
 For additional information, such as for debugging purposes,
-the \texttt{as.character} method provides a more complete ASCII
+the \code{as.character} method provides a more complete ASCII
 representation of the contents of a message.
@@
 binary wire-format representation.  
-The \texttt{serialize} method is implemented for
+The \code{serialize} method is implemented for
 Protocol Buffer messages to serialize a message into a sequence of
@@
 
-\texttt{serialize} can also be called in a more traditional
+\code{serialize} can also be called in a more traditional
 object oriented fashion using the dollar operator.
@@
 
-The \texttt{input} argument of \texttt{read} can also be a binary
+The \code{input} argument of \code{read} can also be a binary
 readable \proglang{R} connection, such as a binary file connection:
@@
 
-\texttt{read} can also be used as a pseudo-method of the descriptor
+\code{read} can also be used as a pseudo-method of the descriptor
 object:
@@
 allows the package to dispatch methods that are not
-generic in the S3 sense, such as \texttt{new} and
-\texttt{serialize}.
+generic in the S3 sense, such as \code{new} and
+\code{serialize}.
 Table~\ref{class-summary-table} lists the six
@@
 contains an external pointer to an object managed by the
-\texttt{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
+\code{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
 than 100 \proglang{C++} functions that provide the
@@
 object.  This functionality is implemented with the
-\texttt{.DollarNames} S3 generic function defined in the \pkg{utils}
+\code{.DollarNames} S3 generic function defined in the \pkg{utils}
 package that is included with \proglang{R} \citep{r}.
@@
 
-The \texttt{Message} S4 class represents Protocol Buffer Messages and
-is the core abstraction of \pkg{RProtoBuf}. Each \texttt{Message}
-contains a pointer to a \texttt{Descriptor} which defines the schema
+The \code{Message} S4 class represents Protocol Buffer Messages and
+is the core abstraction of \pkg{RProtoBuf}. Each \code{Message}
+contains a pointer to a \code{Descriptor} which defines the schema
 of the data defined in the Message, as well as a number of
-\texttt{FieldDescriptors} for the individual fields of the message.  A
-complete list of the slots and methods for \texttt{Messages}
+\code{FieldDescriptors} for the individual fields of the message.  A
+complete list of the slots and methods for \code{Messages}
 is available in Table~\ref{Message-methods-table}.
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{Message} object of the \proglang{C++} protobuf library. Documentation for the
-\texttt{Message} class is available from the Protocol Buffer project page. \\
+\code{pointer} & External pointer to the \code{Message} object of the \proglang{C++} protobuf library. Documentation for the
+\code{Message} class is available from the Protocol Buffer project page. \\
 %(\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message}) \\
-\texttt{type} & Fully qualified name of the message. For example a \texttt{Person} message
-has its \texttt{type} slot set to \texttt{tutorial.Person} \\[.3cm]
+\code{type} & Fully qualified name of the message. For example a \code{Person} message
+has its \code{type} slot set to \code{tutorial.Person} \\[.3cm]
 
@@
 \cmidrule(r){2-2}
-\texttt{has} & Indicates if a message has a given field.   \\
-\texttt{clone} & Creates a clone of the message \\
-\texttt{isInitialized} & Indicates if a message has all its required fields set\\
-\texttt{serialize} & serialize a message to a file, binary connection, or raw vector\\
-\texttt{clear} & Clear one or several fields of a message, or the entire message\\
-\texttt{size} & The number of elements in a message field\\
-\texttt{bytesize} & The number of bytes the message would take once serialized\\[3mm]
+\code{has} & Indicates if a message has a given field.   \\
+\code{clone} & Creates a clone of the message \\
+\code{isInitialized} & Indicates if a message has all its required fields set\\
+\code{serialize} & serialize a message to a file, binary connection, or raw vector\\
+\code{clear} & Clear one or several fields of a message, or the entire message\\
+\code{size} & The number of elements in a message field\\
+\code{bytesize} & The number of bytes the message would take once serialized\\[3mm]
 %
-\texttt{swap} & swap elements of a repeated field of a message\\
-\texttt{set} & set elements of a repeated field\\
-\texttt{fetch} & fetch elements of a repeated field\\
-\texttt{setExtension} & set an extension of a message\\
-\texttt{getExtension} & get the value of an extension of a message\\
-\texttt{add} & add elements to a repeated field \\[3mm]
+\code{swap} & swap elements of a repeated field of a message\\
+\code{set} & set elements of a repeated field\\
+\code{fetch} & fetch elements of a repeated field\\
+\code{setExtension} & set an extension of a message\\
+\code{getExtension} & get the value of an extension of a message\\
+\code{add} & add elements to a repeated field \\[3mm]
 %
-\texttt{str} & the \proglang{R} structure of the message\\
-\texttt{as.character} & character representation of a message\\
-\texttt{toString} & character representation of a message (same as \texttt{as.character}) \\
-\texttt{as.list} & converts message to a named \proglang{R} list\\
-\texttt{update} & updates several fields of a message at once\\
-\texttt{descriptor} & get the descriptor of the message type of this message\\
-\texttt{fileDescriptor} & get the file descriptor of this message's descriptor\\
+\code{str} & the \proglang{R} structure of the message\\
+\code{as.character} & character representation of a message\\
+\code{toString} & character representation of a message (same as \code{as.character}) \\
+\code{as.list} & converts message to a named \proglang{R} list\\
+\code{update} & updates several fields of a message at once\\
+\code{descriptor} & get the descriptor of the message type of this message\\
+\code{fileDescriptor} & get the file descriptor of this message's descriptor\\
 \hline
@@
 \end{small}
-\caption{\label{Message-methods-table}Description of slots and methods for the \texttt{Message} S4 class.}
+\caption{\label{Message-methods-table}Description of slots and methods for the \code{Message} S4 class.}
 \end{table}
@@
 descriptors are represented in \proglang{R} by the \emph{Descriptor} S4
-class. The class contains the slots \texttt{pointer} and
-\texttt{type}.  Similarly to messages, the \verb|$| operator can be
+class. The class contains the slots \code{pointer} and
+\code{type}.  Similarly to messages, the \verb|$| operator can be
 used to retrieve descriptors that are contained in the descriptor, or
@@
 When \pkg{RProtoBuf} is first loaded it calls
-\texttt{readProtoFiles} to read in the example \texttt{addressbook.proto} file
-included with the package.  The \texttt{tutorial.Person} descriptor
-and all other descriptors defined in the loaded \texttt{.proto} files are
+\code{readProtoFiles} to read in the example \code{addressbook.proto} file
+included with the package.  The \code{tutorial.Person} descriptor
+and all other descriptors defined in the loaded \code{.proto} files are
 then available on the search path\footnote{This explains why the example in
 Table~\ref{tab:proto} lacked an explicit call to
-\texttt{readProtoFiles}.}.
+\code{readProtoFiles}.}.
 
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{Descriptor} object of the \proglang{C++} proto library. Documentation for the
-\texttt{Descriptor} class is available from the Protocol Buffer project page.\\
+\code{pointer} & External pointer to the \code{Descriptor} object of the \proglang{C++} proto library. Documentation for the
+\code{Descriptor} class is available from the Protocol Buffer project page.\\
 %\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
-\texttt{type} & Fully qualified path of the message type. \\[.3cm]
+\code{type} & Fully qualified path of the message type. \\[.3cm]
 %
@@
 \cmidrule(r){2-2}
-\texttt{new} & Creates a prototype of a message described by this descriptor.\\
-\texttt{read} & Reads a message from a file or binary connection.\\
-\texttt{readASCII} & Read a message in ASCII format from a file or
+\code{new} & Creates a prototype of a message described by this descriptor.\\
+\code{read} & Reads a message from a file or binary connection.\\
+\code{readASCII} & Read a message in ASCII format from a file or
 text connection.\\
-\texttt{name} & Retrieve the name of the message type associated with
+\code{name} & Retrieve the name of the message type associated with
 this descriptor.\\
-\texttt{as.character} & character representation of a descriptor\\
-\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}) \\
-\texttt{as.list} & return a named
+\code{as.character} & character representation of a descriptor\\
+\code{toString} & character representation of a descriptor (same as \code{as.character}) \\
+\code{as.list} & return a named
 list of the field, enum, and nested descriptors included in this descriptor.\\
-\texttt{asMessage} & return DescriptorProto message. \\
-\texttt{fileDescriptor} & Retrieve the file descriptor of this
+\code{asMessage} & return DescriptorProto message. \\
+\code{fileDescriptor} & Retrieve the file descriptor of this
 descriptor.\\
-\texttt{containing\_type} & Retrieve the descriptor describing the message type containing this descriptor.\\
-\texttt{field\_count} & Return the number of fields in this descriptor.\\
-\texttt{field} & Return the descriptor for the specified field in this descriptor.\\
-\texttt{nested\_type\_count} & The number of nested types in this descriptor.\\
-\texttt{nested\_type} & Return the descriptor for the specified nested 
+\code{containing\_type} & Retrieve the descriptor describing the message type containing this descriptor.\\
+\code{field\_count} & Return the number of fields in this descriptor.\\
+\code{field} & Return the descriptor for the specified field in this descriptor.\\
+\code{nested\_type\_count} & The number of nested types in this descriptor.\\
+\code{nested\_type} & Return the descriptor for the specified nested 
 type in this descriptor.\\
-\texttt{enum\_type\_count} & The number of enum types in this descriptor.\\
-\texttt{enum\_type} & Return the descriptor for the specified enum
+\code{enum\_type\_count} & The number of enum types in this descriptor.\\
+\code{enum\_type} & Return the descriptor for the specified enum
 type in this descriptor.\\
@@
 \end{small}
-\caption{\label{Descriptor-methods-table}Description of slots and methods for the \texttt{Descriptor} S4 class.}
+\caption{\label{Descriptor-methods-table}Description of slots and methods for the \code{Descriptor} S4 class.}
 \end{table}
@@
 descriptors in \proglang{R}. This is a wrapper S4 class around the
-\texttt{google::protobuf::FieldDescriptor} \proglang{C++} class.
+\code{google::protobuf::FieldDescriptor} \proglang{C++} class.
 Table~\ref{fielddescriptor-methods-table} describes the methods
-defined for the \texttt{FieldDescriptor} class.
+defined for the \code{FieldDescriptor} class.
 
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{FieldDescriptor} \proglang{C++} variable \\
-\texttt{name} & Simple name of the field \\
-\texttt{full\_name} & Fully qualified name of the field \\
-\texttt{type} & Name of the message type where the field is declared \\[.3cm]
+\code{pointer} & External pointer to the \code{FieldDescriptor} \proglang{C++} variable \\
+\code{name} & Simple name of the field \\
+\code{full\_name} & Fully qualified name of the field \\
+\code{type} & Name of the message type where the field is declared \\[.3cm]
 %
@@
 \cmidrule(r){2-2}
-\texttt{as.character} & Character representation of a descriptor\\
-\texttt{toString} & Character representation of a descriptor (same as \texttt{as.character}) \\
-\texttt{asMessage} & Return FieldDescriptorProto message. \\
-\texttt{name} & Return the name of the field descriptor.\\
-\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
-\texttt{containing\_type} & Return the containing descriptor of this field.\\
-\texttt{is\_extension} & Return TRUE if this field is an extension.\\
-\texttt{number} & Gets the declared tag number of the field.\\
-\texttt{type} & Gets the type of the field.\\
-\texttt{cpp\_type} & Gets the \proglang{C++} type of the field.\\
-\texttt{label} & Gets the label of a field (optional, required, or repeated).\\
-\texttt{is\_repeated} & Return TRUE if this field is repeated.\\
-\texttt{is\_required} & Return TRUE if this field is required.\\
-\texttt{is\_optional} & Return TRUE if this field is optional.\\
-\texttt{has\_default\_value} & Return TRUE if this field has a default value.\\
-\texttt{default\_value} & Return the default value.\\
-\texttt{message\_type} & Return the message type if this is a message type field.\\
-\texttt{enum\_type} & Return the enum type if this is an enum type field.\\
+\code{as.character} & Character representation of a descriptor\\
+\code{toString} & Character representation of a descriptor (same as \code{as.character}) \\
+\code{asMessage} & Return FieldDescriptorProto message. \\
+\code{name} & Return the name of the field descriptor.\\
+\code{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
+\code{containing\_type} & Return the containing descriptor of this field.\\
+\code{is\_extension} & Return TRUE if this field is an extension.\\
+\code{number} & Gets the declared tag number of the field.\\
+\code{type} & Gets the type of the field.\\
+\code{cpp\_type} & Gets the \proglang{C++} type of the field.\\
+\code{label} & Gets the label of a field (optional, required, or repeated).\\
+\code{is\_repeated} & Return TRUE if this field is repeated.\\
+\code{is\_required} & Return TRUE if this field is required.\\
+\code{is\_optional} & Return TRUE if this field is optional.\\
+\code{has\_default\_value} & Return TRUE if this field has a default value.\\
+\code{default\_value} & Return the default value.\\
+\code{message\_type} & Return the message type if this is a message type field.\\
+\code{enum\_type} & Return the enum type if this is an enum type field.\\
 \bottomrule
@@
 \caption{\label{fielddescriptor-methods-table}Description of slots and
-  methods for the \texttt{FieldDescriptor} S4 class.}
+  methods for the \code{FieldDescriptor} S4 class.}
 \end{table}
@@
 This is a wrapper S4 class around the
-\texttt{google::protobuf::EnumDescriptor} \proglang{C++} class.
+\code{google::protobuf::EnumDescriptor} \proglang{C++} class.
 Table~\ref{enumdescriptor-methods-table} describes the methods
-defined for the \texttt{EnumDescriptor} class.
+defined for the \code{EnumDescriptor} class.
 
@@
 
-The \texttt{EnumDescriptor} contains information about what values this type
-defines, while the \texttt{EnumValueDescriptor} describes an
+The \code{EnumDescriptor} contains information about what values this type
+defines, while the \code{EnumValueDescriptor} describes an
 individual enum constant of a particular type.
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{EnumDescriptor} \proglang{C++} variable \\
-\texttt{name} & Simple name of the enum \\
-\texttt{full\_name} & Fully qualified name of the enum \\
-\texttt{type} & Name of the message type where the enum is declared \\[.3cm]
+\code{pointer} & External pointer to the \code{EnumDescriptor} \proglang{C++} variable \\
+\code{name} & Simple name of the enum \\
+\code{full\_name} & Fully qualified name of the enum \\
+\code{type} & Name of the message type where the enum is declared \\[.3cm]
 %
@@
 \cmidrule(r){2-2}
-\texttt{as.list} & return a named
+\code{as.list} & return a named
 integer vector with the values of the enum and their names.\\
-\texttt{as.character} & character representation of a descriptor\\
-\texttt{toString} & character
-representation of a descriptor (same as \texttt{as.character}) \\
-\texttt{asMessage} & return EnumDescriptorProto message. \\
-\texttt{name} & Return the name of the enum descriptor.\\
-\texttt{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
-\texttt{containing\_type} & Return the containing descriptor of this field.\\
-\texttt{length} & Return the number of constants in this enum.\\
-\texttt{has} & Return TRUE if this enum contains the specified named constant string.\\
-\texttt{value\_count} & Return the number of constants in this enum (same as \texttt{length}).\\
-\texttt{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
+\code{as.character} & character representation of a descriptor\\
+\code{toString} & character
+representation of a descriptor (same as \code{as.character}) \\
+\code{asMessage} & return EnumDescriptorProto message. \\
+\code{name} & Return the name of the enum descriptor.\\
+\code{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
+\code{containing\_type} & Return the containing descriptor of this field.\\
+\code{length} & Return the number of constants in this enum.\\
+\code{has} & Return TRUE if this enum contains the specified named constant string.\\
+\code{value\_count} & Return the number of constants in this enum (same as \code{length}).\\
+\code{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
 \bottomrule
@@
 \caption{\label{enumdescriptor-methods-table}Description of slots and methods
-  for the \texttt{EnumDescriptor} S4 class.} 
+  for the \code{EnumDescriptor} S4 class.} 
 \end{table}
@@
 descriptors in \proglang{R}.  This is a wrapper S4 class around the
-\texttt{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
+\code{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
 Table~\ref{EnumValueDescriptor-methods-table} describes the methods
-defined for the \texttt{EnumValueDescriptor} class.
+defined for the \code{EnumValueDescriptor} class.
 
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & External pointer to the \texttt{EnumValueDescriptor} \proglang{C++} variable \\
-\texttt{name} & simple name of the enum value \\
-\texttt{full\_name} & fully qualified name of the enum value \\[.3cm]
+\code{pointer} & External pointer to the \code{EnumValueDescriptor} \proglang{C++} variable \\
+\code{name} & simple name of the enum value \\
+\code{full\_name} & fully qualified name of the enum value \\[.3cm]
 %
@@
 \cmidrule(r){2-2}
-\texttt{number} & return the number of this EnumValueDescriptor. \\
-\texttt{name} & Return the name of the enum value descriptor.\\
-\texttt{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
-\texttt{as.character} & character representation of a descriptor. \\
-\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
-\texttt{asMessage} & return EnumValueDescriptorProto message. \\
+\code{number} & return the number of this EnumValueDescriptor. \\
+\code{name} & Return the name of the enum value descriptor.\\
+\code{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
+\code{as.character} & character representation of a descriptor. \\
+\code{toString} & character representation of a descriptor (same as \code{as.character}). \\
+\code{asMessage} & return EnumValueDescriptorProto message. \\
 \bottomrule
@@
 \caption{\label{EnumValueDescriptor-methods-table}Description of slots
-  and methods for the \texttt{EnumValueDescriptor} S4 class.}
+  and methods for the \code{EnumValueDescriptor} S4 class.}
 \end{table}
@@
 \cmidrule(r){2-2}
-\texttt{pointer} & external pointer to the \texttt{FileDescriptor} object of the \proglang{C++} proto library. Documentation for the
-\texttt{FileDescriptor} class is available from the Protocol Buffer project page:
+\code{pointer} & external pointer to the \code{FileDescriptor} object of the \proglang{C++} proto library. Documentation for the
+\code{FileDescriptor} class is available from the Protocol Buffer project page:
 \url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
-\texttt{filename} & fully qualified pathname of the \texttt{.proto} file.\\
-\texttt{package} & package name defined in this \texttt{.proto} file.\\[.3cm]
+\code{filename} & fully qualified pathname of the \code{.proto} file.\\
+\code{package} & package name defined in this \code{.proto} file.\\[.3cm]
 
@@
 \cmidrule(r){2-2}
-\texttt{name} & Return the filename for this FileDescriptorProto.\\
-\texttt{package} & Return the file-level package name specified in this FileDescriptorProto.\\
-\texttt{as.character} & character representation of a descriptor. \\
-\texttt{toString} & character representation of a descriptor (same as \texttt{as.character}). \\
-\texttt{asMessage} & return FileDescriptorProto message. \\
-\texttt{as.list} & return named list of descriptors defined in this file descriptor.\\
+\code{name} & Return the filename for this FileDescriptorProto.\\
+\code{package} & Return the file-level package name specified in this FileDescriptorProto.\\
+\code{as.character} & character representation of a descriptor. \\
+\code{toString} & character representation of a descriptor (same as \code{as.character}). \\
+\code{asMessage} & return FileDescriptorProto message. \\
+\code{as.list} & return named list of descriptors defined in this file descriptor.\\
 \bottomrule
@@
 \end{small}
-\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \texttt{FileDescriptor} S4 class.}
+\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \code{FileDescriptor} S4 class.}
 \end{table}
@@
 This is a wrapper S4 class around the
-\texttt{google::protobuf::FileDescriptor} \proglang{C++} class.
+\code{google::protobuf::FileDescriptor} \proglang{C++} class.
 Table~\ref{filedescriptor-methods-table} describes the methods
-defined for the \texttt{FileDescriptor} class.
+defined for the \code{FileDescriptor} class.
 
@@
 \cmidrule(r){2-3}
-double	& \texttt{double} vector & \texttt{double} vector \\
-float	& \texttt{double} vector & \texttt{double} vector \\[3mm]
-uint32	  & \texttt{double} vector & \texttt{double} vector \\
-fixed32	  & \texttt{double} vector & \texttt{double} vector \\[3mm]
-int32	  & \texttt{integer} vector & \texttt{integer} vector \\
-sint32	  & \texttt{integer} vector & \texttt{integer} vector \\
-sfixed32  & \texttt{integer} vector & \texttt{integer} vector \\[3mm]
-int64	  & \texttt{integer} or \texttt{character}
-vector    & \texttt{integer} or \texttt{character} vector \\
-uint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-sint64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-fixed64	  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\
-sfixed64  & \texttt{integer} or \texttt{character} vector & \texttt{integer} or \texttt{character} vector \\[3mm]
-bool	& \texttt{logical} vector & \texttt{logical} vector \\[3mm]
-string	& \texttt{character} vector & \texttt{character} vector \\
-bytes	& \texttt{character} vector & \texttt{character} vector \\[3mm]
-enum & \texttt{integer} vector & \texttt{integer} vector \\[3mm]
-message & \texttt{S4} object of class \texttt{Message} & \texttt{list} of \texttt{S4} objects of class \texttt{Message} \\
+double	& \code{double} vector & \code{double} vector \\
+float	& \code{double} vector & \code{double} vector \\[3mm]
+uint32	  & \code{double} vector & \code{double} vector \\
+fixed32	  & \code{double} vector & \code{double} vector \\[3mm]
+int32	  & \code{integer} vector & \code{integer} vector \\
+sint32	  & \code{integer} vector & \code{integer} vector \\
+sfixed32  & \code{integer} vector & \code{integer} vector \\[3mm]
+int64	  & \code{integer} or \code{character}
+vector    & \code{integer} or \code{character} vector \\
+uint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
+sint64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
+fixed64	  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\
+sfixed64  & \code{integer} or \code{character} vector & \code{integer} or \code{character} vector \\[3mm]
+bool	& \code{logical} vector & \code{logical} vector \\[3mm]
+string	& \code{character} vector & \code{character} vector \\
+bytes	& \code{character} vector & \code{character} vector \\[3mm]
+enum & \code{integer} vector & \code{integer} vector \\[3mm]
+message & \code{S4} object of class \code{Message} & \code{list} of \code{S4} objects of class \code{Message} \\
 \bottomrule
@@
 
-\proglang{R} booleans can accept three values: \texttt{TRUE}, \texttt{FALSE}, and
-\texttt{NA}.  However, most other languages, including the Protocol
-Buffer schema, only accept \texttt{TRUE} or \texttt{FALSE}.  This means
+\proglang{R} booleans can accept three values: \code{TRUE}, \code{FALSE}, and
+\code{NA}.  However, most other languages, including the Protocol
+Buffer schema, only accept \code{TRUE} or \code{FALSE}.  This means
 that we simply can not store \proglang{R} logical vectors that include all three
 possible values as booleans.  The library will refuse to store
-\texttt{NA}s in Protocol Buffer boolean fields, and users must instead
+\code{NA}s in Protocol Buffer boolean fields, and users must instead
 choose another type (such as enum or integer) capable of storing three
@@
 will be returned if the \code{RProtoBuf.int64AsString} option is set
-to \texttt{TRUE}.  The character values are useful because they can
+to \code{TRUE}.  The character values are useful because they can
 accurately be used as unique identifiers and can easily be passed to \proglang{R}
@@
 
-In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \texttt{proto}
+In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \code{proto}
 schema used by \pkg{RHIPE} for exchanging \proglang{R} data with Hadoop \citep{rhipe}. This 
-schema, which we will refer to as \texttt{rexp.proto}, is printed in
+schema, which we will refer to as \code{rexp.proto}, is printed in
 %appendix \ref{rexp.proto}.
@@
 
-The \texttt{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
-These include \texttt{NULL}, \texttt{list} and vectors of type \texttt{logical}, 
-\texttt{character}, \texttt{double}, \texttt{integer}, and \texttt{complex}. In addition,
-every type can contain a named set of attributes, as is the case in \proglang{R}. The \texttt{rexp.proto}
-schema does not support some of the special \proglang{R} specific storage types, such as \texttt{function},
-\texttt{language} or \texttt{environment}. Such objects have no native equivalent 
+The \code{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
+These include \code{NULL}, \code{list} and vectors of type \code{logical}, 
+\code{character}, \code{double}, \code{integer}, and \code{complex}. In addition,
+every type can contain a named set of attributes, as is the case in \proglang{R}. The \code{rexp.proto}
+schema does not support some of the special \proglang{R} specific storage types, such as \code{function},
+\code{language} or \code{environment}. Such objects have no native equivalent 
 type in Protocol Buffers, and have little meaning outside the context of \proglang{R}.
-When serializing \proglang{R} objects using \texttt{serialize\_pb}, values or attributes of
+When serializing \proglang{R} objects using \code{serialize\_pb}, values or attributes of
 unsupported types are skipped with a warning. If the user really wishes to serialize these 
 objects, they need to be converted into a supported type. For example, the  can use 
-\texttt{deparse} to convert functions or language objects into strings, or \texttt{as.list}
+\code{deparse} to convert functions or language objects into strings, or \code{as.list}
 for environments.
@@
 package included with \proglang{R}. These data sets include data frames, matrices, time series, tables lists,
-and some more exotic data classes. The \texttt{can\_serialize\_pb} method is 
-used to determine which of those can fully be converted to the \texttt{rexp.proto}
+and some more exotic data classes. The \code{can\_serialize\_pb} method is 
+used to determine which of those can fully be converted to the \code{rexp.proto}
 Protocol Buffer representation. This method simply checks if any of the values or
@@
 without loss of information (\Sexpr{format(100*m/n,digits=1)}\%). Upon closer
-inspection, all other data sets are objects of class \texttt{nfnGroupedData}.
+inspection, all other data sets are objects of class \code{nfnGroupedData}.
 This class represents a special type of data frame that has some additional 
@@
 Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
-other systems, and are not supported by the \texttt{rexp.proto} descriptor.
-When \texttt{serialize\_pb} is used on objects of this class, it will serialize
+other systems, and are not supported by the \code{rexp.proto} descriptor.
+When \code{serialize\_pb} is used on objects of this class, it will serialize
 the data frame and all attributes, except for the formula.
@@
 messages in Python for later consumption by \proglang{R}, we first compile the 
-\texttt{histogram.proto} descriptor into a python module using the
-\texttt{protoc} compiler:
+\code{histogram.proto} descriptor into a python module using the
+\code{protoc} compiler:
 
@@
 \end{verbatim}
-This generates a Python module called \texttt{histogram\_pb2.py}, containing both the 
+This generates a Python module called \code{histogram\_pb2.py}, containing both the 
 descriptor information as well as methods to read and manipulate the histogram 
@@
 structures into well defined messages, which can then be exchanged using any
-protocol. The descriptors (\texttt{.proto} files) are used to formally define
+protocol. The descriptors (\code{.proto} files) are used to formally define
 the interface of a remote API or network application. Libraries to parse and
@@
 using several data interchange formats, such as Protocol Buffers.  
-OpenCPU uses the \texttt{serialize\_pb} and \texttt{unserialize\_pb} functions
+OpenCPU uses the \code{serialize\_pb} and \code{unserialize\_pb} functions
 from the \pkg{RProtoBuf} package to convert between \proglang{R} objects and protobuf
-messages. Therefore, clients need the \texttt{rexp.proto} descriptor mentioned
+messages. Therefore, clients need the \code{rexp.proto} descriptor mentioned
 earlier to parse and generate protobuf messages when interacting with OpenCPU.
@@
 
-The \texttt{HTTP GET} method is used to read a resource from OpenCPU. For example,
-to access the data set \texttt{Animals} from the package \texttt{MASS}, a 
+The \code{HTTP GET} method is used to read a resource from OpenCPU. For example,
+to access the data set \code{Animals} from the package \code{MASS}, a 
 client performs the following HTTP request:
@@
 \end{verbatim}
-The postfix \texttt{/pb} in the URL tells the server to send this
+The postfix \code{/pb} in the URL tells the server to send this
 object in the form of a protobuf message. Alternative formats include 
-\texttt{/json}, \texttt{/csv}, \texttt{/rds} and others. If the request
+\code{/json}, \code{/csv}, \code{/rds} and others. If the request
 is successful, OpenCPU returns the serialized object with HTTP status 
-code 200 and HTTP response header \texttt{Content-Type: application/x-protobuf}. 
+code 200 and HTTP response header \code{Content-Type: application/x-protobuf}. 
 The latter is the conventional MIME type that formally notifies the client to
@@
 OpenCPU using a protobuf message. In \proglang{R}, we use the HTTP client from 
-the \texttt{httr} package \citep{httr}. In this example we
+the \code{httr} package \citep{httr}. In this example we
 download a data set which is part of the base \proglang{R} distribution, so we can
@@
 data set in a Python client. To parse messages in Python, we first compile the 
-\texttt{rexp.proto} descriptor into a python module using the \texttt{protoc} compiler:
+\code{rexp.proto} descriptor into a python module using the \code{protoc} compiler:
 
@@
 \end{verbatim}
-This generates Python module called \texttt{rexp\_pb2.py}, containing both the 
+This generates Python module called \code{rexp\_pb2.py}, containing both the 
 descriptor information as well as methods to read and manipulate the \proglang{R} object 
-message. In the example below we use the HTTP client from the \texttt{urllib2}
+message. In the example below we use the HTTP client from the \code{urllib2}
 module.
@@
 \end{verbatim}
-The \texttt{msg} object contains all data from the Animals data set. From here we
+The \code{msg} object contains all data from the Animals data set. From here we
 can easily extract the desired fields for further use in Python.
@@
 
-The example above shows how the \texttt{HTTP GET} method retrieves a 
-resource from OpenCPU, for example an \proglang{R} object. The \texttt{HTTP POST} 
+The example above shows how the \code{HTTP GET} method retrieves a 
+resource from OpenCPU, for example an \proglang{R} object. The \code{HTTP POST} 
 method on the other hand is used for calling functions and running scripts, 
-which is the primary purpose of the framework. As before, the \texttt{/pb} 
+which is the primary purpose of the framework. As before, the \code{/pb} 
 postfix requests to retrieve the output as a protobuf message, in this
@@
 function arguments is not required, and for simple (scalar) arguments 
-the standard \texttt{application/x-www-form-urlencoded} format might be sufficient.
+the standard \code{application/x-www-form-urlencoded} format might be sufficient.
 However, with Protocol Buffers the client can perform function calls with
@@
 The following example \proglang{R} client code performs the remote function call 
-\texttt{stats::rnorm(n=42, mean=100)}. The function arguments (in this
-case \texttt{n} and \texttt{mean}) as well as the return value (a vector
+\code{stats::rnorm(n=42, mean=100)}. The function arguments (in this
+case \code{n} and \code{mean}) as well as the return value (a vector
 with 42 random numbers) are transferred using a protobuf message. RPC in
-OpenCPU works like the \texttt{do.call} function in \proglang{R}, hence all arguments
+OpenCPU works like the \code{do.call} function in \proglang{R}, hence all arguments
 are contained within a list.
@@
 strengths and weaknesses.  
-Text based formats such as \texttt{CSV} and \texttt{JSON} are easy to use, and will likely 
+Text based formats such as \code{CSV} and \code{JSON} are easy to use, and will likely 
 remain popular among statisticians for many years to come. However, in the 
@@
 
-Below a print of the \texttt{rexp.proto} schema (originally designed by \cite{rhipe})
-that is included with the \pkg{RProtoBuf} package and used by \texttt{serialize\_pb} and
-\texttt{unserialize\_pb}.
+Below a print of the \code{rexp.proto} schema (originally designed by \cite{rhipe})
+that is included with the \pkg{RProtoBuf} package and used by \code{serialize\_pb} and
+\code{unserialize\_pb}.
 

### 2014-03-24 papers/jss/article.Rnw
Replace sweave code with manual CodeInput/CodeOutput for now so this is buildable again with protobuf-2.5 library which doesn't like our unittest.proto.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@880 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<echo=FALSE,print=FALSE>>=
+<<echo=FALSE,print=FALSE,eval=FALSE>>=
 if (!exists("protobuf_unittest.TestAllTypes",
@@
 
-<<>>=
-test <- new(protobuf_unittest.TestAllTypes)
-test$repeated_int64 <- c(2^53, 2^53+1)
-length(unique(test$repeated_int64))
-@
+% We want a cleaner error message here.
+\begin{CodeChunk}
+\begin{CodeInput}
+R> test <- new(protobuf_unittest.TestAllTypes)
+R> test$repeated_int64 <- c(2^53, 2^53+1)
+R> length(unique(test$repeated_int64))
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 1
+\end{CodeOutput}
+\end{CodeChunk}
 
@@
 
-<<>>=
-test$repeated_int64 <- c("9007199254740992", "9007199254740993")
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> test$repeated_int64 <- c("9007199254740992", "9007199254740993")
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-<<>>=
-options("RProtoBuf.int64AsString" = FALSE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-options("RProtoBuf.int64AsString" = TRUE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> options("RProtoBuf.int64AsString" = FALSE)
+R> test$repeated_int64
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 9.007e+15 9.007e+15
+\end{CodeOutput}
+\begin{CodeInput}
+R> length(unique(test$repeated_int64))
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 1
+\end{CodeOutput}
+\begin{CodeInput}
+R> options("RProtoBuf.int64AsString" = TRUE)
+R> test$repeated_int64
+\end{CodeInput}
+\begin{CodeOutput}
+[1] "9007199254740992" "9007199254740993"
+\end{CodeOutput}
+\begin{CodeInput}
+R> length(unique(test$repeated_int64))
+\end{CodeInput}
+\begin{CodeOutput}
+[1] 2
+\end{CodeOutput}
+\end{CodeChunk}
 

### 2014-03-24 papers/jss/article.Rnw
Better fix.  Add a new proto with only the examples we need for this paper rather than pulling in all fo the complexity of TestAllTypes or faking it with CodeInput/CodeOutput.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@881 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-03-23 papers/jss/article.Rnw
@@
 <<echo=FALSE,print=FALSE>>=
-if (!exists("protobuf_unittest.TestAllTypes",
-            "RProtoBuf:DescriptorPool")) {
-    unittest.proto.file <- system.file("unitTests", "data",
-                                       "unittest.proto",
-                                       package="RProtoBuf")
-    readProtoFiles(file=unittest.proto.file)
+if (!exists("JSSPaper.Example1", "RProtoBuf:DescriptorPool")) {
+    readProtoFiles(file="int64.proto")
 }
@@
 \begin{CodeInput}
-R> a <- new(protobuf_unittest.TestAllTypes)
+R> a <- new(JSSPaper.Example1)
 R> a$optional_bool <- TRUE
@@
 <<>>=
-test <- new(protobuf_unittest.TestAllTypes)
+test <- new(JSSPaper.Example1)
 test$repeated_int64 <- c(2^53, 2^53+1)

### 2014-03-25 papers/jss/article.Rnw
Ditch the non-portable int64 examples with a description of what the option does.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@885 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-If we try to set an int64 field in \proglang{R} to double values, we lose
-precision:
-
-<<>>=
-test <- new(JSSPaper.Example1)
-test$repeated_int64 <- c(2^53, 2^53+1)
-length(unique(test$repeated_int64))
-@
-
-But when the values are specified as character strings, \pkg{RProtoBuf}
-will automatically coerce them into a true 64-bit integer types 
-before storing them in the Protocol Buffer message:
-
-<<>>=
-test$repeated_int64 <- c("9007199254740992", "9007199254740993")
-@
-
-When reading the value back into \proglang{R}, numeric types are returned by
-default, but when the full precision is required a character value
-will be returned if the \code{RProtoBuf.int64AsString} option is set
-to \code{TRUE}.  The character values are useful because they can
+On 64-bit platforms, character strings representing large decimal
+numbers will be coerced to int64 during assignment to 64-bit protocol
+buffer types to work around the lack of native 64-bit types in \proglang{R}.  The
+values are stored as distinct int64 values in memory but when accessed
+from \proglang{R} language code they will be coerced into numeric values.  If the
+full 64-bit precision is required, the \code{RProtoBuf.int64AsString}
+option can be set to \code{TRUE} to return int64 values from messages as character
+strings.  The character values are useful because they can
 accurately be used as unique identifiers and can easily be passed to \proglang{R}
@@
 
-<<>>=
-options("RProtoBuf.int64AsString" = FALSE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-options("RProtoBuf.int64AsString" = TRUE)
-test$repeated_int64
-length(unique(test$repeated_int64))
-@
-
-<<echo=FALSE,print=FALSE>>=
-options("RProtoBuf.int64AsString" = FALSE)
-@
+% test <- new(JSSPaper.Example1)
+% test$repeated_int64 <- c(2^53, 2^53+1)
+% length(unique(test$repeated_int64))
+
+% test$repeated_int64 <- c("9007199254740992", "9007199254740993")
+
+% options("RProtoBuf.int64AsString" = FALSE)
+% test$repeated_int64
+% length(unique(test$repeated_int64))
+% options("RProtoBuf.int64AsString" = TRUE)
+% test$repeated_int64
+% length(unique(test$repeated_int64))
+
+% options("RProtoBuf.int64AsString" = FALSE)
 

### 2014-03-25 papers/jss/article.Rnw
minor pass over Murray's excellent edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@886 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \proglang{R} also does not support the native 64-bit integer type. Numeric vectors
-with values $\geq 2^{31}$ can only be stored as doubles, which have
-limited precision. Thereby \proglang{R} loses the ability to distinguish some
-distinct integers:
+with integer values greater or equal to $2^{31}$ can only be stored as
+floating-point double precision variables. This conversion incurs a loss of
+precision, and \proglang{R} loses the ability to distinguish between some
+distinct integer variables:
 
@@
 
-However, most modern languages do have support for 64-bit integers, 
+Most modern languages do have support for 64-bit integer values, 
 which becomes problematic when \pkg{RProtoBuf} is used to exchange data 
@@
 On 64-bit platforms, character strings representing large decimal
-numbers will be coerced to int64 during assignment to 64-bit protocol
-buffer types to work around the lack of native 64-bit types in \proglang{R}.  The
-values are stored as distinct int64 values in memory but when accessed
-from \proglang{R} language code they will be coerced into numeric values.  If the
+numbers will be coerced to \code{int64} during assignment to 64-bit Protocol
+Buffer types to work around the lack of native 64-bit types in \proglang{R} itself.  The
+values are stored as distinct \code{int64} values in memory. But when accessed
+from \proglang{R} language code, they will be coerced into numeric
+(floating-point) values.  If the
 full 64-bit precision is required, the \code{RProtoBuf.int64AsString}
-option can be set to \code{TRUE} to return int64 values from messages as character
-strings.  The character values are useful because they can
-accurately be used as unique identifiers and can easily be passed to \proglang{R}
+option can be set to \code{TRUE} to return \code{int64} values from messages as character
+strings.  Such character values are useful because they can
+accurately be used as unique identifiers, and can easily be passed to \proglang{R}
 packages such as \pkg{int64} \citep{int64} or \pkg{bit64}

### 2014-09-25 papers/jss/article.Rnw
Address some reviewer feedback about why we use RObjectTables and some of the pitfalls, with some of this text provided verbatim from Karl Millar, who is already thanked in the acknowledgements section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@897 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 To create or parse a Protocol Buffer Message, one must first read in 
-the message type specification from a \code{.proto} file. The 
-\code{.proto} files are imported using the \code{readProtoFiles}
+the message type specification from a \code{.proto} file. 
+A small number of message types are imported when the package is first
+loaded, including the \code{tutorial.Person} type we saw in the last
+section.
+All other types must be imported from
+\code{.proto} files using the \code{readProtoFiles}
 function, which can either import a single file, all files in a directory,
@@
 dynamically queries the in-memory database of loaded descriptors
-during normal variable lookup.
+during normal variable lookup.  This allows new descriptors to be
+parsed from \code{.proto} files and added to the global
+namespace.\footnote{Note that there is a significant performance
+  overhead with this RObjectTable implementation.  Because the table
+  is on the search path and isn't cacheable, lookups of symbols that
+  are behind it in the search path can't be added to the global object
+  cache, and R must perform an expensive lookup through all of the
+  attached environments and the protocol buffer definitions to find common
+  symbols (most notably those in base) from the global environment.
+  Fortunately, proper use of namespaces and package imports reduces
+  the impact of this for code in packages.}
 

### 2014-11-15 papers/jss/article.Rnw
Make section 3 a little more concise and address referee feedback by not having so many small subsections, only using one style and sticking to it in this section, and not talking about ill defined 'pseudo-methods'.
Sorry I didn't build this yet about to go home wanted to easily
capture some edits on work computer.  hopefully it doesn't break the
build. will be easy to fix if i did.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@898 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \code{.proto} files}
+% Note: We comment out subsections in favor of textbf blocks to save
+% space and shrink down this section a little bit.
+%\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \code{.proto} files}
+
+\textbf{Importing message descriptors from \code{.proto} files}
 
@@
 
-<<>>=
+% Commented out for now because its too detailed.  Lets shorten
+% section 3 per referee feedback.
+
+<<echo=FALSE,print=FALSE>>=
 ls("RProtoBuf:DescriptorPool")
@@
 
-\subsection{Creating a message}
+% \subsection{Creating a message}
+\textbf{Creating, accessing, and modifying a message.}
 
@@
 <<>>=
-p1 <- new(tutorial.Person)
 p <- new(tutorial.Person, name = "Murray", id = 1)
@@
 
-\subsection{Access and modify fields of a message}
+%\subsection{Access and modify fields of a message}
 
@@
 
-\subsection{Display messages}
+%\subsection{Display messages}
+\textbf{Printing, Reading, and Writing Messages}
 
@@
 
-\subsection{Serializing messages}
+% \subsection{Serializing messages}
 
@@
 
-\code{serialize} can also be called in a more traditional
-object oriented fashion using the dollar operator.
-
-<<>>=
-p$serialize(tf1)
-con <- file(tf2, open = "wb")
-p$serialize(con)
-close(con)
-@
-
-Here, we first serialize to a file \code{tf1} before we serialize to a binary
-connection to file \code{tf2}.
-
-\subsection{Parsing messages}
+% TODO(mstokely): commentd out per referee feedback, but see if this is
+% covered in the package documentation well.
+%
+%\code{serialize} can also be called in a more traditional
+%object oriented fashion using the dollar operator.
+%
+%<<>>=
+%p$serialize(tf1)
+%con <- file(tf2, open = "wb")
+%p$serialize(con)
+%close(con)
+%@
+%
+%Here, we first serialize to a file \code{tf1} before we serialize to a binary
+%connection to file \code{tf2}.
+
+%\subsection{Parsing messages}
 
@@
 
-\code{read} can also be used as a pseudo-method of the descriptor
-object:
-
-<<>>=
-message <- tutorial.Person$read(tf1)
-con <- file(tf2, open = "rb")
-message <- tutorial.Person$read(con)
-close(con)
-message <- tutorial.Person$read(payload)
-@
-
-Here we read first from a file, then from a binary connection and lastly from
-a message payload.
+% TODO(mstokely): comment out and use only one style, not both per
+% referee feedback.  Also avoid using the term 'pseudo-method' which
+% is unclear.
+%
+%\code{read} can also be used as a pseudo-method of the descriptor
+%object:
+%
+%<<>>=
+%message <- tutorial.Person$read(tf1)
+%con <- file(tf2, open = "rb")
+%message <- tutorial.Person$read(con)
+%close(con)
+%message <- tutorial.Person$read(payload)
+%@
+%
+%Here we read first from a file, then from a binary connection and lastly from
+%a message payload.
 

### 2014-11-15 papers/jss/article.Rnw
Comment out additional redundant information to make section 3 a more concise introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@899 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<echo=FALSE,print=FALSE>>=
-ls("RProtoBuf:DescriptorPool")
-@
+%<<echo=FALSE,print=FALSE>>=
+%ls("RProtoBuf:DescriptorPool")
+%@
 
 % \subsection{Creating a message}
-\textbf{Creating, accessing, and modifying a message.}
+
+\\
+
+\textbf{Creating, accessing, and modifying messages.}
 
@@
 %\subsection{Display messages}
+
+\\
+
 \textbf{Printing, Reading, and Writing Messages}
@@
 
-For additional information, such as for debugging purposes,
-the \code{as.character} method provides a more complete ASCII
+%For additional information, such as for debugging purposes,
+The \code{as.character} method provides a more complete ASCII
 representation of the contents of a message.
@@
 
-The same method can be used to serialize messages to files:
+The same method can be used to serialize messages to files or arbitrary binary connections:
 
@@
 
-Or to arbitrary binary connections:
-
-<<>>=
-tf2 <- tempfile()
-con <- file(tf2, open = "wb")
-serialize(p, con)
-close(con)
-readBin(tf2, raw(0), 500)
-@
+% TODO(mstokely): Comment out, combined with last statement. make this
+% shorter, more succinct summary of the key features of RProtoBuf.
+
+%Or to arbitrary binary connections:
+%
+%<<>>=
+%tf2 <- tempfile()
+%con <- file(tf2, open = "wb")
+%serialize(p, con)
+%close(con)
+%readBin(tf2, raw(0), 500)
+%@
 
@@
 The \code{input} argument of \code{read} can also be a binary
-readable \proglang{R} connection, such as a binary file connection:
-
-<<>>=
-con <- file(tf2, open = "rb")
-message <- read(tutorial.Person, con)
-close(con)
-writeLines(as.character(message))
-@
-
-Finally, the raw vector payload of the message can be used:
-
-<<>>=
-payload <- readBin(tf1, raw(0), 5000)
-message <- read(tutorial.Person, payload)
-@
+readable \proglang{R} connection, such as a binary file connection, or a raw vector of serialized bytes.
+
+% <<>>=
+% con <- file(tf2, open = "rb")
+% message <- read(tutorial.Person, con)
+% close(con)
+% writeLines(as.character(message))
+% @
+
+% Finally, the raw vector payload of the message can be used:
+%
+%<<>>=
+%payload <- readBin(tf1, raw(0), 5000)
+%message <- read(tutorial.Person, payload)
+%@
 

### 2014-11-19 papers/jss/article.Rnw
Address referee #1 feedback about section 4.
Answer many of the specific questions from the referee: Why use S4 and
not RC (answer: written before RC availalbe, acknowledge that RC would
be better). objects made mutable by usual functional copy on modify
semantics in mutation methods, avoid pseudo-method.

Fix typographical issue with subheadings.  Make it more concise with
\subsection* and combine a few.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@900 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-\textbf{Importing message descriptors from \code{.proto} files}
+\subsection*{Importing message descriptors from \code{.proto} files}
 
@@
 
-\\
-
-\textbf{Creating, accessing, and modifying messages.}
+% \\
+
+\subsection*{Creating, accessing, and modifying messages.}
 
@@
 
-%\subsection{Access and modify fields of a message}
+% \subsection*{Access and modify fields of a message}
 
@@
 
-%\subsection{Display messages}
-
-\\
-
-\textbf{Printing, Reading, and Writing Messages}
+\subsection*{Printing, reading, and writing Messages}
+
+%\\
+
+% \textbf{Printing, Reading, and Writing Messages}
 
@@
 %
-%\code{read} can also be used as a pseudo-method of the descriptor
+%\code{read} can also be used as a method of the descriptor
 %object:
@@
 
-\section{Under the hood: S4 classes, methods, and pseudo methods}
+\section{Under the hood: S4 classes and methods}
 \label{sec:rprotobuf-classes}
@@
 The \pkg{RProtoBuf} package uses the S4 system to store
-information about descriptors and messages.  Using the S4 system
-allows the package to dispatch methods that are not
-generic in the S3 sense, such as \code{new} and
-\code{serialize}.
-Table~\ref{class-summary-table} lists the six
-primary Message and Descriptor classes in \pkg{RProtoBuf}.  Each \proglang{R} object
+information about descriptors and messages.
+Each \proglang{R} object
 contains an external pointer to an object managed by the
@@
 classes.
+S4 objects are immutable, and so the methods that modify field values of a message return a new copy of the object with R's usual functional copy on modify semantics\footnote{RProtoBuf was designed and implemented before Reference Classes were introduced to offer a new class system with mutable objects.
+If RProtoBuf were
+implemented today Reference Classes would almost certainly be a better
+design choice than S4 classes.}.
+Using the S4 system
+allows the package to dispatch methods that are not
+generic in the S3 sense, such as \code{new} and
+\code{serialize}.
+
+The \pkg{Rcpp} package
+\citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
+facilitate this integration of the \proglang{R} and \proglang{C++} code for these objects.
+Each method is wrapped individually which allows us to add 
+user-friendly custom error handling, type coercion, and performance
+improvements at the cost of a more verbose implementation.
+The \pkg{RProtoBuf} package in many ways motivated
+the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
+which provide a more concise way of wrapping \proglang{C++} functions and classes
+in a single entity.
+
+Since \pkg{RProtoBuf} users are most often switching between two or
+more different languages as part of a larger data analysis pipeline,
+both generic function and message passing OO style calling conventions
+are supported:
+
+\begin{itemize}
+\item The functional dispatch mechanism of the the form
+  \verb|method(object, arguments)| (common to \proglang{R}), and
+\item The traditional object oriented notation
+  \verb|object$method(arguments)|.
+\end{itemize}
+
+Additionally, \pkg{RProtoBuf} supports tab completion for all
+classes.  Completion possibilities include method names for all
+classes, plus \emph{dynamic dispatch} on names or types specific to a given
+object.  This functionality is implemented with the
+\code{.DollarNames} S3 generic function defined in the \pkg{utils}
+package that is included with \proglang{R} \citep{r}.
+
+Table~\ref{class-summary-table} lists the six primary Message and
+Descriptor classes in \pkg{RProtoBuf}.
+% Please see the package
+%documentation for a complete description of the slots and methods for
+%each class.
 
@@
 \end{table}
-
-The \pkg{Rcpp} package
-\citep{eddelbuettel2011rcpp,eddelbuettel2013seamless} is used to 
-facilitate this integration of the \proglang{R} and \proglang{C++} code for these objects.
-Each method is wrapped individually which allows us to add 
-user-friendly custom error handling, type coercion, and performance
-improvements at the cost of a more verbose implementation.
-The \pkg{RProtoBuf} package in many ways motivated
-the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
-which provide a more concise way of wrapping \proglang{C++} functions and classes
-in a single entity.
-
-The \pkg{RProtoBuf} package supports two forms for calling
-functions with these S4 classes:
-\begin{itemize}
-\item The functional dispatch mechanism of the the form
-  \verb|method(object, arguments)| (common to \proglang{R}), and
-\item The traditional object oriented notation
-  \verb|object$method(arguments)|.
-\end{itemize}
-
-Additionally, \pkg{RProtoBuf} supports tab completion for all
-classes.  Completion possibilities include pseudo-method names for all
-classes, plus \emph{dynamic dispatch} on names or types specific to a given
-object.  This functionality is implemented with the
-\code{.DollarNames} S3 generic function defined in the \pkg{utils}
-package that is included with \proglang{R} \citep{r}.
 
@@
 used to retrieve descriptors that are contained in the descriptor, or
-invoke pseudo-methods.
+invoke methods.
 
@@
 constants contained in the EnumDescriptor, or to invoke
-pseudo-methods.
+methods.
 
@@
 
-The \verb|$| operator can be used to invoke pseudo-methods.
+The \verb|$| operator can be used to invoke methods.
 
@@
 The \verb|$| operator can be used to retrieve named fields defined in
-the FileDescriptor, or to invoke pseudo-methods.
+the FileDescriptor, or to invoke methods.
 

### 2014-11-19 papers/jss/article.Rnw
Remove tables 3-8 with full slot and method descriptions for each class in RProtoBuf.  This belongs in the documentation, not the paper. Refer the reader to the full docs once at the beginning of this section and remove all the references to tables in each subsection. Further minor edits are needed here.  This was suggested by referee #1.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@901 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Descriptor classes in \pkg{RProtoBuf}.
-% Please see the package
-%documentation for a complete description of the slots and methods for
-%each class.
-
-\begin{table}[bp]
+Please see the package
+documentation for a complete description of the slots and methods for
+each class.
+
+% [bp]
+\begin{table}
 \centering
@@
 of the data defined in the Message, as well as a number of
-\code{FieldDescriptors} for the individual fields of the message.  A
-complete list of the slots and methods for \code{Messages}
-is available in Table~\ref{Message-methods-table}.
+\code{FieldDescriptors} for the individual fields of the message.
 
@@
 @
-
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & External pointer to the \code{Message} object of the \proglang{C++} protobuf library. Documentation for the
-\code{Message} class is available from the Protocol Buffer project page. \\
-%(\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.message.html#Message}) \\
-\code{type} & Fully qualified name of the message. For example a \code{Person} message
-has its \code{type} slot set to \code{tutorial.Person} \\[.3cm]
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{has} & Indicates if a message has a given field.   \\
-\code{clone} & Creates a clone of the message \\
-\code{isInitialized} & Indicates if a message has all its required fields set\\
-\code{serialize} & serialize a message to a file, binary connection, or raw vector\\
-\code{clear} & Clear one or several fields of a message, or the entire message\\
-\code{size} & The number of elements in a message field\\
-\code{bytesize} & The number of bytes the message would take once serialized\\[3mm]
-%
-\code{swap} & swap elements of a repeated field of a message\\
-\code{set} & set elements of a repeated field\\
-\code{fetch} & fetch elements of a repeated field\\
-\code{setExtension} & set an extension of a message\\
-\code{getExtension} & get the value of an extension of a message\\
-\code{add} & add elements to a repeated field \\[3mm]
-%
-\code{str} & the \proglang{R} structure of the message\\
-\code{as.character} & character representation of a message\\
-\code{toString} & character representation of a message (same as \code{as.character}) \\
-\code{as.list} & converts message to a named \proglang{R} list\\
-\code{update} & updates several fields of a message at once\\
-\code{descriptor} & get the descriptor of the message type of this message\\
-\code{fileDescriptor} & get the file descriptor of this message's descriptor\\
-\hline
-\end{tabular}
-\end{small}
-\caption{\label{Message-methods-table}Description of slots and methods for the \code{Message} S4 class.}
-\end{table}
 
@@
 
-Table~\ref{Descriptor-methods-table} provides a complete list of the
-slots and available methods for Descriptors.
-
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & External pointer to the \code{Descriptor} object of the \proglang{C++} proto library. Documentation for the
-\code{Descriptor} class is available from the Protocol Buffer project page.\\
-%\url{http://code.google.com/apis/protocolbuffers/docs/reference/cpp/google.protobuf.descriptor.html#Descriptor} \\
-\code{type} & Fully qualified path of the message type. \\[.3cm]
-%
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{new} & Creates a prototype of a message described by this descriptor.\\
-\code{read} & Reads a message from a file or binary connection.\\
-\code{readASCII} & Read a message in ASCII format from a file or
-text connection.\\
-\code{name} & Retrieve the name of the message type associated with
-this descriptor.\\
-\code{as.character} & character representation of a descriptor\\
-\code{toString} & character representation of a descriptor (same as \code{as.character}) \\
-\code{as.list} & return a named
-list of the field, enum, and nested descriptors included in this descriptor.\\
-\code{asMessage} & return DescriptorProto message. \\
-\code{fileDescriptor} & Retrieve the file descriptor of this
-descriptor.\\
-\code{containing\_type} & Retrieve the descriptor describing the message type containing this descriptor.\\
-\code{field\_count} & Return the number of fields in this descriptor.\\
-\code{field} & Return the descriptor for the specified field in this descriptor.\\
-\code{nested\_type\_count} & The number of nested types in this descriptor.\\
-\code{nested\_type} & Return the descriptor for the specified nested 
-type in this descriptor.\\
-\code{enum\_type\_count} & The number of enum types in this descriptor.\\
-\code{enum\_type} & Return the descriptor for the specified enum
-type in this descriptor.\\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{Descriptor-methods-table}Description of slots and methods for the \code{Descriptor} S4 class.}
-\end{table}
-
 \subsection{Field descriptors}
@@
 \code{google::protobuf::FieldDescriptor} \proglang{C++} class.
-Table~\ref{fielddescriptor-methods-table} describes the methods
-defined for the \code{FieldDescriptor} class.
-
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & External pointer to the \code{FieldDescriptor} \proglang{C++} variable \\
-\code{name} & Simple name of the field \\
-\code{full\_name} & Fully qualified name of the field \\
-\code{type} & Name of the message type where the field is declared \\[.3cm]
-%
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{as.character} & Character representation of a descriptor\\
-\code{toString} & Character representation of a descriptor (same as \code{as.character}) \\
-\code{asMessage} & Return FieldDescriptorProto message. \\
-\code{name} & Return the name of the field descriptor.\\
-\code{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
-\code{containing\_type} & Return the containing descriptor of this field.\\
-\code{is\_extension} & Return TRUE if this field is an extension.\\
-\code{number} & Gets the declared tag number of the field.\\
-\code{type} & Gets the type of the field.\\
-\code{cpp\_type} & Gets the \proglang{C++} type of the field.\\
-\code{label} & Gets the label of a field (optional, required, or repeated).\\
-\code{is\_repeated} & Return TRUE if this field is repeated.\\
-\code{is\_required} & Return TRUE if this field is required.\\
-\code{is\_optional} & Return TRUE if this field is optional.\\
-\code{has\_default\_value} & Return TRUE if this field has a default value.\\
-\code{default\_value} & Return the default value.\\
-\code{message\_type} & Return the message type if this is a message type field.\\
-\code{enum\_type} & Return the enum type if this is an enum type field.\\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{fielddescriptor-methods-table}Description of slots and
-  methods for the \code{FieldDescriptor} S4 class.}
-\end{table}
+
+<<>>=
+tutorial.Person$email
+tutorial.Person$email$is_required()
+tutorial.Person$email$type()
+tutorial.Person$email$as.character()
+@
 
@@
 \code{google::protobuf::EnumDescriptor} \proglang{C++} class.
-Table~\ref{enumdescriptor-methods-table} describes the methods
-defined for the \code{EnumDescriptor} class.
 
@@
 
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & External pointer to the \code{EnumDescriptor} \proglang{C++} variable \\
-\code{name} & Simple name of the enum \\
-\code{full\_name} & Fully qualified name of the enum \\
-\code{type} & Name of the message type where the enum is declared \\[.3cm]
-%
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{as.list} & return a named
-integer vector with the values of the enum and their names.\\
-\code{as.character} & character representation of a descriptor\\
-\code{toString} & character
-representation of a descriptor (same as \code{as.character}) \\
-\code{asMessage} & return EnumDescriptorProto message. \\
-\code{name} & Return the name of the enum descriptor.\\
-\code{fileDescriptor} & Return the fileDescriptor where this field is defined.\\
-\code{containing\_type} & Return the containing descriptor of this field.\\
-\code{length} & Return the number of constants in this enum.\\
-\code{has} & Return TRUE if this enum contains the specified named constant string.\\
-\code{value\_count} & Return the number of constants in this enum (same as \code{length}).\\
-\code{value} & Return the EnumValueDescriptor of an enum value of specified index, name, or number.\\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{enumdescriptor-methods-table}Description of slots and methods
-  for the \code{EnumDescriptor} S4 class.} 
-\end{table}
-
 \subsection{Enum value descriptors}
@@
 \code{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
-Table~\ref{EnumValueDescriptor-methods-table} describes the methods
-defined for the \code{EnumValueDescriptor} class.
 
@@
 
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & External pointer to the \code{EnumValueDescriptor} \proglang{C++} variable \\
-\code{name} & simple name of the enum value \\
-\code{full\_name} & fully qualified name of the enum value \\[.3cm]
-%
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{number} & return the number of this EnumValueDescriptor. \\
-\code{name} & Return the name of the enum value descriptor.\\
-\code{enum\_type} & return the EnumDescriptor type of this EnumValueDescriptor. \\
-\code{as.character} & character representation of a descriptor. \\
-\code{toString} & character representation of a descriptor (same as \code{as.character}). \\
-\code{asMessage} & return EnumValueDescriptorProto message. \\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{EnumValueDescriptor-methods-table}Description of slots
-  and methods for the \code{EnumValueDescriptor} S4 class.}
-\end{table}
-
 \subsection{File descriptors}
 \label{subsec-file-descriptor}
-
-\begin{table}[tbp]
-\centering
-\begin{small}
-\begin{tabular}{lp{10cm}}
-\toprule
-Slot & Description \\
-\cmidrule(r){2-2}
-\code{pointer} & external pointer to the \code{FileDescriptor} object of the \proglang{C++} proto library. Documentation for the
-\code{FileDescriptor} class is available from the Protocol Buffer project page:
-\url{http://developers.google.com/protocol-buffers/docs/reference/cpp/google.protobuf.descriptor.html#FileDescriptor} \\
-\code{filename} & fully qualified pathname of the \code{.proto} file.\\
-\code{package} & package name defined in this \code{.proto} file.\\[.3cm]
-
-Method & Description \\
-\cmidrule(r){2-2}
-\code{name} & Return the filename for this FileDescriptorProto.\\
-\code{package} & Return the file-level package name specified in this FileDescriptorProto.\\
-\code{as.character} & character representation of a descriptor. \\
-\code{toString} & character representation of a descriptor (same as \code{as.character}). \\
-\code{asMessage} & return FileDescriptorProto message. \\
-\code{as.list} & return named list of descriptors defined in this file descriptor.\\
-\bottomrule
-\end{tabular}
-\end{small}
-\caption{\label{filedescriptor-methods-table}Description of slots and methods for the \code{FileDescriptor} S4 class.}
-\end{table}
 
@@
 \code{google::protobuf::FileDescriptor} \proglang{C++} class.
-Table~\ref{filedescriptor-methods-table} describes the methods
-defined for the \code{FileDescriptor} class.
 

### 2014-11-19 papers/jss/article.Rnw
Make section 4 more concise and less manual-like per referee #1 feedback.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@902 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-<<>>=
-tutorial.Person$email
-
-tutorial.Person$PhoneType
-
-tutorial.Person$PhoneNumber
-
-tutorial.Person.PhoneNumber
-@
-
-\subsection{Field descriptors}
+\subsubsection*{Field descriptors}
 \label{subsec-field-descriptor}
-
-The class \emph{FieldDescriptor} represents field
-descriptors in \proglang{R}. This is a wrapper S4 class around the
-\code{google::protobuf::FieldDescriptor} \proglang{C++} class.
 
@@
 tutorial.Person$email$as.character()
-@
-
-\subsection{Enum descriptors}
+class(tutorial.Person$email)
+@
+
+\subsubsection*{Enum and EnumValue descriptors}
 \label{subsec-enum-descriptor}
 
-The class \emph{EnumDescriptor} represents enum descriptors in \proglang{R}.
-This is a wrapper S4 class around the
-\code{google::protobuf::EnumDescriptor} \proglang{C++} class.
-
-The \verb|$| operator can be used to retrieve the value of enum
-constants contained in the EnumDescriptor, or to invoke
-methods.
-
-The \code{EnumDescriptor} contains information about what values this type
-defines, while the \code{EnumValueDescriptor} describes an
-individual enum constant of a particular type.
+The \code{EnumDescriptor} contains information about what values a
+type defines, while the \code{EnumValueDescriptor} describes an
+individual enum constant of a particular type.  The \verb|$| operator
+can be used to retrieve the value of enum constants contained in the
+EnumDescriptor, or to invoke methods.
 
@@
 tutorial.Person$PhoneType$WORK
-@
-
-\subsection{Enum value descriptors}
-\label{subsec-enumvalue-descriptor}
-
-The class \emph{EnumValueDescriptor} represents enumeration value
-descriptors in \proglang{R}.  This is a wrapper S4 class around the
-\code{google::protobuf::EnumValueDescriptor} \proglang{C++} class.
-
-The \verb|$| operator can be used to invoke methods.
-
-<<>>=
+class(tutorial.Person$PhoneType)
 tutorial.Person$PhoneType$value(1)
@@
 tutorial.Person$PhoneType$value(number=1)
-@
-
-\subsection{File descriptors}
+class(tutorial.Person$PhoneType$value(1))
+@
+
+\subsubsection*{File descriptors}
 \label{subsec-file-descriptor}
@@
 The class \emph{FileDescriptor} represents file descriptors in \proglang{R}.
-This is a wrapper S4 class around the
-\code{google::protobuf::FileDescriptor} \proglang{C++} class.
-
 The \verb|$| operator can be used to retrieve named fields defined in

### 2014-11-24 papers/jss/article.Rnw
Remove a duplicate word in the last paragraph of the introduction, add a word to fix a linewrap in section 4, and add a paragraph to the beginning of section 7 to make it more interesting.  More work on sections 6,7, and 8 is sorely needed.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@903 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
-details the implementation details of the main S4 classes and methods.  
+details the implementation of the main S4 classes and methods.  
 Section~\ref{sec:types} describes the challenges of type coercion
@@
 
-The \code{EnumDescriptor} contains information about what values a
+The \code{EnumDescriptor} type contains information about what values a
 type defines, while the \code{EnumValueDescriptor} describes an
@@
 
+Protocol Buffers have been used extensively at Google for almost all
+RPC protocols, and for storing structured information in a variety of
+persistent storage systems since 2000 \citep{dean2009designs}.  The
+\pkg{RProtoBuf} package has been in widespread use by hundreds of
+analysts at Google since 2010.
+
 Many large data sets in fields such as particle physics and information

### 2014-11-24 papers/jss/article.Rnw
Improve section 7 and update the conclusions to note we've been using this package at Google for 5 years, not 3 years.  Done with section 7 updates.  6 and 8 need work.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@904 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \pkg{RProtoBuf} package has been in widespread use by hundreds of
-analysts at Google since 2010.
+statisticians and software engineers at Google since 2010.  This
+section describes a simplified example of a common design pattern of
+collecting a large structured data set in one language for later
+analysis in \proglang{R}.
 
@@
 
-One of the authors has used this design pattern for several
-large-scale studies of distributed storage systems
+One of the authors has used this design pattern with large-scale \proglang{C++}
+MapReduces over very large data sets to write out histogram protocol
+buffers for several large-scale studies of distributed storage systems
 \citep{sciencecloud,janus}.
@@
 Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
-for the past three years by statisticians, analysts, and software engineers.
+for the past five years by statisticians, analysts, and software engineers.
 At the time of this writing there are over 300 active

### 2014-11-25 papers/jss/article.Rnw
Use listings package to add line numbers so we can explain an example in section 7 better per referee #1 feedback.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@905 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage{booktabs}
+\usepackage{listings}
 \usepackage[toc,page]{appendix}
@@
 %\code{serialize} can also be called in a more traditional
-%object oriented fashion using the dollar operator.
+%object-oriented fashion using the dollar operator.
 %
@@
   \verb|method(object, arguments)| (common to \proglang{R}), and
-\item The traditional object oriented notation
+\item The message passing object-oriented notation of the form
   \verb|object$method(arguments)|.
@@
 
-The Protocol Buffer can then be read into \proglang{R} and converted to a native
-\proglang{R} histogram object for plotting. Here, the schema is read first,
-then the (serialized) histogram is read into the variable \code{hist} which
-is then converted a histogram object which is display as a plot.
-
-\begin{Code}
-library("RProtoBuf")
-library("HistogramTools")
-
-readProtoFiles(package="HistogramTools")
-
-hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
-hist
+The Protocol Buffer created from this \proglang{Python} script can then be read into \proglang{R} and converted to a native
+\proglang{R} histogram object for plotting.
+Line~1 in the listing below attaches the \pkg{HistogramTools} package which imports \pkg{RProtoBuf}.
+Line~2 then reads all of the \code{.proto} descriptor definitions provided by \pkg{HistogramTools} and adds them to the environment as described in Section~\ref{sec:rprotobuf-basic}.
+Line~3 parses the serialized protocol buffer using the \code{HistogramTools.HistogramState} schema.
+Line~8 converts the protocol buffer representation of the histogram to a native \proglang{R} histogram object with \code{as.histogram} and passes the result to \code{plot}.
+
+% Here, the schema is read first,
+%then the (serialized) histogram is read into the variable \code{hist} which
+%is then converted a histogram object which is display as a plot.
+
+\lstdefinelanguage{jss}
+  {sensitive=false,
+   morecomment=[l]{R>}}
+
+\lstset{language=jss, basicstyle=\ttfamily, numbers=left, numberstyle=\tiny, stepnumber=2, numbersep=5pt, columns=fullflexible, keepspaces=true, showstringspaces=false, commentstyle=\textsl}
+%\begin{Code}
+\begin{lstlisting}
+R> library("HistogramTools")
+R> readProtoFiles(package="HistogramTools")
+R> hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
+R> hist
+
 [1] "message of type 'HistogramTools.HistogramState' with 3 fields set"
 
-plot(as.histogram(hist))
-\end{Code}
+R> plot(as.histogram(hist))
+\end{lstlisting}
+%\end{Code}
 
@@
 <<echo=FALSE,fig=TRUE,width=8,height=4>>=
-require(RProtoBuf)
 require(HistogramTools)

### 2014-11-25 papers/jss/article.Rnw
three small edits
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@907 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 extensible mechanism for sharing and storing structured data.  Some of
-the key features provided by Protocol Buffers for data analysis include:
+the key features provided by Protocol Buffers for data analysis are:
 
@@
   is on the search path and isn't cacheable, lookups of symbols that
-  are behind it in the search path can't be added to the global object
+  are behind it in the search path cannot be added to the global object
   cache, and R must perform an expensive lookup through all of the
@@
 Descriptor classes in \pkg{RProtoBuf}.
-Please see the package
-documentation for a complete description of the slots and methods for
+The package documentation provides a complete description of the slots and methods for
 each class.

### 2014-11-25 papers/jss/article.Rnw
Improve example in section 7 using some of the specific advantages suggested by referee #2 and point out why we've given the user a simplified example and how it differs from the real MapReduce context where this would be more useful.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@908 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 written in other languages and only the resulting output histograms
-need to be manipulated in \proglang{R}.  For example, to create HistogramState
+need to be manipulated in \proglang{R}.
+
+\subsection*{A trivial single-machine example for Python to R serialization}
+
+To create HistogramState
 messages in Python for later consumption by \proglang{R}, we first compile the 
@@
 
-One of the authors has used this design pattern with large-scale \proglang{C++}
+This simple example uses a constant histogram generated in
+\proglang{Python} to illustrate the serialization concepts without
+requiring the reader to be familiar with the interface of any
+particular MapReduce implementation.  In practice, using Protocol
+Buffers to pass histograms between another programming language and R
+would provide a much greater benefit in a distributed context.
+For example, a first-class data type to represent histograms would
+prevent individual histograms from being split up and would allow the
+use of combiners on Map workers to process large data sets more
+efficiently than simply passing around lists of counts and buckets.
+
+One of the authors has used this design pattern with \proglang{C++}
 MapReduces over very large data sets to write out histogram protocol

### 2014-11-25 papers/jss/article.Rnw
Add acknowledgements for Tim Hesterberg and two anonymous referees.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@909 edb9625f-4e0d-4859-8d74-9fd3b1da38cb
Updates to section 7:
Add a better transition to start section 7 reminding the user what the
application in section 6 was about and how/why this one is different.

Remove several duplicate sentences about the basics of protocol buffer
.proto files and such which are explained earlier in the paper.

Remove a few sentences that provide unnecessary level of detail about
OpenCPU.  In this section it is an example web service and

@@
 
-As described earlier, the primary application of Protocol Buffers is data
-interchange in the context of inter-system communications.  Network protocols
+The previous section described an application where data from a
+program written in another language was output to persistent storage
+and then read into \proglang{R} for further analysis.  This section
+describes another common use case where Protocol Buffers are used as
+the interchange format for client-server communication.
+Network protocols
 such as HTTP provide mechanisms for client-server communication, i.e., how to
@@
 structures into well defined messages, which can then be exchanged using any
-protocol. The descriptors (\code{.proto} files) are used to formally define
-the interface of a remote API or network application. Libraries to parse and
-generate protobuf messages are available for many programming languages,
-making it relatively straightforward to implement clients and servers.
+protocol.
+%The descriptors (\code{.proto} files) are used to formally define
+%the interface of a remote API or network application.
+%Libraries to parse and
+%generate protobuf messages are available for many programming languages,
+%making it relatively straightforward to implement clients and servers.
 
@@
 One example of a system that supports Protocol Buffers to interact
-with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework for embedded statistical 
-computation and reproducible research based on \proglang{R} and \LaTeX. It exposes a 
-HTTP(S) API to access and manipulate \proglang{R} objects and allows for performing 
-remote \proglang{R} function calls. Clients do not need to understand 
-or generate any \proglang{R} code: HTTP requests are automatically mapped to 
-function calls, and arguments/return values can be posted/retrieved
-using several data interchange formats, such as Protocol Buffers.  
-OpenCPU uses the \code{serialize\_pb} and \code{unserialize\_pb} functions
-from the \pkg{RProtoBuf} package to convert between \proglang{R} objects and protobuf
-messages. Therefore, clients need the \code{rexp.proto} descriptor mentioned
-earlier to parse and generate protobuf messages when interacting with OpenCPU.
+with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework
+for embedded statistical computation and reproducible research based
+on \proglang{R} and \LaTeX. It exposes a HTTP(S) API to access and
+manipulate \proglang{R} objects and execute remote \proglang{R}
+function calls. Clients do not need to understand or generate any
+\proglang{R} code: HTTP requests are automatically mapped to function
+calls, and arguments/return values can be posted/retrieved using
+several data interchange formats, such as Protocol Buffers.  OpenCPU
+uses the \code{rexp.proto} descriptor and the \code{serialize\_pb} and
+\code{unserialize\_pb} functions described in
+Section~\ref{sec:evaluation} to convert between \proglang{R} objects
+and protocol buffer messages.
 
@@
 The postfix \code{/pb} in the URL tells the server to send this
-object in the form of a protobuf message. Alternative formats include 
-\code{/json}, \code{/csv}, \code{/rds} and others. If the request
+object in the form of a protocol buffer message.
+% Alternative formats include \code{/json}, \code{/csv}, \code{/rds} and others.
+If the request
 is successful, OpenCPU returns the serialized object with HTTP status
@@
 The latter is the conventional MIME type that formally notifies the client to
-interpret the response as a protobuf message.
+interpret the response as a protocol buffer.
 
@@
 languages, clients can be implemented in just a few lines of code. Below
-is example code for both \proglang{R} and Python that retrieves a data set from \proglang{R} with 
-OpenCPU using a protobuf message. In \proglang{R}, we use the HTTP client from 
+is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a protocol buffer message from OpenCPU.
+In \proglang{R}, we use the HTTP client from
 the \code{httr} package \citep{httr}. In this example we
@@
 
-This code suggests a method for exchanging objects between \proglang{R} servers, however this might as 
-well be done without Protocol Buffers. The main advantage of using an inter-operable format 
-is that we can actually access \proglang{R} objects from within another
-programming language. For example, in a very similar fashion we can retrieve the same
-data set in a Python client. To parse messages in Python, we first compile the 
-\code{rexp.proto} descriptor into a python module using the \code{protoc} compiler:
+Similarly, to retrieve the same data set in a Python client, we first
+compile the \code{rexp.proto} descriptor into a python module
+using the \code{protoc} compiler:
 
@@
 \end{verbatim}
-This generates Python module called \code{rexp\_pb2.py}, containing both the 
-descriptor information as well as methods to read and manipulate the \proglang{R} object 
-message. In the example below we use the HTTP client from the \code{urllib2}
-module.
+
+This generates Python module called \code{rexp\_pb2.py}, containing
+both the descriptor information as well as methods to read and
+manipulate the \proglang{R} object message. We use the
+HTTP client from the \code{urllib2} module in our example to retrieve the
+encoded protocol buffer from the remote server then parse and print it
+from Python.
 
@@
 
-The example above shows how the \code{HTTP GET} method retrieves a 
-resource from OpenCPU, for example an \proglang{R} object. The \code{HTTP POST} 
-method on the other hand is used for calling functions and running scripts, 
-which is the primary purpose of the framework. As before, the \code{/pb} 
-postfix requests to retrieve the output as a protobuf message, in this
-case the function return value. However, OpenCPU allows us to supply the
-arguments of the function call in the form of protobuf messages as well.
-This is a bit more work, because clients needs to both generate messages 
-containing \proglang{R} objects to post to the server, as well as retrieve and parse
-protobuf messages returned by the server. Using Protocol Buffers to post
-function arguments is not required, and for simple (scalar) arguments 
-the standard \code{application/x-www-form-urlencoded} format might be sufficient.
-However, with Protocol Buffers the client can perform function calls with
-more complex arguments such as \proglang{R} vectors or lists. The result is a complete
-RPC system to do arbitrary \proglang{R} function calls from within 
-any programming language.
-
-The following example \proglang{R} client code performs the remote function call 
-\code{stats::rnorm(n=42, mean=100)}. The function arguments (in this
-case \code{n} and \code{mean}) as well as the return value (a vector
-with 42 random numbers) are transferred using a protobuf message. RPC in
-OpenCPU works like the \code{do.call} function in \proglang{R}, hence all arguments
-are contained within a list.
+The previous example used a simple \code{HTTP GET} method to retrieve
+an \proglang{R} object from a remote service (OpenCPU) encoded as a
+protocol buffer.
+In many cases simple \code{HTTP GET} methods are insufficient, and a
+more complete RPC system may need to create compact protocol buffers
+for each request to send to the remote server in addition to parsing
+the response protocol buffers.
+
+The OpenCPU framework allows us to do arbitrary \proglang{R} function
+calls from within any programming language by encoding the arguments
+in the request protocol buffer.  The following example \proglang{R}
+client code performs the remote function call \code{stats::rnorm(n=42,
+mean=100)}. The function arguments (in this case \code{n} and
+\code{mean}) as well as the return value (a vector with 42 random
+numbers) are transferred using protocol buffer messages. RPC in OpenCPU
+works like the \code{do.call} function in \proglang{R}, hence all
+arguments are contained within a list.
 
@@
 Kenton Varda was generous with his time in reviewing code and explaining
-obscure Protocol Buffer semantics.  Karl Millar was very
+obscure Protocol Buffer semantics.  Karl Millar and Tim Hesterberg were very
 helpful in reviewing code and offering suggestions.  Saptarshi Guha's
@@
 methods for turning arbitrary R objects into Protocol Buffers without
-a specialized pre-defined schema.
+a specialized pre-defined schema.  Feedback from two anonymous
+referees greatly improved both the presentation of this paper and the
+package contents.
 

### 2014-11-26 papers/jss/article.Rnw
s/protocol buffers/Protocol Buffers/
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@911 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Section~\ref{sec:evaluation} to convert between \proglang{R} objects
-and protocol buffer messages.
+and Protocol Buffer messages.
 
@@
 The postfix \code{/pb} in the URL tells the server to send this
-object in the form of a protocol buffer message.
+object in the form of a Protocol Buffer message.
 % Alternative formats include \code{/json}, \code{/csv}, \code{/rds} and others.
@@
 The latter is the conventional MIME type that formally notifies the client to
-interpret the response as a protocol buffer.
+interpret the response as a Protocol Buffer.
 
@@
 languages, clients can be implemented in just a few lines of code. Below
-is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a protocol buffer message from OpenCPU.
+is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a Protocol Buffer message from OpenCPU.
 In \proglang{R}, we use the HTTP client from
@@
 HTTP client from the \code{urllib2} module in our example to retrieve the
-encoded protocol buffer from the remote server then parse and print it
+encoded Protocol Buffer from the remote server then parse and print it
 from Python.
@@
 an \proglang{R} object from a remote service (OpenCPU) encoded as a
-protocol buffer.
+Protocol Buffer.
 In many cases simple \code{HTTP GET} methods are insufficient, and a
-more complete RPC system may need to create compact protocol buffers
+more complete RPC system may need to create compact Protocol Buffers
 for each request to send to the remote server in addition to parsing
-the response protocol buffers.
+the response Protocol Buffers.
 
@@
 calls from within any programming language by encoding the arguments
-in the request protocol buffer.  The following example \proglang{R}
+in the request Protocol Buffer.  The following example \proglang{R}
 client code performs the remote function call \code{stats::rnorm(n=42,
@@
 \code{mean}) as well as the return value (a vector with 42 random
-numbers) are transferred using protocol buffer messages. RPC in OpenCPU
+numbers) are transferred using Protocol Buffer messages. RPC in OpenCPU
 works like the \code{do.call} function in \proglang{R}, hence all

### 2014-11-26 papers/jss/article.Rnw
Cut section 6 in half by removing the section explaining the caveats about formulas and types not supported by serialized_pb, and just explain now that we serialize everything, but in one sentence explain the caveat that we fall back to base::serialize for R-specific types like language,function,and environment.  This basically removes the need for 6.1 at all, so remove that section, and move a tiny bit of the top text about which datasets we are using into the top of section 6.2 which explains the compresison performance.
Next step: Replace table with a plot as hadley and one of the referees

@@
 
-The \code{rexp.proto} schema supports all main \proglang{R} storage types holding \emph{data}.
-These include \code{NULL}, \code{list} and vectors of type \code{logical}, 
-\code{character}, \code{double}, \code{integer}, and \code{complex}. In addition,
-every type can contain a named set of attributes, as is the case in \proglang{R}. The \code{rexp.proto}
-schema does not support some of the special \proglang{R} specific storage types, such as \code{function},
-\code{language} or \code{environment}. Such objects have no native equivalent 
-type in Protocol Buffers, and have little meaning outside the context of \proglang{R}.
-When serializing \proglang{R} objects using \code{serialize\_pb}, values or attributes of
-unsupported types are skipped with a warning. If the user really wishes to serialize these 
-objects, they need to be converted into a supported type. For example, the  can use 
-\code{deparse} to convert functions or language objects into strings, or \code{as.list}
-for environments.
-
-\subsection[Evaluation: Converting R data sets]{Evaluation: Converting \proglang{R} data sets}
-
-To illustrate how this method works, we attempt to convert all of the built-in 
-data sets from \proglang{R} into this serialized Protocol Buffer representation.
-
-<<echo=TRUE>>=
+The \code{rexp.proto} schema natively supports all main \proglang{R}
+storage types holding \emph{data}.  These include \code{NULL},
+\code{list} and vectors of type \code{logical}, \code{character},
+\code{double}, \code{integer}, and \code{complex}. In addition, every
+type can contain a named set of attributes, as is the case in
+\proglang{R}. The storage types \code{function}, \code{language}, and
+\code{environment} are specific to \proglang{R} and have no equivalent
+native type in Protocol Buffers.  These three types are supported by
+first serializing with \code{base::serialize} in \proglang{R} and
+then stored in a raw bytes field.
+
+\subsection[Evaluation: Serializing R data sets]{Evaluation: Serializing \proglang{R} data sets}
+\label{sec:compression}
+
+<<echo=FALSE>>=
 datasets <- as.data.frame(data(package="datasets")$results)
@@
 
-There are \Sexpr{n} standard data sets included in the \pkg{datasets}
-package included with \proglang{R}. These data sets include data frames, matrices, time series, tables lists,
-and some more exotic data classes. The \code{can\_serialize\_pb} method is 
-used to determine which of those can fully be converted to the \code{rexp.proto}
-Protocol Buffer representation. This method simply checks if any of the values or
-attributes in an object is of an unsupported type:
-
-<<echo=TRUE>>=
-m <- sum(sapply(datasets$name, function(x) can_serialize_pb(get(x))))
-@
-
-\Sexpr{m} data sets can be converted to Protocol Buffers
-without loss of information (\Sexpr{format(100*m/n,digits=1)}\%). Upon closer
-inspection, all other data sets are objects of class \code{nfnGroupedData}.
-This class represents a special type of data frame that has some additional 
-attributes (such as a \emph{formula} object) used by the \pkg{nlme} package \citep{nlme}.
-Because formulas are \proglang{R} \emph{language} objects, they have little meaning to
-other systems, and are not supported by the \code{rexp.proto} descriptor.
-When \code{serialize\_pb} is used on objects of this class, it will serialize
-the data frame and all attributes, except for the formula.
-
-<<>>=
-attr(CO2, "formula")
-msg <- serialize_pb(CO2, NULL)
-object <- unserialize_pb(msg)
-identical(CO2, object)
-identical(class(CO2), class(object))
-identical(dim(CO2), dim(object))
-attr(object, "formula")
-@
-
-\subsection{Compression performance}
-\label{sec:compression}
-
-This section compares how many bytes are used to store data sets
-using four different methods:
+This section evaluates the effectiveness of serializing arbitrary
+\proglang{R} data structures into Protocol Buffers.  We use the
+\Sexpr{n} standard data sets included in the \pkg{datasets} package
+included with \proglang{R} as our evaluation data. These data sets
+include data frames, matrices, time series, tables, lists, and some
+more exotic data classes.  For each data set, we compare how many
+bytes are used to store the data set using four different methods:
 
@@
 Table~\ref{tab:compression} shows the sizes of 50 sample \proglang{R} data sets as
-returned by object.size() compared to the serialized sizes.
+returned by \code{object.size()} compared to the serialized sizes.
 %The summary compression sizes are listed below, and a full table for a
@@
 package rexp;
-
 message REXP {
@@
     NULLTYPE = 7;
+    LANGUAGE = 8;
+    ENVIRONMENT = 9;
+    FUNCTION = 10;
   }
@@
   }
-
-  required RClass rclass = 1 ; 
+  required RClass rclass = 1;
   repeated double realValue = 2 [packed=true];
@@
   repeated REXP attrValue = 12;
+  optional bytes languageValue = 13;
+  optional bytes environmentValue = 14;
+  optional bytes functionValue = 14;
 }

### 2014-11-27 papers/jss/article.Rnw
Improve section 6 to address referee feedback:
Replace massive full page 50 row table with a more succinct plot of
the relevant data points, and label the outliers of this plot so we
can talk about the interesting cases in the text.

Add a small 3-row table beneath the plot showing the data for the two
outliers plus the aggregate of all datasets.

Explain why protocol buffers are so much more space-efficient for one
dataset, and slightly less efficient for another.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@916 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
                        "gzipped RProtoBuf"=datasets$RProtoBuf.serialize.size.gz,
+		       "ratio.serialized" = datasets$R.serialize.size / datasets$object.size,
+		       "ratio.rprotobuf" = datasets$RProtoBuf.serialize.size / datasets$object.size,
+		       "ratio.serialized.gz" = datasets$R.serialize.size.gz / datasets$object.size,
+		       "ratio.rprotobuf.gz" = datasets$RProtoBuf.serialize.size.gz / datasets$object.size,
+		       "savings.serialized" = 1-(datasets$R.serialize.size / datasets$object.size),
+		       "savings.rprotobuf" = 1-(datasets$RProtoBuf.serialize.size / datasets$object.size),
+		       "savings.serialized.gz" = 1-(datasets$R.serialize.size.gz / datasets$object.size),
+		       "savings.rprotobuf.gz" = 1-(datasets$RProtoBuf.serialize.size.gz / datasets$object.size),
                        check.names=FALSE)
-@
-
-Table~\ref{tab:compression} shows the sizes of 50 sample \proglang{R} data sets as
-returned by \code{object.size()} compared to the serialized sizes.
-%The summary compression sizes are listed below, and a full table for a
-%sample of 50 data sets is included on the next page.
+
+all.df<-data.frame(dataset="TOTAL", object.size=sum(datasets$object.size),
+				    "serialized"=sum(datasets$R.serialize.size),
+                       "gzipped serialized"=sum(datasets$R.serialize.size.gz),
+                       "RProtoBuf"=sum(datasets$RProtoBuf.serialize.size),
+                       "gzipped RProtoBuf"=sum(datasets$RProtoBuf.serialize.size.gz),
+		       "ratio.serialized" = sum(datasets$R.serialize.size) / sum(datasets$object.size),
+		       "ratio.rprotobuf" = sum(datasets$RProtoBuf.serialize.size) / sum(datasets$object.size),
+		       "ratio.serialized.gz" = sum(datasets$R.serialize.size.gz) / sum(datasets$object.size),
+		       "ratio.rprotobuf.gz" = sum(datasets$RProtoBuf.serialize.size.gz) / sum(datasets$object.size),
+		       "savings.serialized" = 1-(sum(datasets$R.serialize.size) / sum(datasets$object.size)),
+		       "savings.rprotobuf" = 1-(sum(datasets$RProtoBuf.serialize.size) / sum(datasets$object.size)),
+		       "savings.serialized.gz" = 1-(sum(datasets$R.serialize.size.gz) / sum(datasets$object.size)),
+		       "savings.rprotobuf.gz" = 1-(sum(datasets$RProtoBuf.serialize.size.gz) / sum(datasets$object.size)),
+                       check.names=FALSE)
+clean.df<-rbind(clean.df, all.df)
+@
+
+Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Uncompressed Size}}{\textrm{Compressed Size}}\right)$ for each of the data sets using each of these four methods.
+The associated table shows the exact data sizes for two outliers and the aggregate of all \Sexpr{n} data sets.
 Note that Protocol Buffer serialization results in slightly
@@
 %serialization format. % \citep{serialization}.
-One takeaway from this table is that the universal \proglang{R} object schema
-included in \pkg{RProtoBuf} does not in general provide
+
+The \code{crimtab} dataset of anthropometry measurements of British
+prisoners \citep{garson1900metric}
+shows the greatest difference in the space savings when
+using Protocol Buffers compared to \proglang{R} native serialization.
+This dataset is a 42x22 table of integers, most equal to 0.  Small
+integer values like this can be very efficiently encoded by the
+\emph{Varint} integer encoding scheme used by Protocol Buffers which
+use a variable number of bytes for each value.
+
+The other extreme is represented by the \code{faithful} dataset of
+waiting time and eruptions of the Old Faithful geyser in Yellowstone
+National Park, Wyoming, USA \citep{azzalini1990look}.  This dataset is
+a data frame with 272 observations of 2 numeric variables.  The
+\proglang{R} native serialization of repeated numeric values is more
+space-efficient, resulting in a slightly smaller object size compared
+to the serialized Protocol Buffer equivalent.
+
+This evaluation shows that the \code{rexp.proto} universal
+\proglang{R} object schema included in \pkg{RProtoBuf} does not in
+general provide
 any significant saving in file size compared to the normal serialization
@@
 
-% latex table generated in \proglang{R} 3.0.2 by xtable 1.7-0 package
-% Fri Dec 27 17:00:03 2013
-\begin{table}[h!]
+\begin{figure}[t!]
 \begin{center}
-  \small
-\scalebox{0.9}{
-\begin{tabular}{lrrrrr}
+<<echo=FALSE,fig=TRUE,width=8,height=4>>=
+plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings")
+points(clean.df$savings.serialized.gz, clean.df$savings.rprotobuf.gz,pch=2, col="blue")
+# grey dotted diagonal
+abline(a=0,b=1, col="grey",lty=3)
+
+# find point furthest off the X axis.
+clean.df$savings.diff <- clean.df$savings.serialized - clean.df$savings.rprotobuf
+clean.df$savings.diff.gz <- clean.df$savings.serialized.gz - clean.df$savings.rprotobuf.gz
+
+# The one to label.
+tmp.df <- clean.df[which(clean.df$savings.diff == min(clean.df$savings.diff)),]
+# This minimum means most to the left of our line, so pos=2 is label to the left
+text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2)
+text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2)
+
+tmp.df <- clean.df[which(clean.df$savings.diff == max(clean.df$savings.diff)),]
+# This minimum means most to the right of the diagonal, so pos=4 is label to the right
+text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4)
+text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4)
+
+#outlier.dfs <- clean.df[c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
+
+legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"))
+
+interesting.df <- clean.df[unique(c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
+                             which(clean.df$savings.diff == max(clean.df$savings.diff)),
+                             which(clean.df$savings.diff.gz == max(clean.df$savings.diff.gz)),
+			     which(clean.df$dataset == "TOTAL"))),c("dataset", "object.size", "serialized", "gzipped serialized", "RProtoBuf", "gzipped RProtoBuf", "savings.serialized", "savings.serialized.gz", "savings.rprotobuf", "savings.rprotobuf.gz")]
+# Print without .00 in xtable
+interesting.df$object.size <- as.integer(interesting.df$object.size)
+@
+
+% latex table generated in R 3.0.2 by xtable 1.7-0 package
+% Wed Nov 26 15:31:30 2014
+%\begin{table}[ht]
+%\begin{center}
+\begin{tabular}{rlrrrrr}
   \toprule
   Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
-  \multicolumn{2}{c}{RProtoBuf Serial.} \\
+  \multicolumn{2}{c}{RProtoBuf Serialization} \\
   & & default & gzipped & default & gzipped \\
   \cmidrule(r){2-6}
-  uspop & 584 & 268 & 172 & 211 & 148 \\
-  Titanic & 1960 & 633 & 257 & 481 & 249 \\
-  volcano & 42656 & 42517 & 5226 & 42476 & 4232 \\
-  euro.cross & 2728 & 1319 & 910 & 1207 & 891 \\
-  attenu & 14568 & 8234 & 2165 & 7771 & 2336 \\
-  ToothGrowth & 2568 & 1486 & 349 & 1239 & 391 \\
-  lynx & 1344 & 1028 & 429 & 971 & 404 \\
-  nottem & 2352 & 2036 & 627 & 1979 & 641 \\
-  sleep & 2752 & 746 & 282 & 483 & 260 \\
-  co2 & 4176 & 3860 & 1473 & 3803 & 1453 \\
-  austres & 1144 & 828 & 439 & 771 & 410 \\
-  ability.cov & 1944 & 716 & 357 & 589 & 341 \\
-  EuStockMarkets & 60664 & 59785 & 21232 & 59674 & 19882 \\
-  treering & 64272 & 63956 & 17647 & 63900 & 17758 \\
-  freeny.x & 1944 & 1445 & 1311 & 1372 & 1289 \\
-  Puromycin & 2088 & 813 & 306 & 620 & 320 \\
-  warpbreaks & 2768 & 1231 & 310 & 811 & 343 \\
-  BOD & 1088 & 334 & 182 & 226 & 168 \\
-  sunspots & 22992 & 22676 & 6482 & 22620 & 6742 \\
-  beaver2 & 4184 & 3423 & 751 & 3468 & 840 \\
-  anscombe & 2424 & 991 & 375 & 884 & 352 \\
-  esoph & 5624 & 3111 & 548 & 2240 & 665 \\
-  PlantGrowth & 1680 & 646 & 303 & 459 & 314 \\
-  infert & 15848 & 14328 & 1172 & 13197 & 1404 \\
-  BJsales & 1632 & 1316 & 496 & 1259 & 465 \\
-  stackloss & 1688 & 917 & 293 & 844 & 283 \\
-  crimtab & 7936 & 4641 & 713 & 1655 & 576 \\
-  LifeCycleSavings & 6048 & 3014 & 1420 & 2825 & 1407 \\
-  Harman74.cor & 9144 & 6056 & 2045 & 5861 & 2070 \\
-  nhtemp & 912 & 596 & 240 & 539 & 223 \\
-  faithful & 5136 & 4543 & 1339 & 4936 & 1776 \\
-  freeny & 5296 & 2465 & 1518 & 2271 & 1507 \\
-  discoveries & 1232 & 916 & 199 & 859 & 180 \\
-  state.x77 & 7168 & 4251 & 1754 & 4068 & 1756 \\
-  pressure & 1096 & 498 & 277 & 427 & 273 \\
-  fdeaths & 1008 & 692 & 291 & 635 & 272 \\
-  euro & 976 & 264 & 186 & 202 & 161 \\
-  LakeHuron & 1216 & 900 & 420 & 843 & 404 \\
-  mtcars & 6736 & 3798 & 1204 & 3633 & 1206 \\
-  precip & 4992 & 1793 & 813 & 1615 & 815 \\
-  state.area & 440 & 422 & 246 & 405 & 235 \\
-  attitude & 3024 & 1990 & 544 & 1920 & 561 \\
-  randu & 10496 & 9794 & 8859 & 10441 & 9558 \\
-  state.name & 3088 & 844 & 408 & 724 & 415 \\
-  airquality & 5496 & 4551 & 1241 & 2874 & 1294 \\
-  airmiles & 624 & 308 & 170 & 251 & 148 \\
-  quakes & 33112 & 32246 & 9898 & 29063 & 11595 \\
-  islands & 3496 & 1232 & 563 & 1098 & 561 \\
-  OrchardSprays & 3600 & 2164 & 445 & 1897 & 483 \\
-  WWWusage & 1232 & 916 & 274 & 859 & 251 \\
-  \bottomrule
-%  Total & 391176 & 327537 & 99161 & 313456 & 100308 \\
-  Relative Size & 100\% & 83.7\% & 25.3\% & 80.1\% & 25.6\%\\
-  \bottomrule
+ crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
+ faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
+   \hline
+ All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
+\hline
 \end{tabular}
-}
-\caption{Serialization sizes for default serialization in \proglang{R} and
-  \pkg{RProtoBuf} for 50 \proglang{R} data sets.}
-\label{tab:compression}
 \end{center}
-\end{table}
+\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package.
+Points to the left of the dotted $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
+(Bottom) Absolute space savings of two outlier datasets and the aggregate performance of all datasets.}
+\label{fig:compression}
+\end{figure}
 
@@
 
-R> plot(as.histogram(hist))
+R> plot(as.histogram(hist), main="")
 \end{lstlisting}
@@
 hist <- HistogramTools.HistogramState$read("hist.pb")
-plot(as.histogram(hist))
+plot(as.histogram(hist), main="")
 @

### 2014-12-01 papers/jss/article.Rnw
Update citations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@919 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   E-mail: \email{jeroen.ooms@stat.ucla.edu}\\
-  URL: \url{http://jeroenooms.github.io}
+  URL: \url{https://jeroenooms.github.io}
 }

### 2014-12-01 papers/jss/article.Rnw
Use shorter ocpu URL
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@920 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \begin{verbatim}
-  GET https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb
+  GET https://demo.ocpu.io/MASS/data/Animals/pb
 \end{verbatim}
@@
 
-req <- GET('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+req <- GET('https://demo.ocpu.io/MASS/data/Animals/pb')
 output <- unserialize_pb(req$content)
@@
 
-req = urllib2.Request('https://public.opencpu.org/ocpu/library/MASS/data/Animals/pb')
+req = urllib2.Request('https://demo.ocpu.io/MASS/data/Animals/pb')
 res = urllib2.urlopen(req)
@@
 req <- POST (
-  url = "https://public.opencpu.org/ocpu/library/stats/R/rnorm/pb",
+  url = "https://demo.ocpu.io/stats/R/rnorm/pb",
   body = payload,

### 2014-12-01 papers/jss/article.Rnw
Rewrite mapreduce introduction.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@921 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Protocol Buffers have been used extensively at Google for almost all
-RPC protocols, and for storing structured information in a variety of
-persistent storage systems since 2000 \citep{dean2009designs}.  The
-\pkg{RProtoBuf} package has been in widespread use by hundreds of
-statisticians and software engineers at Google since 2010.  This
-section describes a simplified example of a common design pattern of
-collecting a large structured data set in one language for later
-analysis in \proglang{R}.
-
-Many large data sets in fields such as particle physics and information
-processing are stored in binned or histogram form in order to reduce
-the data storage requirements \citep{scott2009multivariate}.  In the
-last decade, the MapReduce programming model \citep{dean2008mapreduce}
-has emerged as a popular design pattern that enables the processing of
-very large data sets on large compute clusters.
-
-Many types of data analysis over large data sets may involve very rare
+Protocol Buffers are used extensively at Google for almost all
+RPC protocols, and to store structured information on a variety of
+persistent storage systems \citep{dean2009designs}. Since the 
+initial release in 2010, hundreds of Google's statisticians and
+software engineers use the \pkg{RProtoBuf} package on daily basis 
+to interact with these systems from within \proglang{R}.
+The current section illustrates the power of Protocol Buffers to
+collect and manage large structured data in one language 
+before analyzing it in \proglang{R}. Our example uses MapReduce
+\citep{dean2008mapreduce}, which has emerged in the last
+decade as a popular design pattern to facilitate parallel 
+processing of big data using distributed computing clusters.
+
+Big data sets in fields such as particle physics and information
+processing are often stored in binned (histogram) form in order 
+to reduce storage requirements \citep{scott2009multivariate}. 
+Because analysis over such large data sets may involve very rare
 phenomenon or deal with highly skewed data sets or inflexible
-raw data storage systems from which unbiased sampling is not feasible.
-In such situations, MapReduce and binning may be combined as a
+raw data storage systems, unbiased sampling is often not feasible.
+In these situations, MapReduce and binning may be combined as a
 pre-processing step for a wide range of statistical and scientific
@@
 more reducer tasks which merge the intermediate histograms from the
-mappers.
-
-In the second method, illustrated in
+mappers. In the second method, illustrated in
 Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data

### 2014-12-01 papers/jss/article.Rnw
Make the dotted y=x line in figure 2 dashed with a bigger width to make it more visible.  Suggested by Steve Scott.  Also add some commented out code to add line numbers to make review copies for folks that have offered to do a final review before our resubmit.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@922 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \usepackage[toc,page]{appendix}
+
+% Line numbers for drafts.
+%\usepackage[switch, modulo]{lineno}
+%\linenumbers
 
@@
 # grey dotted diagonal
-abline(a=0,b=1, col="grey",lty=3)
+abline(a=0,b=1, col="grey",lty=2,lwd=3)
 
@@
 \caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package.
-Points to the left of the dotted $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
+Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
 (Bottom) Absolute space savings of two outlier datasets and the aggregate performance of all datasets.}

### 2014-12-01 papers/jss/article.Rnw
Add a missing article to Jeroen's nice rewording of this section.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@923 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 initial release in 2010, hundreds of Google's statisticians and
-software engineers use the \pkg{RProtoBuf} package on daily basis 
+software engineers use the \pkg{RProtoBuf} package on a daily basis
 to interact with these systems from within \proglang{R}.

### 2014-12-02 papers/jss/article.Rnw
Grammatical improvements throughout the paper suggested by Tim Hesterberg.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@925 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 conventions on which characters is used as separator or decimal point vary by
-country.  \emph{Extensible Markup Language} (\code{XML}) is another
+country.  \emph{Extensible Markup Language} (\code{XML}) is a
 well-established and widely-supported format with the ability to define just
@@
 for this complexity with comparatively large and verbose messages, and added
-complexity at the parsing side (which are somewhat mitigated by the
+complexity at the parsing side (these problems are somewhat mitigated by the
 availability of mature libraries and parsers). Because \code{XML} is
@@
 between \proglang{R} and other languages.  Section~\ref{sec:evaluation} introduces a
-general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and evaluates
-it against the serialization capabilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
+general \proglang{R} language schema for serializing arbitrary \proglang{R} objects and compares it to
+the serialization capabilities built directly into \proglang{R}.  Sections~\ref{sec:mapreduce}
 and \ref{sec:opencpu} provide real-world use cases of \pkg{RProtoBuf}
@@
 Protocol Buffers are a modern, language-neutral, platform-neutral,
-extensible mechanism for sharing and storing structured data.  Some of
-the key features provided by Protocol Buffers for data analysis are:
+extensible mechanism for sharing and storing structured data.  Key
+features provided by Protocol Buffers for data analysis include:
 
@@
 Protocol Buffer in \proglang{R} that is then serialized and sent over the network to a
-remote server.  The server would then deserialize the message, act on the
-request, and respond with a new Protocol Buffer over the network. 
+remote server.  The server deserializes the message, acts on the
+request, and responds with a new Protocol Buffer over the network.
 The key difference to, say, a request to an \pkg{Rserve}
@@
 To create or parse a Protocol Buffer Message, one must first read in
-the message type specification from a \code{.proto} file. 
+the message descriptor (\emph{message type}) from a \code{.proto} file.
 A small number of message types are imported when the package is first
@@
 
-One of the primary benefits of Protocol Buffers is the efficient
+A primary benefit of Protocol Buffers is an efficient
 binary wire-format representation.
@@
 point to a bucket width and outputs that bucket as a key and '1' as a
-value.  Reducers then sum up all of the values with the same key and
-output to a data store.
+value.  Reducers count how many times each key occurs and outputs a
+histogram to a data store.
 
@@
 from histogram_pb2 import HistogramState;
-
 hist = HistogramState()
-
 hist.counts.extend([2, 6, 2, 4, 6])
@@
 hist.name="Example Histogram Created in Python"
-
 outfile = open("/tmp/hist.pb", "wb")
@@
 Line~3 parses the serialized protocol buffer using the \code{HistogramTools.HistogramState} schema.
-Line~8 converts the protocol buffer representation of the histogram to a native \proglang{R} histogram object with \code{as.histogram} and passes the result to \code{plot}.
+The last line converts the protocol buffer representation of the histogram to a native \proglang{R} histogram object with \code{as.histogram} and passes the result to \code{plot}.
 
@@
 The previous section described an application where data from a
-program written in another language was output to persistent storage
+program written in another language was saved to persistent storage
 and then read into \proglang{R} for further analysis.  This section
@@
 protocol that prescribes how these structures are to be represented in
-messages (buffers) on the network. Protocol Buffers solve exactly this
+messages (buffers) on the network. Protocol Buffers solve this
 problem by providing a cross-platform method for serializing arbitrary
@@
 from rexp_pb2 import REXP
-
 req = urllib2.Request('https://demo.ocpu.io/MASS/data/Animals/pb')
 res = urllib2.urlopen(req)
-
 msg = REXP()
@@
 other languages. We hope that making Protocol Buffers available to the
-\proglang{R} community will contribute towards better software integration
+\proglang{R} community will contribute to better software integration
 and allow for building even more advanced applications and analysis pipelines 
@@
   optional bytes environmentValue = 14;
-  optional bytes functionValue = 14;
+  optional bytes functionValue = 15;
 }

### 2014-12-03 papers/jss/article.Rnw
Improve the plot and point out 3 outliers now and explain them in the text.  Correct an error in the space savings definition.  Change trivial example to simple example.
Suggestions from: Andy Chu


git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@927 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 
-Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Uncompressed Size}}{\textrm{Compressed Size}}\right)$ for each of the data sets using each of these four methods.
-The associated table shows the exact data sizes for two outliers and the aggregate of all \Sexpr{n} data sets.
+Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Compressed Size}}{\textrm{Uncompressed Size}}\right)$ for each of the data sets using each of these four methods.
+The associated table shows the exact data sizes for some outliers and the aggregate of all \Sexpr{n} data sets.
 Note that Protocol Buffer serialization results in slightly
-smaller byte streams compared to native \proglang{R} serialization in most cases,
-but this difference disappears if the results are compressed with gzip.
+smaller byte streams compared to native \proglang{R} serialization in most cases (red dots),
+but this difference disappears if the results are compressed with gzip (blue triangles).
 %Sizes are comparable but Protocol Buffers provide simple getters and setters
@@
 The \code{crimtab} dataset of anthropometry measurements of British
-prisoners \citep{garson1900metric}
-shows the greatest difference in the space savings when
+prisoners \citep{garson1900metric} and the \code{airquality} dataset of air quality measurements in New York show the
+greatest difference in the space savings when
 using Protocol Buffers compared to \proglang{R} native serialization.
-This dataset is a 42x22 table of integers, most equal to 0.  Small
-integer values like this can be very efficiently encoded by the
+The \code{crimtab} dataset is a 42x22 table of integers, most equal to 0, and the \code{airquality} dataset is a data frame of 154 observations of 1 numeric and 5 integer variables.
+In both data sets, the large number of small
+integer values can be very efficiently encoded by the
 \emph{Varint} integer encoding scheme used by Protocol Buffers which
@@
 
-\begin{figure}[t!]
+\begin{figure}[hbt!]
 \begin{center}
-<<echo=FALSE,fig=TRUE,width=8,height=4>>=
-plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings")
+<<label=SER,echo=FALSE,include=FALSE,fig=TRUE>>=
+old.mar<-par("mar")
+new.mar<-old.mar
+new.mar[3]<-0
+new.mar[4]<-0
+my.cex<-1.3
+par("mar"=new.mar)
+plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings", xlim=c(0,1),ylim=c(0,1),cex.lab=my.cex, cex.axis=my.cex)
 points(clean.df$savings.serialized.gz, clean.df$savings.rprotobuf.gz,pch=2, col="blue")
@@
 # This minimum means most to the left of our line, so pos=2 is label to the left
-text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2)
-text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2)
+text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)
+
+# Some gziped version
+# text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2, cex=my.cex)
+
+# Second one is also an outlier
+tmp.df <- clean.df[which(clean.df$savings.diff == sort(clean.df$savings.diff)[2]),]
+# This minimum means most to the left of our line, so pos=2 is label to the left
+text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)
+#text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=my.cex)
 
@@
 # This minimum means most to the right of the diagonal, so pos=4 is label to the right
-text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4)
-text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4)
+# Only show the gziped one.
+#text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4, cex=my.cex)
+text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4, cex=my.cex)
 
@@
 
-legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"))
+legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"), cex=my.cex)
 
@@
 interesting.df$object.size <- as.integer(interesting.df$object.size)
-@
+par("mar"=old.mar)
+@
+\includegraphics[width=0.45\textwidth]{figures/fig-SER}
 
@@
  crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
+ airquality & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
  faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
@@
 Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
-(Bottom) Absolute space savings of two outlier datasets and the aggregate performance of all datasets.}
+(Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.}
 \label{fig:compression}
@@
 
-\subsection*{A trivial single-machine example for Python to R serialization}
+\subsection*{A simple single-machine example for Python to R serialization}
 

### 2014-12-15 papers/jss/article.Rnw
one "isn't" replaced with "is not"; one sentence reworked
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@930 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 Protocol Buffers are a modern, language-neutral, platform-neutral,
-extensible mechanism for sharing and storing structured data.  Key
-features provided by Protocol Buffers for data analysis include:
+extensible mechanism for sharing and storing structured data. Some of their
+features, particularly in the context of data analysis, are:
 
@@
   overhead with this RObjectTable implementation.  Because the table
-  is on the search path and isn't cacheable, lookups of symbols that
+  is on the search path and is not cacheable, lookups of symbols that
   are behind it in the search path cannot be added to the global object

### 2014-12-15 papers/jss/response-to-reviewers.tex
added referee report; started a point-by-point reply --- which will need a lot of work still
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@931 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\documentclass[10pt]{article}
\usepackage{url}
\usepackage{vmargin}
\setpapersize{USletter}
% left top right bottom -- headheight headsep footheight footskop
\setmarginsrb{1in}{1in}{1in}{0.5in}{0pt}{0mm}{10pt}{0.5in}
\usepackage{charter}

\setlength{\parskip}{1ex plus1ex minus1ex}
\setlength{\parindent}{0pt}

\newcommand{\proglang}[1]{\textsf{#1}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

\newcommand{\pointRaised}[2]{\smallskip %\hrule 
  \textsl{{\fontseries{b}\selectfont #1}: #2}\newline}
\newcommand{\simplePointRaised}[1]{\bigskip \hrule\textsl{#1} }
\newcommand{\reply}[1]{\textbf{Reply}:\ #1 \smallskip } %\hrule \smallskip}

\begin{document}

\author{Dirk Eddelbuettel\\Debian Project \and 
        Murray Stokely\\Google, Inc \and
        Jeroen Ooms\\UCLA}
\title{Submission JSS 1313: \\ Response to Reviewers' Comments}
\maketitle 
\thispagestyle{empty}

Thank you for reviewing our manuscript, and for giving us an opportunity to
rewrite, extend and and tighten both the paper and the underlying package.

\smallskip
We truly appreciate the comments and suggestions. Below, we have regrouped the sets
of comments, and have provided detailed point-by-point replies.
%
We hope that this satisfies the request for changes necessary to proceed with
the publication of the revised and updated manuscript, along with the revised
and updated package (which was recently resubmitted to CRAN as version 0.4.2).

\section*{Response to Reviewer \#1}

\pointRaised{Comment 1}{Overall, I think this is a strong paper. Cross-language communication
  is a challenging problem, and good solutions for R are important to
  establish R as a well-behaved member of a data analysis pipeline. The
  paper is well written, and I recommend that it be accepted subject to
  the suggestions below.}
\reply{Thank you. We are providing a point-by-point reply below.}

\subsubsection*{More big picture, less details}

\pointRaised{Comment 2}{Overall, I think the paper provides too much detail on
  relatively unimportant topics and not enough on the reasoning behind
  important design decisions. I think you could comfortably reduce the paper
  by 5-10 pages, referring the interested reader to the documentation for
  more detail.}
\reply{The paper was rewritten throughout and is now much tighter at just 23 pages.}

\pointRaised{Comment 3}{I'd recommend shrinking section 3 to ~2 pages, and removing the
  subheadings. This section should quickly orient the reader to the
  RProtobuf API so they understand the big picture before learning more
  details in the subsequent sections. I'd recommend picking one OO style
  and sticking to it in this section - two is confusing.}
\reply{We followed this recommendation and reduced section 3 to about 2 1/2 pages.}

\pointRaised{Comment 3}{Section 4 dives into the details without giving a good overview and
  motivation. Why use S4 and not RC? How are the objects made mutable?
  Why do you provide both generic function and message passing OO
  styles? What does \$ do in this context? What the heck is a
  pseudo-method? Spend more time on those big issues rather than
  describing each class in detail. Reduce class descriptions to a
  bulleted list giving a high-level overview, then encourage the reader
  to refer to the documentation for further details. Similarly, Tables
  3-5 belong in the documentation, not in a vignette/paper.}
\reply{Done. TO BE EXPANDED}

\pointRaised{Comment 4}{Section 7 is weak. I think the important message is that RProtobuf is
  being used in practice at large scale for for large data, and is
  useful for communicating between R and Python. How can you make that
  message stronger while avoiding (for the purposes of this paper) the
  relatively unimportant details of the map-reduce setup?}
\reply{TBD}

\subsubsection*{R to/from Protobuf translation}

\pointRaised{Comment 5}{The discussion of R to/from Protobuf could be improved. Table 9 would be
  much simpler if instead of Message, you provided a "vectorised"
  Messages class (this would also make the interface more consistent and
  hence the package easier to use).}
\reply{TBD}

\pointRaised{Comment 6}{Along these lines, I think it would make sense to combine sections 5
  and 6 and discuss translation challenges in both direction
  simultaneously. At the minimum, add the equivalent for Table 9 that
  shows how important R classes are converted to their protobuf
  equivalents.}
\reply{TBD}

\pointRaised{Comment 7}{You should discuss how missing values are handled for strings and
  integers, and why enums are not equivalent to factors. I think you
  could make explicit how coercion of factors, dates, times and matrices
  occurs, and the implications of this on sharing data structures
  between programming languages. For example, how do you share date/time
  data between R and python using RProtoBuf?}
\reply{TBD}

\pointRaised{Comment 8}{Table 10 is dying to be a plot, and a natural companion would be to
  show how long it takes to serialise data frames using both RProtoBuf
  and R's native serialisation. Is there a performance penalty to using
  protobufs?}
\reply{TBD}

\subsubsection*{RObjectTables magic}

\pointRaised{Comment 9}{The use of RObjectTables magic makes me uneasy. It doesn't seem like a
  good fit for an infrastructure package and it's not clear what
  advantages it has over explicitly loading a protobuf definition into
  an object.}
\reply{TBD}

\pointRaised{Comment 10}{Using global state makes understanding code much harder. In Table 1,
  it's not obvious where \texttt{tutorial.Person} comes from. Is it loaded by
  default by RProtobuf? This need some explanation. In Section 7, what
  does \texttt{readProtoFiles()} do? Why does \texttt{RProtobuf} need to be attached
  as well as \texttt{HistogramTools}? This needs more explanation, and a
  comment on the implications of this approach on CRAN packages and
  namespaces.}
\reply{TBD}

\pointRaised{Comment 11}{
  I'd prefer you eliminate this magic from the magic, but failing that,
  you need a good explanation of why.}
\reply{TBD}

\subsubsection*{Code comments}

\pointRaised{Comment 12}{Using \texttt{file.create()} to determine the absolute path seems like a bad idea.}
\reply{TBD}

\subsubsection*{Minor niggles}

\pointRaised{Comment 13}{Don't refer to the message passing style of OO as traditional.}
\reply{TBD}

\pointRaised{Comment 14}{In Section 3.4, if messages isn't a vectorised class, the default
   print method should use \texttt{cat()} to eliminate the confusing \texttt{[1]}.}
\reply{TBD}

\pointRaised{Comment 15}{The REXP definition would have been better defined using an enum that
   matches R's SEXPTYPE "enum". But I guess that ship has sailed.}
\reply{TBD}

\pointRaised{Comment 16}{Why does \texttt{serialize\_pb(CO2, NULL)} fail silently? Shouldn't it at least
   warn that the serialization is partial?}
\reply{TBD}

\section*{Response to Reviewer \#2}

\pointRaised{Comment 1}{The paper gives an overview of the RProtoBuf package which implements an 
  R interface to the Protocol Buffers library for an efficient 
  serialization of objects. The paper is well written and easy to read. 
  Introductory code is clear and the package provides objects to play with 
  immediately without the need to jump through hoops, making it appealing. 
  The software implementation is executed well.}
\reply{Thank you.}

\pointRaised{Comment 2}{There are, however, a few inconsistencies in the implementation and some 
  issues with specific sections in the paper. In the following both issues 
  will be addressed sequentially by their occurrence in the paper.}
\reply{TBD}

\pointRaised{Comment 3}{p.4 illustrates the use of messages. The class implements list-like 
  access via \texttt{[[} and \$, but strangely \texttt{names()} return NULL and \texttt{length() }
  doesn't correspond to the number of fields leading to startling results like
the following:}

\begin{verbatim}
 > p
[1] "message of type 'tutorial.Person' with 2 fields set"
 > length(p)
[1] 2
 > p[[3]]
[1] ""
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 3 cont.}{The inconsistencies get even more bizarre with descriptors (p.9):}

\begin{verbatim}
 > tutorial.Person$email
[1] "descriptor for field 'email' of type 'tutorial.Person' "
 > tutorial.Person[["email"]]
Error in tutorial.Person[["email"]] : this S4 class is not subsettable
 > names(tutorial.Person)
NULL
 > length(tutorial.Person)
[1] 1
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 3 cont.}{It appears that there is no way to find out the fields of a descriptor 
  directly (although the low-level object methods seem to be exposed as 
  \texttt{\$field\_count()} and \texttt{\$fields()} - but that seems extremely cumbersome). 
  Again, implementing names() and subsetting may help here.}
\reply{TBD}

\pointRaised{Comment 4}{Another inconsistency concerns the \texttt{as.list()} method which by design 
  coerces objects to lists (see \texttt{?as.list}), but the implementation for 
  EnumDescriptor breaks that contract and returns a vector instead:}

\begin{verbatim}
 > is.list(as.list(tutorial.Person$PhoneType))
[1] FALSE
 > str(as.list(tutorial.Person$PhoneType))
  Named int [1:3] 0 1 2
  - attr(*, "names")= chr [1:3] "MOBILE" "HOME" "WORK"
\end{verbatim}

\pointRaised{Comment 4 cont}{As with the other interfaces, names() returns NULL so it is again quite
  difficult to perform even simple operations such as finding out the 
  values. It may be natural use some of the standard methods like names(), 
  levels() or similar. As with the previous cases, the lack of [[ support
  makes it impossible to map named enum values to codes and vice-versa.}
\reply{TBD}

\pointRaised{Comment 5}{In general, the package would benefit from one pass of checks to assess
  the consistency of the API. Since the authors intend direct interaction
  with the objects via basic standard R methods, the classes should behave 
  consistently.}
\reply{TBD}

\pointRaised{Comment 6}{Finally, most classes implement coercion to characters, which is not 
  mentioned and is not quite intuitive for some objects. For example, one
  may think that as.character() on a file descriptor returns let's say the 
  filename, but we get:}

\begin{verbatim}
 > cat(as.character(tutorial.Person$fileDescriptor()))
syntax = "proto2";

package tutorial;

option java_package = "com.example.tutorial";
option java_outer_classname = "AddressBookProtos";
[...]
\end{verbatim}
\reply{TBD}

\pointRaised{Comment 7}{It is not necessary clear what java\_package has to do with a file 
  descriptor in R. Depending on the intention here, it may be useful to 
  explain this feature.
}
\reply{TBD}

\subsubsection*{Other comments:}

\pointRaised{Comment 8}{p.17: "does not support ... function, language or environment. Such 
  objects have no native equivalent type in Protocol Buffers, and have 
  little meaning outside the context or R"
  That is certainly false. Native mirror of environments are hash tables - 
  a very useful type indeed. Language objects are just lists, so there is
  no reason to not include them - they can be useful to store expressions
  that may not be necessary specific to R. Further on p. 18 your run into
  the same problem that could be fixed so easily.}
\reply{TBD}

\pointRaised{Comment 9}{The examples in sections 7 and 8 are somewhat weak. It does not seem 
  clear why one would wish to unleash the power of PB just to transfer 
  breaks and counts for plotting - even a simple ASCII file would do that
  just fine. The main point in the example is presumably that there are 
  code generation methods for Hadoop based on PB IDL such that Hadoop can
  be made aware of the data types, thus making a histogram a proper record 
  that won't be split, can be combined etc. -- yet that is not mentioned 
  nor a way presented how that can be leveraged in practice. The Python 
  example code simply uses a static example with constants to simulate the 
  output of a reducer so it doesn't illustrate the point - the reader is 
  left confused why something as trivial would require PB while a savvy 
  reader is not able to replicate the illustrated process. Possibly 
  explaining the benefits and providing more details on how one would 
  write such a job would make it much more relevant.}
\reply{TBD}

\pointRaised{Comment 10}{Section 8 is not very well motivated. It is much easier to use other 
  formats for HTTP exchange - JSON is probably the most popular, but even
  CSV works in simple settings. PB is a much less common standard. The 
  main advantage of PB is the performance over the alternatives, but HTTP
  services are not necessarily known for their high-throughput so why one
  would sacrifice interoperability by using PB (they are still more hassle 
  and require special installations)? It would be useful if the reason 
  could be made explicit here or a better example chosen.}
\reply{TBD}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2014-12-15 papers/jss/response-to-reviewers.tex
Add more point to point replies.  Still working.  I can mostly finish this up today.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@932 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   3-5 belong in the documentation, not in a vignette/paper.}
-\reply{Done. TO BE EXPANDED}
+\reply{Done. RProtoBuf was designed and implemented before RC were
+  available, and this is noted in a footnote now.  Explanation of how
+  they are made mutable haas been added.  Better explanation of the
+  two styles and '\$' as been added, while no longer using the
+  confusing term
+  'pseudo-method' anywhere.  Moved Tables 3-5 into the documentation
+  and out of the paper, as suggested.}
 
@@
   relatively unimportant details of the map-reduce setup?}
-\reply{TBD}
+\reply{Done.  Rewritten with more motivation taking into account this feedback.}
 
@@
   hence the package easier to use).}
-\reply{TBD}
+\reply{This is an area for future work and is a space explored in
+  another package called Motobuf by other authors.}
 
@@
   equivalents.}
-\reply{TBD}
+
+\reply{We have updated these sections to make it clearer that the main
+  distinction is between schema-based datastructures (section 5) and
+  schema-less use where a catch-all .proto is used (section 6).
+  Neither section is meant to focus on only a single direction of the
+  conversion, but how conversion works when you have a schema or not.
+  How important R classes are converted to their protobuf equivalents
+  isn't super useful as a C++, Java, or Python program is unlikely to
+  want to read in an R data.frame exactly as it is defined.  Much more
+  likely is an application-specific message format is defined between the
+  two services, such as the HistogramTools example in the next section.
+  Much more detail has been added to an interesting part of section 6 --
+  which datasets exactly are better served with RProtoBuf than
+  base::serialize and why?}
 
@@
   protobufs?}
-\reply{TBD}
+\reply{Table 10 has been replaced with a plot, the outliers are
+  labeled, and the text now includes some interesting explanation
+  about the outliers.  Page 4 explains that the R implementation of
+  protocol buffers uses reflection to make operations slower but makes
+  it more convenient for interactive data analysis.  None of the
+  built-in datasets are large enough for performance to really come up
+  as an issue, and for any serialization method examples could be
+  found that significantly favor one over another, so we don't think
+  there will be benefit to adding anything here.
+}
 
@@
   an object.}
-\reply{TBD}
+\reply{More information about the advantages and disadvantages of this
+  approach have been added.}
 
@@
   namespaces.}
-\reply{TBD}
+\reply{We followed this recommendation and added explanation for how
+\texttt{tutorial.Person} is loaded, specifically : \emph{A small number of message types are imported when the
+package is first loaded, including the tutorial.Person type we saw in
+the last section.}  We removed the superfluous attach of \texttt{RProtoBuf}.}
 
@@
   you need a good explanation of why.}
-\reply{TBD}
+\reply{We've added more explanation about this.}
 
@@
 \pointRaised{Comment 12}{Using \texttt{file.create()} to determine the absolute path seems like a bad idea.}
-\reply{TBD}
+\reply{We followed this recommendation and removed two instances of
+  \texttt{file.create()} for this purpose with calls to
+  \texttt{normalizePath} with \texttt{mustWork=FALSE}.}
 

### 2014-12-15 papers/jss/response-to-reviewers.tex
More point-by-point responses.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@933 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 package is first loaded, including the tutorial.Person type we saw in
-the last section.}  We removed the superfluous attach of \texttt{RProtoBuf}.}
+the last section.}  Thank you also for spotting the superfluous attach
+of \texttt{RProtoBuf}, it has been removed from the example.}
 
@@
 \pointRaised{Comment 13}{Don't refer to the message passing style of OO as traditional.}
-\reply{TBD}
+\reply{Done, we don't refer to this style as traditional anywhere in
+  the manuscript anymore.}
 
@@
    print method should use \texttt{cat()} to eliminate the confusing \texttt{[1]}.}
-\reply{TBD}
+\reply{Done}
 
@@
    matches R's SEXPTYPE "enum". But I guess that ship has sailed.}
-\reply{TBD}
+\reply{Acknowledged.  We chose to maintain compatibility with RHIPE here.  The main
+use of RProtoBuf is not with rexp.proto however -- it with
+application-specific schemas in .proto files for sending data between
+applications.  Users that want to do something very R-specific are
+welcome to use their own \texttt{.proto} files with an enum to represent R SEXPTYPEs.}
 
@@
    warn that the serialization is partial?}
-\reply{TBD}
+\reply{Fixed, \texttt{serialize\_pb} now works for all built-in datatypes in R
+  and no longer fails silently if it encounters something it can't serialize.}
 
@@
   will be addressed sequentially by their occurrence in the paper.}
-\reply{TBD}
+\reply{These and others have been identified and addressed.  Thank you
+  for taking the time to enumerate these issues.}
 
@@
 \end{verbatim}
-\reply{TBD}
+\reply{We've corrected the list-like accessor, fixed \texttt{length()} to
+  correspond to the number of set fields, and added \texttt{names()}:}
+\begin{verbatim}
+> p
+message of type 'tutorial.Person' with 0 fields set
+> length(p)
+[1] 0
+> p[[3]]
+[1] ""
+> p$id <- 1
+> length(p)
+[1] 1
+> names(p)
+[1] "name"  "id"    "email" "phone"
+\end{verbatim}
 
@@
 \end{verbatim}
-\reply{TBD}
+\reply{We agree, and have addressed this inconsistency.  Thank you:}
+\begin{verbatim}
+> tutorial.Person$email
+descriptor for field 'email' of type 'tutorial.Person' 
+> tutorial.Person[["email"]]
+descriptor for field 'email' of type 'tutorial.Person' 
+> names(tutorial.Person)
+[1] "name"        "id"          "email"       "phone"       "PhoneNumber"
+[6] "PhoneType"  
+> length(tutorial.Person)
+[1] 6
+\end{verbatim}
 
@@
   Again, implementing names() and subsetting may help here.}
-\reply{TBD}
+\reply{\texttt{names} and subsetting implemented.  Thank you for the
+  suggestion.:}
+\begin{verbatim}
+> tutorial.Person[[1]]
+descriptor for field 'name' of type 'tutorial.Person' 
+> tutorial.Person[[2]]
+descriptor for field 'id' of type 'tutorial.Person' 
+\end{verbatim}
 
@@
 
+\reply{Fixed, thank you. New output:}
+\begin{verbatim}
+> is.list(as.list(tutorial.Person$PhoneType))
+[1] TRUE
+> str(as.list(tutorial.Person$PhoneType))
+List of 3
+ $ MOBILE: int 0
+ $ HOME  : int 1
+ $ WORK  : int 2
+\end{verbatim}
+
 \pointRaised{Comment 4 cont}{As with the other interfaces, names() returns NULL so it is again quite
@@
   makes it impossible to map named enum values to codes and vice-versa.}
-\reply{TBD}
+\reply{Fixed, thank you.  New output:}
+\begin{verbatim}
+> names(tutorial.Person$PhoneType)
+[1] "MOBILE" "HOME"   "WORK"  
+> tutorial.Person$PhoneType[["HOME"]]
+[1] 1
+\end{verbatim}
 
@@
   consistently.}
-\reply{TBD}
+\reply{We made several passes, correcting issues as documented in
+  \texttt{ChangeLog} and now present in our latest 0.4.2 release on CRAN.}
 
@@
 \end{verbatim}
-\reply{TBD}
+\reply{In choosing the debug output for a file descriptor we agree
+  that \texttt{filename} is a reasonable thing to expect, but we also
+  think that the contents of the \texttt{.proto} file is also
+  reasonable, and also more useful.  We document this in
+  ``FileDescriptor-class'', the vignette, and other sources.
+  \texttt{@filename} is one of the slots of the FileDescriptor class
+  and so very easy to find.  The contents of the \texttt{.proto} are
+  not as easily accessible in a slot, however, and so we find it much
+  more useful to be output with \texttt{as.character()}.}
 
@@
 }
-\reply{TBD}
+\reply{This snippet has been removed as part of the general move of
+  less relevant details to the package documentation, but for
+  reference the \texttt{.proto} file syntax is defined in the Protocol Buffers
+  language guide which is referenced earlier. It is a cross platform
+  library and so this syntax specifies some parameters when Java code
+  is used to access the structures defined in this file.  No such
+  special syntax is required in the \texttt{.proto} files for R
+  language code and so this line about java\_package was not relevant
+  or needed in any way for RProtoBuf and is documented elsewhere.}
 
@@
   the same problem that could be fixed so easily.}
-\reply{TBD}
+\reply{You are right.  Environments are more than just hash
+  tables because they include other configuration parameters that are
+  necessary to serialize as well to make sure
+  serialization/unserialization is indempotent, but we agree it is
+  cleaner and the package and the exposition in the paper to just make
+  sure we serialize everything.  We can now fall back to
+  \texttt{base::serialize} and storing the bits in a rawString type of
+  RProtoBuf to make the R schema-less serialization more complete.}
 

### 2014-12-16 papers/jss/response-to-reviewers.tex
Address remaining points in referee feedback.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@934 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   more detail.}
-\reply{The paper was rewritten throughout and is now much tighter at just 23 pages.}
+\reply{The paper is now 6-pages much tighter at just 23 pages.
+  Sections 3 - 8 (all but sec 1 introduction, sec 2 protocol buffers,
+  and sec 9 conclusion have been rewritten to address the specific and
+  general feedback in these reviews)}
 
@@
   and sticking to it in this section - two is confusing.}
-\reply{We followed this recommendation and reduced section 3 to about 2 1/2 pages.}
+\reply{We followed this recommendation and reduced section 3 to about $2\frac{1}{2}$ pages.}
 
@@
   available, and this is noted in a footnote now.  Explanation of how
-  they are made mutable haas been added.  Better explanation of the
-  two styles and '\$' as been added, while no longer using the
+  they are made mutable has been added.  Better explanation of the
+  two styles and '\$' as been added.  We are no longer using the
   confusing term
-  'pseudo-method' anywhere.  Moved Tables 3-5 into the documentation
+  'pseudo-method' anywhere.  We moved Tables 3-5 into the documentation
   and out of the paper, as suggested.}
@@
   hence the package easier to use).}
-\reply{This is an area for future work and is a space explored in
-  another package called Motobuf by other authors.}
+\reply{This is a good observation that only became clear to us after
+  significant usage of \texttt{RProtoBuf}.  Providing a full ``vectorized'' Messages class would require slicing
+  operators that let you quickly extract a given field from each
+  element of the message vector in order to be really useful.  This
+  would require significant amounts of C++ code for efficient
+  manipulation on the order of data.table or other similar large C++ R
+  packages on CRAN.  There is another package called Motobuf by other authors
+  that takes this approach but in practice, at Google at least, the
+  ease-of-use provided by the simple Message interface of RProtoBuf
+  has won with users.  It is still future work to keep the simple
+  interactive interface of RProtoBuf with the vectorized efficiency of
+  Motobuf.  For now, users typically do their slicing of vectors like
+  this through a distributed database (NewSQL is the term of the day?)
+  like Dremel or other system and then just get the response Protocol
+  Buffers in return to the request.}
 
@@
   equivalents.}
-
 \reply{We have updated these sections to make it clearer that the main
@@
   data between R and python using RProtoBuf?}
-\reply{TBD}
+\reply{All of these details are application-specific, whereas
+  RProtoBuf is an infrastructure package.  Distributed systems define
+  their own interfaces, with their own date/time fields, usually as
+  int64s of fractional seconds since the unix epoch for the systems I
+  have worked on.  An example is given for Histograms in the next
+  section.  Factors could be represented as repeated enums in protocol
+  buffers, certainly, if that is how one wanted to define a schema.}
 
@@
   as an issue, and for any serialization method examples could be
-  found that significantly favor one over another, so we don't think
-  there will be benefit to adding anything here.
-}
+  found that significantly favor one over another in runtime, so we
+  don't think there will be benefit to adding anything here.  }
 
@@
    print method should use \texttt{cat()} to eliminate the confusing \texttt{[1]}.}
-\reply{Done}
+\reply{Done, thanks.}
 
@@
 \reply{Acknowledged.  We chose to maintain compatibility with RHIPE here.  The main
-use of RProtoBuf is not with rexp.proto however -- it with
-application-specific schemas in .proto files for sending data between
+use of RProtoBuf is not with \texttt{rexp.proto} however -- it with
+application-specific schemas in \texttt{.proto} files for sending data between
 applications.  Users that want to do something very R-specific are
@@
   mentioned and is not quite intuitive for some objects. For example, one
-  may think that as.character() on a file descriptor returns let's say the 
+  may think that \texttt{as.character()} on a file descriptor returns let's say the 
   filename, but we get:}
@@
 \end{verbatim}
-\reply{In choosing the debug output for a file descriptor we agree
+\reply{The behavior is documented in the package documentation but
+  seemed like a minor detail not important for an already-long paper.
+  In choosing the debug output for a file descriptor we agree
   that \texttt{filename} is a reasonable thing to expect, but we also
   think that the contents of the \texttt{.proto} file is also
-  reasonable, and also more useful.  We document this in
+  reasonable, but more useful.  We document this in
   ``FileDescriptor-class'', the vignette, and other sources.
@@
   write such a job would make it much more relevant.}
-\reply{TBD}
+\reply{Yes, we added more detail about the advantages of using a
+  proper data type for the histograms in this example that you mentioned here -- the
+  ability to write combiners, prevent arbitrary splitting of the
+  records, etc that can greatly improve performance.  We agree with
+  the other reviewer that we don't want to get bogged down in details
+  about a particular MapReduce implementation (such as Hadoop) and so
+  now we specifically mention that goal here.
+  I think we make a better connection now between the
+  abstract MapReduce example given, and then the simpler Python
+  example code with a static example.}
 
@@
   could be made explicit here or a better example chosen.}
-\reply{TBD}
+\reply{This section has been reworded to make it shorter and more
+  crisp, with fewer extraneous details about OpenCPU.
+Protocol
+  Buffers is an efficient protocol used between distributed systems at
+  many of the world's largest internet companies (Twitter, Sony,
+  Google, etc.) but the design and implementation of a large
+  enterprise-scale distributed system with a complex RPC system and
+  serialization needs is well beyond the scope of what we can add to a
+  paper about RProtoBuf.  We chose this example because it is a much
+  more accessible example that any reader can use to easily
+  send/receive RPCs and parse the results with RProtoBuf.}
 

### 2014-12-17 papers/jss/response-to-reviewers.tex
Halfway done another pass. This is coming together very well too.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@936 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
   more detail.}
-\reply{The paper is now 6-pages much tighter at just 23 pages.
-  Sections 3 - 8 (all but sec 1 introduction, sec 2 protocol buffers,
-  and sec 9 conclusion have been rewritten to address the specific and
-  general feedback in these reviews)}
+\reply{The paper is now six pages shorter at just 23 pages.
+  Sections 3 - 8 (all but Section 1 (``Introduction''), Section 2 (``Protocol Buffers''),
+  and Section 9 (``Conclusion'') have been thoroughly rewritten to address the specific and
+  general feedback in these reviews.}
 
@@
   and sticking to it in this section - two is confusing.}
-\reply{We followed this recommendation and reduced section 3 to about $2\frac{1}{2}$ pages.}
+\reply{We followed this recommendation, reduced section 3 to about
+  $2\frac{1}{2}$ pages, removed the subheadings and tightened the exposition.}
 
@@
 \reply{Done. RProtoBuf was designed and implemented before RC were
-  available, and this is noted in a footnote now.  Explanation of how
+  available, and this is now noted explicitly in a new footnote.  Explanation of how
   they are made mutable has been added.  Better explanation of the
   two styles and '\$' as been added.  We are no longer using the
-  confusing term
-  'pseudo-method' anywhere.  We moved Tables 3-5 into the documentation
-  and out of the paper, as suggested.}
+  confusing term 'pseudo-method' anywhere.  We also moved Tables 3-5 into the
+  documentation and out of the paper, as suggested.}
 
@@
   packages on CRAN.  There is another package called Motobuf by other authors
-  that takes this approach but in practice, at Google at least, the
-  ease-of-use provided by the simple Message interface of RProtoBuf
+  that takes this approach but in practice (at least for the several hundred
+  users at Google), the ease-of-use provided by the simple Message interface of RProtoBuf
   has won with users.  It is still future work to keep the simple
@@
   equivalents.}
-\reply{We have updated these sections to make it clearer that the main
-  distinction is between schema-based datastructures (section 5) and
-  schema-less use where a catch-all .proto is used (section 6).
+\reply{Done. We have updated these sections to make it clearer that the main
+  distinction is between schema-based datastructures (Section 5) and
+  schema-less use where a catch-all \texttt{.proto} is used (Section 6).
   Neither section is meant to focus on only a single direction of the
@@
   which datasets exactly are better served with RProtoBuf than
-  base::serialize and why?}
+  \texttt{base::serialize} and why?}
 
@@
   their own interfaces, with their own date/time fields, usually as
-  int64s of fractional seconds since the unix epoch for the systems I
+  a double of fractional seconds since the unix epoch for the systems I
   have worked on.  An example is given for Histograms in the next
-  section.  Factors could be represented as repeated enums in protocol
-  buffers, certainly, if that is how one wanted to define a schema.}
+  section.  Factors could be represented as repeated enums in Protocol
+  Buffers, certainly, if that is how one wanted to define a schema.}
 
@@
   protobufs?}
-\reply{Table 10 has been replaced with a plot, the outliers are
+\reply{Done. Table 10 has been replaced with a plot, the outliers are
   labeled, and the text now includes some interesting explanation
   about the outliers.  Page 4 explains that the R implementation of
-  protocol buffers uses reflection to make operations slower but makes
+  Protocol Buffers uses reflection to make operations slower but makes
   it more convenient for interactive data analysis.  None of the
@@
   an object.}
-\reply{More information about the advantages and disadvantages of this
+\reply{Done. More information about the advantages and disadvantages of this
   approach have been added.}
@@
   namespaces.}
-\reply{We followed this recommendation and added explanation for how
+\reply{Done. We followed this recommendation and added explanation for how
   \texttt{tutorial.Person} is loaded, specifically : \emph{A small number of message types are imported when the
@@
   you need a good explanation of why.}
-\reply{We've added more explanation about this.}
+\reply{Done. We've added more explanation about this.}
 
@@
 \pointRaised{Comment 12}{Using \texttt{file.create()} to determine the absolute path seems like a bad idea.}
-\reply{We followed this recommendation and removed two instances of
+\reply{Done. We followed this recommendation and removed two instances of
   \texttt{file.create()} for this purpose with calls to
@@
 \pointRaised{Comment 13}{Don't refer to the message passing style of OO as traditional.}
-\reply{Done, we don't refer to this style as traditional anywhere in
+\reply{Done. We don't refer to this style as traditional anywhere in
   the manuscript anymore.}
@@
    warn that the serialization is partial?}
-\reply{Fixed, \texttt{serialize\_pb} now works for all built-in datatypes in R
+\reply{Done. We fixed this and \texttt{serialize\_pb} now works for all built-in datatypes in R
   and no longer fails silently if it encounters something it can't serialize.}

### 2015-04-13 papers/jss/article.Rnw
today's changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@940 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \item The functional dispatch mechanism of the the form
-  \verb|method(object, arguments)| (common to \proglang{R}), and
+  \verb|method(object, arguments)| (common to \proglang{R}).
 \item The message passing object-oriented notation of the form
@@
 
-<<>>=
-f <- tutorial.Person$fileDescriptor()
-f
-f$Person
-@
+% < < > > =
+% f <- tutorial.Person$fileDescriptor()
+% f
+% f$Person
+% @
+
+\begin{Schunk}
+\begin{Sinput}
+R> f <- tutorial.Person$fileDescriptor()
+R> f
+\end{Sinput}
+\begin{Soutput}
+file descriptor for package tutorial \
+    (/usr/local/lib/R/site-library/RProtoBuf/proto/addressbook.proto)
+\end{Soutput}
+\begin{Sinput}
+R> f$Person
+\end{Sinput}
+\begin{Soutput}
+descriptor for type 'tutorial.Person' 
+\end{Soutput}
+\end{Schunk}
 
@@
 \begin{small}
-\begin{tabular}{lp{5cm}p{5cm}}
+\begin{tabular}{lp{5cm}p{5.5cm}}
 \toprule
@@
 \caption{\label{table-get-types}Correspondence between field type and
-  \proglang{R} type retrieved by the extractors. Note that \proglang{R} lacks native
+  \proglang{R} type retrieved by the extractors. \proglang{R} lacks native
   64-bit integers, so the \code{RProtoBuf.int64AsString} option is
   available to return large integers as characters to avoid losing
-  precision.  This option is described in Section~\ref{sec:int64}.}
+  precision; see Section~\ref{sec:int64} below. 
+  All but the \code{Message} type can be represented in vectors of one or
+  more elements; for the latter a list is used.}
 \end{table}
@@
   \toprule
-  Data Set & object.size & \multicolumn{2}{c}{\proglang{R} Serialization} &
-  \multicolumn{2}{c}{RProtoBuf Serialization} \\
+  Data set & object.size & \multicolumn{2}{c}{\proglang{R} serialization} &
+  \multicolumn{2}{c}{RProtoBuf serialization} \\
   & & default & gzipped & default & gzipped \\

### 2015-04-13 papers/jss/JSSemails/response-to-reviewers.tex
some file cleanups, part two
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@946 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2014-12-17 papers/jss/response-to-reviewers.tex
@@
   will be addressed sequentially by their occurrence in the paper.}
-\reply{These and others have been identified and addressed.  Thank you
+\reply{Done. These and others have been identified and addressed.  Thank you
   for taking the time to enumerate these issues.}
@@
 \end{verbatim}
-\reply{We've corrected the list-like accessor, fixed \texttt{length()} to
+\reply{Done. We have corrected the list-like accessor, fixed \texttt{length()} to
   correspond to the number of set fields, and added \texttt{names()}:}
@@
 \end{verbatim}
-\reply{We agree, and have addressed this inconsistency.  Thank you:}
+\reply{Done. We agree, and have addressed this inconsistency.  Thank you for
+  catching this.}
 \begin{verbatim}
@@
   Again, implementing names() and subsetting may help here.}
-\reply{\texttt{names} and subsetting implemented.  Thank you for the
-  suggestion.:}
+\reply{Done. We have implemented \texttt{names} and subsetting.  Thank you for the
+  suggestion.}
 \begin{verbatim}
@@
 
-\reply{Fixed, thank you. New output:}
+\reply{Done, thank you. New output below:}
 \begin{verbatim}
@@
   makes it impossible to map named enum values to codes and vice-versa.}
-\reply{Fixed, thank you.  New output:}
+\reply{Done, thank you.  New output:}
 \begin{verbatim}
@@
   consistently.}
-\reply{We made several passes, correcting issues as documented in
+\reply{We made several passes, correcting issues as documented in the
   \texttt{ChangeLog} and now present in our latest 0.4.2 release on CRAN.}
@@
   think that the contents of the \texttt{.proto} file is also
-  reasonable, but more useful.  We document this in
+  reasonable, but more useful.  We document this in the help for
   ``FileDescriptor-class'', the vignette, and other sources.
@@
 }
-\reply{This snippet has been removed as part of the general move of
+\reply{Done. This snippet has been removed as part of the general move of
   less relevant details to the package documentation, but for
@@
   the same problem that could be fixed so easily.}
-\reply{You are right.  Environments are more than just hash
+\reply{Acknowledged.  Environments are more than just hash
   tables because they include other configuration parameters that are
@@
   sure we serialize everything.  We can now fall back to
-  \texttt{base::serialize} and storing the bits in a rawString type of
+  \texttt{base::serialize()} and storing the bits in a rawString type of
   RProtoBuf to make the R schema-less serialization more complete.}
@@
   could be made explicit here or a better example chosen.}
-\reply{This section has been reworded to make it shorter and more
-  crisp, with fewer extraneous details about OpenCPU.
-Protocol
+\reply{Done. This section has been reworded to make it shorter and more
+  crisp, with fewer extraneous details about OpenCPU. Protocol
   Buffers is an efficient protocol used between distributed systems at

### 2016-07-10 papers/jss-finalising/v71i02.orig.tex
finalising: content of emailed zip file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@952 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2015-04-13 papers/jss/article.Rnw
@@
 \documentclass[article]{jss}
-\usepackage{booktabs}
-\usepackage{listings}
-\usepackage[toc,page]{appendix}
-
-% Line numbers for drafts.
-%\usepackage[switch, modulo]{lineno}
-%\linenumbers
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Spelling Standardization:
-% Protocol Buffers, not protocol buffers
-% large-scale, not large scale
-% Oxford comma: foo, bar, and baz.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%
-% Local helpers to make this more compatible with R Journal style.
-%
-\RequirePackage{fancyvrb}
-\RequirePackage{alltt}
+\usepackage{thumbpdf,listings} 
+\graphicspath{{Figures/}}
+\shortcites{sciencecloud,janus,dremel,nlme}
+
 \DefineVerbatimEnvironment{example}{Verbatim}{}
-% Articles with many authors we should shorten to FirstAuthor, et al.
-\shortcites{sciencecloud,janus,dremel,nlme}
+
 \author{Dirk Eddelbuettel\\Debian Project \And 
-        Murray Stokely\\Google, Inc \And
-        Jeroen Ooms\\UCLA}
+        Murray Stokely\\Google, Inc. \And
+        Jeroen Ooms\\University of California,\\Los Angeles}
+\Plainauthor{Dirk Eddelbuettel, Murray Stokely, Jeroen Ooms}
+
 \title{\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}}
-
-%% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Dirk Eddelbuettel, Murray Stokely, Jeroen Ooms} %% comma-separated
 \Plaintitle{RProtoBuf: Efficient Cross-Language Data Serialization in R}
-\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in \proglang{R}} %% a short title (if necessary)
-
-%% an abstract and keywords
+\Shorttitle{\pkg{RProtoBuf}: Protocol Buffers in \proglang{R}}
+
 \Abstract{
@@
   a sophisticated mix of applications written in general purpose and
-  specialized programming languages.  
-  Many formats commonly used to import and export data between
-  different programs or systems, such as \code{CSV} or \code{JSON}, are
-  verbose, inefficient, not type-safe, or tied to a specific programming language.
-  Protocol Buffers are a popular
-  method of serializing structured data between applications---while remaining
-  independent of programming languages or operating systems.
-  They offer a unique combination of features, performance, and maturity that seems
-  particularly well suited for data-driven applications and numerical
-  computing.
-  The \pkg{RProtoBuf} package provides a complete interface to Protocol
-  Buffers from the
-  \proglang{R} environment for statistical computing.
+  specialized programming languages.  Many formats commonly used to
+  import and export data between different programs or systems, such
+  as CSV or JSON, are verbose, inefficient, not
+  type-safe, or tied to a specific programming language.  Protocol
+  Buffers are a popular method of serializing structured data between
+  applications -- while remaining independent of programming languages
+  or operating systems.  They offer a unique combination of features,
+  performance, and maturity that seems particularly well suited for
+  data-driven applications and numerical computing.  The
+  \pkg{RProtoBuf} package provides a complete interface to Protocol
+  Buffers from the \proglang{R} environment for statistical computing.
   This paper outlines the general class of data serialization
   requirements for statistical computing, describes the implementation
-  of the \pkg{RProtoBuf} package, and illustrates its use with
-  example applications in large-scale data collection pipelines and web
+  of the \pkg{RProtoBuf} package, and illustrates its use with example
+  applications in large-scale data collection pipelines and web
   services.
-  %% TODO(ms) keep it less than 150 words. -- I think this may be 154,
-  %% depending how emacs is counting.
 }
+
 \Keywords{\proglang{R}, \pkg{Rcpp}, Protocol Buffers, serialization, cross-platform}
-\Plainkeywords{R, Rcpp, Protocol Buffers, serialization, cross-platform} %% without formatting
-%% at least one keyword must be supplied
-
-%% publication information
-%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
-%% \Volume{50}
-%% \Issue{9}
-%% \Month{June}
-%% \Year{2012}
-%% \Submitdate{2012-06-04}
-%% \Acceptdate{2012-06-04}
-
-%% The address of (at least) one author should be given
-%% in the following format:
+\Plainkeywords{R, Rcpp, Protocol Buffers, serialization, cross-platform}
+
+\Volume{71}
+\Issue{2}
+\Month{July}
+\Year{2016}
+\Submitdate{2014-02-05}
+\Acceptdate{2015-04-13}
+\DOI{10.18637/jss.v071.i02}
+
 \Address{
@@
   Debian Project \\
-  River Forest, IL, USA\\
+  River Forest, IL, United States of America\\
   E-mail: \email{edd@debian.org}\\
-  URL: \url{http://dirk.eddelbuettel.com}\\
-  \\
+  URL: \url{http://dirk.eddelbuettel.com/}\\
+
   Murray Stokely\\
@@
   1600 Amphitheatre Parkway\\
-  Mountain View, CA, USA\\
+  Mountain View, CA, United States of America\\
   E-mail: \email{mstokely@google.com}\\
   URL: \url{http://www.stokely.org/}\\
-  \\
+
   Jeroen Ooms\\
   UCLA Department of Statistics\\
-  University of California\\
-  Los Angeles, CA, USA\\
+  University of California, Los Angeles\\
+  Los Angeles, CA, United States of America\\
   E-mail: \email{jeroen.ooms@stat.ucla.edu}\\
-  URL: \url{https://jeroenooms.github.io}
+  URL: \url{https://jeroenooms.github.io/}
 }
-%% It is also possible to add a telephone and fax number
-%% before the e-mail in the following format:
-%% Telephone: +43/512/507-7103
-%% Fax: +43/512/507-2851
-
-%% for those who use Sweave please include the following line (with % symbols):
-%% need no \usepackage{Sweave.sty}
-
-%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\SweaveOpts{concordance=FALSE,prefix.string=figures/fig}
-
-%% include your article here, just as usual
-%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.
-
-% We don't want a left margin for Sinput or Soutput for our table 1.
-%\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=0em}
-%\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=0em}
-%\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}
-% Setting the topsep to 0 reduces spacing from input to output and
-% improves table 1.
-\fvset{listparameters={\setlength{\topsep}{0pt}}}
-\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}
-
-%% DE: I tend to have wider option(width=...) so this
-%%     guarantees better line breaks
-<<echo=FALSE,print=FALSE>>=
-## cf http://www.jstatsoft.org/style#q12
-options(prompt = "R> ", 
-        continue = "+  ", 
-        width = 70, 
-        useFancyQuotes = FALSE, 
-        digits = 4)
-@
-
-\maketitle
+
+\vspace*{-0.3cm}
 
@@
 These pipelines are frequently built using different programming 
-languages for the different phases of data analysis --- collection,
+languages for the different phases of data analysis -- collection,
 cleaning, modeling, analysis, post-processing, and
-presentation --- in order to take advantage of the unique combination of
+presentation -- in order to take advantage of the unique combination of
 performance, speed of development, and library support offered by
@@
 
-Given these requirements, how do we safely and efficiently share intermediate results
-between different applications, possibly written in different
-languages, and possibly running on different computer systems?
-In computer programming, \emph{serialization} is the process of
-translating data structures, variables, and session state into a
-format that can be stored or transmitted and then reconstructed in the
-original form later \citep{clinec++}.
-Programming
-languages such as \proglang{R}, \proglang{Julia}, \proglang{Java}, and \proglang{Python} include built-in
-support for serialization, but the default formats 
-are usually language-specific and thereby lock the user into a single
-environment.
-
-Data analysts and researchers often use character-separated text formats such
-as \code{CSV} \citep{shafranovich2005common} to export and import
-data. However, anyone who has ever used \code{CSV} files will have noticed
-that this method has many limitations: it is restricted to tabular data,
-lacks type-safety, and has limited precision for numeric values.  Moreover,
-ambiguities in the format itself frequently cause problems.  For example,
-conventions on which characters is used as separator or decimal point vary by
-country.  \emph{Extensible Markup Language} (\code{XML}) is a
-well-established and widely-supported format with the ability to define just
-about any arbitrarily complex schema \citep{nolan2013xml}. However, it pays
-for this complexity with comparatively large and verbose messages, and added
-complexity at the parsing side (these problems are somewhat mitigated by the
-availability of mature libraries and parsers). Because \code{XML} is
-text-based and has no native notion of numeric types or arrays, it usually not a
-very practical format to store numeric data sets as they appear in statistical
-applications.
-
-A more modern format is \emph{JavaScript ObjectNotation}
-(\code{JSON}), which is derived from the object literals of
+Given these requirements, how do we safely and efficiently share
+intermediate results between different applications, possibly written
+in different languages, and possibly running on different computer
+systems?  In computer programming, \emph{serialization} is the process
+of translating data structures, variables, and session states into a
+format that can be stored or transmitted and then later reconstructed
+in the original form \citep{clinec++}.  Programming languages such as
+\proglang{R} \citep{r}, \proglang{Julia} \citep{julia},
+\proglang{Java}, and \proglang{Python} \citep{python} include built-in
+support for serialization, but the default formats are usually
+language-specific and thereby lock the user into a single environment.
+
+Data analysts and researchers often use character-separated text
+formats such as CSV \citep{shafranovich2005common} to export
+and import data. However, anyone who has ever used CSV files
+will have noticed that this method has many limitations: It is
+restricted to tabular data, lacks type-safety, and has limited
+precision for numeric values.  Moreover, ambiguities in the format
+itself frequently cause problems.  For example, conventions on which
+characters are used as separator or decimal point vary by country.
+\emph{Extensible markup language} (XML) is a well-established
+and widely-supported format with the ability to define just about any
+arbitrarily complex schema \citep{nolan2013xml}. However, it pays for
+this complexity with comparatively large and verbose messages, and
+added complexity at the parsing side (these problems are somewhat
+mitigated by the availability of mature libraries and
+parsers). Because XML is text-based and has no native notion of
+numeric types or arrays, it is usually not a very practical format to
+store numeric data sets as they appear in statistical applications.
+
+A more modern format is \emph{\proglang{JavaScript} object notation}
+(JSON), which is derived from the object literals of
 \proglang{JavaScript}, and already widely-used on the world wide web.
-Several \proglang{R} packages implement functions to parse and generate
-\code{JSON} data from \proglang{R} objects \citep{rjson,RJSONIO,jsonlite}.
-\code{JSON} natively supports arrays and four primitive types: numbers, strings,
-booleans, and null. However, as it too is a text-based format, numbers are
-stored as human-readable decimal notation which is inefficient and
-leads to loss of type (double versus integer) and precision. 
-A number of binary formats based on \code{JSON} have been proposed
-that reduce the parsing cost and improve efficiency, but these formats
-are not widely supported.  Furthermore, such formats lack a separate
-schema for the serialized data and thus still duplicate field names
-with each message sent over the network or stored in a file.
+Several \proglang{R} packages implement functions to parse and
+generate JSON data from \proglang{R} objects
+\citep{rjson,RJSONIO,jsonlite}.  JSON natively supports arrays
+and four primitive types: numbers, strings, booleans, and
+null. However, as it also is a text-based format, numbers are stored
+in human-readable decimal notation which is inefficient and leads to
+loss of type (double versus integer) and precision.  A number of
+binary formats based on JSON have been proposed that reduce the
+parsing cost and improve efficiency, but these formats are not widely
+supported.  Furthermore, such formats lack a separate schema for the
+serialized data and thus still duplicate field names with each message
+sent over the network or stored in a file.
 
@@
 %\citet{Sumaray:2012:CDS:2184751.2184810} for one such comparison.
-Protocol Buffers performs well in the comparison of such formats by
+Protocol Buffers perform well in the comparison of such formats by
 \citet{Sumaray:2012:CDS:2184751.2184810}.
@@
 \emph{Messages} and \emph{Descriptors}.  Section~\ref{sec:rprotobuf-classes}
-details the implementation of the main S4 classes and methods.
+details the implementation of the main \proglang{S}4 classes and methods.
 Section~\ref{sec:types} describes the challenges of type coercion
@@
 features, particularly in the context of data analysis, are:
-
+%
 \begin{itemize}
@@
 \end{itemize}
-
-%\begin{figure}[bp]
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/protobuf-distributed-system-crop.pdf}
-\end{center}
+%
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.9\textwidth]{protobuf-distributed-system-crop.pdf}
 \caption{Example usage of Protocol Buffers.}
@@
 \end{figure}
-
+%
 Figure~\ref{fig:protobuf-distributed-usecase} illustrates an example
@@
 
-While traditional IDLs have at times been criticized for code bloat and
-complexity, Protocol Buffers are based on a simple list and records
-model that is flexible and easy to use.  The schema for structured
-Protocol Buffer data is defined in \code{.proto} files, which may
-contain one or more message types.  Each message type has one or more
-fields.  A field is specified with a unique number (called a \emph{tag number}), a name, a value
-type, and a field rule specifying whether the field is optional,
-required, or repeated.  The supported value types are numbers,
-enumerations, booleans, strings, raw bytes, or other nested message
-types.  The \code{.proto} file syntax for defining the structure of Protocol
-Buffer data is described comprehensively on Google Code\footnote{See 
-\url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.}.
-Table~\ref{tab:proto} shows an example \code{.proto} file that
-defines the \code{tutorial.Person} type\footnote{The compound name
-  \code{tutorial.Person} in R is derived from the name of the
-  message (\emph{Person}) and the name of the package defined at the top of the
-  \code{.proto} file in which it is defined (\emph{tutorial}).}.  The \proglang{R} code in the right
-column shows an example of creating a new message of this type and
-populating its fields.
-
-\noindent
-\begin{table}
+While traditional IDLs have at times been criticized for code bloat
+and complexity, Protocol Buffers are based on a simple list and
+records model that is flexible and easy to use.  The schema for
+structured Protocol Buffer data is defined in \code{.proto} files,
+which may contain one or more message types.  Each message type has
+one or more fields.  A field is specified with a unique number (called
+a \emph{tag number}), a name, a value type, and a field rule
+specifying whether the field is optional, required, or repeated.  The
+supported value types are numbers, enumerations, booleans, strings,
+raw bytes, or other nested message types.  The \code{.proto} file
+syntax for defining the structure of Protocol Buffer data is described
+comprehensively on Google Code at
+  \url{http://code.google.com/apis/protocolbuffers/docs/proto.html}.
+Table~\ref{tab:proto} shows an example \code{.proto} file that defines
+the \code{tutorial.Person} type\footnote{The compound name
+  \code{tutorial.Person} in \proglang{R} is derived from the name of
+  the message (\emph{Person}) and the name of the package defined at
+  the top of the \code{.proto} file in which it is defined
+  (\emph{tutorial}).}.  The \proglang{R} code in the right column
+shows an example of creating a new message of this type and populating
+its fields.
+
+\begin{table}[t!]
 \begin{tabular}{p{0.45\textwidth}p{0.5\textwidth}}
-\toprule
+\hline
 Schema : \code{addressbook.proto} & Example \proglang{R} session\\
-\cmidrule{1-2}
+\hline\\[-10pt]
 \begin{minipage}{.40\textwidth}
@@
 \end{minipage} & \begin{minipage}{.55\textwidth}
-<<echo=TRUE>>=
-library("RProtoBuf")
-p <- new(tutorial.Person, id=1,
-         name="Dirk")
-p$name
-p$name <- "Murray"
-cat(as.character(p))
-serialize(p, NULL)
-class(p)
-@
+\begin{Schunk}
+\begin{Sinput}
+R> library("RProtoBuf")
+R> p <- new(tutorial.Person, id = 1,
++    name = "Dirk")
+R> p$name
+\end{Sinput}
+\begin{Soutput}
+[1] "Dirk"
+\end{Soutput}
+\begin{Sinput}
+R> p$name <- "Murray"
+R> cat(as.character(p))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray"
+id: 1
+\end{Soutput}
+\begin{Sinput}
+R> serialize(p, NULL)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 06 4d 75 72 72 61 79 10 01
+\end{Soutput}
+\begin{Sinput}
+R> class(p)
+\end{Sinput}
+\begin{Soutput}
+[1] "Message"
+attr(,"package")
+[1] "RProtoBuf"
+\end{Soutput}
+\end{Schunk}
+\vspace{0.5mm}
 \end{minipage} \\
-\bottomrule
+\hline
 \end{tabular}
 \caption{The schema representation from a \code{.proto} file for the
-  \code{tutorial.Person} class (left) and simple \proglang{R} code for creating
+  \code{tutorial.Person} type (left) and simple \proglang{R} code for creating
   an object of this class and accessing its fields (right).}
@@
 
-% Note: We comment out subsections in favor of textbf blocks to save
-% space and shrink down this section a little bit.
-%\subsection[Importing message descriptors from .proto files]{Importing message descriptors from \code{.proto} files}
-
-\subsection*{Importing message descriptors from \code{.proto} files}
-
-To create or parse a Protocol Buffer Message, one must first read in
-the message descriptor (\emph{message type}) from a \code{.proto} file.
+\subsection[Importing Message Descriptors from .proto files]{Importing
+  Message Descriptors from \code{.proto} files}
+
+To create or parse a Protocol Buffer message, one must first read in
+the Message Descriptor (\emph{message type}) from a \code{.proto} file.
 A small number of message types are imported when the package is first
@@
 
-After importing proto files, the corresponding message descriptors are
-available by name from the \code{RProtoBuf:DescriptorPool} environment in 
-the \proglang{R} search path.  This environment is implemented with the 
-user-defined tables framework from the \pkg{RObjectTables} package
+After importing proto files, the corresponding Message Descriptors are
+available by name from the \code{RProtoBuf:DescriptorPool} environment
+in the \proglang{R} search path.  This environment is implemented with
+the user-defined tables framework from the \pkg{RObjectTables} package
 available from the OmegaHat project \citep{RObjectTables}.  Instead of
@@
 namespace.\footnote{Note that there is a significant performance
-  overhead with this RObjectTable implementation.  Because the table
-  is on the search path and is not cacheable, lookups of symbols that
-  are behind it in the search path cannot be added to the global object
-  cache, and R must perform an expensive lookup through all of the
-  attached environments and the protocol buffer definitions to find common
-  symbols (most notably those in base) from the global environment.
-  Fortunately, proper use of namespaces and package imports reduces
-  the impact of this for code in packages.}
-
-% Commented out for now because its too detailed.  Lets shorten
-% section 3 per referee feedback.
-
-%<<echo=FALSE,print=FALSE>>=
-%ls("RProtoBuf:DescriptorPool")
-%@
-
-% \subsection{Creating a message}
-
-% \\
-
-\subsection*{Creating, accessing, and modifying messages.}
+  overhead with this \pkg{RObjectTables} implementation.  Because the
+  table is on the search path and is not cacheable, lookups of symbols
+  that are behind it in the search path cannot be added to the global
+  object cache, and \proglang{R} must perform an expensive lookup
+  through all of the attached environments and the Protocol Buffer
+  definitions to find common symbols (most notably those in base) from
+  the global environment.  Fortunately, proper use of namespaces and
+  package imports reduces the impact of this for code in packages.}
+
+\subsection{Creating, accessing, and modifying messages}
 
@@
 %function to create messages.
-
-<<>>=
-p <- new(tutorial.Person, name = "Murray", id = 1)
-@
-
-% \subsection*{Access and modify fields of a message}
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> p <- new(tutorial.Person, name = "Murray", id = 1)
+\end{Sinput}
+\end{Schunk}
+%
 Once the message is created, its fields can be queried
@@
 Buffer messages seem like lists.
-
-<<>>=
-p$name
-p$id
-p$email <- "murray@stokely.org"
-@
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> p$name
+\end{Sinput}
+\begin{Soutput}
+[1] "Murray"
+\end{Soutput}
+\begin{Sinput}
+R> p$id
+\end{Sinput}
+\begin{Soutput}
+[1] 1
+\end{Soutput}
+\begin{Sinput}
+R> p$email <- "murray@stokely.org"
+\end{Sinput}
+\end{Schunk}
+%
 As opposed to \proglang{R} lists, no partial matching is performed
@@
 The \verb|[[| operator can also be used to query and set fields
-of a messages, supplying either their name or their tag number:
-
-<<>>=
-p[["name"]] <- "Murray Stokely"
-p[[ 2 ]] <- 3
-p[["email"]]
-@
-
+of a message, supplying either their name or their tag number:
+%
+\begin{Schunk}
+\begin{Sinput}
+R> p[["name"]] <- "Murray Stokely"
+R> p[[2]] <- 3
+R> p[["email"]]
+\end{Sinput}
+\begin{Soutput}
+[1] "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
+%
 Protocol Buffers include a 64-bit integer type, but \proglang{R} lacks native
@@
 
-\subsection*{Printing, reading, and writing Messages}
-
-%\\
-
-% \textbf{Printing, Reading, and Writing Messages}
+\subsection{Printing, reading, and writing messages}
 
@@
 methods that provide basic information about the message:
-
-<<>>=
-p
-@
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> p
+\end{Sinput}
+\begin{Soutput}
+message of type 'tutorial.Person' with 3 fields set
+\end{Soutput}
+\end{Schunk}
+%
 %For additional information, such as for debugging purposes,
@@
 representation of the contents of a message.
-
-<<>>=
-writeLines(as.character(p))
-@
-
-% \subsection{Serializing messages}
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> writeLines(as.character(p))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray Stokely"
+id: 3
+email: "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
+%
 A primary benefit of Protocol Buffers is an efficient
@@
 as long as the message type is known and its descriptor is available.
-
-<<>>=
-serialize(p, NULL)
-@
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> serialize(p, NULL)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12 6d 75
+[23] 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
+\end{Soutput}
+\end{Schunk}
+%
 The same method can be used to serialize messages to files or arbitrary binary connections:
-
-<<>>=
-tf1 <- tempfile()
-serialize(p, tf1)
-readBin(tf1, raw(0), 500)
-@
-
-% TODO(mstokely): Comment out, combined with last statement. make this
-% shorter, more succinct summary of the key features of RProtoBuf.
-
-%Or to arbitrary binary connections:
-%
-%<<>>=
-%tf2 <- tempfile()
-%con <- file(tf2, open = "wb")
-%serialize(p, con)
-%close(con)
-%readBin(tf2, raw(0), 500)
-%@
-
-% TODO(mstokely): commentd out per referee feedback, but see if this is
-% covered in the package documentation well.
-%
-%\code{serialize} can also be called in a more traditional
-%object-oriented fashion using the dollar operator.
-%
-%<<>>=
-%p$serialize(tf1)
-%con <- file(tf2, open = "wb")
-%p$serialize(con)
-%close(con)
-%@
-%
-%Here, we first serialize to a file \code{tf1} before we serialize to a binary
-%connection to file \code{tf2}.
-
-%\subsection{Parsing messages}
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> tf1 <- tempfile()
+R> serialize(p, tf1)
+R> readBin(tf1, raw(0), 500)
+\end{Sinput}
+\begin{Soutput}
+ [1] 0a 0e 4d 75 72 72 61 79 20 53 74 6f 6b 65 6c 79 10 03 1a 12 6d 75
+[23] 72 72 61 79 40 73 74 6f 6b 65 6c 79 2e 6f 72 67
+\end{Soutput}
+\end{Schunk}
+%
 The \pkg{RProtoBuf} package defines the \code{read} and
@@
 to the \code{read} function in the form of a descriptor:
-
-<<>>=
-msg <- read(tutorial.Person, tf1)
-writeLines(as.character(msg))
-@
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> msg <- read(tutorial.Person, tf1)
+R> writeLines(as.character(msg))
+\end{Sinput}
+\begin{Soutput}
+name: "Murray Stokely"
+id: 3
+email: "murray@stokely.org"
+\end{Soutput}
+\end{Schunk}
+%
 The \code{input} argument of \code{read} can also be a binary
@@
 
-% <<>>=
-% con <- file(tf2, open = "rb")
-% message <- read(tutorial.Person, con)
-% close(con)
-% writeLines(as.character(message))
-% @
-
-% Finally, the raw vector payload of the message can be used:
-%
-%<<>>=
-%payload <- readBin(tf1, raw(0), 5000)
-%message <- read(tutorial.Person, payload)
-%@
-
-% TODO(mstokely): comment out and use only one style, not both per
-% referee feedback.  Also avoid using the term 'pseudo-method' which
-% is unclear.
-%
-%\code{read} can also be used as a method of the descriptor
-%object:
-%
-%<<>>=
-%message <- tutorial.Person$read(tf1)
-%con <- file(tf2, open = "rb")
-%message <- tutorial.Person$read(con)
-%close(con)
-%message <- tutorial.Person$read(payload)
-%@
-%
-%Here we read first from a file, then from a binary connection and lastly from
-%a message payload.
-
-\section{Under the hood: S4 classes and methods}
+\section[Under the hood: S4 classes and methods]{Under the hood: \proglang{S}4 classes and methods}
 \label{sec:rprotobuf-classes}
 
-The \pkg{RProtoBuf} package uses the S4 system to store
-information about descriptors and messages.
-Each \proglang{R} object
+The \pkg{RProtoBuf} package uses the \proglang{S}4 system to store
+information about descriptors and messages.  Each \proglang{R} object
 contains an external pointer to an object managed by the
-\code{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
-than 100 \proglang{C++} functions that provide the
-glue code between the \proglang{R} language classes and the underlying \proglang{C++}
-classes.
-S4 objects are immutable, and so the methods that modify field values of a message return a new copy of the object with R's usual functional copy on modify semantics\footnote{RProtoBuf was designed and implemented before Reference Classes were introduced to offer a new class system with mutable objects.
-If RProtoBuf were
-implemented today Reference Classes would almost certainly be a better
-design choice than S4 classes.}.
-Using the S4 system
-allows the package to dispatch methods that are not
-generic in the S3 sense, such as \code{new} and
-\code{serialize}.
+\pkg{protobuf} \proglang{C++} library, and the \proglang{R} objects
+make calls into more than 100 \proglang{C++} functions that provide
+the glue code between the \proglang{R} language classes and the
+underlying \proglang{C++} classes.  \proglang{S}4 objects are
+immutable, and so the methods that modify field values of a message
+return a new copy of the object with \proglang{R}'s usual functional
+copy on modify semantics\footnote{\pkg{RProtoBuf} was designed and
+  implemented before Reference Classes were introduced to offer a new
+  class system with mutable objects.  If \pkg{RProtoBuf} were implemented
+  today Reference Classes would almost certainly be a better design
+  choice than \proglang{S}4 classes.}.  Using the \proglang{S}4 system
+allows the package to dispatch methods that are not generic in the
+\proglang{S}3 sense, such as \code{new} and \code{serialize}.
 
@@
 The \pkg{RProtoBuf} package in many ways motivated
-the development of \pkg{Rcpp} Modules \citep{eddelbuettel2013exposing},
+the development of \pkg{Rcpp} modules \citep{eddelbuettel2013exposing},
 which provide a more concise way of wrapping \proglang{C++} functions and classes
@@
 are supported:
-
+%
 \begin{itemize}
-\item The functional dispatch mechanism of the the form
+\item The functional dispatch mechanism of the form
   \verb|method(object, arguments)| (common to \proglang{R}).
@@
 
-Additionally, \pkg{RProtoBuf} supports tab completion for all
-classes.  Completion possibilities include method names for all
-classes, plus \emph{dynamic dispatch} on names or types specific to a given
-object.  This functionality is implemented with the
-\code{.DollarNames} S3 generic function defined in the \pkg{utils}
-package that is included with \proglang{R} \citep{r}.
-
-Table~\ref{class-summary-table} lists the six primary Message and
-Descriptor classes in \pkg{RProtoBuf}.
-The package documentation provides a complete description of the slots and methods for
-each class.
-
-% [bp]
-\begin{table}
+Additionally, \pkg{RProtoBuf} supports tab completion for all classes.
+Completion possibilities include method names for all classes, plus
+\emph{dynamic dispatch} on names or types specific to a given object.
+This functionality is implemented with the \code{.DollarNames}
+\proglang{S}3 generic function defined in the \pkg{utils} package that
+is included with \proglang{R} \citep{r}.
+
+Table~\ref{class-summary-table} lists the six primary message and
+descriptor classes in \pkg{RProtoBuf}.  The package documentation
+provides a complete description of the slots and methods for each
+class.
+
+\begin{table}[t!]
 \centering
 \begin{tabular}{lccl}
-\toprule
+\hline
 Class               & Slots & Methods & Dynamic dispatch\\
-\cmidrule{2-4}
-Message             & 2 & 20 & yes (field names)\\
-Descriptor          & 2 & 16 & yes (field names, enum types, nested types)\\
-FieldDescriptor     & 4 & 18 & no\\
-EnumDescriptor      & 4 & 11 & yes (enum constant names)\\
-EnumValueDescriptor & 3 & \phantom{1}6 & no\\
-FileDescriptor      & 3 & \phantom{1}6 & yes (message/field definitions)\\
-\bottomrule
+\hline
+`\code{Message}'             & 2 & 20 & yes (field names)\\
+`\code{Descriptor}'          & 2 & 16 & yes (field names, enum types, nested types)\\
+`\code{FieldDescriptor}'     & 4 & 18 & no\\
+`\code{EnumDescriptor}'      & 4 & 11 & yes (enum constant names)\\
+`\code{EnumValueDescriptor}' & 3 & \phantom{1}6 & no\\
+`\code{FileDescriptor}'      & 3 & \phantom{1}6 & yes (message/field definitions)\\
+\hline
 \end{tabular}
@@
 
-The \code{Message} S4 class represents Protocol Buffer Messages and
-is the core abstraction of \pkg{RProtoBuf}. Each \code{Message}
-contains a pointer to a \code{Descriptor} which defines the schema
-of the data defined in the Message, as well as a number of
-\code{FieldDescriptors} for the individual fields of the message.
-
-<<>>=
-new(tutorial.Person)
-@
+The `\code{Message}' \proglang{S}4 class represents Protocol Buffer
+messages and is the core abstraction of \pkg{RProtoBuf}. Each
+`\code{Message}' contains a pointer to a `\code{Descriptor}' which
+defines the schema of the data defined in the message, as well as a
+number of `\code{FieldDescriptor}'s for the individual fields of the
+message.
+%
+\begin{Schunk}
+\begin{Sinput}
+R> new(tutorial.Person)
+\end{Sinput}
+\begin{Soutput}
+message of type 'tutorial.Person' with 0 fields set
+\end{Soutput}
+\end{Schunk}
 
@@
 
-Descriptors describe the type of a Message.  This includes what fields
+Descriptors describe the type of a message.  This includes what fields
 a message contains and what the types of those fields are.  Message
-descriptors are represented in \proglang{R} by the \emph{Descriptor} S4
-class. The class contains the slots \code{pointer} and
-\code{type}.  Similarly to messages, the \verb|$| operator can be
-used to retrieve descriptors that are contained in the descriptor, or
+descriptors are represented in \proglang{R} by the `\code{Descriptor}'
+\proglang{S}4 class. The class contains the slots \code{pointer} and
+\code{type}.  Similarly to messages, the \verb|$| operator can be used
+to retrieve descriptors that are contained in the descriptor, or
 invoke methods.
@@
 When \pkg{RProtoBuf} is first loaded it calls
-\code{readProtoFiles} to read in the example \code{addressbook.proto} file
+\code{readProtoFiles} which reads in the example file \code{addressbook.proto} 
 included with the package.  The \code{tutorial.Person} descriptor
@@
 
-\subsubsection*{Field descriptors}
+\subsubsection{Field descriptors}
 \label{subsec-field-descriptor}
 
-<<>>=
-tutorial.Person$email 
-tutorial.Person$email$is_required()
-tutorial.Person$email$type()
-tutorial.Person$email$as.character()
-class(tutorial.Person$email)
-@
-
-\subsubsection*{Enum and EnumValue descriptors}
+\begin{Schunk}
+\begin{Sinput}
+R> tutorial.Person$email 
+\end{Sinput}
+\begin{Soutput}
+descriptor for field 'email' of type 'tutorial.Person' 
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$email$is_required()
+\end{Sinput}
+\begin{Soutput}
+[1] FALSE
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$email$type()
+\end{Sinput}
+\begin{Soutput}
+[1] 9
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$email$as.character()
+\end{Sinput}
+\begin{Soutput}
+[1] "optional string email = 3;\n"
+\end{Soutput}
+\begin{Sinput}
+R> class(tutorial.Person$email)
+\end{Sinput}
+\begin{Soutput}
+[1] "FieldDescriptor"
+attr(,"package")
+[1] "RProtoBuf"
+\end{Soutput}
+\end{Schunk}
+
+\subsubsection{Enum and EnumValue descriptors}
 \label{subsec-enum-descriptor}
 
-The \code{EnumDescriptor} type contains information about what values a
-type defines, while the \code{EnumValueDescriptor} describes an
-individual enum constant of a particular type.  The \verb|$| operator
-can be used to retrieve the value of enum constants contained in the
-EnumDescriptor, or to invoke methods.
-
-<<>>=
-tutorial.Person$PhoneType
-tutorial.Person$PhoneType$WORK
-class(tutorial.Person$PhoneType)
-tutorial.Person$PhoneType$value(1)
-tutorial.Person$PhoneType$value(name="HOME")
-tutorial.Person$PhoneType$value(number=1)
-class(tutorial.Person$PhoneType$value(1))
-@
-
-\subsubsection*{File descriptors}
+The `\code{EnumDescriptor}' class contains information about what
+values a type defines, while the `\code{EnumValueDescriptor}'
+describes an individual enum constant of a particular type.  The
+\verb|$| operator can be used to retrieve the value of enum constants
+contained in the `\code{EnumDescriptor}', or to invoke methods.
+%
+\begin{Schunk}
+\begin{Sinput}
+R> tutorial.Person$PhoneType
+\end{Sinput}
+\begin{Soutput}
+descriptor for enum 'PhoneType' with 3 values
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$WORK
+\end{Sinput}
+\begin{Soutput}
+[1] 2
+\end{Soutput}
+\begin{Sinput}
+R> class(tutorial.Person$PhoneType)
+\end{Sinput}
+\begin{Soutput}
+[1] "EnumDescriptor"
+attr(,"package")
+[1] "RProtoBuf"
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(1)
+\end{Sinput}
+\begin{Soutput}
+enum value descriptor tutorial.Person.MOBILE
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(name = "HOME")
+\end{Sinput}
+\begin{Soutput}
+enum value descriptor tutorial.Person.HOME
+\end{Soutput}
+\begin{Sinput}
+R> tutorial.Person$PhoneType$value(number = 1)
+\end{Sinput}
+\begin{Soutput}
+enum value descriptor tutorial.Person.HOME
+\end{Soutput}
+\begin{Sinput}
+R> class(tutorial.Person$PhoneType$value(1))
+\end{Sinput}
+\begin{Soutput}
+[1] "EnumValueDescriptor"
+attr(,"package")
+[1] "RProtoBuf"
+\end{Soutput}
+\end{Schunk}
+
+\subsubsection{File descriptors}
 \label{subsec-file-descriptor}
 
-The class \emph{FileDescriptor} represents file descriptors in \proglang{R}.
-The \verb|$| operator can be used to retrieve named fields defined in
-the FileDescriptor, or to invoke methods.
-
-% < < > > =
-% f <- tutorial.Person$fileDescriptor()
-% f
-% f$Person
-% @
-
+The class `\code{FileDescriptor}' represents file descriptors in
+\proglang{R}.  The \verb|$| operator can be used to retrieve named
+fields defined in the `\code{FileDescriptor}', or to invoke methods.
+%
 \begin{Schunk}
@@
 
-One of the benefits of using an Interface Definition Language (IDL)
+One of the benefits of using an interface definition language (IDL)
 like Protocol Buffers is that it provides a highly portable basic type
@@
 
-\begin{table}[h]
+\begin{table}[t!]
 \centering
-\begin{small}
-\begin{tabular}{lp{5cm}p{5.5cm}}
-\toprule
+\begin{tabular}{lp{5cm}p{6cm}}
+\hline
 Field type & \proglang{R} type (non repeated) & \proglang{R} type (repeated) \\
-\cmidrule(r){2-3}
+\hline
 double	& \code{double} vector & \code{double} vector \\
@@
 enum & \code{integer} vector & \code{integer} vector \\[3mm]
-message & \code{S4} object of class \code{Message} & \code{list} of \code{S4} objects of class \code{Message} \\
-\bottomrule
+message & \proglang{S}4 object of class `\code{Message}' & \code{list} of \proglang{S}4 objects of class `\code{Message}' \\
+\hline
 \end{tabular}
-\end{small}
 \caption{\label{table-get-types}Correspondence between field type and
@@
   precision; see Section~\ref{sec:int64} below. 
-  All but the \code{Message} type can be represented in vectors of one or
+  All but the `\code{Message}' type can be represented in vectors of one or
   more elements; for the latter a list is used.}
@@
 Buffer schema, only accept \code{TRUE} or \code{FALSE}.  This means
-that we simply can not store \proglang{R} logical vectors that include all three
-possible values as booleans.  The library will refuse to store
+that we simply cannot store \proglang{R} logical vectors that include all three
+possible values as booleans.  The package will refuse to store
 \code{NA}s in Protocol Buffer boolean fields, and users must instead
@@
 distinct values.
-
-<<echo=FALSE,print=FALSE>>=
-if (!exists("JSSPaper.Example1", "RProtoBuf:DescriptorPool")) {
-    readProtoFiles(file="int64.proto")
-}
-@
-
-% We want a cleaner error message here.
+%
 \begin{CodeChunk}
@@
 stored as doubles in \proglang{R}.
-
-<<>>=
-as.integer(2^31-1)
-as.integer(2^31 - 1) + as.integer(1)
-2^31
-class(2^31)
-@
+%
+\begin{Schunk}
+\begin{Sinput}
+R> as.integer(2^31 - 1)
+\end{Sinput}
+\begin{Soutput}
+[1] 2147483647
+\end{Soutput}
+\begin{Sinput}
+R> as.integer(2^31 - 1) + as.integer(1)
+\end{Sinput}
+\begin{Soutput}
+[1] NA
+\end{Soutput}
+\begin{Sinput}
+R> 2^31
+\end{Sinput}
+\begin{Soutput}
+[1] 2.147e+09
+\end{Soutput}
+\begin{Sinput}
+R> class(2^31)
+\end{Sinput}
+\begin{Soutput}
+[1] "numeric"
+\end{Soutput}
+\end{Schunk}
 
@@
 distinct integer variables:
-
-<<>>=
-2^53 == (2^53 + 1)
-@
-
+%
+\begin{Schunk}
+\begin{Sinput}
+R> 2^53 == (2^53 + 1)
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\end{Schunk}
+%
 Most modern languages do have support for 64-bit integer values, 
@@
 
-% test <- new(JSSPaper.Example1)
-% test$repeated_int64 <- c(2^53, 2^53+1)
-% length(unique(test$repeated_int64))
-
-% test$repeated_int64 <- c("9007199254740992", "9007199254740993")
-
-% options("RProtoBuf.int64AsString" = FALSE)
-% test$repeated_int64
-% length(unique(test$repeated_int64))
-% options("RProtoBuf.int64AsString" = TRUE)
-% test$repeated_int64
-% length(unique(test$repeated_int64))
-
-% options("RProtoBuf.int64AsString" = FALSE)
-
-\section[Converting R data structures into Protocol Buffers]{Converting \proglang{R} data structures into Protocol Buffers}
+\section[Converting R data structures into Protocol
+Buffers]{Converting \proglang{R} data structures into Protocol
+  Buffers}
 \label{sec:evaluation}
@@
 message:
-
-<<>>=
-msg <- serialize_pb(iris, NULL)
-identical(iris, unserialize_pb(msg))
-@
-
-In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all \code{proto}
-schema used by \pkg{RHIPE} for exchanging \proglang{R} data with Hadoop \citep{rhipe}. This
-schema, which we will refer to as \code{rexp.proto}, is printed in
-%appendix \ref{rexp.proto}.
-the appendix.
-The Protocol Buffer messages generated by \pkg{RProtoBuf} and
-\pkg{RHIPE} are naturally compatible between the two systems because they use the
-same schema. This shows the power of using a schema-based cross-platform format such
-as Protocol Buffers: interoperability is achieved without effort or close coordination.
+%
+\begin{Schunk}
+\begin{Sinput}
+R> msg <- serialize_pb(iris, NULL)
+R> identical(iris, unserialize_pb(msg))
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\end{Schunk}
+%
+In order to accomplish this, \pkg{RProtoBuf} uses the same catch-all
+\code{proto} schema used by \pkg{RHIPE} for exchanging \proglang{R}
+data with \pkg{Hadoop} \citep{rhipe}. This schema, which we will refer to as
+\code{rexp.proto}, is printed in Appendix~\ref{rexp.proto}.  The
+Protocol Buffer messages generated by \pkg{RProtoBuf} and \pkg{RHIPE}
+are naturally compatible between the two systems because they use the
+same schema. This shows the power of using a schema-based
+cross-platform format such as Protocol Buffers: Interoperability is
+achieved without effort or close coordination.
 
@@
 
-<<echo=FALSE>>=
-datasets <- as.data.frame(data(package="datasets")$results)
-datasets$name <- sub("\\s+.*$", "", datasets$Item)
-n <- nrow(datasets)
-@
-
 This section evaluates the effectiveness of serializing arbitrary
 \proglang{R} data structures into Protocol Buffers.  We use the
-\Sexpr{n} standard data sets included in the \pkg{datasets} package
+104 standard data sets included in the \pkg{datasets} package
 included with \proglang{R} as our evaluation data. These data sets
@@
 bytes are used to store the data set using four different methods:
-
+%
 \begin{itemize}
@@
 \end{itemize}
-
-<<echo=FALSE,print=FALSE>>=
-datasets$object.size <- unname(sapply(datasets$name, function(x) object.size(eval(as.name(x)))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize(eval(as.name(x)), NULL))))
-
-datasets$R.serialize.size.gz <- unname(sapply(datasets$name, function(x) length(memCompress(serialize(eval(as.name(x)), NULL), "gzip"))))
-
-datasets$RProtoBuf.serialize.size <- unname(sapply(datasets$name, function(x) length(serialize_pb(eval(as.name(x)), NULL))))
-
-datasets$RProtoBuf.serialize.size.gz <- unname(sapply(datasets$name, function(x) length(memCompress(serialize_pb(eval(as.name(x)), NULL), "gzip"))))
-
-clean.df <- data.frame(dataset=datasets$name,
-                       object.size=datasets$object.size,
-                       "serialized"=datasets$R.serialize.size,
-                       "gzipped serialized"=datasets$R.serialize.size.gz,
-                       "RProtoBuf"=datasets$RProtoBuf.serialize.size,
-                       "gzipped RProtoBuf"=datasets$RProtoBuf.serialize.size.gz,
-		       "ratio.serialized" = datasets$R.serialize.size / datasets$object.size,
-		       "ratio.rprotobuf" = datasets$RProtoBuf.serialize.size / datasets$object.size,
-		       "ratio.serialized.gz" = datasets$R.serialize.size.gz / datasets$object.size,
-		       "ratio.rprotobuf.gz" = datasets$RProtoBuf.serialize.size.gz / datasets$object.size,
-		       "savings.serialized" = 1-(datasets$R.serialize.size / datasets$object.size),
-		       "savings.rprotobuf" = 1-(datasets$RProtoBuf.serialize.size / datasets$object.size),
-		       "savings.serialized.gz" = 1-(datasets$R.serialize.size.gz / datasets$object.size),
-		       "savings.rprotobuf.gz" = 1-(datasets$RProtoBuf.serialize.size.gz / datasets$object.size),
-                       check.names=FALSE)
-
-all.df<-data.frame(dataset="TOTAL", object.size=sum(datasets$object.size),
-				    "serialized"=sum(datasets$R.serialize.size),
-                       "gzipped serialized"=sum(datasets$R.serialize.size.gz),
-                       "RProtoBuf"=sum(datasets$RProtoBuf.serialize.size),
-                       "gzipped RProtoBuf"=sum(datasets$RProtoBuf.serialize.size.gz),
-		       "ratio.serialized" = sum(datasets$R.serialize.size) / sum(datasets$object.size),
-		       "ratio.rprotobuf" = sum(datasets$RProtoBuf.serialize.size) / sum(datasets$object.size),
-		       "ratio.serialized.gz" = sum(datasets$R.serialize.size.gz) / sum(datasets$object.size),
-		       "ratio.rprotobuf.gz" = sum(datasets$RProtoBuf.serialize.size.gz) / sum(datasets$object.size),
-		       "savings.serialized" = 1-(sum(datasets$R.serialize.size) / sum(datasets$object.size)),
-		       "savings.rprotobuf" = 1-(sum(datasets$RProtoBuf.serialize.size) / sum(datasets$object.size)),
-		       "savings.serialized.gz" = 1-(sum(datasets$R.serialize.size.gz) / sum(datasets$object.size)),
-		       "savings.rprotobuf.gz" = 1-(sum(datasets$RProtoBuf.serialize.size.gz) / sum(datasets$object.size)),
-                       check.names=FALSE)
-clean.df<-rbind(clean.df, all.df)
-@
-
+%
 Figure~\ref{fig:compression} shows the space savings $\left(1 - \frac{\textrm{Compressed Size}}{\textrm{Uncompressed Size}}\right)$ for each of the data sets using each of these four methods.
-The associated table shows the exact data sizes for some outliers and the aggregate of all \Sexpr{n} data sets.
+The associated table shows the exact data sizes for some outliers and the aggregate of all 104 data sets.
 Note that Protocol Buffer serialization results in slightly
@@
 but this difference disappears if the results are compressed with gzip (blue triangles).
-%Sizes are comparable but Protocol Buffers provide simple getters and setters
-%in multiple languages instead of requiring other programs to parse the \proglang{R}
-%serialization format. % \citep{serialization}.
+% Sizes are comparable but Protocol Buffers provide simple getters and
+% setters in multiple languages instead of requiring other programs to
+% parse the \proglang{R} serialization
+% format. % \citep{serialization}.
 
 The \code{crimtab} dataset of anthropometry measurements of British
-prisoners \citep{garson1900metric} and the \code{airquality} dataset of air quality measurements in New York show the
-greatest difference in the space savings when
-using Protocol Buffers compared to \proglang{R} native serialization.
-The \code{crimtab} dataset is a 42x22 table of integers, most equal to 0, and the \code{airquality} dataset is a data frame of 154 observations of 1 numeric and 5 integer variables.
-In both data sets, the large number of small
-integer values can be very efficiently encoded by the
+prisoners \citep{garson1900metric} and the \code{airquality} dataset
+of air quality measurements in New York show the greatest difference
+in the space savings when using Protocol Buffers compared to
+\proglang{R} native serialization.  The \code{crimtab} dataset is a 42
+$\times$ 22 table of integers, most equal to 0, and the
+\code{airquality} dataset is a data frame of 154 observations of 1
+numeric and 5 integer variables.  In both data sets, the large number
+of small integer values can be very efficiently encoded by the
 \emph{Varint} integer encoding scheme used by Protocol Buffers which
@@
 
-\begin{figure}[hbt!]
-\begin{center}
-<<label=SER,echo=FALSE,include=FALSE,fig=TRUE>>=
-old.mar<-par("mar")
-new.mar<-old.mar
-new.mar[3]<-0
-new.mar[4]<-0
-my.cex<-1.3
-par("mar"=new.mar)
-plot(clean.df$savings.serialized, clean.df$savings.rprotobuf, pch=1, col="red", las=1, xlab="Serialization Space Savings", ylab="Protocol Buffer Space Savings", xlim=c(0,1),ylim=c(0,1),cex.lab=my.cex, cex.axis=my.cex)
-points(clean.df$savings.serialized.gz, clean.df$savings.rprotobuf.gz,pch=2, col="blue")
-# grey dotted diagonal
-abline(a=0,b=1, col="grey",lty=2,lwd=3)
-
-# find point furthest off the X axis.
-clean.df$savings.diff <- clean.df$savings.serialized - clean.df$savings.rprotobuf
-clean.df$savings.diff.gz <- clean.df$savings.serialized.gz - clean.df$savings.rprotobuf.gz
-
-# The one to label.
-tmp.df <- clean.df[which(clean.df$savings.diff == min(clean.df$savings.diff)),]
-# This minimum means most to the left of our line, so pos=2 is label to the left
-text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)
-
-# Some gziped version
-# text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=2, cex=my.cex)
-
-# Second one is also an outlier
-tmp.df <- clean.df[which(clean.df$savings.diff == sort(clean.df$savings.diff)[2]),]
-# This minimum means most to the left of our line, so pos=2 is label to the left
-text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=2, cex=my.cex)
-#text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=my.cex)
-
-tmp.df <- clean.df[which(clean.df$savings.diff == max(clean.df$savings.diff)),]
-# This minimum means most to the right of the diagonal, so pos=4 is label to the right
-# Only show the gziped one.
-#text(tmp.df$savings.serialized, tmp.df$savings.rprotobuf, labels=tmp.df$dataset, pos=4, cex=my.cex)
-text(tmp.df$savings.serialized.gz, tmp.df$savings.rprotobuf.gz, labels=tmp.df$dataset, pos=4, cex=my.cex)
-
-#outlier.dfs <- clean.df[c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
-
-legend("topleft", c("Raw", "Gzip Compressed"), pch=1:2, col=c("red", "blue"), cex=my.cex)
-
-interesting.df <- clean.df[unique(c(which(clean.df$savings.diff == min(clean.df$savings.diff)),
-                             which(clean.df$savings.diff == max(clean.df$savings.diff)),
-                             which(clean.df$savings.diff.gz == max(clean.df$savings.diff.gz)),
-			     which(clean.df$dataset == "TOTAL"))),c("dataset", "object.size", "serialized", "gzipped serialized", "RProtoBuf", "gzipped RProtoBuf", "savings.serialized", "savings.serialized.gz", "savings.rprotobuf", "savings.rprotobuf.gz")]
-# Print without .00 in xtable
-interesting.df$object.size <- as.integer(interesting.df$object.size)
-par("mar"=old.mar)
-@
-\includegraphics[width=0.45\textwidth]{figures/fig-SER}
-
-% latex table generated in R 3.0.2 by xtable 1.7-0 package
-% Wed Nov 26 15:31:30 2014
-%\begin{table}[ht]
-%\begin{center}
-\begin{tabular}{rlrrrrr}
-  \toprule
-  Data set & object.size & \multicolumn{2}{c}{\proglang{R} serialization} &
-  \multicolumn{2}{c}{RProtoBuf serialization} \\
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.45\textwidth]{fig-SER}
+
+\vspace*{0.3cm}
+
+\small
+\begin{tabular}{rrrrrrr}
+  \hline
+
+  Data set & \code{object.size} & \multicolumn{2}{c}{\proglang{R} serialization} &
+  \multicolumn{2}{c}{\pkg{RProtoBuf} serialization} \\
   & & default & gzipped & default & gzipped \\
-  \cmidrule(r){2-6}
- crimtab & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
- airquality & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
- faithful & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
+  \hline
+ \code{crimtab} & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
+ \code{airquality} & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
+ \code{faithful} & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
    \hline
@@
 \end{tabular}
-\end{center}
-\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the \Sexpr{n} data sets in the \pkg{datasets} package.
+\caption{(Top) Relative space savings of Protocol Buffers and native \proglang{R} serialization over the raw object sizes of each of the 104 data sets in the \pkg{datasets} package.
 Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
@@
 Figure~\ref{fig:mr-histogram-pattern1}, each mapper rounds a data
-point to a bucket width and outputs that bucket as a key and '1' as a
+point to a bucket width and outputs that bucket as a key and ``1'' as a
 value.  Reducers count how many times each key occurs and outputs a
@@
 
-\begin{figure}[h!]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{figures/histogram-mapreduce-diag1.pdf}
-\end{center}
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.9\textwidth]{histogram-mapreduce-diag1.pdf}
 \caption{Diagram of MapReduce histogram generation pattern.}
@@
 The \pkg{HistogramTools} package \citep{histogramtools} enhances
-\pkg{RProtoBuf} by providing a concise schema for \proglang{R} histogram objects:
-
+\pkg{RProtoBuf} by providing a concise schema for \proglang{R}
+histogram objects:
+%
 \begin{example}
@@
 \end{example}
-
-This HistogramState message type is designed to be helpful if some of
-the Map or Reduce tasks are written in \proglang{R}, or if those components are
-written in other languages and only the resulting output histograms
-need to be manipulated in \proglang{R}.
-
-\subsection*{A simple single-machine example for Python to R serialization}
-
-To create HistogramState
-messages in Python for later consumption by \proglang{R}, we first compile the 
+%
+This \code{HistogramState} message type is designed to be helpful if
+some of the Map or Reduce tasks are written in \proglang{R}, or if
+those components are written in other languages and only the resulting
+output histograms need to be manipulated in \proglang{R}.
+
+\subsection[A simple single-machine example for Python to R
+serialization]{A simple single-machine example for \proglang{Python}
+  to \proglang{R} serialization}
+
+To create \code{HistogramState}
+messages in \proglang{Python} for later consumption by \proglang{R}, we first compile the 
 \code{histogram.proto} descriptor into a python module using the
 \code{protoc} compiler:
-
-\begin{verbatim}
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 protoc histogram.proto --python_out=.
-\end{verbatim}
-This generates a Python module called \code{histogram\_pb2.py}, containing both the 
+\end{CodeInput}
+\end{CodeChunk}
+%
+This generates a \proglang{Python} module called \code{histogram\_pb2.py}, containing both the 
 descriptor information as well as methods to read and manipulate the histogram 
-message data.  The following simple Python script uses this generated
+message data.  The following simple \proglang{Python} script uses this generated
 module to create a histogram (to which breakpoints and binned data are
@@
 representation to a file:
-
+%
 \begin{Code}
@@
 hist.breaks.extend(range(6))
-hist.name="Example Histogram Created in Python"
+hist.name = "Example Histogram Created in Python"
 outfile = open("/tmp/hist.pb", "wb")
@@
 \end{Code}
-
-The Protocol Buffer created from this \proglang{Python} script can then be read into \proglang{R} and converted to a native
-\proglang{R} histogram object for plotting.
-Line~1 in the listing below attaches the \pkg{HistogramTools} package which imports \pkg{RProtoBuf}.
-Line~2 then reads all of the \code{.proto} descriptor definitions provided by \pkg{HistogramTools} and adds them to the environment as described in Section~\ref{sec:rprotobuf-basic}.
-Line~3 parses the serialized protocol buffer using the \code{HistogramTools.HistogramState} schema.
-The last line converts the protocol buffer representation of the histogram to a native \proglang{R} histogram object with \code{as.histogram} and passes the result to \code{plot}.
-
-% Here, the schema is read first,
-%then the (serialized) histogram is read into the variable \code{hist} which
-%is then converted a histogram object which is display as a plot.
-
-\lstdefinelanguage{jss}
-  {sensitive=false,
-   morecomment=[l]{R>}}
-
-\lstset{language=jss, basicstyle=\ttfamily, numbers=left, numberstyle=\tiny, stepnumber=2, numbersep=5pt, columns=fullflexible, keepspaces=true, showstringspaces=false, commentstyle=\textsl}
-%\begin{Code}
-\begin{lstlisting}
+%
+The Protocol Buffer created from this \proglang{Python} script can
+then be read into \proglang{R} and converted to a native \proglang{R}
+histogram object for plotting. The code below first attaches the
+\pkg{HistogramTools} package which imports \pkg{RProtoBuf}.  Then
+reads all of the \code{.proto} descriptor definitions provided by
+\pkg{HistogramTools} and adds them to the environment as described in
+Section~\ref{sec:rprotobuf-basic}.  Next the serialized Protocol
+Buffer is parsed using the \code{HistogramTools.HistogramState}
+schema.  Finally the Protocol Buffer representation of the histogram
+is converted to a native \proglang{R} histogram object with
+\code{as.histogram} and passes the result to \code{plot} (see
+Figure~\ref{figure}).
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 R> library("HistogramTools")
-R> readProtoFiles(package="HistogramTools")
+R> readProtoFiles(package = "HistogramTools")
 R> hist <- HistogramTools.HistogramState$read("/tmp/hist.pb")
 R> hist
-
+\end{CodeInput}
+\begin{CodeOutput}
 [1] "message of type 'HistogramTools.HistogramState' with 3 fields set"
-
-R> plot(as.histogram(hist), main="")
-\end{lstlisting}
-%\end{Code}
-
-\begin{center}
-<<echo=FALSE,fig=TRUE,width=8,height=4>>=
-require(HistogramTools)
-readProtoFiles(package="HistogramTools")
-hist <- HistogramTools.HistogramState$read("hist.pb")
-plot(as.histogram(hist), main="")
-@
-\end{center}
+\end{CodeOutput}
+\begin{CodeInput}
+R> plot(as.histogram(hist), main = "")
+\end{CodeInput}
+\end{CodeChunk}
+%
+\begin{figure}[t!]
+  \centering
+  \includegraphics{fig-021}
+  \caption{Histogram converted to an \proglang{R} object from \proglang{Python}.}\label{figure}
+\end{figure}
 
@@
 particular MapReduce implementation.  In practice, using Protocol
-Buffers to pass histograms between another programming language and R
-would provide a much greater benefit in a distributed context.
-For example, a first-class data type to represent histograms would
-prevent individual histograms from being split up and would allow the
-use of combiners on Map workers to process large data sets more
-efficiently than simply passing around lists of counts and buckets.
+Buffers to pass histograms between another programming language and
+\proglang{R} would provide a much greater benefit in a distributed
+context.  For example, a first-class data type to represent histograms
+would prevent individual histograms from being split up and would
+allow the use of combiners on Map workers to process large data sets
+more efficiently than simply passing around lists of counts and
+buckets.
 
@@
 initiate requests, authenticate, send messages, etc.  However, network
-protocols generally do not regulate the \emph{content} of messages: they
+protocols generally do not regulate the \emph{content} of messages: They
 allow transfer of any media type, such as web pages, static files or
@@
 protocol.
-%The descriptors (\code{.proto} files) are used to formally define
-%the interface of a remote API or network application.
-%Libraries to parse and
-%generate protobuf messages are available for many programming languages,
-%making it relatively straightforward to implement clients and servers.
 
@@
 One example of a system that supports Protocol Buffers to interact
-with \proglang{R} is OpenCPU \citep{opencpu}. OpenCPU is a framework
+with \proglang{R} is \pkg{OpenCPU} \citep{opencpu}. \pkg{OpenCPU} is a framework
 for embedded statistical computation and reproducible research based
@@
 calls, and arguments/return values can be posted/retrieved using
-several data interchange formats, such as Protocol Buffers.  OpenCPU
+several data interchange formats, such as Protocol Buffers.  \pkg{OpenCPU}
 uses the \code{rexp.proto} descriptor and the \code{serialize\_pb} and
@@
 
-\subsection[HTTP GET: Retrieving an R object]{HTTP GET: Retrieving an \proglang{R} object}
-
-The \code{HTTP GET} method is used to read a resource from OpenCPU. For example,
-to access the data set \code{Animals} from the package \code{MASS}, a
-client performs the following HTTP request:
-
-\begin{verbatim}
+\subsection[HTTP GET: Retrieving an R object]{\code{HTTP GET}: Retrieving an \proglang{R} object}
+
+The \code{HTTP GET} method is used to read a resource from
+\pkg{OpenCPU}. For example, to access the data set \code{Animals} from the
+package \pkg{MASS} \citep{MASS}, a client performs the following HTTP
+request:
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 GET https://demo.ocpu.io/MASS/data/Animals/pb
-\end{verbatim}
+\end{CodeInput}
+\end{CodeChunk}
+%
 The postfix \code{/pb} in the URL tells the server to send this
@@
 If the request
-is successful, OpenCPU returns the serialized object with HTTP status
+is successful, \pkg{OpenCPU} returns the serialized object with HTTP status
 code 200 and HTTP response header \code{Content-Type: application/x-protobuf}.
@@
 
-Because both HTTP and Protocol Buffers have libraries available for many
-languages, clients can be implemented in just a few lines of code. Below
-is example code for both \proglang{R} and Python that retrieves an \proglang{R} data set encoded as a Protocol Buffer message from OpenCPU.
-In \proglang{R}, we use the HTTP client from
-the \code{httr} package \citep{httr}. In this example we
-download a data set which is part of the base \proglang{R} distribution, so we can
-verify that the object was transferred without loss of information.
-
-<<eval=FALSE>>=
-library("RProtoBuf")
-library("httr")
-
-req <- GET('https://demo.ocpu.io/MASS/data/Animals/pb')
-output <- unserialize_pb(req$content)
-
-identical(output, MASS::Animals)
-@
-
-Similarly, to retrieve the same data set in a Python client, we first
+Because both HTTP and Protocol Buffers have libraries available for
+many languages, clients can be implemented in just a few lines of
+code. Below is example code for both \proglang{R} and
+\proglang{Python} that retrieves an \proglang{R} data set encoded as a
+Protocol Buffer message from \pkg{OpenCPU}.  In \proglang{R}, we use the
+HTTP client from the \pkg{httr} package \citep{httr}. In this example
+we download a data set which is part of the base \proglang{R}
+distribution, so we can verify that the object was transferred without
+loss of information.
+%
+\begin{Schunk}
+\begin{Sinput}
+R> library("RProtoBuf")
+R> library("httr")
+R> req <- GET("https://demo.ocpu.io/MASS/data/Animals/pb")
+R> output <- unserialize_pb(req$content)
+R> identical(output, MASS::Animals)
+\end{Sinput}
+\end{Schunk}
+%
+Similarly, to retrieve the same data set in a \proglang{Python} client, we first
 compile the \code{rexp.proto} descriptor into a python module
 using the \code{protoc} compiler:
-
-\begin{verbatim}
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 protoc rexp.proto --python_out=.
-\end{verbatim}
-
-This generates Python module called \code{rexp\_pb2.py}, containing
+\end{CodeInput}
+\end{CodeChunk}
+%
+This generates \proglang{Python} module called \code{rexp\_pb2.py}, containing
 both the descriptor information as well as methods to read and
@@
 encoded Protocol Buffer from the remote server then parse and print it
-from Python.
-
-\begin{verbatim}
+from \proglang{Python}.
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 import urllib2
 from rexp_pb2 import REXP
-req = urllib2.Request('https://demo.ocpu.io/MASS/data/Animals/pb')
+req = urllib2.Request("https://demo.ocpu.io/MASS/data/Animals/pb")
 res = urllib2.urlopen(req)
@@
 print(msg)
-\end{verbatim}
-The \code{msg} object contains all data from the Animals data set. From here we
-can easily extract the desired fields for further use in Python.
-
-\subsection[HTTP POST: Calling an R function]{HTTP POST: Calling an \proglang{R} function}
+\end{CodeInput}
+\end{CodeChunk}
+%
+The \code{msg} object contains all data from the \code{Animals} data
+set. From here we can easily extract the desired fields for further
+use in \proglang{Python}.
+
+\subsection[HTTP POST: Calling an R function]{\code{HTTP POST}: Calling an \proglang{R} function}
 
 The previous example used a simple \code{HTTP GET} method to retrieve
-an \proglang{R} object from a remote service (OpenCPU) encoded as a
+an \proglang{R} object from a remote service (\pkg{OpenCPU}) encoded as a
 Protocol Buffer.
@@
 
-The OpenCPU framework allows us to do arbitrary \proglang{R} function
+The \pkg{OpenCPU} framework allows us to do arbitrary \proglang{R} function
 calls from within any programming language by encoding the arguments
 in the request Protocol Buffer.  The following example \proglang{R}
-client code performs the remote function call \code{stats::rnorm(n=42,
-mean=100)}. The function arguments (in this case \code{n} and
+client code performs the remote function call \code{stats::rnorm(n =
+  42, mean = 100)}. The function arguments (in this case \code{n} and
 \code{mean}) as well as the return value (a vector with 42 random
-numbers) are transferred using Protocol Buffer messages. RPC in OpenCPU
-works like the \code{do.call} function in \proglang{R}, hence all
-arguments are contained within a list.
-
-<<eval=FALSE>>=
-library("httr")
-library("RProtoBuf")
-
-args <- list(n=42, mean=100)
-payload <- serialize_pb(args, NULL)
-
-req <- POST (
-  url = "https://demo.ocpu.io/stats/R/rnorm/pb",
-  body = payload,
-  add_headers (
-    "Content-Type" = "application/x-protobuf"
-  )
-)
-
-output <- unserialize_pb(req$content)
-print(output)
-@
-The OpenCPU server basically performs the following steps to process the above RPC request:
-
-<<eval=FALSE>>=
-fnargs <- unserialize_pb(inputmsg)
-val <- do.call(stats::rnorm, fnargs)
-outputmsg <- serialize_pb(val)
-@
+numbers) are transferred using Protocol Buffer messages. RPC in
+\pkg{OpenCPU} works like the \code{do.call} function in \proglang{R}, hence
+all arguments are contained within a list.
+%
+\begin{Schunk}
+\begin{Sinput}
+R> library("httr")
+R> library("RProtoBuf")
+R> args <- list(n = 42, mean = 100)
+R> payload <- serialize_pb(args, NULL)
+R> req <- POST(url = "https://demo.ocpu.io/stats/R/rnorm/pb", body = payload,
++    add_headers("Content-Type" = "application/x-protobuf"))
+R> output <- unserialize_pb(req$content)
+R> print(output)
+\end{Sinput}
+\end{Schunk}
+%
+The \pkg{OpenCPU} server basically performs the following steps to process the above RPC request:
+%
+\begin{Schunk}
+\begin{Sinput}
+R> fnargs <- unserialize_pb(inputmsg)
+R> val <- do.call(stats::rnorm, fnargs)
+R> outputmsg <- serialize_pb(val)
+\end{Sinput}
+\end{Schunk}
 
@@
 Over the past decade, many formats for interoperable
-data exchange have become available, each with their unique features,
+data exchange have become available, each with its unique features,
 strengths and weaknesses.  
-Text based formats such as \code{CSV} and \code{JSON} are easy to use, and will likely 
+Text based formats such as CSV and JSON are easy to use, and will likely 
 remain popular among statisticians for many years to come. However, in the 
@@
 
-The \pkg{RProtoBuf} package builds on the Protocol Buffers \proglang{C++} library, 
-and extends the \proglang{R} system with the ability to create, read,
-write, parse, and manipulate Protocol
-Buffer messages. \pkg{RProtoBuf} has been used extensively inside Google 
-for the past five years by statisticians, analysts, and software engineers.
-At the time of this writing there are over 300 active
-users of \pkg{RProtoBuf} using it to read data from and otherwise interact
-with distributed systems written in \proglang{C++}, \proglang{Java}, \proglang{Python}, and 
-other languages. We hope that making Protocol Buffers available to the
-\proglang{R} community will contribute to better software integration
-and allow for building even more advanced applications and analysis pipelines 
-with \proglang{R}.
+The \pkg{RProtoBuf} package builds on the Protocol Buffers
+\proglang{C++} library, and extends the \proglang{R} system with the
+ability to create, read, write, parse, and manipulate Protocol Buffer
+messages. \pkg{RProtoBuf} has been used extensively inside Google for
+the past five years by statisticians, analysts, and software
+engineers.  At the time of this writing there are over 300 active
+users of \pkg{RProtoBuf} using it to read data from and otherwise
+interact with distributed systems written in \proglang{C++},
+\proglang{Java}, \proglang{Python}, and other languages. We hope that
+making Protocol Buffers available to the \proglang{R} community will
+contribute to better software integration and allow for building even
+more advanced applications and analysis pipelines with \proglang{R}.
 
@@
 
-The first versions of \pkg{RProtoBuf} were written during 2009 - 2010.
+The first versions of \pkg{RProtoBuf} were written during 2009--2010.
 Very significant contributions, both in code and design, were made by
 Romain Fran\c{c}ois whose continued influence on design and code is
-greatly appreciated. Several features of the package reflect
-the design of the \pkg{rJava} package by Simon Urbanek.
-The user-defined table mechanism, implemented by Duncan Temple Lang for the
-purpose of the \pkg{RObjectTables} package, allows for the dynamic symbol lookup.
-Kenton Varda was generous with his time in reviewing code and explaining
-obscure Protocol Buffer semantics.  Karl Millar and Tim Hesterberg were very
-helpful in reviewing code and offering suggestions.  Saptarshi Guha's
-work on RHIPE and implementation of a universal message type for \proglang{R}
-language objects allowed us to add the \code{serialize_pb} and \code{unserialize_pb}
-methods for turning arbitrary R objects into Protocol Buffers without
-a specialized pre-defined schema.  Feedback from two anonymous
+greatly appreciated. Several features of the package reflect the
+design of the \pkg{rJava} package by Simon Urbanek \citep{rjava}.  The
+user-defined table mechanism, implemented by Duncan Temple Lang for
+the purpose of the \pkg{RObjectTables} package, allows for the dynamic
+symbol lookup.  Kenton Varda was generous with his time in reviewing
+code and explaining obscure Protocol Buffer semantics.  Karl Millar
+and Tim Hesterberg were very helpful in reviewing code and offering
+suggestions.  Saptarshi Guha's work on \pkg{RHIPE} and implementation
+of a universal message type for \proglang{R} language objects allowed
+us to add the \code{serialize_pb} and \code{unserialize_pb} methods
+for turning arbitrary \proglang{R} objects into Protocol Buffers
+without a specialized pre-defined schema.  Feedback from two anonymous
 referees greatly improved both the presentation of this paper and the
@@
 
-\newpage
-\appendix
-\setcounter{secnumdepth}{0}
-%\begin{appendices}
-
-\section*{Appendix: The rexp.proto schema descriptor}
-
+\bibliography{v71i02}
+
+\begin{appendix}
+
+\section[The rexp.proto schema descriptor]{The \code{rexp.proto} schema descriptor}
 \label{rexp.proto}
 
-Below a print of the \code{rexp.proto} schema (originally designed by \cite{rhipe})
-that is included with the \pkg{RProtoBuf} package and used by \code{serialize\_pb} and
-\code{unserialize\_pb}.
-
-\begin{verbatim}
+Below a print of the \code{rexp.proto} schema (originally designed by
+\citealt{rhipe}) that is included with the \pkg{RProtoBuf} package and
+used by \code{serialize\_pb} and \code{unserialize\_pb}.
+%
+\begin{CodeChunk}
+\begin{CodeInput}
 package rexp;
@@
   enum RBOOLEAN {
-    F=0;
-    T=1;
-    NA=2;
+    F = 0;
+    T = 1;
+    NA = 2;
   }
   required RClass rclass = 1;
-  repeated double realValue = 2 [packed=true];
-  repeated sint32 intValue = 3 [packed=true];
+  repeated double realValue = 2 [packed = true];
+  repeated sint32 intValue = 3 [packed = true];
   repeated RBOOLEAN booleanValue = 4;
@@
   optional string strval = 1;
-  optional bool isNA = 2 [default=false];
+  optional bool isNA = 2 [default = false];
 }
 message CMPLX {
-  optional double real = 1 [default=0];
+  optional double real = 1 [default = 0];
   required double imag = 2;
 }
-\end{verbatim}
-% \end{appendices}
-\newpage
-\bibliography{article}
+\end{CodeInput}
+\end{CodeChunk}
+
+\end{appendix}
 

### 2016-07-10 papers/jss-finalising/v71i02.tex
finalising: content of emailed zip file
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@952 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

--- 2016-07-10 papers/jss-finalising/v71i02.orig.tex
@@
 \author{Dirk Eddelbuettel\\Debian Project \And 
-        Murray Stokely\\Google, Inc. \And
+        Murray Stokely\\Apple, Inc. \And
         Jeroen Ooms\\University of California,\\Los Angeles}
@@
   Murray Stokely\\
-  Google, Inc.\\
-  1600 Amphitheatre Parkway\\
-  Mountain View, CA, United States of America\\
-  E-mail: \email{mstokely@google.com}\\
+  Apple, Inc.\\
+  1 Infinite Loop\\
+  Cupertino, CA, United States of America\\
+  E-mail: \email{murray@stokely.org}\\
   URL: \url{http://www.stokely.org/}\\
@@
   \hline
- \code{crimtab} & 7,936 & 4,641 (41.5\%) & 713 (91.0\%) & 1,655 (79.2\%) & 576 (92.7\%)\\
- \code{airquality} & 5,496 & 4,551 (17.2\%) & 1,241 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.5\%)\\
- \code{faithful} & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 (3.9\%) & 1,776 (65.5\%)\\
+ \code{crimtab} & 7,936 & 4,641 (41.5\%) & 714 (91.0\%) & 1,655 (79.1\%) & 576 (92.7\%)\\
+ \code{airquality} & 5,496 & 4,551 (17.2\%) & 1,242 (77.4\%) & 2,874 (47.7\%) & 1,294 (76.4\%)\\
+ \code{faithful} & 5,136 & 4,543 (11.5\%) & 1,339 (73.9\%) & 4,936 \phantom{0}(3.9\%) & 1,776 (65.4\%)\\
    \hline
- All & 605,256 & 461,667 (24\%) & 138,937 (77\%) & 435,360 (28\%) & 142,134 (77\%)\\
+ All & 609,024 & 463,833 (24\%) & 139,814 (77\%) & 436,746 (28\%) & 142,783 (77\%)\\
 \hline
@@
 Points to the left of the dashed $y=x$ line represent datasets that are more efficiently encoded with Protocol Buffers.
-(Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.}
+(Bottom) Absolute space savings of three outlier datasets and the aggregate performance of all datasets.
+R version 3.3.1 was used for both the figure and the table. }
 \label{fig:compression}

### 2016-07-10 papers/jss-finalising/v71i02.tex
Summary: Murray at Google not Apple
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@954 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

@@
 \author{Dirk Eddelbuettel\\Debian Project \And 
-        Murray Stokely\\Apple, Inc. \And
+        Murray Stokely\\Google, Inc \And
         Jeroen Ooms\\University of California,\\Los Angeles}
@@
   Murray Stokely\\
-  Apple, Inc.\\
-  1 Infinite Loop\\
-  Cupertino, CA, United States of America\\
+  Google, Inc.\\
+  1600 Amphitheatre Parkway\\
+  Mountain View, CA, United States of America\\
   E-mail: \email{murray@stokely.org}\\
